### 2018-03-26 main.Rmd
init the paper structures

---
title: "Tidy data structures to support exploration and modeling of temporal-context data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: "ATemporal-context data is often rich with information, which is not supported by current time series model objects. The data may have heterogeneous variable types, implicit missings, multiple levels of of temporal structure, many different grouping variables.  The conventional matrix structure that underlies time series models requires numerical variables, and implicit times, which have the effect of discarding some elements of the rich data available. This work addresses the broader issues of better data structures and modern data pipelines for analysing and visualising temporal-context data. We extend the tidy data concept to temporal data, and note that the 'molten' data structure is flexible enough to handle heterogeneity, low time resolutions, and implicit missing values. There are two supporting components required to turn  'molten' data into a valid temporal data: (1) an explicitly declared index variable containing time-stamps; and (2) a unique identifier the multiple groups of measurements, which is referred to as a 'key'. A syntactical approach is introduced to describe nested or crossed data structures, which employ the 'key'. The tidy data structure also supports thinking of operations on the data as part of a data pipeline, which facilitates the workflow for analysis of temporal data. A case study is included to illustrate the tidy structure, the data pipeline ideas and usage. A new representation of tidy temporal data along with the supporting methods has been implemented in the R package **tsibble**."
keywords: "blah, blah"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: no
    number_sections: yes
    citation_package: biblatex
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite"
)
# read_chunk('src/main.R')
```

# Introduction

Data measured or collected at different time points are referred to as temporal-context data or time series. Figure \@ref(fig:data-snapshot) shows a glimpse of on-time performance of domestic flights in U.S.A from 2016 to 2017. It keeps track of on-time records for each flight's at its scheduled time over years. This dataset arrives in a rectangular form, which features:

* heterogeneous data columns, such as numbers, characters and timestamps;
* multiple measured variables including departure/arrival delay in minutes, flight time, distance between origin and destination, and etc.;
* low time resolutions, for example scheduled departure time in minutes;
* multi-levels of grouping like airports nested within cities, and cities further nested within states.

However, there is lack of data management tools for time series involving those characteristics in most statistical software, for example data wrangling, visualisation, analysing and forecasting in time-based context. It becomes more challenging in handling time series as increased data complications. The bottleneck that makes time series data frustrating to work with is the conventional data structure that a vector or matrix underlies.

Although a matrix is a tabular form, it assumes homogeneity (that is, all the columns must be real values.) and time indices implicitly inferred as attributes (or meta-information). This format of data input is model-oriented rather than data-oriented as it proves computationally efficient for many statistical models when doing matrix operations. However, it is has the effect of discarding intersting information in the data, is more opaque in the handling of the temporal components, and less efficient for many other data analysis tasks, such as data transformation, visualisation and model diagnostics, like forcing the fitting of a square peg into a round hole.

In addition, time indices form the basis of temporal data and the source of contextual information to understand temporal dependence. Time as attributes limits the use of time context, since it is difficult to access and manipulate. For example, the components of date-times, such as months, day of week, and time of day, cannot be easily extracted in order to examine seasonal fluctuations. It is also impossible to switch between different time zones and daylight savings to compare time across places when needed.

@wickham2014tidy developed and formulated the conceptual framework of tidy data: (1) each variable forms a column; (2) each observation forms a row; (3) each type of observational unit forms a table. These principles attempt to standardise the mapping from the semantics of a dataset to its structure and facilitate data analysis in a coherent way. We shall extend "tidy data" into the time series domain.

This paper proposes a tidy data structure and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.

# Tidy temporal-context data

A modern reimagining of time series should provide heterogeneous data types and time indices as explicitly declared data column. This can be achieved using a "data frame" in R or other statistical languages to represent a tabular format, instead of "matrix". Tidy temporal-context data at least consists of:

1. index: an explicitly declared data variable contains time indices, such as date-times, year-months, years and etc.
2. key: a set of grouping factors uniquely identifies each unit that measurements take place on over time, which may include single or multiple columns.
3. interval: data with regular time interval results in a common time interval in one table.

In SQL database, a primary key, which uniquely defines each record in a database table, is equivalent to the composition of "index" and "key". In multivariate time series notation as Equation \@ref(eq:mts), $X_{jt}$ represents series $j$, for $j = 1, \dots, p$ and $1 \leq t \leq T$, in the form of a $T \times p$ matrix.

\begin{equation}
  \begin{bmatrix}
  X_{11} & X_{21} & \cdots & X_{p1} \\
  X_{12} & X_{22} & \cdots & X_{p2} \\
  \vdots & \vdots & \ddots & \vdots \\
  X_{1T} & X_{2T} & \cdots & X_{pT}
  \end{bmatrix}
\label{eq:mts}
\end{equation}

## Time index and interval

<!-- note to self: There's a trade-off between data column and attributes. Data column is easily accessible, which gives greater flexibility; however, once the value in the data column is modified, no way to validate for any static approach. Generally, not recommended to modify attributes. -->

Time index forms an integral component and contextual base of temporal data. In temporal data frame, time-based index must be clearly stated as a data column rather than inferred as attributes, and thus can be accessible. This creates flexibility in handling time indices.

(1) temporal elements can be created, and then exploratory data visualisation and analysis (not specialist plotting) and multiple seasonality modelling for sub-daily data
(2) convert to the same time zone and thus compare
(3) join other data tables using the index as common key

For data indexed in regular time space, the time interval is obtained by computing the minimal positive time distance in a data table. This suggests that each observational unit collected at the same interval forms a table.

## Keys

Key variables are usually discrete descriptors, and are typically variables that were created during the data collection to uniquely define the measured values. For instance, to distinguish the performance of each flight in the dataset, the "key" is the `flight` variable, allowing separation of multiple time series in one data table. The "key" not only identifies the unit to be measured over time, but also incorporates structures of data. Without a key, a data table can be considered as a univariate time series; in other words, the key is implicit. With a single key of more than one categories, it lends itself to a collection of time series in a table. But when there are at least two keys in the table, it indicates nested or crossed structures.

In experimental designs, a variable is crossed with another when every category of one variable co-occurs with every category of the other, while a variable is nested within another when each category of the former variable co-occurs with only one category of the latter. Regarding faceted plots in statistical graphics, @wilkinson2006grammar discussed the difference between nesting and crossing structure in depth. The distinction between two graphical layouts may appear subtle but convey completely different meanings of the data. The crossing example could be gender with marital status: married women, married men, single women, and single men. It takes all the possible combinations into account, resulting in 2 by 2 drawing panels in a graphical device. If a panel goes empty, there exists no observation for that category in the dataset, but it would be filled with observations. However, considering pregnancy status and gender, there are only three possible combinations: pregnant women, non-pregnant women, and non-pregnant men, since pregnant status is nested under the gender variable. In this case, any physical layout of such nesting structure must not occur to an impossible category.

It appears more useful in making this distinction in statistical analysis including visualisation and modeling, compared to data manipulation.

# Data pipeline

Tidy data builds a concrete foundation to enable pipeline data analysis, which provides a coherent and fluent framework to work with data. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.

* **row-wise**: `filter()`, `slice()`, `arrange()`, `fill_na()`
* **column-wise**: `mutate()`, `select()`, `summarise()`, `tsummarise()`
* **rolling window**: `slide()`, `tile()`, `stretch()`
* **statistics**: `lag()`, `diff()`, `acf()`

# Application: U.S.A domestic flights on-time performance (2016-2017)

A dataset of on-time performance of domestic flights in U.S.A from 2016 to 2017 is studied and explored for illustration of tidy data and data pipeline.

# Discussion

A tidy representation of time series data, and data pipelines to facilitate data analysis flow have been proposed and discussed. It can be noted that tidy temporal data gains greater flexibility in keeping data richness, making data transformation and visualisation easily. A set of verbs provides a fluent and fluid pipeline to work with tidy time series data in various ways.

The ground of time series modelling or forecasting is left untouched in this paper. The future plan is to bridge the gap between tidy data and model building. Currently, it is required to casting to matrix from tidy data and therefore building a model. But time series models should be directly applied to tidy data as other wrangling tools do, without such an intermediate step. In particular, a univariate time series model, like arima and exponential smoothing, can be applied to multiple time series independently. A tidy format to represent model summaries and forecasting objects will be developed and implemented in the future. Model summaries include coefficients, fitted values, and residuals; forecasting objects include future time path and distributions generating prediction intervals.

### 2018-04-18 main.Rmd
re-worked on abstract

@@
     email : rob.hyndman@monash.edu
-abstract: "ATemporal-context data is often rich with information, which is not supported by current time series model objects.
-The data may have heterogeneous variable types, implicit missings, multiple levels of of temporal structure, many different grouping variables.
-The conventional matrix structure that underlies time series models requires numerical variables, and implicit times, which have the effect of discarding some elements of the rich data available.
-This work addresses the broader issues of better data structures and modern data pipelines for analysing and visualising temporal-context data.
-We extend the tidy data concept to temporal data, and note that the 'molten' data structure is flexible enough to handle heterogeneity, low time resolutions, and implicit missing values.
-There are two supporting components required to turn  'molten' data into a valid temporal data: (1) an explicitly declared index variable containing time-stamps; and (2) a unique identifier the multiple groups of measurements, which is referred to as a 'key'.
-A syntactical approach is introduced to describe nested or crossed data structures, which employ the 'key'.
-The tidy data structure also supports thinking of operations on the data as part of a data pipeline, which facilitates the workflow for analysis of temporal data.
-A case study is included to illustrate the tidy structure, the data pipeline ideas and usage.
-A new representation of tidy temporal data along with the supporting methods has been implemented in the R package **tsibble**."
-keywords: "blah, blah"
+abstract: 'Temporal-context data is often rich with information and time formats, for example multiple units, fine time resolutions, heterogeneous data types, nested and crossed structures and etc.
+This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and modelling routines.
+Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared variable representing time index, (3) single or multiple identifiers separating multiple units over time, which are referred to as a "key".
+A syntax is proposed to describe nested or crossed data structures, which employ the "key".
+The tidy temporal data structure most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline".
+A sound pipeline practice facilitates a transparent and human-readable workflow for analyzing temporal data.
+Applications are included to illustrate the tidy temporal data, data pipeline ideas and usage.
+A new representation of tidy temporal data along with the supporting methods has been implemented in the R package **tsibble**.'
+keywords: "temporal context, time series, data structure, R"
 wpnumber: no/yr
@@
 
-Data measured or collected at different time points are referred to as temporal-context data or time series.
-Figure \@ref(fig:data-snapshot) shows a glimpse of on-time performance of domestic flights in U.S.A from 2016 to 2017.
-It keeps track of on-time records for each flight's at its scheduled time over years.
-This dataset arrives in a rectangular form, which features:
+Data measured or collected at different time points are referred to as temporal-context data or time series. This dataset arrives in a rectangular form, which features:
 
@@
 This format of data input is model-oriented rather than data-oriented as it proves computationally efficient for many statistical models when doing matrix operations.
-However, it is has the effect of discarding intersting information in the data, is more opaque in the handling of the temporal components, and less efficient for many other data analysis tasks, such as data transformation, visualisation and model diagnostics, like forcing the fitting of a square peg into a round hole.
+However, it is has the effect of discarding interesting information in the data, is more opaque in the handling of the temporal components, and less efficient for many other data analysis tasks, such as data transformation, visualisation and model diagnostics, like forcing the fitting of a square peg into a round hole.
 
@@
 
-A modern reimagining of time series should provide heterogeneous data types and time indices as explicitly declared data column.
+A modern re-imagining of time series should provide heterogeneous data types and time indices as explicitly declared data column.
 This can be achieved using a "data frame" in R or other statistical languages to represent a tabular format, instead of "matrix".
@@
 
-In SQL database, a primary key, which uniquely defines each record in a database table, is equivalent to the composition of "index" and "key".
+In SQL database, a primary key [@codd_relational_1970], which uniquely defines each record in a database table, is equivalent to the composition of "index" and "key".
 In multivariate time series notation as Equation \@ref(eq:mts), $X_{jt}$ represents series $j$, for $j = 1, \dots, p$ and $1 \leq t \leq T$, in the form of a $T \times p$ matrix.

### 2018-04-23 main.Rmd
opening paragraph in intro

@@
 ---
-title: "Tidy data structures to support exploration and modeling of temporal-context data"
+title: "Tidy data structure to support exploration and modeling of temporal-context data"
 author:
@@
     email : rob.hyndman@monash.edu
-abstract: 'Temporal-context data is often rich with information and time formats, for example multiple units, fine time resolutions, heterogeneous data types, nested and crossed structures and etc.
+abstract: 'Temporal-context data is often rich with information and time formats, for example multiple observational units, different time lengths, heterogeneous data types, nested and crossed factors and etc.
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and modelling routines.
-Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared variable representing time index, (3) single or multiple identifiers separating multiple units over time, which are referred to as a "key".
-A syntax is proposed to describe nested or crossed data structures, which employ the "key".
-The tidy temporal data structure most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline".
+Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntatical and user-oriented approach in which it imposes nested or crossed structures on the data.
+This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
 A sound pipeline practice facilitates a transparent and human-readable workflow for analyzing temporal data.
-Applications are included to illustrate the tidy temporal data, data pipeline ideas and usage.
-A new representation of tidy temporal data along with the supporting methods has been implemented in the R package **tsibble**.'
+Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage.
+The tidy temporal data along with its supporting methods has been implemented in the R package **tsibble**.'
 keywords: "temporal context, time series, data structure, R"
@@
 
-Data measured or collected at different time points are referred to as temporal-context data or time series. This dataset arrives in a rectangular form, which features:
-
-* heterogeneous data columns, such as numbers, characters and timestamps;
-* multiple measured variables including departure/arrival delay in minutes, flight time, distance between origin and destination, and etc.;
-* low time resolutions, for example scheduled departure time in minutes;
-* multi-levels of grouping like airports nested within cities, and cities further nested within states.
-
-However, there is lack of data management tools for time series involving those characteristics in most statistical software, for example data wrangling, visualisation, analysing and forecasting in time-based context.
-It becomes more challenging in handling time series as increased data complications.
-The bottleneck that makes time series data frustrating to work with is the conventional data structure that a vector or matrix underlies.
+Temporal-context data consist of observational units indexed at different time points $X_{jt}$, where the $j$\textsuperscript{th} unit takes measurements of $X$ on over time $t$, for $j = 1, \cdots, N$ and $1 \le t \le T$.
+Time primarily forms the contextual basis of temporal data, but it could arrive in many possible formats.
+For example, data recorded at fine time resolutions (hours, minutes, and seconds) are typically associated with different time zones and daylight savings.
+Temporal data often carries with rich information other than the time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping variables exhibiting nested or crossed structures, linking to other data tables, and etc.
 
@@
 
+This paper proposes a tidy data structure and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.
+
+# Data semantics
+
 @wickham2014tidy developed and formulated the conceptual framework of tidy data: (1) each variable forms a column; (2) each observation forms a row; (3) each type of observational unit forms a table.
@@
 We shall extend "tidy data" into the time series domain.
-
-This paper proposes a tidy data structure and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.
-
-# Tidy temporal-context data
 

### 2018-04-24 main.Rmd
started to mention time series and panel data

@@
 Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage.
-The tidy temporal data along with its supporting methods has been implemented in the R package **tsibble**.'
+The architecture of tidy temporal data has been implemented in the R package **tsibble**.'
 keywords: "temporal context, time series, data structure, R"
@@
 For example, data recorded at fine time resolutions (hours, minutes, and seconds) are typically associated with different time zones and daylight savings.
-Temporal data often carries with rich information other than the time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping variables exhibiting nested or crossed structures, linking to other data tables, and etc.
+Temporal data often carries with rich information other than the time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors involving nested or crossed structures, linking to other data tables, and etc.
 
-Although a matrix is a tabular form, it assumes homogeneity (that is, all the columns must be real values.) and time indices implicitly inferred as attributes (or meta-information).
-This format of data input is model-oriented rather than data-oriented as it proves computationally efficient for many statistical models when doing matrix operations.
-However, it is has the effect of discarding interesting information in the data, is more opaque in the handling of the temporal components, and less efficient for many other data analysis tasks, such as data transformation, visualisation and model diagnostics, like forcing the fitting of a square peg into a round hole.
+However, such richness in temporal data cannot be easily accommodated by currently available data structures in any statistical computing package.
+In the literature, time series and panel (longitudinal) data are common terms referred to temporal-context data, depending on the research fields.
+Researchers who are concerned with modelling large $T$ and small $N$ would name as "time series" (serial correlation); those who are interested in modelling small $T$ and large $N$ as "panel data" (asymptotic).
+The data format is two-dimensional array, but different modelling focuses lend the data input to column types.
+A matrix is used to represent multivariate time series, assuming homogeneity (that is, all the columns must be of same type.) and time indices implicitly inferred as attributes (or meta-information).
 
-In addition, time indices form the basis of temporal data and the source of contextual information to understand temporal dependence.
-Time as attributes limits the use of time context, since it is difficult to access and manipulate.
-For example, the components of date-times, such as months, day of week, and time of day, cannot be easily extracted in order to examine seasonal fluctuations.
-It is also impossible to switch between different time zones and daylight savings to compare time across places when needed.
-
-This paper proposes a tidy data structure and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.
+This paper proposes a tidy and unified data structure and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.
 

### 2018-04-26 main.Rmd
added panel data str explanations

@@
 The data format is two-dimensional array, but different modelling focuses lend the data input to column types.
-A matrix is used to represent multivariate time series, assuming homogeneity (that is, all the columns must be of same type.) and time indices implicitly inferred as attributes (or meta-information).
+A matrix is used to represent multivariate time series where each row represents observations measured at a time point and each column represents a series.
+This matrix representation requires homogeneity (that is, all the columns must be of same type.), time indices implicitly inferred as attributes or meta-information, series of same time length, and explicit missingness.
+By contrast, panel data are organised in two-dimensional array of heterogeneous column types where multiple study subjects are stacked and repeated for its time indices in a single column, due to commonly arisen unbalanced panels.
+This specification requires explicitly declared panel variable and index, which has been implemented in Stata's **tsset** and R package **plm**.
+This data organisation appears more advantageous than matrix in supporting explicit time index, multiple subjects of different time lengths.
 
@@
 In SQL database, a primary key [@codd_relational_1970], which uniquely defines each record in a database table, is equivalent to the composition of "index" and "key".
-In multivariate time series notation as Equation \@ref(eq:mts), $X_{jt}$ represents series $j$, for $j = 1, \dots, p$ and $1 \leq t \leq T$, in the form of a $T \times p$ matrix.
-
-\begin{equation}
-  \begin{bmatrix}
-  X_{11} & X_{21} & \cdots & X_{p1} \\
-  X_{12} & X_{22} & \cdots & X_{p2} \\
-  \vdots & \vdots & \ddots & \vdots \\
-  X_{1T} & X_{2T} & \cdots & X_{pT}
-  \end{bmatrix}
-\label{eq:mts}
-\end{equation}
 

### 2018-04-28 main.Rmd
added hts introduction

@@
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
-A sound pipeline practice facilitates a transparent and human-readable workflow for analyzing temporal data.
+A sound pipeline practice facilitates a transparent and human readable workflow for analyzing temporal data.
 Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage.
@@
 
-However, such richness in temporal data cannot be easily accommodated by currently available data structures in any statistical computing package.
 In the literature, time series and panel (longitudinal) data are common terms referred to temporal-context data, depending on the research fields.
 Researchers who are concerned with modelling large $T$ and small $N$ would name as "time series" (serial correlation); those who are interested in modelling small $T$ and large $N$ as "panel data" (asymptotic).
-The data format is two-dimensional array, but different modelling focuses lend the data input to column types.
-A matrix is used to represent multivariate time series where each row represents observations measured at a time point and each column represents a series.
+The data format is two-dimensional array, but different modelling focuses lend the data input to different representations.
+A matrix is used to represent multivariate time series where each row represents observations measured at a time point and each column represents a series ("wide" form).
 This matrix representation requires homogeneity (that is, all the columns must be of same type.), time indices implicitly inferred as attributes or meta-information, series of same time length, and explicit missingness.
-By contrast, panel data are organised in two-dimensional array of heterogeneous column types where multiple study subjects are stacked and repeated for its time indices in a single column, due to commonly arisen unbalanced panels.
-This specification requires explicitly declared panel variable and index, which has been implemented in Stata's **tsset** and R package **plm**.
-This data organisation appears more advantageous than matrix in supporting explicit time index, multiple subjects of different time lengths.
+By contrast, panel data are organised in rectangular form of heterogeneous column types where multiple study subjects are stacked and repeated for its time indices in a single column ("long" form), due to commonly arisen unbalanced panels.
+This specification requires explicitly declared panel variable and index, which has been implemented in Stata's time series module and R package **plm**.
+This data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
 
-This paper proposes a tidy and unified data structure and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.
+Temporal data can often be aggregated in a manner that exhibits a nested or crossed structure, also known as hierarchical or grouped time series [@fpp].
+For example, in a manufacturing setting, a company can add up every store’s sales by region, by state and by country, which gives a strictly hierarchical time series; alternatively they can group the sales for each product together based on common attributes such as store, brand, price range and so forth, which leads to a non-hierarchical structure---a grouped time series.
+The R package **hts** is the implementation of this type of time series.
+But it is frustrating to create and work with such data objects due to a machine-oriented interface.
+
+@wickham2014tidy formulated the conceptual framework of tidy data: (1) each variable forms a column; (2) each observation forms a row; (3) each type of observational unit forms a table.
+These principles attempt to standardize the mapping from the semantics of a dataset to its structure and facilitate data analysis in a coherent way.
+A grammar of data manipulation.
+
+This paper proposes a unified data representation of temporal-context data, by leveraging the "tidy data" principles.
+Built on a two-dimensional array of long format, observations and variables find and bridge their meanings in both physical and underlying representations.
+
+This paper proposes a unified data representation and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.
 
 # Data semantics
-
-@wickham2014tidy developed and formulated the conceptual framework of tidy data: (1) each variable forms a column; (2) each observation forms a row; (3) each type of observational unit forms a table.
-These principles attempt to standardise the mapping from the semantics of a dataset to its structure and facilitate data analysis in a coherent way.
-We shall extend "tidy data" into the time series domain.
 
@@
 
-In SQL database, a primary key [@codd_relational_1970], which uniquely defines each record in a database table, is equivalent to the composition of "index" and "key".
+In SQL database, a primary key [@codd_relational_1970], which uniquely defines each record in a database, is equivalent to the composition of "index" and "key".
 
@@
 In experimental designs, a variable is crossed with another when every category of one variable co-occurs with every category of the other, while a variable is nested within another when each category of the former variable co-occurs with only one category of the latter.
-Regarding faceted plots in statistical graphics, @wilkinson2006grammar discussed the difference between nesting and crossing structure in depth.
-The distinction between two graphical layouts may appear subtle but convey completely different meanings of the data.
-The crossing example could be gender with marital status: married women, married men, single women, and single men.
-It takes all the possible combinations into account, resulting in 2 by 2 drawing panels in a graphical device.
-If a panel goes empty, there exists no observation for that category in the dataset, but it would be filled with observations.
-However, considering pregnancy status and gender, there are only three possible combinations: pregnant women, non-pregnant women, and non-pregnant men, since pregnant status is nested under the gender variable.
-In this case, any physical layout of such nesting structure must not occur to an impossible category.
-
 It appears more useful in making this distinction in statistical analysis including visualisation and modeling, compared to data manipulation.

### 2018-04-29 main.Rmd
added 1 paragraph to intro

@@
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
-A sound pipeline practice facilitates a transparent and human readable workflow for analyzing temporal data.
+A sound pipeline practice facilitates a succinct and transparent workflow for analyzing temporal data.
 Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage.
@@
 This data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
+[R]
 
@@
 But it is frustrating to create and work with such data objects due to a machine-oriented interface.
+[R]
 
@@
 A grammar of data manipulation.
+[R]
 
-This paper proposes a unified data representation of temporal-context data, by leveraging the "tidy data" principles.
-Built on a two-dimensional array of long format, observations and variables find and bridge their meanings in both physical and underlying representations.
-
-This paper proposes a unified data representation and a modern data pipeline for storing, managing and analysing time series data, using a collection of fluent and fluid tools to help with exploitation in temporal context.
+This paper proposes a unified data representation of temporal-context data, blending time series of nested and crossed factors into a two-dimensional column-homogeneous array in a long format.
+By leveraging the "tidy data" principles, observations and variables position and bridge their meanings in both physical and underlying structures.
+Data manipulation involves in transforming either observations or variables, or both, which can be described and achieved with a collection of shorthand verbs.
+A chain of data transformations lend itself to a data pipeline.
 
@@
 
+A pipeline exhibits a hierarchy of data operations: (1) atomic (1-dimensional) vectors (`mean(variale)`) --> (2) (2-dimensional) data table (`summarise()`) --> repeat step (1) and (2) to form a chain.
+
 Tidy data builds a concrete foundation to enable pipeline data analysis, which provides a coherent and fluent framework to work with data.
@@
 
-# Discussion
+# Conclusion and future work
 

### 2018-05-01 main.Rmd
added opening paragraph in "data semantics"

@@
 
-Temporal-context data consist of observational units indexed at different time points $X_{jt}$, where the $j$\textsuperscript{th} unit takes measurements of $X$ on over time $t$, for $j = 1, \cdots, N$ and $1 \le t \le T$.
+Temporal-context data consist of observational units indexed at different time points $X_{jt}$, where the $j$\textsuperscript{th} unit takes measurements of $X$ on over time $t$, for $j = 1, \dots, N_{i}$ and $1 \le t \le T$.
 Time primarily forms the contextual basis of temporal data, but it could arrive in many possible formats.
@@
 This specification requires explicitly declared panel variable and index, which has been implemented in Stata's time series module and R package **plm**.
-This data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
+Evidently, this data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
 [R]
@@
 The R package **hts** is the implementation of this type of time series.
-But it is frustrating to create and work with such data objects due to a machine-oriented interface.
+But it is frustrating to create and work with such data objects due to non-extensible and non-modified.
+Variational data aggregation (some groups no need aggregation).
 [R]
 
-@wickham2014tidy formulated the conceptual framework of tidy data: (1) each variable forms a column; (2) each observation forms a row; (3) each type of observational unit forms a table.
+@wickham2014tidy coined the term "tidy data" and formalized the processing from messy data to "tidy data".
 These principles attempt to standardize the mapping from the semantics of a dataset to its structure and facilitate data analysis in a coherent way.
-A grammar of data manipulation.
+Based on the systematic structuring principles, a grammar of data manipulation.
 [R]
@@
 This paper proposes a unified data representation of temporal-context data, blending time series of nested and crossed factors into a two-dimensional column-homogeneous array in a long format.
-By leveraging the "tidy data" principles, observations and variables position and bridge their meanings in both physical and underlying structures.
-Data manipulation involves in transforming either observations or variables, or both, which can be described and achieved with a collection of shorthand verbs.
+By leveraging the "tidy data" principles, observations and variables position and bridge their meanings in both physical and internal structures.
+Data manipulation involves in transforming either observations or variables, or both, which can be described and achieved with a collection of shorthand operators.
 A chain of data transformations lend itself to a data pipeline.
@@
 
-A modern re-imagining of time series should provide heterogeneous data types and time indices as explicitly declared data column.
-This can be achieved using a "data frame" in R or other statistical languages to represent a tabular format, instead of "matrix".
-Tidy temporal-context data at least consists of:
-
-1. index: an explicitly declared data variable contains time indices, such as date-times, year-months, years and etc.
-2. key: a set of grouping factors uniquely identifies each unit that measurements take place on over time, which may include single or multiple columns.
-3. interval: data with regular time interval results in a common time interval in one table.
-
-In SQL database, a primary key [@codd_relational_1970], which uniquely defines each record in a database, is equivalent to the composition of "index" and "key".
+The choice of representation of temporal-context data is made from a data-centric perspective, which is taken in the light of the operations that are to be performed on the data.
+Firstly, a data set must be structured in a "tidy" layout.
+Secondly, declaring the data set to contain temporal observations is determined by "index" that represent time and "key" that uniquely identifies each unit that measurements take place on over time.
+The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
 
@@
 
-A pipeline exhibits a hierarchy of data operations: (1) atomic (1-dimensional) vectors (`mean(variale)`) --> (2) (2-dimensional) data table (`summarise()`) --> repeat step (1) and (2) to form a chain.
+Literature about data pipeline.
+
+A pipeline exhibits a hierarchy of data operations: (1) atomic (1-dimensional) vectors (`mean(variable)`) --> (2) (2-dimensional) data table (`summarise()`) --> repeat step (1) and (2) to form a chain.
+Reversely, a data pipeline is decomposed into rectangular blocks, and then into atomic strips.
+(UNIX pipe)
 
@@
 * **rolling window**: `slide()`, `tile()`, `stretch()`
-* **statistics**: `lag()`, `diff()`, `acf()`
+* **time-wise**: `lag()`, `diff()`, `acf()`
 

### 2018-05-02 main.Rmd
added a couple of more sentence to semantic intro section

@@
 Firstly, a data set must be structured in a "tidy" layout.
-Secondly, declaring the data set to contain temporal observations is determined by "index" that represent time and "key" that uniquely identifies each unit that measurements take place on over time.
+Secondly, declaring the data set to contain temporal observations is determined by an "index" that represent time and a "key" that uniquely identifies each unit that measurements take place on over time.
+The "key" works similarly as the panel variable in the Stata's **tsset** command to define the units or subjects, but it is expanded to include multiple variables rather than a single one.
+A syntax is introduced to express a key consisting of nested and crossed factors.
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
+
+Given the nature of temporal ordering, a temporal data set must be sorted by time index. If a key is explicitly declared, the key will be sorted followed by arranging time in ascending order.
 

### 2018-05-03 main.Rmd
reasoned explicit time index

@@
 The choice of representation of temporal-context data is made from a data-centric perspective, which is taken in the light of the operations that are to be performed on the data.
+This data abstraction reflects most common problems of incoming data and transformation in reality.
 Firstly, a data set must be structured in a "tidy" layout.
@@
 
-Given the nature of temporal ordering, a temporal data set must be sorted by time index. If a key is explicitly declared, the key will be sorted followed by arranging time in ascending order.
+Given the nature of temporal ordering, a temporal data set must be sorted by time index.
+If a key is explicitly declared, the key will be sorted first and followed by arranging time in ascending order.
 
@@
 
-Time index forms an integral component and contextual base of temporal data.
-In temporal data frame, time-based index must be clearly stated as a data column rather than inferred as attributes, and thus can be accessible.
-This creates flexibility in handling time indices.
+Time forms an integral component and a contextual basis of temporal data.
+A variable representing time needs explicitly declared at the process of constructing a temporal data, referred to as "index".
+This accessibility of index promotes transparency and unambiguity while manipulating time.
+For example, subsetting a certain time period of data, extracting time components (like time of day and day of week), or converting time zones are directly dealt with index.
+It is also often to join other data tables based on the common time indices.
+Tools are provided to work with time, rather than wrappers to work with a whole data set.
+This sets the tone on methods design philosophy: well-structured, verbose and unambiguous workflow.
 
-(1) temporal elements can be created, and then exploratory data visualisation and analysis (not specialist plotting) and multiple seasonality modelling for sub-daily data
-(2) convert to the same time zone and thus compare
-(3) join other data tables using the index as common key
+Time representation: year, year-quarter, year-month, year-week, dates, date-times.
 
-For data indexed in regular time space, the time interval is obtained by computing the minimal positive time distance in a data table.
+For data indexed in regular time space, the time interval is obtained by computing the greatest common divisor from positive time distances in a data table.
 This suggests that each observational unit collected at the same interval forms a table.

### 2018-05-23 main.Rmd
added a bit to key

@@
 This data abstraction reflects most common problems of incoming data and transformation in reality.
-Firstly, a data set must be structured in a "tidy" layout.
+Firstly, a data set must be structured in a "tidy" rectangular layout.
 Secondly, declaring the data set to contain temporal observations is determined by an "index" that represent time and a "key" that uniquely identifies each unit that measurements take place on over time.
@@
 If a key is explicitly declared, the key will be sorted first and followed by arranging time in ascending order.
+
+This high-level data abstraction is semantically structured, which shed lights on time index and what we call "key".
 
@@
 Tools are provided to work with time, rather than wrappers to work with a whole data set.
-This sets the tone on methods design philosophy: well-structured, verbose and unambiguous workflow.
+This sets the tone on methods design philosophy: well-structured, expressive and unambiguous workflow.
 
@@
 
-## Keys
+## Key
 
-Key variables are usually discrete descriptors, and are typically variables that were created during the data collection to uniquely define the measured values.
-For instance, to distinguish the performance of each flight in the dataset, the "key" is the `flight` variable, allowing separation of multiple time series in one data table.
-The "key" not only identifies the unit to be measured over time, but also incorporates structures of data.
-Without a key, a data table can be considered as a univariate time series; in other words, the key is implicit.
-With a single key of more than one categories, it lends itself to a collection of time series in a table.
-But when there are at least two keys in the table, it indicates nested or crossed structures.
+The "key" identifies units or subjects that are observed over time in a data table, which are typically known in advance by users.
+The absence of key only occurs to a univariate case where the key is implicit.
+Multiple units must associate with an explicit "key" to allow for identification.
+The key is not constrained to a single column, but is comprised of multiple columns.
+When involving multiple variables, one need to distinguish nesting and crossing data variables.
+"Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter; on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other.
+Nesting is a special case of crossing.
+Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data.
+Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
+The separator `|` is used to indicate nesting variables, while comma `,` for crossing variables.
 
-In experimental designs, a variable is crossed with another when every category of one variable co-occurs with every category of the other, while a variable is nested within another when each category of the former variable co-occurs with only one category of the latter.
-It appears more useful in making this distinction in statistical analysis including visualisation and modeling, compared to data manipulation.
+What I'd like to expand:
+
+1. syntax, user-oriented approach.
+2. structure construction only use observations found in the data. For example, crossing variables will not expand the unique combination that are not found in the data.
 
@@
 * **row-wise**: `filter()`, `slice()`, `arrange()`, `fill_na()`
-* **column-wise**: `mutate()`, `select()`, `summarise()`, `tsummarise()`
+* **column-wise**: `mutate()`, `select()`, `summarise()`
+* **group-wise**: `index_by()`, `group_by()`
+* **reshape**: `gather()`, `spread()`
 * **rolling window**: `slide()`, `tile()`, `stretch()`
-* **time-wise**: `lag()`, `diff()`, `acf()`
+* **time-wise**: `lead()`, `lag()`, `difference()`
 

### 2018-05-30 main.Rmd
added one paragraph for index

@@
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and modelling routines.
-Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntatical and user-oriented approach in which it imposes nested or crossed structures on the data.
+Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntactical and user-oriented approach in which it imposes nested or crossed structures on the data.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
@@
 
-Time representation: year, year-quarter, year-month, year-week, dates, date-times.
-
-For data indexed in regular time space, the time interval is obtained by computing the greatest common divisor from positive time distances in a data table.
-This suggests that each observational unit collected at the same interval forms a table.
+For data indexed in regular time space, the time interval is obtained by computing the greatest common divisor based on the non-negative differences of a given time index, although the data may include implicit time gaps.
+This implies that all the units in the table must share a common interval.
+It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R.
+However these turn out inappropriate for weekly, monthly and quarterly data.
+For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year.
+However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced.
+A genuine year-month representation should avoid associating days with its format.
+Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
 

### 2018-06-01 main.Rmd
added explanations of nesting and crossing

@@
 Nesting is a special case of crossing.
+The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent); whereas the ordering of the latter makes no difference.
 Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data.
 Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
-The separator `|` is used to indicate nesting variables, while comma `,` for crossing variables.
-
-What I'd like to expand:
-
-1. syntax, user-oriented approach.
-2. structure construction only use observations found in the data. For example, crossing variables will not expand the unique combination that are not found in the data.
+The expression of a vertical bar (`|`) is used to indicate nesting variables, while a comma (`,`) for crossing variables.
+@wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.
+A crossing structure displays all unique combinations of crossed variables, including those not found in the data, in a graphical layout.
+However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of missing combinations.
 
@@
 * **group-wise**: `index_by()`, `group_by()`
-* **reshape**: `gather()`, `spread()`
 * **rolling window**: `slide()`, `tile()`, `stretch()`

### 2018-06-02 main.Rmd
mapped out the literature in data pipeline

@@
 Based on the systematic structuring principles, a grammar of data manipulation.
+@r4ds argues that 80% of data analysis tasks can be solved with tidy tools while the remaining 20% requires other tools.
 [R]
@@
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
+
+[A toy example is needed here and will referenced for index and key]
 
@@
 @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.
-A crossing structure displays all unique combinations of crossed variables, including those not found in the data, in a graphical layout.
-However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of missing combinations.
+A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout.
+However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.
 
@@
 
-Literature about data pipeline.
+## Literature review
+
+* A collection of intuitive and consistent programming interfaces empowers users to focus on the primary goal of data analysis and modelling problems, and frees users from computational implementation details.
+* [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R]
+* Unix philosophy [@unix; @unix-philosophy]: each do one simple thing, do it well, and most importantly, work well with each other.
+* "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system" [@eopl]
+
+* Pipelines in interactive settings [@viewing-pipeline; @plumbing; @xie2014reactive]: Model-View-Controller paradigm.
+
+## Time-based pipeline analysis
 
@@
 Reversely, a data pipeline is decomposed into rectangular blocks, and then into atomic strips.
-(UNIX pipe)
 
@@
 * **group-wise**: `index_by()`, `group_by()`
+* **time-wise**: `lead()`, `lag()`, `difference()`
 * **rolling window**: `slide()`, `tile()`, `stretch()`
-* **time-wise**: `lead()`, `lag()`, `difference()`
 

### 2018-06-08 main.Rmd
added two paragraphs to data pipeline

@@
 
+The rest of the paper is structured as follows.
+
 # Data semantics
@@
 This data abstraction reflects most common problems of incoming data and transformation in reality.
-Firstly, a data set must be structured in a "tidy" rectangular layout.
+Firstly, a data set must be structured in a "tidy" rectangular layout each row has the same schema and each field has a single value.
 Secondly, declaring the data set to contain temporal observations is determined by an "index" that represent time and a "key" that uniquely identifies each unit that measurements take place on over time.
@@
 It is also often to join other data tables based on the common time indices.
-Tools are provided to work with time, rather than wrappers to work with a whole data set.
+Tools are provided to work with time itself, rather than wrappers to work with a whole data set.
+This promotes transparency and visibility to make inspection easier.
 This sets the tone on methods design philosophy: well-structured, expressive and unambiguous workflow.
@@
 Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
+
+Extensible. Custom index classes can be supported.
 
@@
 
+This representation choice will prove suitable for almost all conceivable applications of temporal data.
+But the data abstraction cannot live alone with an additional toolchain in order to be able to have insight and knowledge about the data itself.
+A desirable toolchain glues transformation, visualisation, and modelling as a data science pipeline.
+Each component itself is also a self-contained pipeline that can be decomposed into many individual operations.
+This paper will focus on part of the general pipeline---data transformation, for example filtering, subsetting and summarising.
+We shall see how this data abstraction breeds a consistent data process.
+
 ## Literature review
 
-* A collection of intuitive and consistent programming interfaces empowers users to focus on the primary goal of data analysis and modelling problems, and frees users from computational implementation details.
-* [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R]
-* Unix philosophy [@unix; @unix-philosophy]: each do one simple thing, do it well, and most importantly, work well with each other.
-* "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system" [@eopl]
+There has been a long history of pipeline discussions and implementation centering around data in various aspects.
+The Extract, Transform, and Load (ETL) process is most commonly deployed pipeline in data science.
+However, building a sound data pipeline can be technically difficult, which involves many implementation decisions including interface, input/output, functionality and so on to be made.
+
+@unix coined the term "pipelines" in software development, while developing Unix at Bell Labs.
+In Unix-based computer operating systems, a pipeline chains a sequence of operations together on the basis of their standard streams, so that the output of each programme becomes the input to another.
+This shapes the Unix toolbox philosophy: "each do one simple thing, do it well, and most importantly, work well with each other" [@unix-philosophy].
 
@@
 ## Time-based pipeline analysis
+
+* A collection of intuitive and consistent programming interfaces empowers users to focus on the primary goal of data analysis and modelling problems, and frees users from computational implementation details.
+<!-- * [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R] -->
+* "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system" [@eopl]
+
+* error-prone
+* The resulting code is well-structured, unambiguous, easier to read, comprehend and maintain.
+
+Individual operations can talk to each other.
 

### 2018-06-11 main.Rmd
added tb data

@@
 abstract: 'Temporal-context data is often rich with information and time formats, for example multiple observational units, different time lengths, heterogeneous data types, nested and crossed factors and etc.
-This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and modelling routines.
+This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
 Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntactical and user-oriented approach in which it imposes nested or crossed structures on the data.
@@
 toc: false
-bibliography: references.bib
+bibliography: [references.bib, rpkgs.bib]
 biblio-style: authoryear-comp
@@
 )
-# read_chunk('src/main.R')
+read_chunk('R/demo.R')
+```
+
+```{r load-pkg}
 ```
@@
 By contrast, panel data are organised in rectangular form of heterogeneous column types where multiple study subjects are stacked and repeated for its time indices in a single column ("long" form), due to commonly arisen unbalanced panels.
-This specification requires explicitly declared panel variable and index, which has been implemented in Stata's time series module and R package **plm**.
+This specification requires explicitly declared panel variable and index, which has been implemented in Stata's time series module and R package **plm** [@R-plm].
 Evidently, this data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
@@
 For example, in a manufacturing setting, a company can add up every store’s sales by region, by state and by country, which gives a strictly hierarchical time series; alternatively they can group the sales for each product together based on common attributes such as store, brand, price range and so forth, which leads to a non-hierarchical structure---a grouped time series.
-The R package **hts** is the implementation of this type of time series.
+The R package **hts** [@R-hts] is the implementation of this type of time series.
 But it is frustrating to create and work with such data objects due to non-extensible and non-modified.
@@
 These principles attempt to standardize the mapping from the semantics of a dataset to its structure and facilitate data analysis in a coherent way.
-Based on the systematic structuring principles, a grammar of data manipulation.
+Based on the systematic structuring principles, the grammar of data manipulation, as implemented in the R package **dplyr** [@R-dplyr], abstracts an coherent and consistent set of methods to handle a wide range of data transformation tasks.
 @r4ds argues that 80% of data analysis tasks can be solved with tidy tools while the remaining 20% requires other tools.
@@
 
-[A toy example is needed here and will referenced for index and key]
-
 Given the nature of temporal ordering, a temporal data set must be sorted by time index.
@@
 This high-level data abstraction is semantically structured, which shed lights on time index and what we call "key".
+
+```{r load-data}
+```
+
+(ref:tb-sub) A subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012.
 
@@
 This representation choice will prove suitable for almost all conceivable applications of temporal data.
-But the data abstraction cannot live alone with an additional toolchain in order to be able to have insight and knowledge about the data itself.
+But the data abstraction cannot live alone with an additional toolchain in order to be able to acquire insight and knowledge about the data itself.
 A desirable toolchain glues transformation, visualisation, and modelling as a data science pipeline.
@@
 The Extract, Transform, and Load (ETL) process is most commonly deployed pipeline in data science.
-However, building a sound data pipeline can be technically difficult, which involves many implementation decisions including interface, input/output, functionality and so on to be made.
+However, building a sound data pipeline can be technically difficult, which involves many implementation decisions, including interface, input/output, functionality and so on, to be made.
 
 @unix coined the term "pipelines" in software development, while developing Unix at Bell Labs.
-In Unix-based computer operating systems, a pipeline chains a sequence of operations together on the basis of their standard streams, so that the output of each programme becomes the input to another.
+In Unix-based computer operating systems, a pipeline chains together a series of operations on the basis of their standard streams, so that the output of each programme becomes the input to another.
 This shapes the Unix toolbox philosophy: "each do one simple thing, do it well, and most importantly, work well with each other" [@unix-philosophy].
 
-* Pipelines in interactive settings [@viewing-pipeline; @plumbing; @xie2014reactive]: Model-View-Controller paradigm.
+@viewing-pipeline proposed a viewing pipeline for data analysis in interactive statistical graphics and generalised a collection of elements required for the pipeline, which takes control of transformation from data to plot.
+@plumbing and @xie2014reactive implemented the data pipeline in some interactive statistical graphics software, using a reactive programming framework, in which user's interactions trigger a sequence of modules to update views accordingly.
 
 ## Time-based pipeline analysis
+
+Tidy data builds a concrete foundation to enable pipeline data analysis, which provides a coherent and fluent framework to work with data.
+It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.
 
@@
 * error-prone
-* The resulting code is well-structured, unambiguous, easier to read, comprehend and maintain.
+* The resulting code is well-structured, unambiguous, easier to read, comprehend and maintain. A way of chaining together simple operations to perform complex tasks in an efficient way.
 
@@
 
-Tidy data builds a concrete foundation to enable pipeline data analysis, which provides a coherent and fluent framework to work with data.
-It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.
-
 * **row-wise**: `filter()`, `slice()`, `arrange()`, `fill_na()`
@@
 * **rolling window**: `slide()`, `tile()`, `stretch()`
+
+The index and key will automate the update.
 
@@
 Model summaries include coefficients, fitted values, and residuals; forecasting objects include future time path and distributions generating prediction intervals.
+
+```{r write-bib}
+write_bib(c("hts", "plm", "dplyr"), file = "rpkgs.bib")
+```

### 2018-06-11 R/demo.R

@@ -1,0 +1,9 @@
+## ---- load-pkg
+library(tidyverse)
+library(tsibble)
+
+## ---- load-data
+tb_small <- read_rds("data/tb_small.rds")
+tb_small %>% 
+  knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>% 
+  kableExtra::kable_styling(position = "center")

### 2018-06-11 data-raw/subset.R

@@ -1,0 +1,17 @@
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+
+tb <- read_csv("data/tb-2018-06-05.csv")
+
+tb_small <- tb %>% 
+  select(-region, -age) %>% 
+  filter(
+    country %in% c("Australia", "New Zealand", "United States of America"),
+    year > 2010
+  ) %>% 
+  group_by(country, continent, gender, year) %>% 
+  summarise(count = sum(count)) %>% 
+  ungroup()
+
+write_rds(tb_small, "data/tb_small.rds")

### 2018-06-12 main.Rmd
added a bit more to time-based pipeline analysis

@@
 But the data abstraction cannot live alone with an additional toolchain in order to be able to acquire insight and knowledge about the data itself.
-A desirable toolchain glues transformation, visualisation, and modelling as a data science pipeline.
+A desirable toolbox chains transformation, visualisation, and modelling as a data science pipeline.
 Each component itself is also a self-contained pipeline that can be decomposed into many individual operations.
-This paper will focus on part of the general pipeline---data transformation, for example filtering, subsetting and summarising.
+This paper will focus on part of the general pipeline---data transformation, for example filtering observations, selecting and summarising variables.
 We shall see how this data abstraction breeds a consistent data process.
@@
 @viewing-pipeline proposed a viewing pipeline for data analysis in interactive statistical graphics and generalised a collection of elements required for the pipeline, which takes control of transformation from data to plot.
-@plumbing and @xie2014reactive implemented the data pipeline in some interactive statistical graphics software, using a reactive programming framework, in which user's interactions trigger a sequence of modules to update views accordingly.
+@plumbing and @xie2014reactive implemented the data pipeline in interactive statistical software **plumbr** and **cranvas** respectively, using a reactive programming framework, in which user's interactions trigger a sequence of modules to update views accordingly.
 
@@
 
-Tidy data builds a concrete foundation to enable pipeline data analysis, which provides a coherent and fluent framework to work with data.
+What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers them to focus on the statistical analysis tasks at hand without concerning the details of computational implementation.
+A fluent and coherent pipeline glues the grammar of data manipulation and tidy data as a fundamental unit of data analysis together.
 It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.
 
-* A collection of intuitive and consistent programming interfaces empowers users to focus on the primary goal of data analysis and modelling problems, and frees users from computational implementation details.
-<!-- * [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R] -->
-* "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system" [@eopl]
-
-* error-prone
-* The resulting code is well-structured, unambiguous, easier to read, comprehend and maintain. A way of chaining together simple operations to perform complex tasks in an efficient way.
-
-Individual operations can talk to each other.
-
-A pipeline exhibits a hierarchy of data operations: (1) atomic (1-dimensional) vectors (`mean(variable)`) --> (2) (2-dimensional) data table (`summarise()`) --> repeat step (1) and (2) to form a chain.
-Reversely, a data pipeline is decomposed into rectangular blocks, and then into atomic strips.
+An individual operation that performs on a specific task can be phrased as a verb.
+This verb is self-explanatory enough to express what the specific task is.
+We shall apply **dplyr** key verbs to time domain and expand its vocabulary a little more to handle time-related analysis problems.
 
@@
 * **rolling window**: `slide()`, `tile()`, `stretch()`
+
+<!-- * [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R] -->
+* "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system" [@eopl]
+
+* The resulting code is well-structured, easier to read and comprehend. A way of chaining together simple operations to perform complex tasks in an efficient way.
+
+Individual operations can talk to each other.
+
+A pipeline exhibits a hierarchy of data operations: (1) atomic (1-dimensional) vectors (`mean(variable)`) --> (2) (2-dimensional) data table (`summarise()`) --> repeat step (1) and (2) to form a chain.
+Reversely, a data pipeline is decomposed into rectangular blocks, and then into atomic strips.
 

### 2018-06-14 main.Rmd
added explanations to tb data

@@
 
-# Introduction
+# Introduction {#sec:intro}
 
@@
 
-# Data semantics
+# Data semantics {#sec:semantics}
 
@@
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
+Table \@ref(tab:tb-sub) shows a "tidy" temporal data example using a subset of tuberculosis cases estimated by @tb-data.
+It contains 12 observations and 5 variables arranged in a "long" tabular form.
+Each observation uniquely records the number of people who are diagnosed tuberculosis for each gender at three selected countries in the years of 2011 and 2012, where column `year` is declared as the index variable, columns `country`, `continent` and `gender` as key.
+To be more specific, the key is passed via `id(country | continent, gender)`: a geographic hierarchy of country nested within continent, crossed with a demographic factor.
+Detailed discussions about index and key specification will be given in Section \@ref(sec:index) and \@ref(sec:key).
+Although column `count` is a single measure in this case, it is sufficiently flexible to add other measured variables, for example, adding the corresponding population to calibrate the count.
 
@@
 
-```{r load-data}
+```{r tb-sub}
 ```
 
-(ref:tb-sub) A subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012.
+(ref:tb-sub) A subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012, with 12 observations and 5 variables.
+The index refers to column `year`, the key to multiple columns: `contry` nested under `continent` crossed with `gender`, and the measured variable to column `count`.
 
-## Time index and interval
+## Time index and interval {#sec:index}
 
@@
 Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
+(coercion)
 
@@
 
-## Key
+## Key {#sec:key}
 
@@
 
-# Data pipeline
+# Data pipeline {#sec:pipeline}
 
@@
 
-## Literature review
+## Literature review (take this heading out)
 

### 2018-06-14 R/demo.R

@@ -2,7 +2,7 @@
 library(tidyverse)
 library(tsibble)
 
-## ---- load-data
+## ---- tb-sub
 tb_small <- read_rds("data/tb_small.rds")
 tb_small %>% 
   knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>% 

### 2018-07-03 main.Rmd
added a bit more on how to arrange the toy example

--- 2018-06-12 main.Rmd
@@
 Evidently, this data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
+less defaults to be made.
 [R]
@@
 Firstly, a data set must be structured in a "tidy" rectangular layout each row has the same schema and each field has a single value.
-Secondly, declaring the data set to contain temporal observations is determined by an "index" that represent time and a "key" that uniquely identifies each unit that measurements take place on over time.
-The "key" works similarly as the panel variable in the Stata's **tsset** command to define the units or subjects, but it is expanded to include multiple variables rather than a single one.
+Secondly, declaring the data set to contain temporal observations is determined by an "index" that represents time and a "key" that uniquely identifies each unit that measurements take place on over time.
+The "key" serves a similar purpose as the panel variable in the Stata's **tsset** command to define the units or subjects, but it is expanded to include multiple variables rather than a single one.
 A syntax is introduced to express a key consisting of nested and crossed factors.
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
+Table \@ref(tab:tb-sub) shows a "tidy" temporal data example using a subset of tuberculosis cases estimated by @tb-data.
+It contains 12 observations and 5 variables arranged in a "long" tabular form.
+Each observation uniquely records the number of people, who are diagnosed tuberculosis for each gender at three selected countries in the years of 2011 and 2012, where column `year` is declared as the index variable, columns `country`, `continent` and `gender` as key.
+To be more specific, the key is passed via `id(country | continent, gender)` or `id(gender, country | continent)`: a geographic hierarchy of country nested within continent, crossed with a demographic factor.
+Detailed discussions about index and key specifications will be given in Section \@ref(sec:index) and \@ref(sec:key).
+Although column `count` is the only measure in this case, it is sufficiently flexible to hold other measured variables, for example, adding the corresponding population size (if known) in order to calibrate the count later.
 
-Given the nature of temporal ordering, a temporal data set must be sorted by time index.
-If a key is explicitly declared, the key will be sorted first and followed by arranging time in ascending order.
+Given the nature of temporal ordering, a temporal data set must be sorted by time index as a result.
+If a key is explicitly declared, the key will be sorted first depending on the ordering of factor levels and the ordering of variables, and followed by arranging time in ascending order.
+The tuberculosis example shows the key is sorted in alphabetic order, followed by ascending years by default, as is Table \@ref(tab:tb-sub).
+Since time-based operations, for example leading, lagging, or differencing series, require a vector in its time order, the arranged data format assures its validity for these operations without returning unexpected results.
+There are other possible arrangements in the key, but the baseline is the index (i.e.
+`year`) ascends within each key element.
+If any operations on the tidy temporal data results a deviation from this baseline, it starts issuing a warning to the users.
 
@@
 
-```{r load-data}
+```{r tb-sub}
 ```
 
-(ref:tb-sub) A subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012.
+(ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012, with 12 observations and 5 variables.
+The index refers to column `year`, the key to multiple columns: `contry` nested under `continent` crossed with `gender`, and the measured variable to column `count`.
 
-## Time index and interval
+## Time index and interval {#sec:index}
 
@@
 Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
+This argument is also applicable to the creation of year-week and year-quarter.
+(coercion)
+
+Since Table \@ref(tab:tb-sub) possesses complete observations for each category, the interval obtained from the greatest common factor is 1 year.
 
@@
 When involving multiple variables, one need to distinguish nesting and crossing data variables.
-"Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter; on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other.
+"Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`).
 Nesting is a special case of crossing.
-The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent); whereas the ordering of the latter makes no difference.
+The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference.
 Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data.
@@
 
-# Data pipeline
-
-This representation choice will prove suitable for almost all conceivable applications of temporal data.
-But the data abstraction cannot live alone with an additional toolchain in order to be able to acquire insight and knowledge about the data itself.
-A desirable toolbox chains transformation, visualisation, and modelling as a data science pipeline.
-Each component itself is also a self-contained pipeline that can be decomposed into many individual operations.
-This paper will focus on part of the general pipeline---data transformation, for example filtering observations, selecting and summarising variables.
-We shall see how this data abstraction breeds a consistent data process.
-
-## Literature review
+# Data pipeline {#sec:pipeline}
 
@@
 
-## Time-based pipeline analysis
-
 What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers them to focus on the statistical analysis tasks at hand without concerning the details of computational implementation.
@@
 
+## Time-based pipeline
+
+This "tidy" data representation choice will prove suitable for almost all conceivable applications of temporal data.
+But the data abstraction cannot live alone with an additional toolchain in order to be able to acquire insight and knowledge about the data itself.
+A desirable toolbox chains transformation, visualisation, and modelling as a data science pipeline.
+Each component itself is also a self-contained pipeline that can be decomposed into many individual operations.
+This paper will focus on part of the general pipeline---data transformation, for example filtering observations, selecting, and summarising variables.
+We shall see how this data abstraction breeds a consistent data process.
+
+## Expressive verbs
+
 An individual operation that performs on a specific task can be phrased as a verb.
-This verb is self-explanatory enough to express what the specific task is.
-We shall apply **dplyr** key verbs to time domain and expand its vocabulary a little more to handle time-related analysis problems.
+This verb is self-explanatory enough to express what it is supposed to do or fail.
+We shall adapt **dplyr** key verbs to time domain and expand its vocabulary a little more to handle time-related analysis problems.
 

### 2018-08-08 main.Rmd
updated some thoughts

@@
 Evidently, this data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
-less defaults to be made.
+less restrictions to be made.
 [R]
@@
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
+
 Table \@ref(tab:tb-sub) shows a "tidy" temporal data example using a subset of tuberculosis cases estimated by @tb-data.
@@
 Each observation uniquely records the number of people, who are diagnosed tuberculosis for each gender at three selected countries in the years of 2011 and 2012, where column `year` is declared as the index variable, columns `country`, `continent` and `gender` as key.
-To be more specific, the key is passed via `id(country | continent, gender)` or `id(gender, country | continent)`: a geographic hierarchy of country nested within continent, crossed with a demographic factor.
+To be more specific, the key is passed via `id(country | continent, gender)`, which produces a geographic hierarchy of country nested within continent, crossed with a demographic factor.
 Detailed discussions about index and key specifications will be given in Section \@ref(sec:index) and \@ref(sec:key).
@@
 Given the nature of temporal ordering, a temporal data set must be sorted by time index as a result.
-If a key is explicitly declared, the key will be sorted first depending on the ordering of factor levels and the ordering of variables, and followed by arranging time in ascending order.
-The tuberculosis example shows the key is sorted in alphabetic order, followed by ascending years by default, as is Table \@ref(tab:tb-sub).
+If a key is explicitly declared, the key will be sorted first depending on the ordering of factor levels and the ordering of variables, and followed by arranging time in past-to-future order.
+The tuberculosis example shows the key is sorted in alphabetic order, followed by ascending years by default, as in Table \@ref(tab:tb-sub).
 Since time-based operations, for example leading, lagging, or differencing series, require a vector in its time order, the arranged data format assures its validity for these operations without returning unexpected results.
@@
 `year`) ascends within each key element.
-If any operations on the tidy temporal data results a deviation from this baseline, it starts issuing a warning to the users.
+If any operations on the tidy temporal data results in a deviation from this baseline, it will issue a warning to the users.
 
@@
 Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
-The expression of a vertical bar (`|`) is used to indicate nesting variables, while a comma (`,`) for crossing variables.
+The expression of a vertical bar (`|`) (conditional) is used to indicate nesting variables, while a comma (`,`) for crossing variables.
 @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.
@@
 Each component itself is also a self-contained pipeline that can be decomposed into many individual operations.
-This paper will focus on part of the general pipeline---data transformation, for example filtering observations, selecting, and summarising variables.
+This paper will focus on part of the general pipeline---data transformation, for example filtering observations, selecting, and summarising variables, by contrast to data cleaning.
 We shall see how this data abstraction breeds a consistent data process.
 
-## Expressive verbs
+All the individual operations strive to keep the time context so that they will implicitly preserve the index variable.
+
+## Modularity and clarity
 
@@
 * **rolling window**: `slide()`, `tile()`, `stretch()`
+
+## Composition and transparency
 
@@
 
-# Application: U.S.A domestic flights on-time performance (2016-2017)
+# Case studies
+
+## U.S.A domestic flights on-time performance (2016-2017)
 
@@
 
+A new visualization framework for handling time series with nesting and crossing structure is under development.
+The proposed data structure proves a natural input that can be piped into a visualization program.
+As long as the program understands the embedded structure, it will produce sensible plots.
+
 The ground of time series modelling or forecasting is left untouched in this paper.
@@
 But time series models should be directly applied to tidy data as other wrangling tools do, without such an intermediate step.
-In particular, a univariate time series model, like arima and exponential smoothing, can be applied to multiple time series independently.
+In particular, a univariate time series model, like ARIMA and Exponential Smoothing, can be applied to multiple time series independently.
 A tidy format to represent model summaries and forecasting objects will be developed and implemented in the future.

### 2018-08-12 main.Rmd
rewrite the intro to ts

@@
 
-Temporal-context data consist of observational units indexed at different time points $X_{jt}$, where the $j$\textsuperscript{th} unit takes measurements of $X$ on over time $t$, for $j = 1, \dots, N_{i}$ and $1 \le t \le T$.
+Temporal-context data consist of observational units indexed at different time points $X_{jt}$, where the $j$\textsuperscript{th} in advance unit takes measurements of $X$ on over time $t$, for $j = 1, \dots, N_{i}$ and $1 \le t \le T$.
 Time primarily forms the contextual basis of temporal data, but it could arrive in many possible formats.
@@
 
-In the literature, time series and panel (longitudinal) data are common terms referred to temporal-context data, depending on the research fields.
-Researchers who are concerned with modelling large $T$ and small $N$ would name as "time series" (serial correlation); those who are interested in modelling small $T$ and large $N$ as "panel data" (asymptotic).
-The data format is two-dimensional array, but different modelling focuses lend the data input to different representations.
-A matrix is used to represent multivariate time series where each row represents observations measured at a time point and each column represents a series ("wide" form).
-This matrix representation requires homogeneity (that is, all the columns must be of same type.), time indices implicitly inferred as attributes or meta-information, series of same time length, and explicit missingness.
+## Time series and longitudinal data
+
+Time series and longitudinal data are the common terms when referring to temporal-context data.
+Despite of exactly the same data input, its representation diverges in these two areas due to different modelling focuses.
+Time series can be univariate or multivariate, but usually involves observations of relatively large in length possessed by each series.
+Time series researchers and analysts who are concerned with this large $T$ property, are mostly interested in modelling temporal dependency.
+Time series are represented as vectors or matrices in statistical software, for example MATLAB [@Matlab] and R [@R-base] (including `ts`, **zoo** [@zoo2005], and **xts** [@R-xts]).
+In the matrix case, each row is assumed to contain observations measured at a time point and each column to contain a single time series.
+This implies that data columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric cases.
+In addition to homogeneity, time indices must be stripped off the data and implicitly inferred as attributes or meta-information; the number of observations must be the same across all the series.
+Data wrangling from "raw" form to this specialist format can be frustrating and difficult, in order to perform a variety of downstream tasks such as analytics.
+
+On the other hand, those who are interested in modelling small $T$ and large $N$ as "panel data" (asymptotic).
 By contrast, panel data are organised in rectangular form of heterogeneous column types where multiple study subjects are stacked and repeated for its time indices in a single column ("long" form), due to commonly arisen unbalanced panels.
@@
 
+## Hierarchical and grouped time series
+
 Temporal data can often be aggregated in a manner that exhibits a nested or crossed structure, also known as hierarchical or grouped time series [@fpp].
 For example, in a manufacturing setting, a company can add up every store’s sales by region, by state and by country, which gives a strictly hierarchical time series; alternatively they can group the sales for each product together based on common attributes such as store, brand, price range and so forth, which leads to a non-hierarchical structure---a grouped time series.
+Additionally, hierarchical structure can be derived from statistical models that classifies the time series.
 The R package **hts** [@R-hts] is the implementation of this type of time series.
@@
 
-@wickham2014tidy coined the term "tidy data" and formalized the processing from messy data to "tidy data".
+## Tidy data
+
+@wickham2014tidy coined the term "tidy data", which is a rephrasing of the second and third normal forms in relational databases.
+and formalized the processing from messy data to "tidy data".
 These principles attempt to standardize the mapping from the semantics of a dataset to its structure and facilitate data analysis in a coherent way.
@@
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
+(The matrix representation also implies that every observation is determined by column variable and time-indexed row.)
 
@@
 
-The "key" identifies units or subjects that are observed over time in a data table, which are typically known in advance by users.
+The "key" identifies units or subjects that are observed over time in a data table, which is typically a priori known structure by users.
 The absence of key only occurs to a univariate case where the key is implicit.
@@
 Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
-The expression of a vertical bar (`|`) (conditional) is used to indicate nesting variables, while a comma (`,`) for crossing variables.
+The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables, while a comma (`,`) for crossing variables.
 @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.
@@
 
-What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers them to focus on the statistical analysis tasks at hand without concerning the details of computational implementation.
+What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the statistical analysis tasks at hand without concerning the details of computational implementation.
 A fluent and coherent pipeline glues the grammar of data manipulation and tidy data as a fundamental unit of data analysis together.
@@
 A new visualization framework for handling time series with nesting and crossing structure is under development.
-The proposed data structure proves a natural input that can be piped into a visualization program.
-As long as the program understands the embedded structure, it will produce sensible plots.
+The proposed data structure proves a natural input (unit time series and parent-children structure) that can be piped into a visualization program.
+As long as the program understands the embedded structure, it will produce sensible layout and interactivity.
 
@@
 ```{r write-bib}
-write_bib(c("hts", "plm", "dplyr"), file = "rpkgs.bib")
+write_bib(c("hts", "plm", "dplyr", "xts", "zoo"), file = "rpkgs.bib")
 ```

### 2018-08-14 main.Rmd
added explanations on longitudinal data

@@
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
-Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntactical and user-oriented approach in which it imposes nested or crossed structures on the data.
+Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntax-based and user-oriented approach in which it imposes nested or crossed structures on the data.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
@@
 Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage.
-The architecture of tidy temporal data has been implemented in the R package **tsibble**.'
+The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
 keywords: "temporal context, time series, data structure, R"
@@
 
-## Time series and longitudinal data
+## Representing time series and longitudinal data
 
@@
 Despite of exactly the same data input, its representation diverges in these two areas due to different modelling focuses.
-Time series can be univariate or multivariate, but usually involves observations of relatively large in length possessed by each series.
-Time series researchers and analysts who are concerned with this large $T$ property, are mostly interested in modelling temporal dependency.
+Time series can be univariate or multivariate, but usually involves observations of relatively large in length (i.e.
+large $T$) possessed by each series.
+Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic process, for the primary purpose of forecasting, temporal dynamics, and etc.
 Time series are represented as vectors or matrices in statistical software, for example MATLAB [@Matlab] and R [@R-base] (including `ts`, **zoo** [@zoo2005], and **xts** [@R-xts]).
-In the matrix case, each row is assumed to contain observations measured at a time point and each column to contain a single time series.
+The structure the multivariate time series are in is called **wide format** where each row is assumed to hold observations at a time point and each column to contain a single time series.
 This implies that data columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric cases.
@@
 
-On the other hand, those who are interested in modelling small $T$ and large $N$ as "panel data" (asymptotic).
-By contrast, panel data are organised in rectangular form of heterogeneous column types where multiple study subjects are stacked and repeated for its time indices in a single column ("long" form), due to commonly arisen unbalanced panels.
-This specification requires explicitly declared panel variable and index, which has been implemented in Stata's time series module and R package **plm** [@R-plm].
+On the other hand, those who are engaged in longitudinal design, are primarily interested in explaining variations among individuals in change process and making inference about the underlying population.
+Longitudinal data or panel data typically refers to fewer measurements (small $T$) over a large number of individuals (large $N$).
+It often occurs to longitudinal design that some individuals are not observed at certain points in time, resulting in an unbalanced panel.
+This data issue leads representing longitudinal data much more commonly in **long format** where each row holds an observational unit per time and each column contains measurements across the individuals for all points in time.
 Evidently, this data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
 less restrictions to be made.
-[R]
+This specification requires explicitly declared panel variable and index, which has been implemented in Stata's time series module and R package **plm** [@R-plm].
 

### 2018-08-16 main.Rmd
done with sec 1.1

@@
 
-## Representing time series and longitudinal data
+## Time series and longitudinal data
 
@@
 It often occurs to longitudinal design that some individuals are not observed at certain points in time, resulting in an unbalanced panel.
-This data issue leads representing longitudinal data much more commonly in **long format** where each row holds an observational unit per time and each column contains measurements across the individuals for all points in time.
-Evidently, this data organisation appears more flexible than matrix in supporting explicit time index, multiple subjects of different time lengths, and implicit missing values.
-less restrictions to be made.
-This specification requires explicitly declared panel variable and index, which has been implemented in Stata's time series module and R package **plm** [@R-plm].
+This data issue leads representing longitudinal data much more commonly in **long format** where each row corresponds to an observational unit per time and each column contains measurements across the individuals for all points in time.
+Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it in that wide format.
+Another appealing feature for a long format is data structured in a concise and semantic manner with reference to observations and variables.
+This subsequently supports perceiving time index as an explicit variable instead of an implicit attribute.
+In general, fewer restrictions are made if representing temporal-context data in long format.
 
-## Hierarchical and grouped time series
+Longitudinal data representation has been implemented in Stata's time series module [@Stata] and R package **plm** [@plm2008].
+The underpinning data structure is a two-dimensional column-homogeneous array.
+Specifying a longitudinal data set from a generic array needs explicitly declaring time and individuals (or panel variable in Stata's `tsset` command).
+Therefore, each row of the data can be identified by a specific individual and time point.
+Individuals, however, can be only declared through a single variable, not multiple.
+
+## Time series of nested and crossed factors
 
@@
 
-## Tidy data
+## Tidy data and the grammar of data manipulation
 

### 2018-08-17 main.Rmd
rewrite hts

@@
 Another appealing feature for a long format is data structured in a concise and semantic manner with reference to observations and variables.
-This subsequently supports perceiving time index as an explicit variable instead of an implicit attribute.
+This supports perceiving time index as an explicit variable instead of an implicit attribute, and subsequently opens the door to manipulating time such as extracting time components.
 In general, fewer restrictions are made if representing temporal-context data in long format.
@@
 Longitudinal data representation has been implemented in Stata's time series module [@Stata] and R package **plm** [@plm2008].
-The underpinning data structure is a two-dimensional column-homogeneous array.
+The underpinning data structure is a two-dimensional column-homogeneous array, as other tabular data.
 Specifying a longitudinal data set from a generic array needs explicitly declaring time and individuals (or panel variable in Stata's `tsset` command).
@@
 
-Temporal data can often be aggregated in a manner that exhibits a nested or crossed structure, also known as hierarchical or grouped time series [@fpp].
-For example, in a manufacturing setting, a company can add up every store’s sales by region, by state and by country, which gives a strictly hierarchical time series; alternatively they can group the sales for each product together based on common attributes such as store, brand, price range and so forth, which leads to a non-hierarchical structure---a grouped time series.
-Additionally, hierarchical structure can be derived from statistical models that classifies the time series.
+A collection of time series can often be arranged along multivariate data of nested and crossed factors, which is also known as hierarchical and grouped time series [@fpp].
+For example, from an operational point of view, a manufacturer can add up every store’s sales by region, by state, and by country along a geographical tree; alternatively, one can group the sales for each product together based on common attributes such as store, brand, price range and so forth, resulting in a non-hierarchical structure.
+The former gives an example of hierarchical time series, and the latter shows grouped time series.
+Hierarchical time series can also be derived using hierarchical clustering.
+
 The R package **hts** [@R-hts] is the implementation of this type of time series.

### 2018-08-19 main.Rmd
done with sec 1.2

@@
 
-A collection of time series can often be arranged along multivariate data of nested and crossed factors, which is also known as hierarchical and grouped time series [@fpp].
+A collection of time series can often be intrinsically arranged along multivariate data of nested and crossed factors, which is also known as hierarchical and grouped time series [@fpp].
 For example, from an operational point of view, a manufacturer can add up every store’s sales by region, by state, and by country along a geographical tree; alternatively, one can group the sales for each product together based on common attributes such as store, brand, price range and so forth, resulting in a non-hierarchical structure.
 The former gives an example of hierarchical time series, and the latter shows grouped time series.
-Hierarchical time series can also be derived using hierarchical clustering.
+And even when collections of time series are absent from hierarchical structure, it is possible to arrange them empirically into a hierarchy such as regression tree [@sankaran_interactive_2017].
 
-The R package **hts** [@R-hts] is the implementation of this type of time series.
-But it is frustrating to create and work with such data objects due to non-extensible and non-modified.
-Variational data aggregation (some groups no need aggregation).
-[R]
+The R package **hts** [@R-hts] is the implementation of structured collections of time series, and provides tools for visualizing, analyzing, and forecasting such data.
+But it is a pain to configure node list or grouping matrix as the increase in hierarchic depth and grouping factors.
+This is resulted from the fact that the structure specification is separated from data itself and represented numerically, which is rather counterintuitive from the user's perspective.
 
@@
 Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
-The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables, while a comma (`,`) for crossing variables.
+The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables (reflect depths), while a comma (`,`) for crossing variables.
 @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.

### 2018-08-20 main.Rmd
abstract update

@@
     email : rob.hyndman@monash.edu
-abstract: 'Temporal-context data is often rich with information and time formats, for example multiple observational units, different time lengths, heterogeneous data types, nested and crossed factors and etc.
+abstract: 'Mining temporal-context data for information is often inhibited by a multitude of time formats: irregular or multiple time intervals, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types, nested and crossed factors indicating hierarchical sub-groups.
+Time series models, in particular, the software supporting time series forecasting makes strict assumptions on data that needs to be provided, typically a matrix of numeric data with an implicit time index.
+Going from raw data to model-ready data is painful.
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
-Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntax-based and user-oriented approach in which it imposes nested or crossed structures on the data.
+Tidy data principles are applied, and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntax-based and user-oriented approach in which it imposes nested or crossed structures on the data.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
-A sound pipeline practice facilitates a succinct and transparent workflow for analyzing temporal data.
-Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage.
+A sound data pipeline facilitates a succinct and transparent workflow for analyzing temporal data.
+Applications are included to illustrate tidy temporal data structure, data pipeline structure and usage.
 The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'

### 2018-08-20 main.Rmd
tweaking first few sections

@@
 The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
-keywords: "temporal context, time series, data structure, R"
+keywords: "temporal data, time series, data structures, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis"
 wpnumber: no/yr
@@
 
-Temporal-context data consist of observational units indexed at different time points $X_{jt}$, where the $j$\textsuperscript{th} in advance unit takes measurements of $X$ on over time $t$, for $j = 1, \dots, N_{i}$ and $1 \le t \le T$.
-Time primarily forms the contextual basis of temporal data, but it could arrive in many possible formats.
-For example, data recorded at fine time resolutions (hours, minutes, and seconds) are typically associated with different time zones and daylight savings.
-Temporal data often carries with rich information other than the time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors involving nested or crossed structures, linking to other data tables, and etc.
+Temporal-context data, $X_{jt}$, consist of $j = 1, \dots, N_{i}$ observational units indexed at different time points, $1 \le t \le T$.
+Time forms the contextual basis of temporal data, but it can arrive in many possible formats, which makes preparing it for modeling and forecasting can be painful.
+For example, data can be recorded at various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
+It could be irregularly recorded, which is particularly true with longitudinal measurements like patient visits to a doctor's office.
+Temporal data also often contains with rich information other than the time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors involving nested or crossed structures.
 
@@
 
-Time series and longitudinal data are the common terms when referring to temporal-context data.
-Despite of exactly the same data input, its representation diverges in these two areas due to different modelling focuses.
-Time series can be univariate or multivariate, but usually involves observations of relatively large in length (i.e.
-large $T$) possessed by each series.
-Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic process, for the primary purpose of forecasting, temporal dynamics, and etc.
-Time series are represented as vectors or matrices in statistical software, for example MATLAB [@Matlab] and R [@R-base] (including `ts`, **zoo** [@zoo2005], and **xts** [@R-xts]).
-The structure the multivariate time series are in is called **wide format** where each row is assumed to hold observations at a time point and each column to contain a single time series.
-This implies that data columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric cases.
-In addition to homogeneity, time indices must be stripped off the data and implicitly inferred as attributes or meta-information; the number of observations must be the same across all the series.
-Data wrangling from "raw" form to this specialist format can be frustrating and difficult, in order to perform a variety of downstream tasks such as analytics.
+Despite of exactly the same data input, the representation of time series and longitudinal data diverges due to different modelling approaches.
 
-On the other hand, those who are engaged in longitudinal design, are primarily interested in explaining variations among individuals in change process and making inference about the underlying population.
-Longitudinal data or panel data typically refers to fewer measurements (small $T$) over a large number of individuals (large $N$).
-It often occurs to longitudinal design that some individuals are not observed at certain points in time, resulting in an unbalanced panel.
-This data issue leads representing longitudinal data much more commonly in **long format** where each row corresponds to an observational unit per time and each column contains measurements across the individuals for all points in time.
-Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it in that wide format.
-Another appealing feature for a long format is data structured in a concise and semantic manner with reference to observations and variables.
-This supports perceiving time index as an explicit variable instead of an implicit attribute, and subsequently opens the door to manipulating time such as extracting time components.
-In general, fewer restrictions are made if representing temporal-context data in long format.
+Time series can be univariate or multivariate, and for modelling requires relatively large in length (i.e.
+large $T$).
+Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic processes, for the primary purpose of forecasting, and characterising temporal dynamics.
+The time series supporting modeling are represented as vectors or matrices in most statistical software, with the standards being provided by the R [@R-base] packages `ts`, **zoo** [@zoo2005], and **xts** [@R-xts] (*Why include MATLAB [@Matlab]??*).
+Multivariate time series are typically assumed to be in the format where each row is assumed to hold observations at a time point and each column to contain a single time series.
+(The tidy data name for this would be **wide format**.) This implies that data are columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric variables.
+In addition, time indexes are stripped off the data and implicitly inferred as attributes or meta-information.
+It strictly requires that the number of observations must be the same across all the series.
+Data wrangling from the form that data arrives in, to this specialist format can be frustrating and difficult, inhibiting the variety of downstream tasks such as analytics.
+
+For longitudinal analysis, researchers and analysts are primarily interested in explaining trends across and variations among individuals, and making inference about a broader population.
+Longitudinal data or panel data typically assumes fewer measurements (small $T$) over a large number of individuals (large $N$).
+It often occurs that measurements for individuals are taken at different time points, resulting in an unbalanced panel.
+Thus, the primary format required for modeling is stacked series, blocks of measurements for each individual, with columns indicating individual, time of measurement and the measurements.
+(The tidy data name for this would be **long format**.) Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it in that wide format which would have missing values in many cells.
+A detriment of this format is that demographic information for subjects is often repeated for each time point.
+However, appealing feature is that data is structured in a concise and semantic manner with reference to observations and variables, with the time index stated explicitly.
+This opens the door to easily operating on time to make calculations and extract different temporal components, such as month and day of the week.
+It is conducive to examining the data in many different ways and leading to more comprehensive exploration and forecasting.
 

### 2018-08-20 main.Rmd
added a bit on tidy data

@@
 
-@wickham2014tidy coined the term "tidy data", which is a rephrasing of the second and third normal forms in relational databases.
-and formalized the processing from messy data to "tidy data".
-These principles attempt to standardize the mapping from the semantics of a dataset to its structure and facilitate data analysis in a coherent way.
-Based on the systematic structuring principles, the grammar of data manipulation, as implemented in the R package **dplyr** [@R-dplyr], abstracts an coherent and consistent set of methods to handle a wide range of data transformation tasks.
+@wickham2014tidy coined the term "tidy data", which is a rephrasing of the second and third normal forms in relational databases but in a way that makes more sense to data scientists.
+A suite of paired verbs such as `gather` and `spread` [@R-tidyr] were used to describe the reshape processing from messy data to tidy data.
+Tidy data, which standardized the mapping from the semantics of a dataset to its physical representation, serves remarkably as a fundamental unit of data analysis.
+R package **ggplot2** [@wickham2009ggplot2] pioneered developing the grammar of graphics based on tidy data, thus making it possible mapping from data space to visual elements.
+**dplyr** [@R-dplyr] generalised a coherent and consistent set of verbs to handle a wide range of data transformation tasks.
 @r4ds argues that 80% of data analysis tasks can be solved with tidy tools while the remaining 20% requires other tools.
@@
 ```{r write-bib}
-write_bib(c("hts", "plm", "dplyr", "xts", "zoo"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr"), file = "rpkgs.bib")
 ```

### 2018-08-22 main.Rmd
tweak wording in abstract

@@
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
-Tidy data principles are applied, and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntax-based and user-oriented approach in which it imposes nested or crossed structures on the data.
+Tidy data principles are applied, and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) including an explicitly declared index variable representing time, (3) incorporating a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntax-based and user-oriented approach in which it imposes nested or crossed structures on the data.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
-A sound data pipeline facilitates a succinct and transparent workflow for analyzing temporal data.
+A sound data pipeline facilitates a fluent and transparent workflow for analyzing temporal data.
 Applications are included to illustrate tidy temporal data structure, data pipeline structure and usage.

### 2018-09-03 main.Rmd
fixed typos in abstract and intro

@@
 abstract: 'Mining temporal-context data for information is often inhibited by a multitude of time formats: irregular or multiple time intervals, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types, nested and crossed factors indicating hierarchical sub-groups.
-Time series models, in particular, the software supporting time series forecasting makes strict assumptions on data that needs to be provided, typically a matrix of numeric data with an implicit time index.
+Time series models, in particular, the software supporting time series forecasting makes strict assumptions on data to be provided, typically a matrix of numeric data with an implicit time index.
 Going from raw data to model-ready data is painful.
@@
 Temporal-context data, $X_{jt}$, consist of $j = 1, \dots, N_{i}$ observational units indexed at different time points, $1 \le t \le T$.
-Time forms the contextual basis of temporal data, but it can arrive in many possible formats, which makes preparing it for modeling and forecasting can be painful.
+Time forms the contextual basis of temporal data, but it can arrive in many possible formats, which makes preparing it for modeling and forecasting painful.
 For example, data can be recorded at various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
 It could be irregularly recorded, which is particularly true with longitudinal measurements like patient visits to a doctor's office.
-Temporal data also often contains with rich information other than the time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors involving nested or crossed structures.
+Temporal data also often contains rich information in addition to time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors involving nested or crossed structures.
+The data problems are grouped into two types of analysis, time series and longitudinal.
 

### 2018-09-04 data-raw/flights.R
added flights data

@@ -1,0 +1,99 @@
+# src: https://github.com/hadley/nycflights13/blob/master/data-raw/flights.R
+# with minor tweaks
+# https://transtats.bts.gov/Prezip/On_Time_On_Time_Performance_2017_1.zip
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+
+flight_url <- function(year = 2017, month) {
+  base_url <- "http://www.transtats.bts.gov/Prezip/"
+  sprintf(paste0(base_url, "On_Time_On_Time_Performance_%d_%d.zip"), year, month)
+}
+
+download_month <- function(year = 2017, month) {
+  url <- flight_url(year, month)
+
+  temp <- tempfile(fileext = ".zip")
+  download.file(url, temp)
+
+  files <- unzip(temp, list = TRUE)
+  # Only extract biggest file
+  csv <- files$Name[order(files$Length, decreasing = TRUE)[1]]
+
+  unzip(temp, exdir = "data-raw/flights", junkpaths = TRUE, files = csv)
+
+  src <- paste0("data-raw/flights/", csv)
+  dst <- paste0("data-raw/flights/", year, "-", month, ".csv")
+  file.rename(src, dst)
+}
+
+download_year <- function(year = 2017) {
+  months <- 1:12
+  needed <- paste0(year, "-", months, ".csv")
+  missing <- months[!(needed %in% dir("data-raw/flights"))]
+  lapply(missing, download_month, year = year)
+}
+
+download_year(year = 2017)
+
+get_all <- function(path) {
+  col_types <- cols(
+    DepTime = col_integer(),
+    ArrTime = col_integer(),
+    CRSDepTime = col_integer(),
+    CRSArrTime = col_integer(),
+    Carrier = col_character(),
+    UniqueCarrier = col_character()
+  )
+  read_csv(path, col_types = col_types) %>%
+    filter(Cancelled == 0) %>% 
+    select(
+      year = Year, month = Month, day = DayofMonth,
+      dep_time = DepTime, sched_dep_time = CRSDepTime, dep_delay = DepDelay,
+      arr_time = ArrTime, sched_arr_time = CRSArrTime, arr_delay = ArrDelay,
+      carrier = Carrier,  flight_num = FlightNum, tailnum = TailNum,
+      origin = Origin, dest = Dest,
+      air_time = AirTime, distance = Distance,
+      origin_city_name = OriginCityName, origin_state = OriginState,
[…]

### 2018-09-04 R/demo.R
renamed some data files and scripts

@@ -3,7 +3,7 @@ library(tidyverse)
 library(tsibble)
 
 ## ---- tb-sub
-tb_small <- read_rds("data/tb_small.rds")
+tb_small <- read_rds("data/tb-small.rds")
 tb_small %>% 
   knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>% 
   kableExtra::kable_styling(position = "center")

### 2018-09-04 data-raw/tb-small.R

@@ -1,8 +1,6 @@
 library(tidyverse)
-library(lubridate)
-library(tsibble)
 
-tb <- read_csv("data/tb-2018-06-05.csv")
+tb <- read_csv("data-raw/tb-2018-06-05.csv")
 
 tb_small <- tb %>% 
   select(-region, -age) %>% 
@@ -14,4 +12,4 @@ tb_small <- tb %>%
   summarise(count = sum(count)) %>% 
   ungroup()
 
-write_rds(tb_small, "data/tb_small.rds")
+write_rds(tb_small, "data/tb-small.rds")

### 2018-09-04 R/flights.R
added airports data

@@ -1,0 +1,217 @@
+## ---- load-pkgs
+library(lubridate)
+library(tidyverse)
+library(tsibble)
+library(forcats)
+
+## ---- load-data
+flights <- read_rds("data/flights.rds")
+
+## ---- map-airlines
+# devtools::install_github("heike/ggmapr")
+library(ggmapr)
+origin_dest <- flights %>% 
+  distinct(origin, origin_state, dest, dest_state)
+airports <- read_rds("data/airports.rds")
+map_dat <- origin_dest %>% 
+  left_join(airports, by = c("origin" = "faa")) %>% 
+  rename(long = lon) %>% 
+  shift(origin_state == "HI", shift_by = c(52.5, 5.5)) %>%
+  scale(origin_state == "AK", scale = 0.3, set_to = c(-117, 27)) %>%
+  rename(origin_lat = lat, origin_lon = long)  %>% 
+  left_join(select(airports, faa, lon, lat), by = c("dest" = "faa")) %>% 
+  rename(long = lon) %>% 
+  shift(dest_state == "HI", shift_by = c(52.5, 5.5)) %>%
+  scale(dest_state == "AK", scale = 0.3, set_to = c(-117, 27)) %>%
+  rename(dest_lat = lat, dest_lon = long)
+
+states <- states %>%
+  shift(NAME == "Hawaii", shift_by = c(52.5, 5.5)) %>%
+  scale(NAME == "Alaska", scale = 0.3, set_to = c(-117, 27)) %>%
+  filter(lat > 20)
+
+ggplot() +
+  geom_polygon(data= states, aes(x = long, y = lat, group = group), 
+    fill = "white", colour = "grey60") +
+  geom_segment(data = map_dat, aes(
+    x = origin_lon, y = origin_lat, xend = dest_lon, yend = dest_lat
+  ), alpha = 0.2, size = 0.5, colour = "#762a83") +
+  geom_point(data = map_dat, aes(x = origin_lon, y = origin_lat), 
+    colour = "#f1a340", size  = 1.5) +
+  coord_map("albers", parameters = c(30, 45)) +
+  ggthemes::theme_map()
+
+## ---- glimpse
+glimpse(flights)
+
+## ---- print
+print(flights, width = 80)
+
+## ---- tsibble
+us_flights <- flights %>% 
+  as_tsibble(
[…]

### 2018-09-04 main.Rmd
added airline map

@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite"
+  fig.show = 'hold', cache = TRUE, external = TRUE
 )
 read_chunk('R/demo.R')
+read_chunk('R/flights.R')
 ```
@@
 
-## U.S.A domestic flights on-time performance (2016-2017)
+## On-time performance flights in U.S.A
 
-A dataset of on-time performance of domestic flights in U.S.A from 2016 to 2017 is studied and explored for illustration of tidy data and data pipeline.
+```{r load-flights-pkgs}
+```
+
+```{r load-flights}
+```
+
+```{r map-airlines}
+```
+
+A dataset of on-time performance of domestic flights in U.S.A in 2017 is studied and explored for illustration of tidy data and data pipeline.
 

### 2018-09-04 R/flights.R

@@ -1,10 +1,10 @@
-## ---- load-pkgs
+## ---- load-flights-pkgs
 library(lubridate)
 library(tidyverse)
 library(tsibble)
 library(forcats)
 
-## ---- load-data
+## ---- load-flights
 flights <- read_rds("data/flights.rds")
 
 ## ---- map-airlines
@@ -35,9 +35,9 @@ ggplot() +
     fill = "white", colour = "grey60") +
   geom_segment(data = map_dat, aes(
     x = origin_lon, y = origin_lat, xend = dest_lon, yend = dest_lat
-  ), alpha = 0.2, size = 0.5, colour = "#762a83") +
+  ), alpha = 0.1, size = 0.4, colour = "#762a83") +
   geom_point(data = map_dat, aes(x = origin_lon, y = origin_lat), 
-    colour = "#f1a340", size  = 1.5) +
+    colour = "#a6dba0", size  = 0.7) +
   coord_map("albers", parameters = c(30, 45)) +
   ggthemes::theme_map()
 

### 2018-09-04 main.Rmd
set appropriate fig size

@@
   fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
-  fig.show = 'hold', cache = TRUE, external = TRUE
+  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
+  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
 )

### 2018-09-04 R/flights.R

@@ -174,7 +174,7 @@ qtl_label <- c(
 hr_qtl %>% 
   filter(hour(dep_datehour) > 4) %>% 
   ggplot(aes(x = hour, y = dep_delay, group = date, colour = qtl)) +
-  geom_hline(yintercept = 0, colour = "#9ecae1", size = 2) +
+  geom_hline(yintercept = 15, colour = "#9ecae1", size = 2) +
   geom_line(alpha = 0.5) +
   facet_grid(
     qtl ~ wday, scales = "free_y", 

### 2018-09-05 main.Rmd
set up tsibble for flights

@@
 * **group-wise**: `index_by()`, `group_by()`
+* **reshape**: `gather()`, `spread()`
 * **time-wise**: `lead()`, `lag()`, `difference()`
@@
 
-## On-time performance flights in U.S.A
+## On-time performance for domestic flights in U.S.A
 
@@
 
-```{r map-airlines}
+```{r map-airlines, fig.cap = "(ref:map-airlines-cap)"}
 ```
 
-A dataset of on-time performance of domestic flights in U.S.A in 2017 is studied and explored for illustration of tidy data and data pipeline.
+(ref:map-airlines-cap) An overview of US airline traffic in 2017.
+Points indicate locations of airports, and lines the plane movements across the country.
+There are a few hubs connecting most airlines and sporadic airports.
+
+The dataset of 2017 on-time performance for US domestic flights represents temporal-context data caught in the wild, sourced from US Bureau of Transportation Statistics (reference).
+The data, arrived in much greater resolution, includes many aspects and measurements of every single flight at its scheduled departure time from performance metrics to flight routes.
+Table ??
+shows the millions of observations and 22 variables ranging from numerics to characters to date-times, which are intrinsically heterogeneous.
+Flights have their own scheduled departure times, resulting irregular space in time.
+This example also characterizes itself as event data, actions and states observed at the point in time.
+
+(Need to format into a table)
+
+```{r glimpse}
+```
+
+The data easily fits into the tsibble framework without the loss of data richness.
+To declare it to a tsibble, the setup is to specify `sched_dep_datetime` as "index", `flight` variable as "key", and `regular = FALSE`.
+A rigorous check is performed for validating the distinct rows determined by key and index.
+
+```{r tsibble, echo = TRUE}
+```
 

### 2018-09-06 main.Rmd
added two paragraphs into flight study

@@
 (ref:map-airlines-cap) An overview of US airline traffic in 2017.
-Points indicate locations of airports, and lines the plane movements across the country.
+Points indicate locations of airports, and lines the connecting routes across the country.
 There are a few hubs connecting most airlines and sporadic airports.
@@
 The data, arrived in much greater resolution, includes many aspects and measurements of every single flight at its scheduled departure time from performance metrics to flight routes.
+An overview of airports and routes is shown in Figure \@ref(fig:map-airlines).
 Table ??
 shows the millions of observations and 22 variables ranging from numerics to characters to date-times, which are intrinsically heterogeneous.
-Flights have their own scheduled departure times, resulting irregular space in time.
+Flights have their own scheduled departure times, resulting in irregular time space.
 This example also characterizes itself as event data, actions and states observed at the point in time.
@@
 
-The data easily fits into the tsibble framework without the loss of data richness.
+The data easily fits into the tsibble framework without the loss of data richness and heterogeneity.
 To declare it to a tsibble, the setup is to specify `sched_dep_datetime` as "index", `flight` variable as "key", and `regular = FALSE`.
@@
 ```
+
+On the basis of the tsibble, what question to be asked leads the way to composing a time-based pipeline in order for exploratory data analysis in time context.
+For example, one would be interested in the number of flights between origins and destinations every day, daily proportions of delayed flights by carriers, and etc.
+In this section, a given pipeline helps explore and demonstrate the distribution of departure delay conditional on time of day and day of week, particularly the upper tail behaviors.
+The accompanying code snippet is shown as follows.
+
+```{r quantile, echo = TRUE}
+```
+
+(Need a flow chart for the code)
+
+This pipeline chains four verbs.
+A set of succinct verbs makes the composition self-explanatory: (1) `index_by()` + `summarise()`: compute 50%, 80% and 95% quantiles for every hour; (2) `mutate()`: add time components that are extracted from the index; (3) `gather()`: melt different quantiles into a long form.
+Step (2) and (3) prepare the data for visualisation later.
+Every verb generates the output of the same type of the input, making chaining possible.
+Once the initial setup is done, the program takes care of updating the index and key, and empowers users concentrate on the substance of data analysis.
+
+Figure \@ref(fig:draw-qtl) is the plot based on the transformed data.
+
+```{r draw-qtl, fig.cap = "(ref:draw-qtl-cap)"}
+```
+
+(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceted by day of week and quantiles.
 

### 2018-09-06 R/flights.R

@@ -171,6 +171,11 @@ qtl_label <- c(
   "qtl95" = "95%" 
 )
 
+min_y <- hr_qtl %>% 
+  filter(hour(dep_datehour) > 4) %>% 
+  pull(dep_delay) %>%
+  min()
+
 hr_qtl %>% 
   filter(hour(dep_datehour) > 4) %>% 
   ggplot(aes(x = hour, y = dep_delay, group = date, colour = qtl)) +
@@ -184,7 +189,7 @@ hr_qtl %>%
   ylab("Depature delay") + 
   scale_x_continuous(limits = c(0, 23), breaks = seq(6, 23, by = 6)) +
   scale_colour_manual(values = break_cols, guide = FALSE) +
-  theme_remark()
+  expand_limits(y = min_y)
 
 ## ---- carrier
 carrier_delay <- us_flights %>% 

### 2018-09-10 main.Rmd
updated flights section

@@
 
-```{r load-flights}
+```{r load-flights, cache.extra = tools::md5sum("data/flights.rds")}
 ```
@@
 shows the millions of observations and 22 variables ranging from numerics to characters to date-times, which are intrinsically heterogeneous.
-Flights have their own scheduled departure times, resulting in irregular time space.
+Flights have their own scheduled departure times in local time zone, resulting in irregular time space.
 This example also characterizes itself as event data, actions and states observed at the point in time.
@@
 The data easily fits into the tsibble framework without the loss of data richness and heterogeneity.
-To declare it to a tsibble, the setup is to specify `sched_dep_datetime` as "index", `flight` variable as "key", and `regular = FALSE`.
+To declare it as a tsibble, the setup is to specify `sched_dep_datetime` as "index", `flight` and `origin` variables as "key", and `regular = FALSE`.
+The observation unit is defined by a flight at its origin.
 A rigorous check is performed for validating the distinct rows determined by key and index.
@@
 
-On the basis of the tsibble, what question to be asked leads the way to composing a time-based pipeline in order for exploratory data analysis in time context.
+On the basis of the tsibble, what question to be asked leads its way to composing a time-based pipeline in order for exploratory data analysis in time context.
 For example, one would be interested in the number of flights between origins and destinations every day, daily proportions of delayed flights by carriers, and etc.
-In this section, a given pipeline helps explore and demonstrate the distribution of departure delay conditional on time of day and day of week, particularly the upper tail behaviors.
+In this section, what time of the day and day of the week is good for travelling in order to avoid severe delay, from passenger's point of view?
+A given pipeline helps explore and demonstrate the distribution of departure delay conditional on time of day and day of week, particularly the upper tail behaviors.
 The accompanying code snippet is shown as follows.
@@
 Once the initial setup is done, the program takes care of updating the index and key, and empowers users concentrate on the substance of data analysis.
+Every one else who do not write the code would find the code fairly easy to comprehend.
 
-Figure \@ref(fig:draw-qtl) is the plot based on the transformed data.
+The data has been transformed in a way that the variables could be mapped to graphical elements.
+Figure \@ref(fig:draw-qtl) shows the plot of the transformed data using the grammar of graphics.
+It suggests to leave early in a day and fly on Tuesdays in a week to avoid severe delay.
 
@@
 (ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceted by day of week and quantiles.
+A blue horizontal line is drawn at 15-minute departure delay, which highlights that the departing flights who delay less than 15 minutes are considered as on time.
 

### 2018-09-10 R/flights.R

@@ -50,8 +50,7 @@ print(flights, width = 80)
 ## ---- tsibble
 us_flights <- flights %>% 
   as_tsibble(
-    index = sched_dep_datetime,
-    key = id(flight), 
+    index = sched_dep_datetime, key = id(flight, origin), 
     regular = FALSE
   )
 

### 2018-09-10 data-raw/flights.R

@@ -90,10 +90,4 @@ flights_full <- map_dfr(dir("data-raw/flights", full.names = TRUE), get_all)
 flights <- flights_full %>% 
   drop_na(air_time) # remove cancelled flights
 
-## rm duplicates due to corrupted data (my reasoning)
-dup <- flights %>% 
-  find_duplicates(key = id(flight), index = sched_dep_datetime)
-dup_df <- flights[dup, 1:2] # remove NK630 from DEN to LGA
-flights <- flights[!dup, ]
-
 write_rds(flights, path = "data/flights.rds", compress = "bz2")

### 2018-10-08 main.Rmd
minor tweak on code reasoning
added more plots to flights data

@@
 
+<!-- 
 ```{r map-airlines, fig.cap = "(ref:map-airlines-cap)"}
@@
 There are a few hubs connecting most airlines and sporadic airports.
+-->
 
@@
 
-```{r glimpse}
-```
-
 The data easily fits into the tsibble framework without the loss of data richness and heterogeneity.
@@
 
-```{r tsibble, echo = TRUE}
+```{r tsibble}
+```
+
+```{r carrier-delayed}
+```
+
+```{r carrier-mosaic}
+```
+
+```{r nyc-flights}
+```
+
+```{r nyc-delay}
+```
+
+```{r nyc-monthly-1}
+```
+
+```{r nyc-monthly-2}
+```
+
+```{r nyc-monthly-plot}
 ```
@@
 
-```{r quantile, echo = TRUE}
+```{r quantile}
 ```
@@
 Once the initial setup is done, the program takes care of updating the index and key, and empowers users concentrate on the substance of data analysis.
-Every one else who do not write the code would find the code fairly easy to comprehend.
+The one who write the code would easily reason about the code and every one else would find the code fairly easy to comprehend.
 
@@
 It suggests to leave early in a day and fly on Tuesdays in a week to avoid severe delay.
+
+```{r draw-qtl-prep}
+```
 

### 2018-10-08 R/flights.R

@@ -51,7 +51,7 @@ print(flights, width = 80)
 us_flights <- flights %>% 
   as_tsibble(
     index = sched_dep_datetime, key = id(flight, origin), 
-    regular = FALSE
+    regular = FALSE, validate = FALSE
   )
 
 ## ---- print-tsibble
@@ -85,48 +85,94 @@ delayed_carrier <- us_flights %>%
     Ontime = sum(delayed == 0),
     Delayed = sum(delayed)
   ) %>% 
-  mutate(carrier = reorder(carrier, - (Ontime + Delayed))) %>% 
   gather(delayed, n_flights, Ontime:Delayed)
 
+## ----- carrier-mosaic
 library(ggmosaic)
-ggplot(data = delayed_carrier) +
-  geom_mosaic(aes(x = product(carrier), fill = delayed, weight = n_flights)) +
-  scale_fill_brewer(palette = "Dark2", name = "Delayed") +
-  theme(legend.position = "bottom") +
-  xlab("Carrier") +
-  ylab("Delayed")
+delayed_carrier %>% 
+  mutate(carrier = fct_reorder(carrier, -n_flights)) %>% 
+  ggplot() +
+    geom_mosaic(aes(x = product(carrier), fill = delayed, weight = n_flights)) +
+    scale_fill_brewer(palette = "Dark2", name = "Delayed") +
+    scale_x_productlist(name = "Carrier") +
+    scale_y_productlist(name = "Delayed") +
+    theme(legend.position = "bottom")
 
-## ----- n-flights
-dep_delay_fct <- as_factor(c("ontime", "15-60 mins", "60+mins"))
-n_flights <- us_flights %>% 
-  mutate(
-    dep_delay_break = case_when(
-      dep_delay <= 15 ~ dep_delay_fct[1],
-      dep_delay <= 60 ~ dep_delay_fct[2],
-      TRUE ~ dep_delay_fct[3])
+## ---- nyc-flights
+nyc_flights <- us_flights %>% 
+  filter(origin %in% c("JFK", "LGA", "EWR"))
+
+## ---- nyc-delay
+nyc_delay <- nyc_flights %>% 
+  mutate(delayed = dep_delay > 15) %>% 
+  group_by(origin) %>% 
+  index_by(sched_dep_date = as_date(sched_dep_datetime)) %>% 
+  summarise(
+    n_flights = n(),
[…]

### 2018-10-11 main.Rmd
re-wrote the intro to the flight data

@@
 
+```{r map-prep}
+```
+
 <!-- 
@@
 
-The dataset of 2017 on-time performance for US domestic flights represents temporal-context data caught in the wild, sourced from US Bureau of Transportation Statistics (reference).
-The data, arrived in much greater resolution, includes many aspects and measurements of every single flight at its scheduled departure time from performance metrics to flight routes.
-An overview of airports and routes is shown in Figure \@ref(fig:map-airlines).
-Table ??
-shows the millions of observations and 22 variables ranging from numerics to characters to date-times, which are intrinsically heterogeneous.
-Flights have their own scheduled departure times in local time zone, resulting in irregular time space.
-This example also characterizes itself as event data, actions and states observed at the point in time.
+The dataset of 2017 on-time performance for US domestic flights represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics (reference).
+It contains `r format(NROW(flights), big.mark = ",")` operating flights (or observations) associated with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format.
+This kind of event describes each flight scheduled for departure at a local timestamp.
+But every single flight is identified by its flight number and origin, from a passenger's point of view.
+For example, the flight number DL771 is tied to multiple origins instead a specific one, shown in Figure \@ref(fig:dl771).
+This dataset is intrinsically heterogeneous, encoded in numbers, strings, and date-times.
 
-(Need to format into a table)
+```{r dl771, fig.cap = "(ref:dl771-cap)"}
+```
+
+(ref:dl771-cap) Multiple flight paths of flight DL771.
 

### 2018-10-11 R/flights.R

@@ -3,13 +3,13 @@ library(lubridate)
 library(tidyverse)
 library(tsibble)
 library(forcats)
+# devtools::install_github("heike/ggmapr")
+library(ggmapr)
 
 ## ---- load-flights
 flights <- read_rds("data/flights.rds")
 
-## ---- map-airlines
-# devtools::install_github("heike/ggmapr")
-library(ggmapr)
+## ---- map-prep
 origin_dest <- flights %>% 
   distinct(origin, origin_state, dest, dest_state)
 airports <- read_rds("data/airports.rds")
@@ -30,6 +30,7 @@ states <- states %>%
   scale(NAME == "Alaska", scale = 0.3, set_to = c(-117, 27)) %>%
   filter(lat > 20)
 
+## ---- map-airlines
 ggplot() +
   geom_polygon(data= states, aes(x = long, y = lat, group = group), 
     fill = "white", colour = "grey60") +
@@ -38,6 +39,27 @@ ggplot() +
   ), alpha = 0.1, size = 0.4, colour = "#762a83") +
   geom_point(data = map_dat, aes(x = origin_lon, y = origin_lat), 
     colour = "#a6dba0", size  = 0.7) +
+  coord_map("albers", parameters = c(30, 45)) +
+  ggthemes::theme_map()
+
+## ---- dl771
+dl771 <- flights %>% 
+  filter(flight_num == "DL771") %>% 
+  left_join(map_dat) %>% 
+  arrange(sched_dep_datetime)
+
+ggplot() +
+  geom_polygon(data= states, aes(x = long, y = lat, group = group), 
+    fill = "white", colour = "grey60") +
+  geom_segment(data = dl771, aes(
+    x = origin_lon, y = origin_lat, xend = dest_lon, yend = dest_lat
+  ), colour = "#762a83", size = 1, alpha = 0.6, 
+    arrow = arrow(length = unit(0.2, "inches"))
+  ) +
+  geom_point(data = dl771, aes(x = origin_lon, y = origin_lat), 
+    colour = "#f1a340", size  = 1.5) +
+  geom_point(data = dl771, aes(x = dest_lon, y = dest_lat), 
+    colour = "#f1a340", size  = 1.5) +
   coord_map("albers", parameters = c(30, 45)) +
   ggthemes::theme_map()
 
@@ -50,7 +72,7 @@ print(flights, width = 80)
 ## ---- tsibble
 us_flights <- flights %>% 
   as_tsibble(
[…]

### 2018-10-11 data-raw/flights.R

@@ -64,7 +64,7 @@ get_all <- function(path) {
       late_aircraft_delay = LateAircraftDelay
     ) %>%
     mutate(
-      flight = paste0(carrier, flight_num),
+      flight_num = paste0(carrier, flight_num),
       origin_city_name = gsub("^(.*?),.*", "\\1", origin_city_name),
       dest_city_name = gsub("^(.*?),.*", "\\1", dest_city_name),
       hour = sched_dep_time %/% 100,
@@ -76,11 +76,11 @@ get_all <- function(path) {
       sched_arr_datetime = make_datetime(year, month, day, hour_arr, minute_arr)
     ) %>% 
     select(
-      -year, -month, -day, -flight_num, -hour, -minute,
-      -dep_time, -sched_dep_time, -origin_state_name, -dest_state_name,
+      -year, -month, -day, -hour, -minute, -dep_time, 
+      -sched_dep_time, -origin_state_name, -dest_state_name,
       -arr_time, -sched_arr_time, -hour_arr, -minute_arr
     ) %>% 
-    select(flight, sched_dep_datetime, sched_arr_datetime, everything()) %>% 
+    select(flight_num, sched_dep_datetime, sched_arr_datetime, everything()) %>% 
     filter(!is.na(dep_delay), !(origin_state %in% c("PR", "TT", "VI")))
 }
 

### 2018-10-27 main.Rmd
rewrote flights tsibble structure

@@
   fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
-  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
+  fig.height = 5, fig.width = 8, out.width = "\\textwidth",
+  comment = "#>"
 )
@@
 The dataset of 2017 on-time performance for US domestic flights represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics (reference).
-It contains `r format(NROW(flights), big.mark = ",")` operating flights (or observations) associated with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format.
-This kind of event describes each flight scheduled for departure at a local timestamp.
-But every single flight is identified by its flight number and origin, from a passenger's point of view.
-For example, the flight number DL771 is tied to multiple origins instead a specific one, shown in Figure \@ref(fig:dl771).
-This dataset is intrinsically heterogeneous, encoded in numbers, strings, and date-times.
+It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format.
+This kind of event describes each flight scheduled for departure at a time point in its local time zone.
+But every single flight can be uniquely identified by its flight number and origin, from a passenger's point of view.
+For example, flight DL771 connects multiple origins and destinations rather than a specific route, shown in Figure \@ref(fig:dl771).
+This dataset is intrinsically heterogeneous, encoding in numbers, strings, and date-times.
 
@@
 
-(ref:dl771-cap) Multiple flight paths of flight DL771.
-
-The data easily fits into the tsibble framework without the loss of data richness and heterogeneity.
-To declare it as a tsibble, the setup is to specify `sched_dep_datetime` as "index", `flight` and `origin` variables as "key", and `regular = FALSE`.
-The observation unit is defined by a flight at its origin.
-A rigorous check is performed for validating the distinct rows determined by key and index.
+(ref:dl771-cap) Flight DL771 operates on multiple routes between different origins and destinations. It suggests that the key required for identification consists of flight number and origin.
 
@@
 
+The tsibble framework, as expected, incorporates this type of data, without the loss of data richness and heterogeneity.
+To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", columns `flight` and `origin` as "key" via `id(flight, origin)`.
+As a result of event data, this data is irregularly spaced, and hence switching `regular` to `FALSE` is necessary.
+The program internally validates if the key and index produce the distinct rows, and then sort the key and the index from past to recent.
+When the tsibble creation is done, the print display is data-oriented and contextually informative, such as dimensions, irregular interval (``r tsibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tsibble::tbl_sum(us_flights)[2]``).
+
+```{r tsibble-header}
+```
+
+Transforming tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can generate sound pipelines.
+From the perspective of a passenger, one need to travel smart, by choosing an efficient carrier to fly with and the time of day to avoid congestion, for example.
+We take a drill-down approach to exploring this data, starting with annual carrier performance and disaggregating to finer time resolutions.
+
 ```{r carrier-delayed}
@@
 
-```{r carrier-mosaic}
+```{r carrier-mosaic, fig.cap = "(ref:carrier-mosaic-cap)"}
 ```
+
+(ref:carrier-mosaic-cap) Mosaic plot showing the association between the size of airline carriers and the delayed proportion of departure in 2017.
+Southwest Airlines is the largest operator, but does not operate as efficient as Delta.
+Hawaiian Airlines, also as a small operator, performs the best.
 
@@
 ```
-
-On the basis of the tsibble, what question to be asked leads its way to composing a time-based pipeline in order for exploratory data analysis in time context.
-For example, one would be interested in the number of flights between origins and destinations every day, daily proportions of delayed flights by carriers, and etc.
-In this section, what time of the day and day of the week is good for travelling in order to avoid severe delay, from passenger's point of view?
-A given pipeline helps explore and demonstrate the distribution of departure delay conditional on time of day and day of week, particularly the upper tail behaviors.
-The accompanying code snippet is shown as follows.
 

### 2018-10-27 R/flights.R

@@ -76,8 +76,8 @@ us_flights <- flights %>%
     regular = FALSE, validate = FALSE
   )
 
-## ---- print-tsibble
-us_flights
+## ---- tsibble-header
+cat(format(us_flights)[1:2], sep = "\n")
 
 ## ---- filter
 us_flights %>% 

### 2018-10-27 main.Rmd
described the mosaic plot and its data pipeline

@@
 To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", columns `flight` and `origin` as "key" via `id(flight, origin)`.
-As a result of event data, this data is irregularly spaced, and hence switching `regular` to `FALSE` is necessary.
+As a result of event data, this data is irregularly spaced, and hence switching to irregular option is necessary.
 The program internally validates if the key and index produce the distinct rows, and then sort the key and the index from past to recent.
@@
 
-Transforming tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can generate sound pipelines.
+Transforming tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can generate well-constructed pipelines.
 From the perspective of a passenger, one need to travel smart, by choosing an efficient carrier to fly with and the time of day to avoid congestion, for example.
-We take a drill-down approach to exploring this data, starting with annual carrier performance and disaggregating to finer time resolutions.
+We take a drill-down approach to exploring this data, starting with annual carrier performance and followed by disaggregating to finer time resolutions.
+
+Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
+This pipeline is initialized defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually.
+To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column.
+The flow chart shown as Figure \@ref(fig:flight-pipeline1) demonstrates the operations undertaking in a data pipeline.
+The input to this pipeline is a tsibble of irregular interval, and the output ends with a tsibble of unknown interval.
+The final data is each carrier along with a single year, in which the interval cannot be determined.
+It also feeds into the mosaic display.
+Southwest Airlines, as the largest carrier, operates less efficiently compared to Delta.
 
@@
 Southwest Airlines is the largest operator, but does not operate as efficient as Delta.
-Hawaiian Airlines, also as a small operator, performs the best.
+Hawaiian Airlines, also as a small operator, outperforms the rest.
+
+```{r flight-pipeline1, fig.cap = "(ref:flight-pipeline1-cap)"}
+include_graphics("img/flight-pipeline1")
+```
+
+(ref:flight-pipeline1-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:carrier-mosaic}.
 
@@
 
-```{r nyc-monthly-plot}
+```{r nyc-monthly-plot, fig.cap = "(ref:nyc-monthly-plot-cap)"}
 ```
+
+(ref:nyc-monthly-plot-cap) Daily departure percentage at New York airports.
 
@@
 ```
-
-(Need a flow chart for the code)
 

### 2018-10-27 R/flights.R

@@ -135,37 +135,6 @@ nyc_delay <- nyc_flights %>%
   ) %>% 
   mutate(pct_delay = n_delayed / n_flights)
 
-## ---- nyc-delay-plot
-nyc_delay %>% 
-  ggplot(aes(x = sched_dep_date, y = pct_delay, colour = origin)) +
-  geom_line() +
-  facet_grid(origin ~ .) +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_colour_brewer(palette = "Dark2") +
-  xlab("Departure date") +
-  ylab("% of flight delay") +
-  theme(legend.position = "bottom")
-
-## ----- nyc-weekly-ma
-nyc_weekly <- nyc_delay %>% 
-  group_by(origin) %>% 
-  mutate(ma_delay = slide_dbl(
-    pct_delay, mean, .size = 7, .align = "center"
-  ))
-nyc_weekly %>% select(origin, ma_delay)
-  
-## ----- nyc-weekly-plot
-nyc_weekly %>% 
-  ggplot(aes(x = sched_dep_date)) +
-  geom_line(aes(y = pct_delay), colour = "grey80", size = 0.8) +
-  geom_line(aes(y = ma_delay, colour = origin), size = 1) +
-  facet_grid(origin ~ .) +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_colour_brewer(palette = "Dark2") +
-  theme(legend.position = "bottom") +
-  xlab("Date") +
-  ylab("Departure delay")
-
 ## ----- nyc-monthly-1
 nyc_lst <- nyc_delay %>% 
   mutate(yrmth = yearmonth(sched_dep_date)) %>% 
@@ -175,9 +144,9 @@ nyc_lst <- nyc_delay %>%
 ## ---- nyc-monthly-2
 nyc_monthly <- nyc_lst %>% 
   group_by(origin) %>% 
-  mutate(monthly_ma = slide_dbl(data, #<<
-    ~ mean(.$pct_delay), .size = 2, .bind = TRUE#<<
-  )) %>% #<<
+  mutate(monthly_ma = slide_dbl(data,
+    ~ mean(.$pct_delay), .size = 2, .bind = TRUE
+  )) %>%
   unnest(key = id(origin))
 
 ## ----- nyc-monthly-plot

### 2018-10-29 main.Rmd
added a paragraph to moving averages

@@
 
+Every verb generates the output of the same type of the input, making chaining possible.
+Once the initial setup is done, the program takes care of updating the index and key, and empowers users concentrate on the substance of data analysis.
+The one who write the code would easily reason about the code and every one else would find the code fairly easy to comprehend.
+
 # Case studies
@@
 
-Transforming tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can generate well-constructed pipelines.
+Transforming tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can result in well-constructed pipelines.
 From the perspective of a passenger, one need to travel smart, by choosing an efficient carrier to fly with and the time of day to avoid congestion, for example.
@@
 The flow chart shown as Figure \@ref(fig:flight-pipeline1) demonstrates the operations undertaking in a data pipeline.
-The input to this pipeline is a tsibble of irregular interval, and the output ends with a tsibble of unknown interval.
+The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval.
 The final data is each carrier along with a single year, in which the interval cannot be determined.
-It also feeds into the mosaic display.
-Southwest Airlines, as the largest carrier, operates less efficiently compared to Delta.
+It in turn feeds into the mosaic display.
+Note that Southwest Airlines, as the largest carrier, operates less efficiently compared to Delta, in Figure \@ref(fig:carrier-mosaic).
 
@@
 
+A closer examination of New York airports will give an indication about how well the busiest airports manage the outflow traffic on a daily basis.
+A subset that contains observations for EWR, JFK and LGA airports is obtained first.
+The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:nyc-monthly-plot).
+LGA fluctuates a lot compared to the other two.
+What superimposes on these lines is two-month moving averages so that a temporal trend is more visible.
+The number of days for each month is variable.
+Moving averages for two months call for computing weighted mean.
+But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without specifying weights.
+The sliding of large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
+
 ```{r nyc-flights}
@@
 
-(ref:nyc-monthly-plot-cap) Daily departure percentage at New York airports.
+(ref:nyc-monthly-plot-cap) Daily delayed percentages for departure with two-month moving averages overlaid at New York airports.
+There are many fluctuations observed at LGA airport.
+The estimates of temporal trend are around 25% across three airports, highlighting relatively less delay in Fall.
 
@@
 ```
-
-This pipeline chains four verbs.
-A set of succinct verbs makes the composition self-explanatory: (1) `index_by()` + `summarise()`: compute 50%, 80% and 95% quantiles for every hour; (2) `mutate()`: add time components that are extracted from the index; (3) `gather()`: melt different quantiles into a long form.
-Step (2) and (3) prepare the data for visualisation later.
-Every verb generates the output of the same type of the input, making chaining possible.
-Once the initial setup is done, the program takes care of updating the index and key, and empowers users concentrate on the substance of data analysis.
-The one who write the code would easily reason about the code and every one else would find the code fairly easy to comprehend.
 

### 2018-10-30 main.Rmd
added a paragragh to delayed quantiles

@@
 The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval.
-The final data is each carrier along with a single year, in which the interval cannot be determined.
+The final data is each carrier along with a single year, thereby the interval undetermined.
 It in turn feeds into the mosaic display.
@@
 Moving averages for two months call for computing weighted mean.
-But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without specifying weights.
-The sliding of large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
+But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without worrying weights specification.
+The sliding with large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
 
@@
 
+What time of day and day of week should we travel to avoid suffering from horrible delay?
+Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples.
+The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown.
+To reduce the likelihood of delay suffering, it is recommended to avoid the peak hour at 18.
+As moving towards the upper extremes, the variations considerably increase, making departure time unpredictable.
+
 ```{r quantile}
 ```
-
-The data has been transformed in a way that the variables could be mapped to graphical elements.
-Figure \@ref(fig:draw-qtl) shows the plot of the transformed data using the grammar of graphics.
-It suggests to leave early in a day and fly on Tuesdays in a week to avoid severe delay.
 
@@
 
-(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceted by day of week and quantiles.
-A blue horizontal line is drawn at 15-minute departure delay, which highlights that the departing flights who delay less than 15 minutes are considered as on time.
+(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceting day of week and 50%, 80% and 95% quantiles.
+A blue horizontal line indicates the 15-minute on-time standard to help grasp the delay severity.
+Passengers are apt to hold up around 18 during a day, and are recommended to travel early.
+The variations increase substantially as the upper tails.
 

### 2018-10-30 R/flights.R

@@ -9,6 +9,16 @@ library(ggmapr)
 ## ---- load-flights
 flights <- read_rds("data/flights.rds")
 
+# dup <- flights %>% 
+#   find_duplicates(key = id(flight_num), index = sched_dep_datetime)
+# dup_data <- flights[dup, ]
+#
+# flights %>% 
+#   filter(
+#     flight_num == dup_data$flight_num,
+#     sched_dep_datetime == dup_data$sched_dep_datetime
+#   ) %>% glimpse()
+
 ## ---- map-prep
 origin_dest <- flights %>% 
   distinct(origin, origin_state, dest, dest_state)

### 2018-11-01 main.Rmd
added duplicated data entry

@@
 
+It encourage users to look into the data as earlier as possible, examining the data quality issues.
+
 # Data pipeline {#sec:pipeline}
@@
 This kind of event describes each flight scheduled for departure at a time point in its local time zone.
-But every single flight can be uniquely identified by its flight number and origin, from a passenger's point of view.
-For example, flight DL771 connects multiple origins and destinations rather than a specific route, shown in Figure \@ref(fig:dl771).
-This dataset is intrinsically heterogeneous, encoding in numbers, strings, and date-times.
+Every single flight would be uniquely identified by the flight number and its scheduled departure time, from a passenger's point of view.
+In fact, it fails to pass the tsibble hurdle due to duplicates in the original data.
+An error is immediately raised when attempting to convert this data into a tsibble, and a closer inspection has to be carried out to locate the issue.
+The **tsibble** package provides the tool to easily find the duplicates in the data with `find_duplicates()`.
+Below shows the problematic entries.
+
+```{r find-duplicate}
+```
+
+The issue is perhaps introduced when updating or entering the data into a system.
+The same flight is scheduled at exactly the same time, together with the same performance statistics but different flight details, which is very unlikely.
+Flight NK630 is usually scheduled at 17:45 from Chicago to New York, searching into the whole records.
+A decision is made on removal of the first row from the duplicated entries before proceeding to the tsibble creation.
+
+<!--
 
@@
 
+-->
+
 ```{r tsibble}
@@
 
+This dataset is intrinsically heterogeneous, encoding in numbers, strings, and date-times.
 The tsibble framework, as expected, incorporates this type of data, without the loss of data richness and heterogeneity.
-To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", columns `flight` and `origin` as "key" via `id(flight, origin)`.
+To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", column `flight_num` as "key" via `id(flight_num)`.
 As a result of event data, this data is irregularly spaced, and hence switching to irregular option is necessary.
@@
 (ref:flight-pipeline1-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:carrier-mosaic}.
+
+[Change New York Airports to others to be able to see seasonal effects]
 

### 2018-11-01 R/flights.R

@@ -9,16 +9,6 @@ library(ggmapr)
 ## ---- load-flights
 flights <- read_rds("data/flights.rds")
 
-# dup <- flights %>% 
-#   find_duplicates(key = id(flight_num), index = sched_dep_datetime)
-# dup_data <- flights[dup, ]
-#
-# flights %>% 
-#   filter(
-#     flight_num == dup_data$flight_num,
-#     sched_dep_datetime == dup_data$sched_dep_datetime
-#   ) %>% glimpse()
-
 ## ---- map-prep
 origin_dest <- flights %>% 
   distinct(origin, origin_state, dest, dest_state)
@@ -73,16 +63,25 @@ ggplot() +
   coord_map("albers", parameters = c(30, 45)) +
   ggthemes::theme_map()
 
-## ---- glimpse
-glimpse(flights)
+## ---- find-duplicate
+dup <- flights %>% 
+  find_duplicates(
+    key = id(flight_num), index = sched_dep_datetime, fromLast = TRUE
+  )
+dup_entry <- flights %>% filter(dup)
 
-## ---- print
-print(flights, width = 80)
+flights %>% 
+  filter(
+    flight_num == dup_entry$flight_num,
+    sched_dep_datetime == dup_entry$sched_dep_datetime
+  ) %>% 
+  as.data.frame()
 
 ## ---- tsibble
 us_flights <- flights %>% 
+  filter(!dup) %>% 
   as_tsibble(
-    index = sched_dep_datetime, key = id(flight_num, origin), 
+    key = id(flight_num), index = sched_dep_datetime,
     regular = FALSE, validate = FALSE
   )
 

### 2018-11-05 main.Rmd
replaced nyc flights with national-wide 5 airports

@@
 
-It encourage users to look into the data as earlier as possible, examining the data quality issues.
+It encourage users to look into the data as earlier as possible, assessing the data quality issues.
 
@@
 
-[Change New York Airports to others to be able to see seasonal effects]
-
 A closer examination of New York airports will give an indication about how well the busiest airports manage the outflow traffic on a daily basis.
 A subset that contains observations for EWR, JFK and LGA airports is obtained first.
-The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:nyc-monthly-plot).
+The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:sel-monthly-plot).
 LGA fluctuates a lot compared to the other two.
@@
 
-```{r nyc-flights}
+```{r sel-flights}
 ```
 
-```{r nyc-delay}
+```{r sel-delay}
 ```
 
-```{r nyc-monthly-1}
+```{r sel-monthly-1}
 ```
 
-```{r nyc-monthly-2}
+```{r sel-monthly-2}
 ```
 
-```{r nyc-monthly-plot, fig.cap = "(ref:nyc-monthly-plot-cap)"}
+```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 6}
 ```
 
-(ref:nyc-monthly-plot-cap) Daily delayed percentages for departure with two-month moving averages overlaid at New York airports.
+(ref:sel-monthly-plot-cap) Daily delayed percentages for departure with two-month moving averages overlaid at five international airports.
 There are many fluctuations observed at LGA airport.
@@
 
+## Smart-Grid Customer Data in Australia
+
+<https://data.gov.au/dataset/smart-grid-smart-city-customer-trial-data>
+
 # Conclusion and future work

### 2018-11-05 R/flights.R

@@ -129,12 +129,12 @@ delayed_carrier %>%
     scale_y_productlist(name = "Delayed") +
     theme(legend.position = "bottom")
 
-## ---- nyc-flights
-nyc_flights <- us_flights %>% 
-  filter(origin %in% c("JFK", "LGA", "EWR"))
+## ---- sel-flights
+sel_flights <- us_flights %>% 
+  filter(origin %in% c("JFK", "SEA", "IAH", "KOA", "LAX"))
 
-## ---- nyc-delay
-nyc_delay <- nyc_flights %>% 
+## ---- sel-delay
+sel_delay <- sel_flights %>% 
   mutate(delayed = dep_delay > 15) %>% 
   group_by(origin) %>% 
   index_by(sched_dep_date = as_date(sched_dep_datetime)) %>% 
@@ -144,22 +144,22 @@ nyc_delay <- nyc_flights %>%
   ) %>% 
   mutate(pct_delay = n_delayed / n_flights)
 
-## ----- nyc-monthly-1
-nyc_lst <- nyc_delay %>% 
+## ----- sel-monthly-1
+sel_lst <- sel_delay %>% 
   mutate(yrmth = yearmonth(sched_dep_date)) %>% 
   group_by(origin, yrmth) %>% 
   nest()
 
-## ---- nyc-monthly-2
-nyc_monthly <- nyc_lst %>% 
+## ---- sel-monthly-2
+sel_monthly <- sel_lst %>% 
   group_by(origin) %>% 
   mutate(monthly_ma = slide_dbl(data,
     ~ mean(.$pct_delay), .size = 2, .bind = TRUE
   )) %>%
   unnest(key = id(origin))
 
-## ----- nyc-monthly-plot
-nyc_monthly %>% 
+## ----- sel-monthly-plot
+sel_monthly %>% 
   ggplot() +
   geom_line(aes(x = sched_dep_date, y = pct_delay), colour = "grey80", size = 0.8) +
   geom_line(aes(x = yrmth, y = monthly_ma, colour = origin), size = 1) +

### 2018-11-05 main.Rmd
switched out nesting and crossing

@@
     email : rob.hyndman@monash.edu
-abstract: 'Mining temporal-context data for information is often inhibited by a multitude of time formats: irregular or multiple time intervals, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types, nested and crossed factors indicating hierarchical sub-groups.
+abstract: 'Mining temporal-context data for information is often inhibited by a multitude of time formats: irregular or multiple time intervals, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
 Time series models, in particular, the software supporting time series forecasting makes strict assumptions on data to be provided, typically a matrix of numeric data with an implicit time index.
@@
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
-Tidy data principles are applied, and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) including an explicitly declared index variable representing time, (3) incorporating a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntax-based and user-oriented approach in which it imposes nested or crossed structures on the data.
+Tidy data principles are applied, and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) including an explicitly declared index variable representing time, (3) incorporating a "key" comprised of single or multiple variables to uniquely identify units over time.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
@@
 
+<!--
 ## Time series of nested and crossed factors
@@
 This is resulted from the fact that the structure specification is separated from data itself and represented numerically, which is rather counterintuitive from the user's perspective.
+-->
 
@@
 The "key" serves a similar purpose as the panel variable in the Stata's **tsset** command to define the units or subjects, but it is expanded to include multiple variables rather than a single one.
-A syntax is introduced to express a key consisting of nested and crossed factors.
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
@@
 Each observation uniquely records the number of people, who are diagnosed tuberculosis for each gender at three selected countries in the years of 2011 and 2012, where column `year` is declared as the index variable, columns `country`, `continent` and `gender` as key.
-To be more specific, the key is passed via `id(country | continent, gender)`, which produces a geographic hierarchy of country nested within continent, crossed with a demographic factor.
 Detailed discussions about index and key specifications will be given in Section \@ref(sec:index) and \@ref(sec:key).

### 2018-11-06 main.Rmd
added new castle smart meter data 2013

@@
 
-## Time index and interval {#sec:index}
+## Index {#sec:index}
 
@@
 
+## Key {#sec:key}
+
+The "key" specification contributes to forming a tsibble alongside the index.
+The concept of "key" is introduced to identify units or subjects that are observed over time in a data table.
+In the wide format, each column holds a series of values, so that the column itself serves for the sake of identification.
+In the long format, all column names are melt to the corresponding "key" values.
+But the "key" is much more flexible than simply column names.
+Because it is not constrained to a single field, but can be comprised of multiple fields.
+The variables remain the same as it is, without additional manipulation to create the identifiers.
+
+It is typically a priori known by users.
+The absence of key only occurs to a univariate case where the key is implicit.
+Multiple units must associate with an explicit "key" to allow for identification.
+
+It encourage users to look into the data as earlier as possible, assessing the data quality issues.
+
+## Interval {#sec:interval}
+
 For data indexed in regular time space, the time interval is obtained by computing the greatest common divisor based on the non-negative differences of a given time index, although the data may include implicit time gaps.
@@
 Extensible. Custom index classes can be supported.
-
-## Key {#sec:key}
-
-The "key" identifies units or subjects that are observed over time in a data table, which is typically a priori known structure by users.
-The absence of key only occurs to a univariate case where the key is implicit.
-Multiple units must associate with an explicit "key" to allow for identification.
-The key is not constrained to a single column, but is comprised of multiple columns.
-When involving multiple variables, one need to distinguish nesting and crossing data variables.
-"Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`).
-Nesting is a special case of crossing.
-The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference.
-Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data.
-Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
-The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables (reflect depths), while a comma (`,`) for crossing variables.
-@wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.
-A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout.
-However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.
-
-It encourage users to look into the data as earlier as possible, assessing the data quality issues.
 
@@
 A tidy format to represent model summaries and forecasting objects will be developed and implemented in the future.
-Model summaries include coefficients, fitted values, and residuals; forecasting objects include future time path and distributions generating prediction intervals.
+Model summaries include coefficients, fitted values, and residuals; forecasting objects.
+
+When involving multiple variables, one need to distinguish nesting and crossing data variables.
+"Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`).
+Nesting is a special case of crossing.
+The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference.
+Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data.
+Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
+The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables (reflect depths), while a comma (`,`) for crossing variables.
+@wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.
+A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout.
+However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.nclude future time path and distributions generating prediction intervals.
 

### 2018-11-08 main.Rmd
new section on "Date structures"

@@
 The data problems are grouped into two types of analysis, time series and longitudinal.
+
+# Data structures {#sec:str}
 
@@
 
+It encourage users to look into the data as earlier as possible, assessing the data quality issues.
+
 ```{r tb-sub}
@@
 
-The "key" specification contributes to forming a tsibble alongside the index.
+The "key" specification contributes to creating a tsibble alongside the index.
 The concept of "key" is introduced to identify units or subjects that are observed over time in a data table.
@@
 Because it is not constrained to a single field, but can be comprised of multiple fields.
-The variables remain the same as it is, without additional manipulation to create the identifiers.
+The identifying variables that the "key" is constituted of, remain the same as they are in the original table, with no additional manipulation.
 
-It is typically a priori known by users.
-The absence of key only occurs to a univariate case where the key is implicit.
-Multiple units must associate with an explicit "key" to allow for identification.
-
-It encourage users to look into the data as earlier as possible, assessing the data quality issues.
+Each tsibble must attach with a "key".
+It is normally a priori known by analysts.
+For example, Table \ref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year.
+This data description suggests that columns `gender` and `country` have to be declared as the key.
+Lacking in either of two will be inadequate for the key and thus fail to construct a tsibble.
+The key is explicit when multiple units exist in the data.
+Key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.
 

### 2018-11-09 main.Rmd
updated data

@@
 The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
-(The matrix representation also implies that every observation is determined by column variable and time-indexed row.)
 
@@
 
-Time forms an integral component and a contextual basis of temporal data.
-A variable representing time needs explicitly declared at the process of constructing a temporal data, referred to as "index".
+Time provides a contextual basis for temporal data. The time could range from date-times (such as `POSIXct` and `Date` in R) to year-months, to years.
+
+A variable representing time is indispensable to a tsibble, referred to as "index".
+The "index" is an intact column instead of an attribute.
 This accessibility of index promotes transparency and unambiguity while manipulating time.

### 2018-11-09 R/smart-meter.R

@@ -1,0 +1,44 @@
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+
+customer <- read_rds("data/customer_newcastle.rds")
+eudm <- read_rds("data/eudm13_newcastle.rds")
+
+eudm_ts <- eudm %>% 
+  build_tsibble(
+    key = id(customer_id), index = reading_datetime,
+    validate = FALSE, ordered = TRUE
+  )
+
+gap_df <- has_gaps(eudm_ts)
+
+gap_df %>% 
+  ggplot(aes(x = factor(1), fill = .gaps)) +
+  geom_bar(position = "fill")
+
+customer_na <- eudm_ts %>% 
+  filter(customer_id %in% (gap_df %>% filter(.gaps) %>% pull(customer_id)))
+
+count_na_df <- customer_na %>% 
+  count_gaps()
+
+count_na_10 <- count_na_df %>% 
+  count(customer_id) %>% 
+  top_n(10) %>% 
+  pull(customer_id)
+
+count_na_df %>% 
+  filter(customer_id %in% count_na_10) %>% 
+  ggplot(aes(x = as.factor(customer_id))) +
+  geom_linerange(aes(ymin = .from, ymax = .to)) +
+  geom_point(aes(y = .from)) +
+  geom_point(aes(y = .to)) +
+  coord_flip() +
+  theme(legend.position = "bottom")
+
+eudm_full <- eudm_ts %>% 
+  fill_na()
+
+eudm_na <- eudm_full %>% 
+  filter(is.na(general_supply_kwh))

### 2018-11-11 main.Rmd
rewrote index section

@@
 (ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012, with 12 observations and 5 variables.
-The index refers to column `year`, the key to multiple columns: `contry` nested under `continent` crossed with `gender`, and the measured variable to column `count`.
+The index refers to column `year`, the key to multiple columns: `country` nested under `continent` crossed with `gender`, and the measured variable to column `count`.
 
@@
 
-Time provides a contextual basis for temporal data. The time could range from date-times (such as `POSIXct` and `Date` in R) to year-months, to years.
+Time provides a contextual basis for temporal data. Time representation could range from date-times (such as `POSIXct` and `Date` in R) to years, to sequential numbers for simulated data for example.
 
 A variable representing time is indispensable to a tsibble, referred to as "index".
-The "index" is an intact column instead of an attribute.
-This accessibility of index promotes transparency and unambiguity while manipulating time.
-For example, subsetting a certain time period of data, extracting time components (like time of day and day of week), or converting time zones are directly dealt with index.
-It is also often to join other data tables based on the common time indices.
-Tools are provided to work with time itself, rather than wrappers to work with a whole data set.
-This promotes transparency and visibility to make inspection easier.
-This sets the tone on methods design philosophy: well-structured, expressive and unambiguous workflow.
+The "index" is an intact data column rather than an attribute, which makes time accessible to users.
+For example, one could easily extract time components, such as time of day and day of week, from the index to visualise seasonal effects of response variables.
+One could also join other data sources to the tsibble based on common time indices.
+The accessibility of tsibble index motivates data analysis towards transparency and human readability.
+But if the "index" is employed as attributes, analysts would write these simple queries in a programmatic manner, which can be avoidable from an analytic point of view.
 

### 2018-11-12 main.Rmd
restarted time-based pipeline

@@
 # Introduction {#sec:intro}
+
+```{r pipeline-ann}
+include_graphics("img/time-series-structure.pdf")
+```
 
@@
 
+Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
+
 ## Key {#sec:key}
@@
 
-Since Table \@ref(tab:tb-sub) possesses complete observations for each category, the interval obtained from the greatest common factor is 1 year.
-
-Extensible. Custom index classes can be supported.
-
 # Data pipeline {#sec:pipeline}
@@
 
+The tidy data abstraction lays the foundation for data science modules including transformation, visualisation and modelling.
+Each module frictionlessly communicates between each other.
+Because a desirable toolbox acquires tidy input, produces tidy output, and consequently chains a series of operations together to accomplish analytic tasks at hand.
+
 What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the statistical analysis tasks at hand without concerning the details of computational implementation.
-A fluent and coherent pipeline glues the grammar of data manipulation and tidy data as a fundamental unit of data analysis together.
+A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together.
 It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.
@@
 
-This "tidy" data representation choice will prove suitable for almost all conceivable applications of temporal data.
-But the data abstraction cannot live alone with an additional toolchain in order to be able to acquire insight and knowledge about the data itself.
-A desirable toolbox chains transformation, visualisation, and modelling as a data science pipeline.
-Each component itself is also a self-contained pipeline that can be decomposed into many individual operations.
-This paper will focus on part of the general pipeline---data transformation, for example filtering observations, selecting, and summarising variables, by contrast to data cleaning.
-We shall see how this data abstraction breeds a consistent data process.
+This tidy data representation choice will prove suitable for almost all conceivable applications of temporal data.
+Time-based pipeline starts with raw data.
+It is advised to scrutinize identical entries of key and index in the right beginning.
+Duplicates signal the data quality issue, which would likely affect the following analyses and hence decision making.
+Analysts are encouraged to gaze at data as early as possible and reason about the process of data cleaning.
+When the data meets the tsibble standard, it flows into the exploration stage.
 
-All the individual operations strive to keep the time context so that they will implicitly preserve the index variable.
+Time series models typically assume that the input is a complete and regularly-spaced series, which most temporal data can barely satisfy.
+Temporal data becomes more granular in time resolution and more disaggregated to individual level.
+This inevitably goes with implicit missing values and noisiness.
+Data needs transformation to some extent for modelling.
+A suite of verbs are introduced to flatten the lumpy path from temporal data to the object that directs to modelling under the umbrella of tsibble structure.
+
+All the individual operations strive to keep the time context so that they will implicitly preserve the index variable, and maintain a valid tsibble all the time.
 
@@
 
-* **row-wise**: `filter()`, `slice()`, `arrange()`, `fill_na()`
+* **implicit missingness**: `has_gaps()`, `count_gaps()`, `fill_na()`
+* **row-wise**: `filter()`, `slice()`, `arrange()`
 * **column-wise**: `mutate()`, `select()`, `summarise()`
 * **group-wise**: `index_by()`, `group_by()`
-* **reshape**: `gather()`, `spread()`
-* **time-wise**: `lead()`, `lag()`, `difference()`
+* **reshape**: `gather()`, `spread()`, `nest()`, `unnest()`
 * **rolling window**: `slide()`, `tile()`, `stretch()`

### 2018-11-12 R/smart-meter.R
started smart meter data analysis

@@ -2,22 +2,23 @@ library(tidyverse)
 library(lubridate)
 library(tsibble)
 
-customer <- read_rds("data/customer_newcastle.rds")
-eudm <- read_rds("data/eudm13_newcastle.rds")
+customer <- read_rds("data/customer.rds")
+elec <- read_rds("data/smart-meter13.rds")
 
-eudm_ts <- eudm %>% 
+elec_ts <- elec %>% 
   build_tsibble(
     key = id(customer_id), index = reading_datetime,
     validate = FALSE, ordered = TRUE
   )
 
-gap_df <- has_gaps(eudm_ts)
+gap_df <- has_gaps(elec_ts)
 
 gap_df %>% 
   ggplot(aes(x = factor(1), fill = .gaps)) +
-  geom_bar(position = "fill")
+  geom_bar(position = "fill") +
+  coord_flip()
 
-customer_na <- eudm_ts %>% 
+customer_na <- elec_ts %>% 
   filter(customer_id %in% (gap_df %>% filter(.gaps) %>% pull(customer_id)))
 
 count_na_df <- customer_na %>% 
@@ -37,8 +38,34 @@ count_na_df %>%
   coord_flip() +
   theme(legend.position = "bottom")
 
-eudm_full <- eudm_ts %>% 
+elec_full <- elec_ts %>% 
   fill_na()
 
-eudm_na <- eudm_full %>% 
+elec_na <- elec_full %>% 
   filter(is.na(general_supply_kwh))
+
+elec_na_mth <- elec_na %>% 
+  mutate(
+    mth = month(reading_datetime, label = TRUE),
+    is_na = 1L
+  ) %>% 
+  group_by(customer_id) %>% 
+  index_by(mth) %>% 
+  summarise(n_na = sum(is_na))
+
+elec_na_mth %>% 
+  ggplot(aes(x = mth)) +
+  geom_bar()
+
+elec_gen <- elec_full %>% 
+  left_join(
+    customer %>% select(customer_key, has_gas, has_aircon),
+    by = c("customer_id" = "customer_key")
+  )
+
+elec_gen %>% 
+  as_tibble() %>% 
+  group_by(has_gas, has_aircon) %>% 
+  summarise(avg_kwh = mean(general_supply_kwh, na.rm = TRUE))
+
+names(customer)

### 2018-11-14 main.Rmd
done with section "time-based pipeline"

@@
 
-The tidy data abstraction lays the foundation for data science modules including transformation, visualisation and modelling.
+The tidy data abstraction lays the foundation for data science modules including transformation, visualization and modelling.
 Each module frictionlessly communicates between each other.
@@
 
-What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the statistical analysis tasks at hand without concerning the details of computational implementation.
+What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the data analysis hand without concerning the details of computational implementation.
 A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together.
@@
 Duplicates signal the data quality issue, which would likely affect the following analyses and hence decision making.
-Analysts are encouraged to gaze at data as early as possible and reason about the process of data cleaning.
+Analysts are encouraged to gaze at data as earlier as possible and reason about the process of data cleaning.
 When the data meets the tsibble standard, it flows into the exploration stage.
@@
 Data needs transformation to some extent for modelling.
-A suite of verbs are introduced to flatten the lumpy path from temporal data to the object that directs to modelling under the umbrella of tsibble structure.
-
-All the individual operations strive to keep the time context so that they will implicitly preserve the index variable, and maintain a valid tsibble all the time.
+A suite of verbs are introduced to flatten the lumpy path from temporal data to the object that directs to modelling under the tsibble framework, as well as to facilitate transforming tsibble in various kinds of shapes for analysis and visualization.
+The principle that underpins every single verb is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood.
+If a tsibble cannot hold, for example swiping index off, an error prompts users to complain and suggest alternatives.
+This warrants users least surprise and reminds them of time awareness.
 

### 2018-11-14 R/flights.R

@@ -66,7 +66,7 @@ ggplot() +
 ## ---- find-duplicate
 dup <- flights %>% 
   find_duplicates(
-    key = id(flight_num), index = sched_dep_datetime, fromLast = TRUE
+    key = id(flight_num), index = sched_dep_datetime, from_last = TRUE
   )
 dup_entry <- flights %>% filter(dup)
 

### 2018-11-14 R/smart-meter.R

@@ -26,7 +26,7 @@ count_na_df <- customer_na %>%
 
 count_na_10 <- count_na_df %>% 
   count(customer_id) %>% 
-  top_n(10) %>% 
+  top_n(100) %>% 
   pull(customer_id)
 
 count_na_df %>% 
@@ -38,10 +38,10 @@ count_na_df %>%
   coord_flip() +
   theme(legend.position = "bottom")
 
-elec_full <- elec_ts %>% 
+elec_ts <- elec_ts %>% 
   fill_na()
 
-elec_na <- elec_full %>% 
+elec_na <- elec_ts %>% 
   filter(is.na(general_supply_kwh))
 
 elec_na_mth <- elec_na %>% 
@@ -57,15 +57,65 @@ elec_na_mth %>%
   ggplot(aes(x = mth)) +
   geom_bar()
 
-elec_gen <- elec_full %>% 
-  left_join(
-    customer %>% select(customer_key, has_gas, has_aircon),
+elec_ts <- elec_ts %>% 
+  mutate(season = if_else(
+    reading_datetime < make_datetime(2013, 4) |
+    reading_datetime >= make_datetime(2013, 10),
+    "Autumn-Winter", "Spring-Summer"
+  ))
+
+qtl_grid <- seq(0.1, 0.9, 0.01)
+gas_aircon <- elec_ts %>% 
+  inner_join(
+    customer %>% 
+      select(customer_key, has_gas, has_aircon) %>% 
+      drop_na(has_gas, has_aircon),
     by = c("customer_id" = "customer_key")
-  )
+  ) %>% 
+  index_by(hms = hms::as.hms(reading_datetime)) %>% 
+  group_by(season, has_gas, has_aircon) %>% 
+  summarise(
+    value = list(quantile(general_supply_kwh, qtl_grid, na.rm = TRUE)),
+    qtl = list(qtl_grid)
+  ) %>% 
+  unnest(key = id(qtl))
 
-elec_gen %>% 
-  as_tibble() %>% 
+gas_aircon %>% 
+  ggplot(aes(x = hms, y = value, colour = qtl, group = qtl)) +
+  geom_line() +
+  facet_grid(has_aircon ~ has_gas + season, labeller = "label_both") +
+  scale_colour_viridis_c()
+
+gas_aircon_avg <- elec_ts %>% 
+  inner_join(
+    customer %>% 
+      select(customer_key, has_gas, has_aircon) %>% 
[…]

### 2018-11-14 main.Rmd
started on section 4.2

@@
 
-```{r pipeline-ann}
-include_graphics("img/time-series-structure.pdf")
-```
-
 Temporal-context data, $X_{jt}$, consist of $j = 1, \dots, N_{i}$ observational units indexed at different time points, $1 \le t \le T$.
@@
 The data problems are grouped into two types of analysis, time series and longitudinal.
+
+```{r pipeline-ann}
+include_graphics("img/time-series-structure.pdf")
+```
+
+The rest of the paper is structured as follows.
 
@@
 A chain of data transformations lend itself to a data pipeline.
-
-The rest of the paper is structured as follows.
 
@@
 This tidy data representation choice will prove suitable for almost all conceivable applications of temporal data.
-Time-based pipeline starts with raw data.
+Time-based pipeline starts with raw data however it lands.
 It is advised to scrutinize identical entries of key and index in the right beginning.
@@
 
-## Modularity and clarity
+## Clarity and transparency
 
-An individual operation that performs on a specific task can be phrased as a verb.
+Transformation assembles a set of modules.
+A unit that makes up a module is function or preferably "verb".
+A tsibble is an object, considered as a conceptualized and materialized noun, an action performed on the object can be phrased as a verb.
+Naming verbs is difficult.
 This verb is self-explanatory enough to express what it is supposed to do or fail.
@@
 
-## Composition and transparency
+## Composition and modularity
 

### 2018-11-17 main.Rmd
added transformation verbs paragraph

@@
 
-What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the data analysis hand without concerning the details of computational implementation.
+What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the data analysis without concerning the details of computational implementation.
 A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together.
@@
 This tidy data representation choice will prove suitable for almost all conceivable applications of temporal data.
-Time-based pipeline starts with raw data however it lands.
+Time-based pipeline starts with raw data however it arrives.
 It is advised to scrutinize identical entries of key and index in the right beginning.
@@
 
-## Clarity and transparency
-
 Transformation assembles a set of modules.
 A unit that makes up a module is function or preferably "verb".
-A tsibble is an object, considered as a conceptualized and materialized noun, an action performed on the object can be phrased as a verb.
-Naming verbs is difficult.
-This verb is self-explanatory enough to express what it is supposed to do or fail.
-We shall adapt **dplyr** key verbs to time domain and expand its vocabulary a little more to handle time-related analysis problems.
+A tsibble is an object, conceptually considered as a noun, and hence an action performed on the object can be phrased as a verb.
+Each verb focuses on one thing and achieve its goal.
+The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, and `select()` picks variables.
+These general-purpose verbs are made available in the **tidyvere** suite.
+When manipulating in temporal context, these verbs are adapted to time domain.
+A perceivable difference is summarizing variables between data frame and tsibble.
+The former will reduce to a single summary, whereas the latter will retain the index and their corresponding summaries.
+New tsibble-specific verbs are proposed to expand the vocabulary in order to facilitate temporal data analysis.
+
+A table to list the verbs and what they do.
 
@@
 * **row-wise**: `filter()`, `slice()`, `arrange()`
-* **column-wise**: `mutate()`, `select()`, `summarise()`
+* **column-wise**: `mutate()`, `select()`, `summarize()`
 * **group-wise**: `index_by()`, `group_by()`
 * **reshape**: `gather()`, `spread()`, `nest()`, `unnest()`
-* **rolling window**: `slide()`, `tile()`, `stretch()`
 
-## Composition and modularity
+* Clarity and transparency
+* Composition and modularity
 
@@
 The one who write the code would easily reason about the code and every one else would find the code fairly easy to comprehend.
+
+## Rolling window in functional programming
+
+* **rolling window**: `slide()`, `tile()`, `stretch()`
 

### 2018-11-17 R/smart-meter.R

@@ -13,16 +13,15 @@ elec_ts <- elec %>%
 
 gap_df <- has_gaps(elec_ts)
 
-gap_df %>% 
-  ggplot(aes(x = factor(1), fill = .gaps)) +
-  geom_bar(position = "fill") +
-  coord_flip()
+# ToDo: use tables to present implicit missingness proportions
 
 customer_na <- elec_ts %>% 
   filter(customer_id %in% (gap_df %>% filter(.gaps) %>% pull(customer_id)))
 
 count_na_df <- customer_na %>% 
   count_gaps()
+
+# ToDo: do a hierarchical clustering for classifying customers
 
 count_na_10 <- count_na_df %>% 
   count(customer_id) %>% 
@@ -59,8 +58,7 @@ elec_na_mth %>%
 
 elec_ts <- elec_ts %>% 
   mutate(season = if_else(
-    reading_datetime < make_datetime(2013, 4) |
-    reading_datetime >= make_datetime(2013, 10),
+    time_in(reading_datetime, ~ "2013-03", "2013-10" ~ .),
     "Autumn-Winter", "Spring-Summer"
   ))
 

### 2018-11-18 main.Rmd
started on tidyverts

@@
 ---
-title: "Tidy data structure to support exploration and modeling of temporal-context data"
+title: "Tidy data structure to support exploration and modeling of temporal data"
 author:
@@
     email : rob.hyndman@monash.edu
-abstract: 'Mining temporal-context data for information is often inhibited by a multitude of time formats: irregular or multiple time intervals, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
+abstract: 'Mining temporal data for information is often inhibited by a multitude of time formats: irregular or multiple time intervals, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
 Time series models, in particular, the software supporting time series forecasting makes strict assumptions on data to be provided, typically a matrix of numeric data with an implicit time index.
@@
 It is advised to scrutinize identical entries of key and index in the right beginning.
-Duplicates signal the data quality issue, which would likely affect the following analyses and hence decision making.
+Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making.
 Analysts are encouraged to gaze at data as earlier as possible and reason about the process of data cleaning.
@@
 A suite of verbs are introduced to flatten the lumpy path from temporal data to the object that directs to modelling under the tsibble framework, as well as to facilitate transforming tsibble in various kinds of shapes for analysis and visualization.
-The principle that underpins every single verb is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood.
+The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood.
 If a tsibble cannot hold, for example swiping index off, an error prompts users to complain and suggest alternatives.
@@
 Each verb focuses on one thing and achieve its goal.
-The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, and `select()` picks variables.
-These general-purpose verbs are made available in the **tidyvere** suite.
+The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picking observations, and `select()` picking variables.
+These general-purpose verbs are made available in the **tidyverse** suite.
 When manipulating in temporal context, these verbs are adapted to time domain.
 A perceivable difference is summarizing variables between data frame and tsibble.
-The former will reduce to a single summary, whereas the latter will retain the index and their corresponding summaries.
-New tsibble-specific verbs are proposed to expand the vocabulary in order to facilitate temporal data analysis.
+The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
+New tsibble-specific verbs are proposed to expand the **tidyverse** vocabulary.
+We believe that users, who are already familiar with **tidyverse**, will experience a gentle learning curve for mastering tsibble verbs and glide into temporal data analysis with low cognitive load.
 
@@
 
-* **implicit missingness**: `has_gaps()`, `count_gaps()`, `fill_na()`
+* **implicit missingness**: `has_gaps()`, `count_gaps()`, `fill_gaps()`
 * **row-wise**: `filter()`, `slice()`, `arrange()`
@@
 
-* Clarity and transparency
-* Composition and modularity
+<!-- * [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R] -->
+@eopl assert "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony.
+The chaining of manipulation verbs prepares tsibble for visualization and modeling in numerous ways.
+Most importantly, the ecosystem for tidy time series analysis has been undertaking on the basis of tsibble in R, know as "tidyverts".
 
-<!-- * [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R] -->
-* "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system" [@eopl]
+As a special case of tabular data, a tsibble pipes into the grammar of graphics straight way, making most use of existing graphical toolkits. Specialist plotting.
 
-* The resulting code is well-structured, easier to read and comprehend. A way of chaining together simple operations to perform complex tasks in an efficient way.
-
-Individual operations can talk to each other.
-
-A pipeline exhibits a hierarchy of data operations: (1) atomic (1-dimensional) vectors (`mean(variable)`) --> (2) (2-dimensional) data table (`summarise()`) --> repeat step (1) and (2) to form a chain.
-Reversely, a data pipeline is decomposed into rectangular blocks, and then into atomic strips.
-
-The index and key will automate the update.
-
-Every verb generates the output of the same type of the input, making chaining possible.
-Once the initial setup is done, the program takes care of updating the index and key, and empowers users concentrate on the substance of data analysis.
-The one who write the code would easily reason about the code and every one else would find the code fairly easy to comprehend.
+Tidy forecasting
 

### 2018-11-19 main.Rmd
added a bit on tidy forecasting

@@
 
-<!-- * [@Unwin99guiand] since programming interfaces are precise and repeatable, they are preferable when we can describe exactly what we want, but a GUI is better when: “Searching for information and interesting structures without fully specified questions.” [R] -->
-@eopl assert "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony.
-The chaining of manipulation verbs prepares tsibble for visualization and modeling in numerous ways.
-Most importantly, the ecosystem for tidy time series analysis has been undertaking on the basis of tsibble in R, know as "tidyverts".
+@eopl asserted "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony.
+This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems.
+Most importantly, the ecosystem for tidy time series analysis has been undertaking on the basis of tsibble in R, known as "tidyverts".
 
-As a special case of tabular data, a tsibble pipes into the grammar of graphics straight way, making most use of existing graphical toolkits. Specialist plotting.
+As a special case of data frame, a tsibble pipes into the grammar of graphics straight way, making most use of this powerful graphical system.
+It should be easy to create and extend some specialist plotting methods based on tsibble structure, such as autocorrelation plots [@tsibblestats] and calendar-based graphics [@sugrrants].
 
-Tidy forecasting
+Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data.
+The verbs listed in Table \ref{tab:??} ease the transition to a tsibble that suits modeling.
+A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation.
+A tsibble usually contains multiple time series.
+Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently.
+This yields a "mable" (short for model table), where each model only associates with each "key" value in tsibble to avoid expensive data copying and reduce model storage.
+It also underlines the advantage of tsibble's "key" in linking between data input and models.
 

### 2018-11-20 main.Rmd
added time-based pipeline summary

@@
 Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently.
-This yields a "mable" (short for model table), where each model only associates with each "key" value in tsibble to avoid expensive data copying and reduce model storage.
-It also underlines the advantage of tsibble's "key" in linking between data input and models.
+This yields a "mable" (short for model table), where each model only tags to each "key" value in tsibble to avoid expensive data copying and reduce model storage.
+The mable is further supplied to forecasting methods, to produce a tsibble in which each "key" along with its future time holds predictions.
+It also underlines the advantage of tsibble's "key" in linking between data inputs, models and forecasts.
+The modeling module will be fulfilled and integrated eventually.
+
+We go through the whole exploration circle, and keep iterating and refining until data insights gained.
+The tsibble data structure substantially lubricates between these modules for time-based pipelines.
+The cohesive and coherent framework results in cleaner, clearer and more expressive code.
 

### 2018-11-20 main.Rmd
mostly done tweaking pipeline

@@
 There has been a long history of pipeline discussions and implementation centering around data in various aspects.
-The Extract, Transform, and Load (ETL) process is most commonly deployed pipeline in data science.
-However, building a sound data pipeline can be technically difficult, which involves many implementation decisions, including interface, input/output, functionality and so on, to be made.
+A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualising the process as it might be applied to a variety of problems.
+The Extract, Transform, and Load (ETL), from the data warehousing literature, outlines the workflow to prepare data for analysis, dates back to (XXX Caserta, Joe, 1965?
+REFS needed) can be considered a data pipeline.
+Building a data pipeline can be technically difficult, to make it sufficiently general for various data, with many implementation decisions on the interface, input and output objects and functionality.
+It is useful to articulate the data pipeline induced by new data tools.
 
@@
 
-@viewing-pipeline proposed a viewing pipeline for data analysis in interactive statistical graphics and generalised a collection of elements required for the pipeline, which takes control of transformation from data to plot.
-@plumbing and @xie2014reactive implemented the data pipeline in interactive statistical software **plumbr** and **cranvas** respectively, using a reactive programming framework, in which user's interactions trigger a sequence of modules to update views accordingly.
+Buja et al.
+(1988) [-@viewing-pipeline] describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot.
+@xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **xgobi**, **ggobi**, **orca**, **plumbr** and **cranvas**, respectively.
+The pipeline is typically described with a one way flow, from data to plot.
+For interactive graphics, where all plots need to be updated when a user interacts with one plot, the events typically trigger the data pipeline to be run.
+@xie2014reactive uses a reactive programming framework, to implement the pipeline, in which user's interactions trigger a sequence of modules to update their views, that is, practically the same as running the data pipeline producing each plot.
 
-The tidy data abstraction lays the foundation for data science modules including transformation, visualization and modelling.
-Each module frictionlessly communicates between each other.
-Because a desirable toolbox acquires tidy input, produces tidy output, and consequently chains a series of operations together to accomplish analytic tasks at hand.
+The tidy data abstraction lays a pipeline infrastructure for data analysis modules, transformation, visualization and modelling.
+Each module communicates between each other, requiring tidy input, producing tidy output, and consequently chains a series of operations together to accomplish the analytic tasks at hand.
 
-What is notable about data pipeline is that it not only frees users from a tedious and error-prone analysis but also empowers a wider audience to focus on the data analysis without concerning the details of computational implementation.
+What is notable about an effective implementation of a data pipeline is that it coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies.
 A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together.

### 2018-11-20 main.Rmd
time-based pipeline needs more organisation

@@
 A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together.
-It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes as many as possible.
+It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
 
@@
 
-This tidy data representation choice will prove suitable for almost all conceivable applications of temporal data.
-Time-based pipeline starts with raw data however it arrives.
+XXX may be good to have a diagram here showing how the time-based pipeline builds on a regular data pipeline
+
+XXX Put comments in above each paragraph, with the topic of each paragraph. its not clear how this section flows yet. the topics would help.
+
+The time-based pipeline shepards raw temporal data through to time series analysis, and plots.
 It is advised to scrutinize identical entries of key and index in the right beginning.

### 2018-11-24 main.Rmd
added the linking of "key"

@@
 
-# Data semantics {#sec:semantics}
+# Contextual semantics {#sec:semantics}
 
@@
 Because it is not constrained to a single field, but can be comprised of multiple fields.
-The identifying variables that the "key" is constituted of, remain the same as they are in the original table, with no additional manipulation.
+The identifying variables that the "key" is constituted of, stay the same as they are in the original table, with no further tweaks.
 
-Each tsibble must attach with a "key".
+Each tsibble must hold a "key".
 It is normally a priori known by analysts.
 For example, Table \ref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year.
-This data description suggests that columns `gender` and `country` have to be declared as the key.
+This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data.
 Lacking in either of two will be inadequate for the key and thus fail to construct a tsibble.
@@
 Key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.
+
+Not only pinpoints a "key" observational units in the tsibble but also provides a solution to seamlessly link between the data, models, and forecasts.
+This neatly separates the data from models and forecasts, leaving more room for compulsory model components, such as coefficients, fitted values and residuals.
+More details are given in the following section.
 
@@
 It also underlines the advantage of tsibble's "key" in linking between data inputs, models and forecasts.
+Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in alike spirit.
 The modeling module will be fulfilled and integrated eventually.
@@
 * **rolling window**: `slide()`, `tile()`, `stretch()`
+
+# Software structure
+
+* Extensibility
 

### 2018-11-24 R/flights.R

@@ -64,22 +64,15 @@ ggplot() +
   ggthemes::theme_map()
 
 ## ---- find-duplicate
-dup <- flights %>% 
-  find_duplicates(
-    key = id(flight_num), index = sched_dep_datetime, from_last = TRUE
-  )
-dup_entry <- flights %>% filter(dup)
-
 flights %>% 
-  filter(
-    flight_num == dup_entry$flight_num,
-    sched_dep_datetime == dup_entry$sched_dep_datetime
-  ) %>% 
+  duplicates(key = id(flight_num), index = sched_dep_datetime) %>% 
   as.data.frame()
+dup_lgl <- are_duplicated(flights, key = id(flight_num), 
+  index = sched_dep_datetime, from_last = TRUE)
 
 ## ---- tsibble
 us_flights <- flights %>% 
-  filter(!dup) %>% 
+  filter(!dup_lgl) %>% 
   as_tsibble(
     key = id(flight_num), index = sched_dep_datetime,
     regular = FALSE, validate = FALSE

### 2018-11-24 R/smart-meter.R

@@ -14,7 +14,6 @@ elec_ts <- elec %>%
 gap_df <- has_gaps(elec_ts)
 
 # ToDo: use tables to present implicit missingness proportions
-
 customer_na <- elec_ts %>% 
   filter(customer_id %in% (gap_df %>% filter(.gaps) %>% pull(customer_id)))
 
@@ -22,7 +21,6 @@ count_na_df <- customer_na %>%
   count_gaps()
 
 # ToDo: do a hierarchical clustering for classifying customers
-
 count_na_10 <- count_na_df %>% 
   count(customer_id) %>% 
   top_n(100) %>% 
@@ -37,8 +35,27 @@ count_na_df %>%
   coord_flip() +
   theme(legend.position = "bottom")
 
+qtl_grid <- seq(0.1, 0.9, 0.01)
+elec_qtl <- elec_ts %>% 
+  summarise(
+    value = list(quantile(general_supply_kwh, qtl_grid, na.rm = TRUE)),
+    qtl = list(qtl_grid)
+  ) %>% 
+  unnest(key = id(qtl)) %>% 
+  mutate(
+    time = hms::as.hms(reading_datetime),
+    date = as_date(reading_datetime)
+  )
+
+elec_qtl %>% 
+  ggplot(aes(x = time, y = value, colour = qtl, group = qtl)) +
+  geom_line() +
+  sugrrants::facet_calendar(~ date, ncol = 3) +
+  scale_colour_viridis_c()
+
+# make complete series
 elec_ts <- elec_ts %>% 
-  fill_na()
+  fill_gaps()
 
 elec_na <- elec_ts %>% 
   filter(is.na(general_supply_kwh))

### 2018-11-25 main.Rmd
rewrote sub-section on "interval"

@@
 
-Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
+How to handle time zone. Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
+
+It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R.
+However these turn out inappropriate for weekly, monthly and quarterly data.
+For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year.
+However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced.
+A genuine year-month representation should avoid associating days with its format.
+Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
+This argument is also applicable to the creation of year-week and year-quarter.
+(coercion)
 
@@
 
-For data indexed in regular time space, the time interval is obtained by computing the greatest common divisor based on the non-negative differences of a given time index, although the data may include implicit time gaps.
-This implies that all the units in the table must share a common interval.
-It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R.
-However these turn out inappropriate for weekly, monthly and quarterly data.
-For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year.
-However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced.
-A genuine year-month representation should avoid associating days with its format.
-Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
-This argument is also applicable to the creation of year-week and year-quarter.
-(coercion)
+One of the cornerstones beneath a tsibble is time interval.
+This information plays a critical role in computing statistics (e.g.
+seasonal unit root tests) and building models (e.g.
+seasonal ARIMA).
+The principal divide is regularly or irregularly spaced in time.
+Since a tsibble permits implicit missing time, it is impossible to distinguish regularity from the index.
+It relies on user's specification by switching the `regular` argument off, when the data involves irregular intervals.
+This type of data can flow into event-based data modelling, not models that expect time series.
+
+As for data indexed in regular time space, the time interval is derived by first computing absolute differences of time indices and then the greatest common divisor, which covers all conceivable cases.
+This implies that all subjects in a tsibble have one and the only interval.
+Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each subject may underly different data generating process.
 
@@
 
-Buja et al.
-(1988) [-@viewing-pipeline] describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot.
+@viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot.
 @xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **xgobi**, **ggobi**, **orca**, **plumbr** and **cranvas**, respectively.

### 2018-11-26 main.Rmd
added one more paragraph to sub-section on "index"

@@
 The "index" is an intact data column rather than an attribute, which makes time accessible to users.
-For example, one could easily extract time components, such as time of day and day of week, from the index to visualise seasonal effects of response variables.
+It is highly advantageous when manipulating time.
+For example, one could easily extract time components, such as time of day and day of week, from the index to visualize seasonal effects of response variables.
 One could also join other data sources to the tsibble based on common time indices.
@@
 
-How to handle time zone. Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
-
-It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R.
-However these turn out inappropriate for weekly, monthly and quarterly data.
-For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year.
-However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced.
-A genuine year-month representation should avoid associating days with its format.
-Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
-This argument is also applicable to the creation of year-week and year-quarter.
-(coercion)
+A number of time classes are supported and easily extensible to new types.
+The base line is that time indices can be potentially ordered in past to future.
+Characters cannot be ordered, so a variable hosting characters will make it unsupported for index.
+Ordered factors, however, can be immediately recognized as valid index.
+For instance, month names have a natural ordering from January to December, and so are weekdays from Monday to Sunday.
 
@@
 * Extensibility
+
+How to handle time zone. Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
+
+It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R.
+However these turn out inappropriate for weekly, monthly and quarterly data.
+For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year.
+However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced.
+A genuine year-month representation should avoid associating days with its format.
+Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
+This argument is also applicable to the creation of year-week and year-quarter.
+(coercion)
 

### 2018-11-27 main.Rmd
roughly done with section on "contextual semantics"

@@
 
+This paper proposes a unified data representation of temporal-context data, blending time series of nested and crossed factors into a two-dimensional column-homogeneous array in a long format.
+By leveraging the "tidy data" principles, observations and variables position and bridge their meanings in both physical and internal structures.
+Data manipulation involves in transforming either observations or variables, or both, which can be described and achieved with a collection of shorthand operators.
+A chain of data transformations lend itself to a data pipeline.
+
 The rest of the paper is structured as follows.
@@
 
-This paper proposes a unified data representation of temporal-context data, blending time series of nested and crossed factors into a two-dimensional column-homogeneous array in a long format.
-By leveraging the "tidy data" principles, observations and variables position and bridge their meanings in both physical and internal structures.
-Data manipulation involves in transforming either observations or variables, or both, which can be described and achieved with a collection of shorthand operators.
-A chain of data transformations lend itself to a data pipeline.
-
 # Contextual semantics {#sec:semantics}
 
-The choice of representation of temporal-context data is made from a data-centric perspective, which is taken in the light of the operations that are to be performed on the data.
-This data abstraction reflects most common problems of incoming data and transformation in reality.
-Firstly, a data set must be structured in a "tidy" rectangular layout each row has the same schema and each field has a single value.
-Secondly, declaring the data set to contain temporal observations is determined by an "index" that represents time and a "key" that uniquely identifies each unit that measurements take place on over time.
-The "key" serves a similar purpose as the panel variable in the Stata's **tsset** command to define the units or subjects, but it is expanded to include multiple variables rather than a single one.
-The composition of index and key uniquely defines each observation in a data table, which is equivalent to a primary key [@codd_relational_1970] in a relational database.
-
-Table \@ref(tab:tb-sub) shows a "tidy" temporal data example using a subset of tuberculosis cases estimated by @tb-data.
-It contains 12 observations and 5 variables arranged in a "long" tabular form.
-Each observation uniquely records the number of people, who are diagnosed tuberculosis for each gender at three selected countries in the years of 2011 and 2012, where column `year` is declared as the index variable, columns `country`, `continent` and `gender` as key.
-Detailed discussions about index and key specifications will be given in Section \@ref(sec:index) and \@ref(sec:key).
-Although column `count` is the only measure in this case, it is sufficiently flexible to hold other measured variables, for example, adding the corresponding population size (if known) in order to calibrate the count later.
-
-Given the nature of temporal ordering, a temporal data set must be sorted by time index as a result.
-If a key is explicitly declared, the key will be sorted first depending on the ordering of factor levels and the ordering of variables, and followed by arranging time in past-to-future order.
-The tuberculosis example shows the key is sorted in alphabetic order, followed by ascending years by default, as in Table \@ref(tab:tb-sub).
-Since time-based operations, for example leading, lagging, or differencing series, require a vector in its time order, the arranged data format assures its validity for these operations without returning unexpected results.
-There are other possible arrangements in the key, but the baseline is the index (i.e.
-`year`) ascends within each key element.
-If any operations on the tidy temporal data results in a deviation from this baseline, it will issue a warning to the users.
-
-This high-level data abstraction is semantically structured, which shed lights on time index and what we call "key".
-
-It encourage users to look into the data as earlier as possible, assessing the data quality issues.
+The choice of tidy representation of temporal data is made from a data-centric perspective, which is taken in the light of the operations that are to be performed on the data.
+Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
+Index, key and time interval are the three stone pillars to this new semantically-structured temporal data, which we name as "tsibble".
 
@@
 (ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012, with 12 observations and 5 variables.
-The index refers to column `year`, the key to multiple columns: `country` nested under `continent` crossed with `gender`, and the measured variable to column `count`.
+The index refers to column `year`, the key to multiple columns: `country` and `gender`, and the measured variable to column `count`.
+
+Table \@ref(tab:tb-sub) presents a subset of tuberculosis cases estimated by @tb-data.
+It contains 12 observations and 5 variables arranged in a "long" tabular form.
+Each observation hosts the number of people who are diagnosed tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
+To turn this data into a tsibble, (1) column `year` is declared as the index variable; (2) the key can be made up of columns `country` and `gender`.
+Despite column `count` is the only measurement in this case, it is sufficiently flexible to hold other measured variables, for example, adding the corresponding population size (if known) in order to calibrate the count later.
 
@@
 A number of time classes are supported and easily extensible to new types.
-The base line is that time indices can be potentially ordered in past to future.
+The baseline is that time indices can be potentially ordered from past to future.
 Characters cannot be ordered, so a variable hosting characters will make it unsupported for index.
@@
 The "key" specification contributes to creating a tsibble alongside the index.
-The concept of "key" is introduced to identify units or subjects that are observed over time in a data table.
+The concept of "key" is introduced to uniquely identify units or subjects that are observed over time in a data table, which is similar to a primary key [@codd_relational_1970] defining each observation in a relational database.
 In the wide format, each column holds a series of values, so that the column itself serves for the sake of identification.
@@
 Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each subject may underly different data generating process.
+
+<!--
+
+## Data first
+
+Given the nature of temporal ordering, a temporal data set must be sorted by time index as a result.
+If a key is explicitly declared, the key will be sorted first depending on the ordering of factor levels and the ordering of variables, and followed by arranging time in past-to-future order.
+The tuberculosis example shows the key is sorted in alphabetic order, followed by ascending years by default, as in Table \@ref(tab:tb-sub).
+Since time-based operations, for example leading, lagging, or differencing series, require a vector in its time order, the arranged data format assures its validity for these operations without returning unexpected results.
+There are other possible arrangements in the key, but the baseline is the index (i.e.
+`year`) ascends within each key element.
+If any operations on the tidy temporal data results in a deviation from this baseline, it will issue a warning to the users.
+
+It encourage users to look into the data as earlier as possible, assessing the data quality issues.
+
+-->
 
@@
 
+data first, function starts with the first argument "data", chained with `%>%`. print display.
+
 * Extensibility
@@
 An error is immediately raised when attempting to convert this data into a tsibble, and a closer inspection has to be carried out to locate the issue.
-The **tsibble** package provides the tool to easily find the duplicates in the data with `find_duplicates()`.
+The **tsibble** package provides tools to easily locate the duplicates in the data with `duplicates()`.
 Below shows the problematic entries.

### 2018-11-28 main.Rmd
added "data first" subsection

@@
 ```
-
-This paper proposes a unified data representation of temporal-context data, blending time series of nested and crossed factors into a two-dimensional column-homogeneous array in a long format.
-By leveraging the "tidy data" principles, observations and variables position and bridge their meanings in both physical and internal structures.
-Data manipulation involves in transforming either observations or variables, or both, which can be described and achieved with a collection of shorthand operators.
-A chain of data transformations lend itself to a data pipeline.
 
@@
 
-# Software structure
+# Software structure and design decisions
 
-data first, function starts with the first argument "data", chained with `%>%`. print display.
+The **tsibble** package highly features modularity and extensibility.
 
-* Extensibility
+## Data first
+
+The prime force that drives the software's design choices is "data".
+All functions in **tsibble** starts with `data` or its variants as the first argument, namely "data first".
+They work naturally with the pipe operator `%>%`, read as "then".
+This not only lays out a consistent interface but also addresses the significance of the data throughout the software.
+
+Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
+Below conveys the pieces of critical and contextual information about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series.
+These details aid users in understanding their data better and manipulate the data with care.
+
+```{r tb-print}
+```
+
+## Modularity
+
+## Extensibility
 

### 2018-11-28 R/demo.R

@@ -7,3 +7,7 @@ tb_small <- read_rds("data/tb-small.rds")
 tb_small %>% 
   knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>% 
   kableExtra::kable_styling(position = "center")
+
+## ---- tb-print
+as_tsibble(tb_small, key = id(country, gender), index = year) %>% 
+  print(n = 5)

### 2018-12-01 main.Rmd
draft section 5.3 extensibility

@@
 
+How to handle time zone. Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
+
 ## Key {#sec:key}
@@
 
-## Time-based pipeline
+## Time-based pipeline {#sec:timepipeline}
 
@@
 Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-Below conveys the pieces of critical and contextual information about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series.
+Below conveys the pieces of critical and contextual picture about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series.
 These details aid users in understanding their data better and manipulate the data with care.
@@
 
-How to handle time zone. Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
+As a fundamental infrastructure, extensibility is highly valued and one of the core design features in the **tsibble** package.
+Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily relies on S3 classes and methods in R [@adv-r].
+It can be extended in two major aspects: custom index and new tsibble class.
 
-It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R.
-However these turn out inappropriate for weekly, monthly and quarterly data.
-For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year.
-However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced.
-A genuine year-month representation should avoid associating days with its format.
-Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
-This argument is also applicable to the creation of year-week and year-quarter.
-(coercion)
+Tsibble supports an extensive range of index types from numerics to nano time, but there might be custom indices used for some occasions, for example school semesters.
+These academic terms vary from one institute to another within an academic year, which is defined differently from a calendar year.
+New index to be valid (defining `index_valid()`), the bottom line is that it can be ordered from past to future.
+The interval regarding semesters is further outlined through `pull_interval()`.
+As a result, the rest software methods such as `has_gaps()`, `count_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
+
+The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble.
+For example, a fable is actually an extension to a tsibble, mentioned in Section \ref{sec:timepipeline}.
+A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass.
+First of all, this new object itself is a tsibble.
+It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable.
+Tsibble verbs are also S3 generics.
+Developers will be able to implement these verbs for the new class if needed.
 

### 2018-12-01 main.Rmd
tweak paragraph in section index

@@
 
-Time provides a contextual basis for temporal data. Time representation could range from date-times (such as `POSIXct` and `Date` in R) to years, to sequential numbers for simulated data for example.
-
+Time provides a contextual basis for temporal data.
 A variable representing time is indispensable to a tsibble, referred to as "index".
@@
 
-A number of time classes are supported and easily extensible to new types.
-The baseline is that time indices can be potentially ordered from past to future.
-Characters cannot be ordered, so a variable hosting characters will make it unsupported for index.
-Ordered factors, however, can be immediately recognized as valid index.
-For instance, month names have a natural ordering from January to December, and so are weekdays from Monday to Sunday.
-
-How to handle time zone. Additional representations are provided, like year-month, year-quarter. Extensible. Custom index classes can be supported.
+Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution [@R-nanotime], and pure numbers in simulations.
+Yet ordered factors can also be a source of time, such as month names from January to December and weekdays from Monday to Sunday.
+New index classes are introduced to accurately represent year-week, year-month and year-quarter.
+Supporting a wide array of time types is a must for a tsibble, and extending for new types is a plus.
 
@@
 These academic terms vary from one institute to another within an academic year, which is defined differently from a calendar year.
-New index to be valid (defining `index_valid()`), the bottom line is that it can be ordered from past to future.
+New index would be immediately recognized by the software once `index_valid()` defined, as long as it can be ordered from past to future.
 The interval regarding semesters is further outlined through `pull_interval()`.

### 2018-12-02 main.Rmd
added Modularity

@@
 
-Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution [@R-nanotime], and pure numbers in simulations.
-Yet ordered factors can also be a source of time, such as month names from January to December and weekdays from Monday to Sunday.
-New index classes are introduced to accurately represent year-week, year-month and year-quarter.
-Supporting a wide array of time types is a must for a tsibble, and extending for new types is a plus.
-
 ## Key {#sec:key}
@@
 
-The **tsibble** package highly features modularity and extensibility.
+data-centric & human-centered.
 
@@
 
+Modular programming is adopted while designing the **tsibble** package. Modularity benefits users with variety and flexibility and developers with easy maintenance.
+
+All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family.
+Each chunk is an independent module, but works interdependently.
+Vector functions in the package mostly deal with time.
+When collapsing a tsibble to less granular interval, these atomic functions can be combined with the `index_by()` table verb to accomplish this.
+A different function results in easily switching to aggregation of different time resolution.
+Since these functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages.
+On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of current tsibble zone.
+It is generally easier to trace back the errors users encounter from separating 1d and 2d functions.
+
 ## Extensibility
 
-As a fundamental infrastructure, extensibility is highly valued and one of the core design features in the **tsibble** package.
+As a fundamental infrastructure, extensibility is a design decision that is focused on from the start of **tsibble**'s development.
 Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily relies on S3 classes and methods in R [@adv-r].
@@
 
+Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution [@R-nanotime], and pure numbers in simulations.
+Yet ordered factors can also be a source of time, such as month names from January to December and weekdays from Monday to Sunday.
 Tsibble supports an extensive range of index types from numerics to nano time, but there might be custom indices used for some occasions, for example school semesters.

### 2018-12-02 main.Rmd
added intro to smart meter data

@@
 read_chunk('R/flights.R')
+read_chunk('R/smart-meter.R')
 ```
@@
 
-## Smart-Grid Customer Data in Australia
+## Smart-grid customer data in Australia
+
+```{r load-elec}
+```
+
+Sensors have been put up to collect data for the project of smart city across major cities in Australia.
+One of the trials is monitoring households' electricity usage through installed smart meters in the area of New Castle over 2010--2014 [@smart-meter].
+Year 2013 has been sliced to examine temporal patterns of customer's energy consumption with **tsibble** in this paper.
+Half-hourly general supply in kwH have been recored for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
+Customer's demographic data provides explanatory variables other than time in a separate data table.
+Two data tables might be joined to explore different sources that contribute to daily electricity use when needed.
 

### 2018-12-02 R/smart-meter.R

@@ -1,3 +1,4 @@
+## ---- load-elec
 library(tidyverse)
 library(lubridate)
 library(tsibble)
@@ -11,6 +12,7 @@ elec_ts <- elec %>%
     validate = FALSE, ordered = TRUE
   )
 
+## ---- gaps-elec
 gap_df <- has_gaps(elec_ts)
 
 # ToDo: use tables to present implicit missingness proportions

### 2018-12-03 main.Rmd
added a little bit to index

@@
 A variable representing time is indispensable to a tsibble, referred to as "index".
-The "index" is an intact data column rather than an attribute, which makes time accessible to users.
+The "index" is an intact data column rather than a masked attribute, which makes time visible and accessible to users.
 It is highly advantageous when manipulating time.
@@
 The accessibility of tsibble index motivates data analysis towards transparency and human readability.
-But if the "index" is employed as attributes, analysts would write these simple queries in a programmatic manner, which can be avoidable from an analytic point of view.
+When the "index" used to be employed as meta information, it created an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.
+
+A variable number of time representation is spotted in the wild.
+Date-time, universally accepted across systems, is the most commonly used type for representing time.
+Date-time also typically associates with a time zone with adjustments like summer time.
+This diversity and time zone is respected and taken into consideration for the *tsibble*'s index.
 
@@
 Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each subject may underly different data generating process.
-
-<!--
-
-## Data first
-
-Given the nature of temporal ordering, a temporal data set must be sorted by time index as a result.
-If a key is explicitly declared, the key will be sorted first depending on the ordering of factor levels and the ordering of variables, and followed by arranging time in past-to-future order.
-The tuberculosis example shows the key is sorted in alphabetic order, followed by ascending years by default, as in Table \@ref(tab:tb-sub).
-Since time-based operations, for example leading, lagging, or differencing series, require a vector in its time order, the arranged data format assures its validity for these operations without returning unexpected results.
-There are other possible arrangements in the key, but the baseline is the index (i.e.
-`year`) ascends within each key element.
-If any operations on the tidy temporal data results in a deviation from this baseline, it will issue a warning to the users.
-
-It encourage users to look into the data as earlier as possible, assessing the data quality issues.
-
--->
 

### 2018-12-04 main.Rmd
restarted conclusion

@@
 @r4ds argues that 80% of data analysis tasks can be solved with tidy tools while the remaining 20% requires other tools.
-[R]
+[R] The prevalence of grammar
 
@@
 The tsibble data structure substantially lubricates between these modules for time-based pipelines.
-The cohesive and coherent framework results in cleaner, clearer and more expressive code.
+The cohesive and coherent framework results in more graceful and expressive code.
 
@@
 It is generally easier to trace back the errors users encounter from separating 1d and 2d functions.
+(lost in a web of functions)
 
@@
 These academic terms vary from one institute to another within an academic year, which is defined differently from a calendar year.
-New index would be immediately recognized by the software once `index_valid()` defined, as long as it can be ordered from past to future.
+New index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future.
 The interval regarding semesters is further outlined through `pull_interval()`.
@@
 
-A tidy representation of time series data, and data pipelines to facilitate data analysis flow have been proposed and discussed.
-It can be noted that tidy temporal data gains greater flexibility in keeping data richness, making data transformation and visualisation easily.
-A set of verbs provides a fluent and fluid pipeline to work with tidy time series data in various ways.
-
-A new visualization framework for handling time series with nesting and crossing structure is under development.
-The proposed data structure proves a natural input (unit time series and parent-children structure) that can be piped into a visualization program.
-As long as the program understands the embedded structure, it will produce sensible layout and interactivity.
-
-The ground of time series modelling or forecasting is left untouched in this paper.
-The future plan is to bridge the gap between tidy data and model building.
-Currently, it is required to casting to matrix from tidy data and therefore building a model.
-But time series models should be directly applied to tidy data as other wrangling tools do, without such an intermediate step.
-In particular, a univariate time series model, like ARIMA and Exponential Smoothing, can be applied to multiple time series independently.
-A tidy format to represent model summaries and forecasting objects will be developed and implemented in the future.
-Model summaries include coefficients, fitted values, and residuals; forecasting objects.
-
-When involving multiple variables, one need to distinguish nesting and crossing data variables.
-"Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`).
-Nesting is a special case of crossing.
-The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference.
-Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data.
-Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics.
-The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables (reflect depths), while a comma (`,`) for crossing variables.
-@wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions.
-A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout.
-However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.nclude future time path and distributions generating prediction intervals.
+A new data abstraction representing temporal data named as "tsibble" has been proposed, spotlighting the "tidy data" principles brought to time domain.
+Tidy data begins to take shape in the state of time with the introduction of contextual semantics: index and key.
+Declared index provides direct support to time variable; variables that comprise the key defines study subjects over time.
+These semantics further determines unique data entries required for a valid tsibble.
+No matter how temporal data arrives, tsibble respects time index and keeps data richness.
+A tsibble frictionlessly pops into transformation, visualization, modelling and smoothly shifts amongst, allowing for rapid iterations in gaining data insights.
 

### 2018-12-06 main.Rmd
added "respect structural missingness" into conclusion

@@
 
-<!--
-## Time series of nested and crossed factors
-
-A collection of time series can often be intrinsically arranged along multivariate data of nested and crossed factors, which is also known as hierarchical and grouped time series [@fpp].
-For example, from an operational point of view, a manufacturer can add up every store’s sales by region, by state, and by country along a geographical tree; alternatively, one can group the sales for each product together based on common attributes such as store, brand, price range and so forth, resulting in a non-hierarchical structure.
-The former gives an example of hierarchical time series, and the latter shows grouped time series.
-And even when collections of time series are absent from hierarchical structure, it is possible to arrange them empirically into a hierarchy such as regression tree [@sankaran_interactive_2017].
-
-The R package **hts** [@R-hts] is the implementation of structured collections of time series, and provides tools for visualizing, analyzing, and forecasting such data.
-But it is a pain to configure node list or grouping matrix as the increase in hierarchic depth and grouping factors.
-This is resulted from the fact that the structure specification is separated from data itself and represented numerically, which is rather counterintuitive from the user's perspective.
--->
-
 ## Tidy data and the grammar of data manipulation
@@
 Not only pinpoints a "key" observational units in the tsibble but also provides a solution to seamlessly link between the data, models, and forecasts.
-This neatly separates the data from models and forecasts, leaving more room for compulsory model components, such as coefficients, fitted values and residuals.
+This neatly decouples the data from models and forecasts, leaving more room for compulsory model components, such as coefficients, fitted values and residuals.
 More details are given in the following section.
@@
 Half-hourly general supply in kwH have been recored for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
-Customer's demographic data provides explanatory variables other than time in a separate data table.
+Customer's demographic data provides explanatory variables other than time in a different data table.
 Two data tables might be joined to explore different sources that contribute to daily electricity use when needed.
@@
 
+A missing piece of the *tsibble* data is to enable user-defined calendars and respect structural missing observations.
+For example, a call center operates only between 9:00am and 5:00pm on week days and stock trading resumes on Monday straight after Friday.
+No data available outside trading hours would be labelled as structural missingess, which *tsibble* currently disregards.
+However, few R packages provide functionality to create and manage many sorts of calendars, including market-specific business calendar.
+This delays the implementation.
+Generally, custom calendars would be easily embedded into the tsibble framework.
+Consequently these tsibble operators, like `fill_gaps()`, would work out of box; forecasts would be generated within its definable time range.
+
 ```{r write-bib}

### 2018-12-07 main.Rmd
restarted introduction

@@
 
-Temporal-context data, $X_{jt}$, consist of $j = 1, \dots, N_{i}$ observational units indexed at different time points, $1 \le t \le T$.
-Time forms the contextual basis of temporal data, but it can arrive in many possible formats, which makes preparing it for modeling and forecasting painful.
+Suppose that temporal data consists of $N$ subjects/observational units.
+Given subject $i$ $(i = 1, 2, \dots, N)$, $X_{ijt}$ denotes the $j$th measurement $(j = 1, 2, \dots, p)$ indexed at a unique sequence of time points $1 \le t \le T_{i}$.
+Temporal data arrives in many possible formats, so does time.
+An overview of data science workflow is given in Figure \ref{fig:pipeline-ann}.
+The divide between data that feeds into the beginning of analysis and the current time series standard that fits into modelling is significant, making data preparation for modeling and forecasting painful.
 For example, data can be recorded at various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
 It could be irregularly recorded, which is particularly true with longitudinal measurements like patient visits to a doctor's office.
-Temporal data also often contains rich information in addition to time: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors involving nested or crossed structures.
-The data problems are grouped into two types of analysis, time series and longitudinal.
+Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
 
-```{r pipeline-ann}
+```{r pipeline-ann, fig.cap = "(ref:pipeline-ann-cap)"}
 include_graphics("img/time-series-structure.pdf")
 ```
+
+(ref:pipeline-ann-cap) An overview of data science program. The existing standards are model-focused, resulting misalliance between data and model.
 
@@
 
-## Time series and longitudinal data
+## Time series and longitudinal data {#sec:time-series}
 
+The data problems are grouped into two types of analysis, time series and longitudinal.
 Despite of exactly the same data input, the representation of time series and longitudinal data diverges due to different modelling approaches.
@@
 large $T$).
-Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic processes, for the primary purpose of forecasting, and characterising temporal dynamics.
-The time series supporting modeling are represented as vectors or matrices in most statistical software, with the standards being provided by the R [@R-base] packages `ts`, **zoo** [@zoo2005], and **xts** [@R-xts] (*Why include MATLAB [@Matlab]??*).
+Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic processes, for the primary purpose of forecasting, and characterizing temporal dynamics.
+The time series supporting modeling are represented as vectors or matrices in most statistical software.
 Multivariate time series are typically assumed to be in the format where each row is assumed to hold observations at a time point and each column to contain a single time series.
@@
 
-Longitudinal data representation has been implemented in Stata's time series module [@Stata] and R package **plm** [@plm2008].
+Longitudinal data representation has been implemented.
 The underpinning data structure is a two-dimensional column-homogeneous array, as other tabular data.
@@
 @viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot.
-@xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **xgobi**, **ggobi**, **orca**, **plumbr** and **cranvas**, respectively.
+@xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **Xgobi**, **Ggobi**, **Orca**, **plumbr** and **cranvas**, respectively.
 The pipeline is typically described with a one way flow, from data to plot.
@@
 
-A table to list the verbs and what they do.
-
-* **implicit missingness**: `has_gaps()`, `count_gaps()`, `fill_gaps()`
-* **row-wise**: `filter()`, `slice()`, `arrange()`
-* **column-wise**: `mutate()`, `select()`, `summarize()`
-* **group-wise**: `index_by()`, `group_by()`
-* **reshape**: `gather()`, `spread()`, `nest()`, `unnest()`
+\begin{table}[!hbt]
+\caption{A test}\label{tab:verb-table}
+\centering
+\begin{tabular}{ lll }
+\toprule
+ & Verb & Description \\ \midrule
+\multirow{3}{*}{Time gaps} & \texttt{has\_gaps()} & Test if a tsibble has gaps in time \\
+ & \texttt{count\_gaps()} & Count and report time gaps \\
+ & \texttt{fill\_gaps()} & Fill in gaps by values and functions \\ \hline
+\multirow{4}{*}{Row-wise} & \textbf{\texttt{filter()}} & Pick rows based on conditions \\
+ & \texttt{filter\_index()} & Provide a shorthand for time subsetting \\
+ & \textbf{\texttt{slice()}} & Select rows based on row positions \\
+ & \textbf{\texttt{arrange()}} & Sort the ordering of row by variables \\ \hline
+\multirow{4}{*}{Column-wise} & \textbf{\texttt{select()}} & Pick columns by variables \\
+ & \textbf{\texttt{mutate()}} & Add new variables \\
+ & \textbf{\texttt{transmute()}} & Add new variables \\
+ & \textbf{\texttt{summarise()}} & Aggregate values over time \\ \hline
+\multirow{3}{*}{Group-wise} & \texttt{index\_by()} & Group by new time index \\
+ & \textbf{\texttt{group\_by()}} & Group by one or more variables \\
+ & \texttt{group\_by\_key()} & Group by key variables \\ \hline
+\multirow{4}{*}{Reshape data} & \textbf{\texttt{gather()}} & David Batty \\
+ & \textbf{\texttt{spread()}} & Eirik Bakke \\
+ & \textbf{\texttt{nest()}} & Eirik Bakke \\
+ & \textbf{\texttt{unnest()}} & Jody Morris \\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 As a special case of data frame, a tsibble pipes into the grammar of graphics straight way, making most use of this powerful graphical system.
-It should be easy to create and extend some specialist plotting methods based on tsibble structure, such as autocorrelation plots [@tsibblestats] and calendar-based graphics [@sugrrants].
+It should be easy to create and extend some specialist plotting methods based on tsibble structure, such as autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
 
@@
 
-data-centric & human-centered.
+Time series are represented as matrices, with standards being provided by native `ts` object in R, extended by **zoo** [@zoo2005], and **xts** [@R-xts].
+As discussed in Section \ref{sec:time-series}, this data organization sets far apart from data origin.
+Due to time isolated from the main data, the supporting functions that intend to deal with time must be applied to whole data.
+That data-centric & human-centered.
 
@@
 
-Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution [@R-nanotime], and pure numbers in simulations.
+Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nantotime** [@R-nanotime], and pure numbers in simulations.
 Yet ordered factors can also be a source of time, such as month names from January to December and weekdays from Monday to Sunday.
@@
 
-```{r map-prep}
-```
-
-<!-- 
-```{r map-airlines, fig.cap = "(ref:map-airlines-cap)"}
-```
-
-(ref:map-airlines-cap) An overview of US airline traffic in 2017.
-Points indicate locations of airports, and lines the connecting routes across the country.
-There are a few hubs connecting most airlines and sporadic airports.
--->
-
-The dataset of 2017 on-time performance for US domestic flights represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics (reference).
+The dataset of 2017 on-time performance for US domestic flights represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics [@flights-data].
 It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format.
@@
 A decision is made on removal of the first row from the duplicated entries before proceeding to the tsibble creation.
-
-<!--
-
-```{r dl771, fig.cap = "(ref:dl771-cap)"}
-```
-
-(ref:dl771-cap) Flight DL771 operates on multiple routes between different origins and destinations. It suggests that the key required for identification consists of flight number and origin.
-
--->
 
@@
 
-<https://data.gov.au/dataset/smart-grid-smart-city-customer-trial-data>
-
 # Conclusion and future work
@@
 
+The **tsibble** package provides an elegant solution to manage and manipulate medium-sized temporal data in memory.
+
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "nanotime"), file = "rpkgs.bib")
 ```

### 2018-12-08 main.Rmd
added paper structure in intro

@@
 Temporal data arrives in many possible formats, so does time.
+For example, time can carry with various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
+It could be irregularly recorded, which is particularly true with longitudinal measurements like patient visits to a doctor's office.
+Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
 An overview of data science workflow is given in Figure \ref{fig:pipeline-ann}.
 The divide between data that feeds into the beginning of analysis and the current time series standard that fits into modelling is significant, making data preparation for modeling and forecasting painful.
-For example, data can be recorded at various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
-It could be irregularly recorded, which is particularly true with longitudinal measurements like patient visits to a doctor's office.
-Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
 
@@
 The rest of the paper is structured as follows.
+Section \ref{sec:str} reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation.
+Section \ref{sec:semantics} proposes contextual semantics for temporal data, built on top of tidy data.
+The concept of data pipelines with respect to the time domain will be discussed in depth in Section \ref{sec:pipeline}, followed by the design choices of software structure in Section \ref{sec:software}.
+Section \ref{sec:cases} presents two case studies for temporal data exploration using the newly implemented infrastructure.
+Section \ref{sec:conclusion} concludes the paper and discusses the future work.
 
@@
 It is conducive to examining the data in many different ways and leading to more comprehensive exploration and forecasting.
-
-Longitudinal data representation has been implemented.
-The underpinning data structure is a two-dimensional column-homogeneous array, as other tabular data.
-Specifying a longitudinal data set from a generic array needs explicitly declaring time and individuals (or panel variable in Stata's `tsset` command).
-Therefore, each row of the data can be identified by a specific individual and time point.
-Individuals, however, can be only declared through a single variable, not multiple.
 
@@
 A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualising the process as it might be applied to a variety of problems.
-The Extract, Transform, and Load (ETL), from the data warehousing literature, outlines the workflow to prepare data for analysis, dates back to (XXX Caserta, Joe, 1965?
-REFS needed) can be considered a data pipeline.
+The Extract, Transform, and Load (ETL), from the data warehousing literature, outlines the workflow to prepare data for analysis, dates back to @kimball2011data can be considered a data pipeline.
 Building a data pipeline can be technically difficult, to make it sufficiently general for various data, with many implementation decisions on the interface, input and output objects and functionality.
@@
 Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data.
-The verbs listed in Table \ref{tab:??} ease the transition to a tsibble that suits modeling.
+The verbs listed in Table \ref{tab:verb-table} ease the transition to a tsibble that suits modeling.
 A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation.
@@
 
-# Software structure and design decisions
+# Software structure and design decisions {#sec:software}
 
@@
 
-# Case studies
+# Case studies {#sec:cases}
 
@@
 
-# Conclusion and future work
+# Conclusion and future work {#sec:conclusion}
 

### 2018-12-09 main.Rmd
restarted tidy data

@@
 Suppose that temporal data consists of $N$ subjects/observational units.
-Given subject $i$ $(i = 1, 2, \dots, N)$, $X_{ijt}$ denotes the $j$th measurement $(j = 1, 2, \dots, p)$ indexed at a unique sequence of time points $1 \le t \le T_{i}$.
+Given subject $i$ $(i = 1, 2, \dots, N)$, $X_{ijt}$ denotes the $j$th measurement $(j = 1, 2, \dots, p)$ indexed at a unique sequence of time points $0 \le t \le T_{i}$.
 Temporal data arrives in many possible formats, so does time.
@@
 Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
-An overview of data science workflow is given in Figure \ref{fig:pipeline-ann}.
-The divide between data that feeds into the beginning of analysis and the current time series standard that fits into modelling is significant, making data preparation for modeling and forecasting painful.
+@r4ds succinctly describes the data science workflow given in Figure \ref{fig:pipeline-ann}.
+It is annotated to show that the divide between data that feeds into the beginning of analysis and the current time series objects that fits into modelling is significant, making data preparation for modeling and forecasting painful.
 
@@
 (ref:pipeline-ann-cap) An overview of data science program. The existing standards are model-focused, resulting misalliance between data and model.
+
+The existing standards, particularly time series objects, are model-oriented with little support for non-numeric values, leaning its data abstraction towards matrix.
+Time series have been isolated from general data problems due to this data choice.
+Analysts who start analysing time series would be pushed to learn new transformation and visualization toolkits that are different from what they used to operate.
 
@@
 
-@wickham2014tidy coined the term "tidy data", which is a rephrasing of the second and third normal forms in relational databases but in a way that makes more sense to data scientists.
-A suite of paired verbs such as `gather` and `spread` [@R-tidyr] were used to describe the reshape processing from messy data to tidy data.
-Tidy data, which standardized the mapping from the semantics of a dataset to its physical representation, serves remarkably as a fundamental unit of data analysis.
-R package **ggplot2** [@wickham2009ggplot2] pioneered developing the grammar of graphics based on tidy data, thus making it possible mapping from data space to visual elements.
-**dplyr** [@R-dplyr] generalised a coherent and consistent set of verbs to handle a wide range of data transformation tasks.
-@r4ds argues that 80% of data analysis tasks can be solved with tidy tools while the remaining 20% requires other tools.
-[R] The prevalence of grammar
+@wickham2014tidy coined the term "tidy data", which is a rephrasing of the second and third normal forms in relational databases but in a way that makes more sense to data scientists by referring rows to observations and columns to variables.
+The principles of "tidy data" attempt to standardize the mapping of the semantics of a dataset to its physical representation.
+This data structure is the fundamental unit of the **tidyverse**, which is a collection of R packages designed for data science; the prevalence of **tidyverse** reversely testifies to the novelty of "tidy data".
+The **tidyverse** provides abstract yet functional grammars to manipulate and visualize data in easier-to-comprehend manners.
+One of the **tidyverse** packages, **dplyr** [@R-dplyr] showcases the value of a grammar as a principled vehicle to transform data for a wide range of data challenges, providing a consistent set of verbs: `mutate()`, `select()`, `filter()`, `summarise()`, and `arrange()`.
+Each verb focuses on a singular task.
+Almost all data tasks can be rephrased and tackled with these five key verbs, by composing them sequentially.
+
+Community effort for building tidy ecosystem
+
+Temporal data is integrated to tidy data and the grammar. @r4ds argues that 80% of data analysis tasks can be solved with tidy tools while the remaining 20% requires other tools.
 
@@
  & Verb & Description \\ \midrule
-\multirow{3}{*}{Time gaps} & \texttt{has\_gaps()} & Test if a tsibble has gaps in time \\
- & \texttt{count\_gaps()} & Count and report time gaps \\
+\multirow{4}{*}{Time gaps} & \texttt{has\_gaps()} & Test if a tsibble has gaps in time \\
+ & \texttt{scan\_gaps()} & Report implicit missing entries \\
+ & \texttt{count\_gaps()} & Summarize time gaps \\
  & \texttt{fill\_gaps()} & Fill in gaps by values and functions \\ \hline
@@
 
+```{r elec-gaps}
+```
+
 # Conclusion and future work {#sec:conclusion}

### 2018-12-09 R/smart-meter.R

@@ -12,32 +12,42 @@ elec_ts <- elec %>%
     validate = FALSE, ordered = TRUE
   )
 
-## ---- gaps-elec
+## ---- elec-gaps
 gap_df <- has_gaps(elec_ts)
+# sum(gap_df$.gaps) / NROW(gap_df)
 
-# ToDo: use tables to present implicit missingness proportions
+## ---- scan-gaps
+elec_gaps <- scan_gaps(elec_ts)
+
+## ---- count-gaps
 customer_na <- elec_ts %>% 
   filter(customer_id %in% (gap_df %>% filter(.gaps) %>% pull(customer_id)))
 
 count_na_df <- customer_na %>% 
   count_gaps()
 
-# ToDo: do a hierarchical clustering for classifying customers
-count_na_10 <- count_na_df %>% 
-  count(customer_id) %>% 
-  top_n(100) %>% 
-  pull(customer_id)
+count_lvl <- count_na_df %>% 
+  group_by(customer_id) %>% 
+  summarise(n_miss = sum(.n)) %>% 
+  mutate(customer_id = fct_reorder(as.character(customer_id), n_miss)) %>% 
+  pull(customer_id) %>% 
+  levels()
 
 count_na_df %>% 
-  filter(customer_id %in% count_na_10) %>% 
-  ggplot(aes(x = as.factor(customer_id))) +
+  mutate(customer_id = factor(customer_id, count_lvl)) %>% 
+  ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to)) +
   geom_point(aes(y = .from)) +
   geom_point(aes(y = .to)) +
   coord_flip() +
-  theme(legend.position = "bottom")
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom"
+  )
 
-qtl_grid <- seq(0.1, 0.9, 0.01)
+## ---- elec-quantiles
+qtl_grid <- seq(0.1, 0.9, 0.002)
 elec_qtl <- elec_ts %>% 
   summarise(
     value = list(quantile(general_supply_kwh, qtl_grid, na.rm = TRUE)),
@@ -49,18 +59,25 @@ elec_qtl <- elec_ts %>%
     date = as_date(reading_datetime)
   )
 
+elec_med <- elec_qtl %>% 
+  filter(qtl == 0.5)
+
 elec_qtl %>% 
[…]

### 2018-12-11 main.Rmd
rewording

@@
     email : rob.hyndman@monash.edu
-abstract: 'Mining temporal data for information is often inhibited by a multitude of time formats: irregular or multiple time intervals, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
-Time series models, in particular, the software supporting time series forecasting makes strict assumptions on data to be provided, typically a matrix of numeric data with an implicit time index.
+abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point event that needs aggregating, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
+Time series models, in particular, the software supporting time series forecasting makes strict assumptions on the data to be provided, typically a matrix of numeric data with implicit time indices.
 Going from raw data to model-ready data is painful.
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
-Tidy data principles are applied, and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) including an explicitly declared index variable representing time, (3) incorporating a "key" comprised of single or multiple variables to uniquely identify units over time.
+Tidy data principles are extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) including an explicitly declared index variable representing time, (3) incorporating a "key" comprised of single or multiple variables to uniquely identify units over time.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
-A sound data pipeline facilitates a fluent and transparent workflow for analyzing temporal data.
-Applications are included to illustrate tidy temporal data structure, data pipeline structure and usage.
+A sound data pipeline facilitates a fluent and fluid workflow for analyzing temporal data.
 The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
@@
 Each verb focuses on a singular task.
-Almost all data tasks can be rephrased and tackled with these five key verbs, by composing them sequentially.
+Most common data tasks can be rephrased and tackled with these five key verbs, by composing them sequentially.
 
-Community effort for building tidy ecosystem
+The **tidyverse** way of thinking largely formalizes exploratory data analysis. Tidy concepts have been extended to other domains in R community.
 
@@
 Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
-Index, key and time interval are the three stone pillars to this new semantically-structured temporal data, which we name as "tsibble".
+Index, key and time interval are the three stone pillars to this new semantically-structured temporal data, which we name as "tsibble", inspired by "ts" and "tibble".
 
@@
 
-The tidy data abstraction lays a pipeline infrastructure for data analysis modules, transformation, visualization and modelling.
-Each module communicates between each other, requiring tidy input, producing tidy output, and consequently chains a series of operations together to accomplish the analytic tasks at hand.
+The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modelling.
+Each module communicates between each other, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.
 
@@
 The time-based pipeline shepards raw temporal data through to time series analysis, and plots.
-It is advised to scrutinize identical entries of key and index in the right beginning.
+It is possible and recommended to check for identical entries of key and index before analysis.
 Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making.
-Analysts are encouraged to gaze at data as earlier as possible and reason about the process of data cleaning.
-When the data meets the tsibble standard, it flows into the exploration stage.
+Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
+When the data meets the tsibble standard, it flows neatly into the exploration stage.
 
@@
 Data needs transformation to some extent for modelling.
-A suite of verbs are introduced to flatten the lumpy path from temporal data to the object that directs to modelling under the tsibble framework, as well as to facilitate transforming tsibble in various kinds of shapes for analysis and visualization.
+A suite of verbs are introduced to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, as well as transformed for visualization.
 The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood.
-If a tsibble cannot hold, for example swiping index off, an error prompts users to complain and suggest alternatives.
-This warrants users least surprise and reminds them of time awareness.
+Attention has been paid to error handling.
+If a tsibble cannot be maintained in the output of a pipeline module, for example the index is dropped by aggregation, an error informs users of the problem and suggests alternatives.
+This avoids negatively surprising users and reminds them of time context.
 
@@
 Each verb focuses on one thing and achieve its goal.
-The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picking observations, and `select()` picking variables.
+The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
 These general-purpose verbs are made available in the **tidyverse** suite.
@@
  & \textbf{\texttt{nest()}} & Eirik Bakke \\
- & \textbf{\texttt{unnest()}} & Jody Morris \\
+ & \textbf{\texttt{unnest()}} & Jody Morris \\ \hline
+\multirow{6}{*}{Join data} & \textbf{\texttt{left\_join()}} & David Batty \\
+ & \textbf{\texttt{right\_join()}} & Eirik Bakke \\
+ & \textbf{\texttt{full\_join()}} & Eirik Bakke \\
+ & \textbf{\texttt{inner\_join()}} & Jody Morris \\
+ & \textbf{\texttt{semi\_join()}} & Jody Morris \\
+ & \textbf{\texttt{anti\_join()}} & Jody Morris \\
 \bottomrule
@@
 
-* **rolling window**: `slide()`, `tile()`, `stretch()`
+A set of verbs have been implemented to conduct temporal operations. There are three kinds of sliding window operations:
+
+* `slide()`/`slide2()`/`pslide()`: sliding window with overlapping observations.
+* `tile()`/`tile2()`/`ptile()`: tiling window without overlapping observations.
+* `stretch()`/`stretch2()`/`pstretch()`: fixing an initial window and expanding to include more observations.
+
+As the name indicates, `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
+This family of window functions features functional programming.
+Any function can be mapped and rolled over a block of elements given a specified window size.
+For type stability, the functions always return lists.
+Other variants including `slide_lgl()`, `slide_int()`, `slide_dbl()`, `slide_chr()` return vectors of the corresponding type, as well as `slide_dfr()` and `slide_dfc()` for row-binding and column-binding data frames respectively.
+Their multiprocessing equivalents prefixed by `future_` enable rolling in parallel.
+This family of functions empowers users to incorporate window-related operations in their workflows.
 

### 2018-12-23 main.Rmd
updated flowcharts

@@
 @r4ds succinctly describes the data science workflow given in Figure \ref{fig:pipeline-ann}.
-It is annotated to show that the divide between data that feeds into the beginning of analysis and the current time series objects that fits into modelling is significant, making data preparation for modeling and forecasting painful.
+It is annotated to emphasize the significant divide between data that feeds into the beginning of analysis and the current time series objects that fits into modelling, making data preparation for modeling and forecasting painful.
 
 ```{r pipeline-ann, fig.cap = "(ref:pipeline-ann-cap)"}
-include_graphics("img/time-series-structure.pdf")
+include_graphics("img/motivation.pdf")
 ```
 
-(ref:pipeline-ann-cap) An overview of data science program. The existing standards are model-focused, resulting misalliance between data and model.
+(ref:pipeline-ann-cap) An overview of the data science program centering around data.
+Temporal data should feed into the import stage, but the existing time series objects are model-focused, thus resulting misalliance between data and model.
 
-The existing standards, particularly time series objects, are model-oriented with little support for non-numeric values, leaning its data abstraction towards matrix.
-Time series have been isolated from general data problems due to this data choice.
-Analysts who start analysing time series would be pushed to learn new transformation and visualization toolkits that are different from what they used to operate.
+Current time series objects, which are model-oriented matrices with little support for non-numeric values, feed into the framework solely at the modelling stage.
+Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.
+The diagram correspondingly would place the model at the centre of the analytical universe, and all the transformations, visualisations would hinge on that format.
 
@@
 
-The **tidyverse** way of thinking largely formalizes exploratory data analysis. Tidy concepts have been extended to other domains in R community.
-
-Temporal data is integrated to tidy data and the grammar. @r4ds argues that 80% of data analysis tasks can be solved with tidy tools while the remaining 20% requires other tools.
+The **tidyverse** largely formalizes exploratory data analysis.
+The whole R community has quickly adopted the **tidyverse** way of thinking to much broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
+Temporal data is left behind in this regard.
+This paper proposes a new tidy representation for temporal data and a unified framework to streamline the workflow from data pre-processing to visualization and forecasting, as an integral part of tidy data analysis.
 
@@
 
-The choice of tidy representation of temporal data is made from a data-centric perspective, which is taken in the light of the operations that are to be performed on the data.
+```{r tsibble-fit, fig.cap = "(ref:tsibble-fit-cap)"}
+include_graphics("img/tsibble-fit.pdf")
+```
+
+(ref:tsibble-fit-cap) Tsibble's abstraction arises from a data-centric point of view, which could neatly embed into the tidyverse ecosystem.
+It should be noted that tsibble positions itself at the "tidy" step rather than "import", which ensures valid tidy data to time series analytics and models.
+
+The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
 Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
@@
 
-A set of verbs have been implemented to conduct temporal operations. There are three kinds of sliding window operations:
+Temporal data often involves moving window calculations. A set of verbs have been implemented to allow for different variations of sliding windows operations:
 
@@
 
-Time series are represented as matrices, with standards being provided by native `ts` object in R, extended by **zoo** [@zoo2005], and **xts** [@R-xts].
-As discussed in Section \ref{sec:time-series}, this data organization sets far apart from data origin.
-Due to time isolated from the main data, the supporting functions that intend to deal with time must be applied to whole data.
-That data-centric & human-centered.
+Matrices are a conventional way of organizing time series, with standards being provided by native `ts` object in R, extended by **zoo** [@zoo2005], and **xts** [@R-xts].
+Time indices are stored as an attribute, separating from the main body of the data.
+
+data-centric & human-centered.
 

### 2018-12-24 main.Rmd
update verb table

@@
 The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
+Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we name as "tsibble", inspired by "ts" and "tibble".
+It bridges the gap between raw temporal data and model inputs.
 Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
-Index, key and time interval are the three stone pillars to this new semantically-structured temporal data, which we name as "tsibble", inspired by "ts" and "tibble".
+Index, key and time interval are the three stone pillars to this new semantically-structured temporal data.
 
@@
 Lacking in either of two will be inadequate for the key and thus fail to construct a tsibble.
+The alternative specification of the key is to include a third variable `continent`.
+Since `country` is nested within `continent`, it is a free variable for use.
+But this variable brings additional information that can be used for forecasting reconciliation [@fpp].
 The key is explicit when multiple units exist in the data.
@@
 \begin{table}[!hbt]
-\caption{A test}\label{tab:verb-table}
+\caption{A list of table verbs working with tsibble. Functions in bold originating from the tidyverse and are adapted for tsibble.}\label{tab:verb-table}
 \centering
@@
  & \textbf{\texttt{mutate()}} & Add new variables \\
- & \textbf{\texttt{transmute()}} & Add new variables \\
+ & \textbf{\texttt{transmute()}} & Drops existing variables \\
  & \textbf{\texttt{summarise()}} & Aggregate values over time \\ \hline
@@
  & \texttt{group\_by\_key()} & Group by key variables \\ \hline
-\multirow{4}{*}{Reshape data} & \textbf{\texttt{gather()}} & David Batty \\
- & \textbf{\texttt{spread()}} & Eirik Bakke \\
- & \textbf{\texttt{nest()}} & Eirik Bakke \\
- & \textbf{\texttt{unnest()}} & Jody Morris \\ \hline
-\multirow{6}{*}{Join data} & \textbf{\texttt{left\_join()}} & David Batty \\
- & \textbf{\texttt{right\_join()}} & Eirik Bakke \\
- & \textbf{\texttt{full\_join()}} & Eirik Bakke \\
- & \textbf{\texttt{inner\_join()}} & Jody Morris \\
- & \textbf{\texttt{semi\_join()}} & Jody Morris \\
- & \textbf{\texttt{anti\_join()}} & Jody Morris \\
+\multirow{4}{*}{Reshape} & \textbf{\texttt{gather()}} & Gather columns into long form \\
+ & \textbf{\texttt{spread()}} & Spread columns into wide form \\
+ & \textbf{\texttt{nest()}} & Nest values in a list-variable \\
+ & \textbf{\texttt{unnest()}} & Unnest a list-variable \\ \hline
+\multirow{6}{*}{Join tables} & \textbf{\texttt{left\_join()}} & \multirow{6}{*}{Join two tables together} \\
+ & \textbf{\texttt{right\_join()}} & \\
+ & \textbf{\texttt{full\_join()}} & \\
+ & \textbf{\texttt{inner\_join()}} & \\
+ & \textbf{\texttt{semi\_join()}} & \\
+ & \textbf{\texttt{anti\_join()}} & \\
 \bottomrule

### 2018-12-25 main.Rmd
added animated window plots

@@
  & \textbf{\texttt{summarise()}} & Aggregate values over time \\ \hline
-\multirow{3}{*}{Group-wise} & \texttt{index\_by()} & Group by new time index \\
+\multirow{3}{*}{Group-wise} & \texttt{index\_by()} & Group by index candidate \\
  & \textbf{\texttt{group\_by()}} & Group by one or more variables \\
@@
 * `stretch()`/`stretch2()`/`pstretch()`: fixing an initial window and expanding to include more observations.
+
+```{r tb-au}
+```
+
+```{r slide-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:slide-animate-cap)"}
+```
+
+(ref:slide-animate-cap) An illustration of window of size 5 sliding over annual tuberculosis cases in Australia. View the animation with Adobe Acrobat Reader.
+
+```{r tile-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:tile-aniamte-cap)"}
+```
+
+(ref:tile-animate-cap) An illustration of window of size 5 tiling over annual tuberculosis cases in Australia. View the animation with Adobe Acrobat Reader.
+
+```{r stretch-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:stretch-animate-cap)"}
+```
+
+(ref:stretch-animate-cap) An illustration of window of size 5 stretching over annual tuberculosis cases in Australia. View the animation with Adobe Acrobat Reader.
 

### 2018-12-25 R/demo.R

@@ -5,9 +5,72 @@ library(tsibble)
 ## ---- tb-sub
 tb_small <- read_rds("data/tb-small.rds")
 tb_small %>% 
+  filter(year > 2010) %>% 
   knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>% 
   kableExtra::kable_styling(position = "center")
 
 ## ---- tb-print
 as_tsibble(tb_small, key = id(country, gender), index = year) %>% 
+  filter(year > 2010) %>% 
   print(n = 5)
+
+## ---- tb-au
+tb_au <- tb_small %>% 
+  filter(country == "Australia") %>% 
+  group_by(year) %>% 
+  summarise(count = sum(count))
+
+## ---- slide-animate
+library(gganimate)
+slide_window <- slider(tb_au$year, .size = 5) %>%
+  map_dfr(function(x) tibble(xmin = min(x), xmax = max(x))) %>%
+  mutate(ymin = -Inf, ymax = Inf, group = row_number())
+ggplot() +
+  geom_line(aes(year, count), data = tb_au) +
+  geom_point(aes(year, count), data = tb_au) +
+  geom_rect(aes(
+    xmin = xmin, xmax = xmax,
+    ymin = ymin, ymax = ymax,
+    group = group
+  ), data = slide_window, fill = "#9ecae1", alpha = 0.6) +
+  xlab("Year") +
+  ylab("Count") +
+  ylim(c(0, max(tb_au$count))) +
+  theme_bw() +
+  transition_manual(group)
+
+## ---- tile-animate
+tile_window <- tiler(tb_au$year, .size = 5) %>%
+  map_dfr(function(x) tibble(xmin = min(x), xmax = max(x))) %>%
+  mutate(ymin = -Inf, ymax = Inf, group = row_number())
+ggplot() +
+  geom_line(aes(year, count), data = tb_au) +
+  geom_point(aes(year, count), data = tb_au) +
+  geom_rect(aes(
+    xmin = xmin, xmax = xmax,
+    ymin = ymin, ymax = ymax,
+    group = group
+  ), data = tile_window, fill = "#9ecae1", alpha = 0.6) +
+  xlab("Year") +
+  ylab("Count") +
+  ylim(c(0, max(tb_au$count))) +
+  theme_bw() +
+  transition_manual(group)
+
+## ---- stretch-animate
[…]

### 2018-12-25 data-raw/tb-small.R

@@ -5,8 +5,7 @@ tb <- read_csv("data-raw/tb-2018-06-05.csv")
 tb_small <- tb %>% 
   select(-region, -age) %>% 
   filter(
-    country %in% c("Australia", "New Zealand", "United States of America"),
-    year > 2010
+    country %in% c("Australia", "New Zealand", "United States of America")
   ) %>% 
   group_by(country, continent, gender, year) %>% 
   summarise(count = sum(count)) %>% 

### 2018-12-25 R/flights.R
changed daily % plot aspect ratio

@@ -159,7 +159,7 @@ sel_monthly %>%
   facet_grid(origin ~ .) +
   scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
   scale_colour_brewer(palette = "Dark2") +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom", aspect.ratio = 0.2) +
   xlab("Date") +
   ylab("Departure delay")
 

### 2018-12-29 main.Rmd
changed daily % plot aspect ratio
added tibbletime

@@
 Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.
-The diagram correspondingly would place the model at the centre of the analytical universe, and all the transformations, visualisations would hinge on that format.
 
@@
 Temporal data is left behind in this regard.
+
+## Existing time series standards in R
+
+Current standards, provided by native `ts` object in R and extended by **zoo** [@zoo2005], and **xts** [@R-xts], organize temporal data in matrices with implicit time indices.
+These objects were designed for modelling methods.
+The diagram correspondingly would place the model at the center of the analytical universe, and all the transformations, visualisations would hinge on that format.
+This is contrary to the tidyverse conceptualisation, which wholistically captures the full data workflow.
+
+A new temporal data class is needed in the upstream of the workflow, which could incorporate all the downstream modules.
+A relatively new R package **tibbletime** proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format.
+It only requires an index variable to declare a temporal data object, thus placing at the import stage.
+However, this is a loosely-defined structure and insufficient for time series analytics and models.
+Because it does not define interested subjects ($i$) over time and time uniqueness.
+
 This paper proposes a new tidy representation for temporal data and a unified framework to streamline the workflow from data pre-processing to visualization and forecasting, as an integral part of tidy data analysis.
@@
 The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
-Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we name as "tsibble", inspired by "ts" and "tibble".
-It bridges the gap between raw temporal data and model inputs.
-Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
-Index, key and time interval are the three stone pillars to this new semantically-structured temporal data.
+Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as "tsibble".
 
@@
 Despite column `count` is the only measurement in this case, it is sufficiently flexible to hold other measured variables, for example, adding the corresponding population size (if known) in order to calibrate the count later.
+
+It bridges the gap between raw temporal data and model inputs.
+Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
+Index, key and time interval are the three stone pillars to this new semantically-structured temporal data.
 
@@
 
-Matrices are a conventional way of organizing time series, with standards being provided by native `ts` object in R, extended by **zoo** [@zoo2005], and **xts** [@R-xts].
-Time indices are stored as an attribute, separating from the main body of the data.
-
 data-centric & human-centered.
@@
 
-```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 6}
+```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 7}
 ```

### 2018-12-30 main.Rmd
two more paragraphs for rolling window

@@
 
-Temporal data often involves moving window calculations. A set of verbs have been implemented to allow for different variations of sliding windows operations:
+Rolling window calculations are widely-used techniques in time series analysis, but are not restricted to temporal applications.
+These operations are dependent on certain ordering, particularly time ordering for temporal data.
+Three common types of variations for sliding window operations are:
 
-* `slide()`/`slide2()`/`pslide()`: sliding window with overlapping observations.
-* `tile()`/`tile2()`/`ptile()`: tiling window without overlapping observations.
-* `stretch()`/`stretch2()`/`pstretch()`: fixing an initial window and expanding to include more observations.
+1. **slide**: sliding window with overlapping observations.
+2. **tile**: tiling window without overlapping observations.
+3. **stretch**: fixing an initial window and expanding to include more observations.
+
+Figure \@ref(fig:slide-animate), \@ref(fig:tile-animate) and \@ref(fig:stretch-animate) show the animations of rolling windows for sliding, tiling and stretching respectively on annual tuberculosis cases in Australia.
+A block of consecutive elements with window size of 5 are initialized and started rolling sequentially till the end of series.
 
@@
 
-```{r tile-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:tile-aniamte-cap)"}
+```{r tile-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:tile-animate-cap)"}
 ```
@@
 
+Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in Table \@ref(tab:verb-table).
+Table verbs expect and return a tsibble, and does something that the function name suggests.
+On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window.
+For example, computing moving averages requires numerics and a function like `mean()`, and produce averaged numerics; however, rolling window regression takes a data frame and a linear regression method like `lm()`, and generate a complex object that contains coefficients, fitted values, and etc.
+In addition, the same function will be called and applied to each window of elements independently, which means programming is done with for-loop statements.
+But functional programming emphasizes on expressions instead of statements.
+
 As the name indicates, `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
-This family of window functions features functional programming.
-Any function can be mapped and rolled over a block of elements given a specified window size.
+An arbitrary function can be mapped and rolled over a block of elements given a specified window size.
 For type stability, the functions always return lists.
-Other variants including `slide_lgl()`, `slide_int()`, `slide_dbl()`, `slide_chr()` return vectors of the corresponding type, as well as `slide_dfr()` and `slide_dfc()` for row-binding and column-binding data frames respectively.
+Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding type, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively.
 Their multiprocessing equivalents prefixed by `future_` enable rolling in parallel.

### 2018-12-31 main.Rmd
mv fun prog to software design

@@
 
-## Rolling window in functional programming
+# Software structure and design decisions {#sec:software}
+
+## Data first
+
+The prime force that drives the software's design choices is "data".
+All functions in **tsibble** starts with `data` or its variants as the first argument, namely "data first".
+They work naturally with the pipe operator `%>%`, read as "then".
+This not only lays out a consistent interface but also addresses the significance of the data throughout the software.
+
+Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
+Below conveys the pieces of critical and contextual picture about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series.
+These details aid users in understanding their data better and manipulate the data with care.
+
+```{r tb-print}
+```
+
+## Functional programming
 
@@
 
-As the name indicates, `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
-An arbitrary function can be mapped and rolled over a block of elements given a specified window size.
+The **purrr** package sets a great example of functional programming in R.
+It provides a complete and consistent set of tools to iterate each element of a vector with a function.
+Rolling window not just iterates but roll over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
+`slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
 For type stability, the functions always return lists.
@@
 This family of functions empowers users to incorporate window-related operations in their workflows.
-
-# Software structure and design decisions {#sec:software}
-
-data-centric & human-centered.
-
-## Data first
-
-The prime force that drives the software's design choices is "data".
-All functions in **tsibble** starts with `data` or its variants as the first argument, namely "data first".
-They work naturally with the pipe operator `%>%`, read as "then".
-This not only lays out a consistent interface but also addresses the significance of the data throughout the software.
-
-Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-Below conveys the pieces of critical and contextual picture about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series.
-These details aid users in understanding their data better and manipulate the data with care.
-
-```{r tb-print}
-```
 
@@
 
+```{r calendar-plot, fig.height = 7, fig.cap = "(ref:calendar-plot-cap)"}
+```
+
+(ref:calendar-plot-cap) Half-hourly average electricity use.
+
 # Conclusion and future work {#sec:conclusion}

### 2018-12-31 R/smart-meter.R

@@ -26,15 +26,10 @@ customer_na <- elec_ts %>%
 count_na_df <- customer_na %>% 
   count_gaps()
 
-count_lvl <- count_na_df %>% 
-  group_by(customer_id) %>% 
-  summarise(n_miss = sum(.n)) %>% 
-  mutate(customer_id = fct_reorder(as.character(customer_id), n_miss)) %>% 
-  pull(customer_id) %>% 
-  levels()
-
 count_na_df %>% 
-  mutate(customer_id = factor(customer_id, count_lvl)) %>% 
+  mutate(
+    customer_id = as.factor(customer_id) %>% fct_lump(50) %>% fct_reorder(.n)
+  ) %>% 
   ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to)) +
   geom_point(aes(y = .from)) +
@@ -46,6 +41,21 @@ count_na_df %>%
     legend.position = "bottom"
   )
 
+## ---- calendar-plot
+library(sugrrants)
+elec_cal <- elec_ts %>% 
+  summarise(avg = mean(general_supply_kwh)) %>% 
+  mutate(
+    date = as_date(reading_datetime), 
+    time = hms::as.hms(reading_datetime, tz = "UTC")
+  ) %>% 
+  frame_calendar(x = time, y = avg, date = date)
+
+p_cal <- elec_cal %>% 
+  ggplot(aes(.time, .avg, group = date)) +
+  geom_line()
+prettify(p_cal)
+
 ## ---- elec-quantiles
 qtl_grid <- seq(0.1, 0.9, 0.002)
 elec_qtl <- elec_ts %>% 
@@ -55,7 +65,7 @@ elec_qtl <- elec_ts %>%
   ) %>% 
   unnest(key = id(qtl)) %>% 
   mutate(
-    time = hms::as.hms(reading_datetime),
+    time = hms::as.hms(reading_datetime, tz = "UTC"),
     date = as_date(reading_datetime)
   )
 

### 2019-01-03 main.Rmd
added a couple sentences to future work

@@
 Each module communicates between each other, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.
+With the pipe operator `%>%`, read from left to right, and omit intermediate values.
 
@@
 
-The **tsibble** package provides an elegant solution to manage and manipulate medium-sized temporal data in memory.
+The **tsibble** package provides the grammar of temporal data manipulation, regardless of how the data is stored.
+Currently, it works for managing and manipulating temporal data frames in memory locally.
+But it is possible to work with remote tables stored in databases, such as SQLite and MySQL, using exactly the same tsibble code.
+This is the future work.
 

### 2019-01-03 R/smart-meter.R

@@ -163,3 +163,33 @@ gas_aircon_avg %>%
   facet_grid(has_gas ~ ., labeller = "label_both") +
   scale_y_log10()
 names(customer)
+
+## ---- forecast
+library(fable)
+elec_mth <- elec_ts %>% 
+  filter_index(~ "2013-01-30") %>% 
+  index_by(datehour = floor_date(reading_datetime, "hour")) %>% 
+  summarise(avg_supply = mean(general_supply_kwh)) %>% 
+  mutate(
+    hour = hour(datehour),
+    date = as_date(datehour)
+  )
+
+elec_fc <- elec_mth %>% 
+  model(ets = ETS(avg_supply)) %>% 
+  forecast(h = 24)
+elec_fc$date <- as_date(elec_fc$datehour)
+
+# elec_fc %>% 
+#   autoplot(data = elec_mth) +
+#   sugrrants::facet_calendar(~ date)
+
+elec_mth %>% 
+  ggplot(aes(x = datehour, y = avg_supply)) +
+  geom_line() +
+  geom_forecast(
+    aes(ymin = lower, ymax = upper, level = level),
+    fortify(elec_fc) %>% mutate(date = as_date(datehour)), stat = "identity"
+  ) +
+  NULL
+  sugrrants::facet_calendar(~ date)

### 2019-01-04 main.Rmd
a bit more on implicit na

@@
 Each module communicates between each other, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.
-With the pipe operator `%>%`, read from left to right, and omit intermediate values.
 
@@
 
-XXX Put comments in above each paragraph, with the topic of each paragraph. its not clear how this section flows yet. the topics would help.
-
 The time-based pipeline shepards raw temporal data through to time series analysis, and plots.
@@
 Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
-When the data meets the tsibble standard, it flows neatly into the exploration stage.
+When the data meets the tsibble standard, it flows neatly into the analysis stage.
 
-Time series models typically assume that the input is a complete and regularly-spaced series, which most temporal data can barely satisfy.
-Temporal data becomes more granular in time resolution and more disaggregated to individual level.
-This inevitably goes with implicit missing values and noisiness.
+Many time operations like lag/lead and time series models, assume an intact vector input ordered in time.
+To prevent inviting these errors into the analysis, it is a good practice to check and inspect any time gaps of the tsibble in the first place.
+Several tools are provided to understand and tackle implicit missing values in: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reports all implicit missing observations; (3) `count_gaps()` summarises the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
+
 Data needs transformation to some extent for modelling.
@@
 This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems.
+Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package.
+A sequence of functions can be composed in a way that can be naturally read from left to right, which incredibly increase the readability of the code.
+It consequently generates a block of code without worrying naming intermediate values.
 Most importantly, the ecosystem for tidy time series analysis has been undertaking on the basis of tsibble in R, known as "tidyverts".

### 2019-01-06 main.Rmd
updated flights case study

@@
 
-## Time-based pipeline {#sec:timepipeline}
+## Time-based transformation {#sec:timepipeline}
 
@@
 
-As a special case of data frame, a tsibble pipes into the grammar of graphics straight way, making most use of this powerful graphical system.
-It should be easy to create and extend some specialist plotting methods based on tsibble structure, such as autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
+## Time series visualization and models {#sec:ts-models}
+
+As a special case of the data frame, a tsibble pipes into the grammar of graphics straight way, making most use of this powerful graphical system.
+It should be easy to create and extend some specialized time-related plotting methods based on tsibble structure, such as autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
 
@@
 
-A closer examination of New York airports will give an indication about how well the busiest airports manage the outflow traffic on a daily basis.
-A subset that contains observations for EWR, JFK and LGA airports is obtained first.
+A closer examination of some big airports across the US will give an indication about how well the busiest airports manage the outflow traffic on a daily basis.
+A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (Hawaii), Los Angeles (LAX) and Seatle (SEA) airports is obtained first.
 The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:sel-monthly-plot).
-LGA fluctuates a lot compared to the other two.
+Winter months tend to fluctuate a lot compared to the summer across all the airports.
 What superimposes on these lines is two-month moving averages so that a temporal trend is more visible.
@@
 But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without worrying weights specification.
-The sliding with large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
+The sliding operation with large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
+The LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady control of delays over time.
+The IAH and JFK airports have undergone more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
 
@@
 (ref:sel-monthly-plot-cap) Daily delayed percentages for departure with two-month moving averages overlaid at five international airports.
-There are many fluctuations observed at LGA airport.
-The estimates of temporal trend are around 25% across three airports, highlighting relatively less delay in Fall.
+There are least fluctuations and relatively fewer delays observed at KOA airport.
+The estimates of temporal trend are around 25% across the other four airports, but highlight different time periods of severe delays.
 
@@
 Sensors have been put up to collect data for the project of smart city across major cities in Australia.
-One of the trials is monitoring households' electricity usage through installed smart meters in the area of New Castle over 2010--2014 [@smart-meter].
+One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter].
 Year 2013 has been sliced to examine temporal patterns of customer's energy consumption with **tsibble** in this paper.
@@
 
+Aggregation over all the individuals helps to grasp a big picture about if the behaviors are affected by the time.
+
 ```{r calendar-plot, fig.height = 7, fig.cap = "(ref:calendar-plot-cap)"}
@@
 
-(ref:calendar-plot-cap) Half-hourly average electricity use.
+(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized in the calendar format.
+Distinctive patterns between summer and winter months occur to daily residential energy consumption.
+Air conditioning drives the peak in the late afternoon in January and December, however, peaks in the morning and evening are more pronounced in the winter months due to heating use.
+
+```{r count-gaps, fig.height = 7, fig.cap = "(ref:count-gaps-cap)"}
+```
 

### 2019-01-06 R/smart-meter.R

@@ -32,13 +32,15 @@ count_na_df %>%
   ) %>% 
   ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to)) +
-  geom_point(aes(y = .from)) +
-  geom_point(aes(y = .to)) +
+  geom_point(aes(y = .from), size = 0.6) +
+  geom_point(aes(y = .to), size = 0.6) +
   coord_flip() +
+  xlab("Customer") +
+  ylab("Time gaps") +
   theme(
     axis.text.y = element_blank(),
     axis.ticks.y = element_blank(),
-    legend.position = "bottom"
+    legend.position = "none"
   )
 
 ## ---- calendar-plot

### 2019-01-08 main.Rmd
added count-gaps plot

@@
 
+During a power outage, electricity usage for some households may become unavailable, thus resulting in implicit missing values in the database.
+Gaps in time occur to `r scales::percent(sum(gap_df$.gaps) / NROW(gap_df))` of the households in this dataset.
+It would be interesting to explore these missing patterns as part of preliminary analysis.
+Since the smart meter has been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for the households, instead of full lengths across the households.
+Figure \@ref(fig:count-gaps) reveals the gaps for the top 49 households and the rest collapsed into a single batch arranged in rows from high to low in tallies.
+Missing values can be found at any time point during the entire time span.
+A small number of customers have undergone energy unavailability in consecutive hours, indicated by a line range in the plot.
+On the other hand, the majority suffer occasional breakdown with more frequent happenings in January.
+
+```{r count-gaps, fig.height = 7, fig.cap = "(ref:count-gaps-cap)"}
+```
+
+(ref:count-gaps-cap) Time gap plots for the top 49 customers with most implicit missing values and the rest grouped into the one.
+Each cross represents the missing in time and a line between two dots shows continuous missingness over time.
+Each row corresponds to one customer describes the incidences of time gaps.
+
 Aggregation over all the individuals helps to grasp a big picture about if the behaviors are affected by the time.
@@
 Air conditioning drives the peak in the late afternoon in January and December, however, peaks in the morning and evening are more pronounced in the winter months due to heating use.
-
-```{r count-gaps, fig.height = 7, fig.cap = "(ref:count-gaps-cap)"}
-```
 

### 2019-01-08 R/smart-meter.R

@@ -28,12 +28,13 @@ count_na_df <- customer_na %>%
 
 count_na_df %>% 
   mutate(
-    customer_id = as.factor(customer_id) %>% fct_lump(50) %>% fct_reorder(.n)
+    customer_id = as.factor(customer_id) %>% 
+      fct_lump(50) %>% fct_reorder(.n, sum)
   ) %>% 
   ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to)) +
-  geom_point(aes(y = .from), size = 0.6) +
-  geom_point(aes(y = .to), size = 0.6) +
+  geom_point(aes(y = .from), size = 0.6, shape = 4) +
+  geom_point(aes(y = .to), size = 0.6, shape = 4) +
   coord_flip() +
   xlab("Customer") +
   ylab("Time gaps") +

### 2019-01-08 main.Rmd
added description to calendar plot

@@
 
-Aggregation over all the individuals helps to grasp a big picture about if the behaviors are affected by the time.
+Aggregation across all individuals helps to sketch a big picture about if the representative behaviors change over time, laid out in the calendar display shown as Figure \@ref(fig:calendar-plot).
+Each glyph represents the daily pattern of residential electricity usage every thirty minutes with higher consumption in the time when human are more active.
+The daily snapshots diverge in traces, depending on the season in the year.
+During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially in the extremely hot days.
+However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.
+This also illustrates how nicely the tsibble data plays with other tools and graphics.
 

### 2019-01-08 main.Rmd
title change

--- 2019-01-08 main.Rmd
@@
 ---
-title: "Tidy data structure to support exploration and modeling of temporal data"
+title: "A new tidy data structure to support exploration and modeling of temporal data"
 author:

### 2019-01-09 main.Rmd
figures merged

@@
 Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
-@r4ds succinctly describes the data science workflow given in Figure \ref{fig:pipeline-ann}.
+@r4ds succinctly describes the data science workflow given in Figure \ref{fig:tsibble-fit}.
 It is annotated to emphasize the significant divide between data that feeds into the beginning of analysis and the current time series objects that fits into modelling, making data preparation for modeling and forecasting painful.
 
+<!--
 ```{r pipeline-ann, fig.cap = "(ref:pipeline-ann-cap)"}
@@
 Temporal data should feed into the import stage, but the existing time series objects are model-focused, thus resulting misalliance between data and model.
+-->
+
+```{r tsibble-fit, fig.cap = "(ref:tsibble-fit-cap)"}
+include_graphics("img/tsibble-fit.pdf")
+```
+
+(ref:tsibble-fit-cap) Illustration of data science workflow, drawn from @r4ds, showing how current time series tools interface with the workflow and how tsibble structure and tools integrate.
+The new data structure, tsibble, makes the connection between temporal data input, and specialist modeling formats.
+It provides elements at the "tidy" step, which produce tidy temporal data for time series visualisation and modeling.
 
@@
 # Contextual semantics {#sec:semantics}
-
-```{r tsibble-fit, fig.cap = "(ref:tsibble-fit-cap)"}
-include_graphics("img/tsibble-fit.pdf")
-```
-
-(ref:tsibble-fit-cap) Tsibble's abstraction arises from a data-centric point of view, which could neatly embed into the tidyverse ecosystem.
-It should be noted that tsibble positions itself at the "tidy" step rather than "import", which ensures valid tidy data to time series analytics and models.
 

### 2019-01-09 main.Rmd
new diagrams

@@
 
+```{r tsibble-semantics, fig.cap = "(ref:tsibble-semantics-cap)"}
+include_graphics("img/tsibble-semantics.pdf")
+```
+
+(ref:tsibble-semantics-cap) The architecture of the tsibble structure is built on top of the data frame---a two-dimensional array in R, with time series semantics: index and key.
+
 The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
@@
 
-XXX may be good to have a diagram here showing how the time-based pipeline builds on a regular data pipeline
+```{r tsibble-pipeline, fig.cap = "(ref:tsibble-pipeline-cap)"}
+include_graphics("img/tsibble-pipeline.pdf")
+```
+
+(ref:tsibble-pipeline-cap) A time series pipeline is different from a regular data pipeline: (1) check if the key and index uniquely identify each observation in the data whist creating a tsibble; (2) check if there are time gaps in the tsibble before analysis.
 

### 2019-01-09 R/flights.R

@@ -166,17 +166,16 @@ sel_monthly %>%
 ## ---- quantile
 hr_qtl <- us_flights %>% 
   index_by(dep_datehour = floor_date(sched_dep_datetime, "hour")) %>% 
-  summarise(    
-    qtl50 = quantile(dep_delay, 0.5),
-    qtl80 = quantile(dep_delay, 0.8),
-    qtl95 = quantile(dep_delay, 0.95)
+  summarise(
+    value = list(quantile(dep_delay, c(0.5, 0.8, 0.95))),
+    qtl = list(paste0("qtl", c(50, 80, 95)))
   ) %>% 
+  unnest(key = id(qtl)) %>% 
   mutate(
     hour = hour(dep_datehour), 
     wday = wday(dep_datehour, label = TRUE, week_start = 1),
     date = as_date(dep_datehour)
-  ) %>% 
-  gather(key = qtl, value = dep_delay, qtl50:qtl95)
+  )
 
 ## ---- draw-qtl-prep
 break_cols <- c(

### 2019-01-12 main.Rmd
added pipeline flowcharts to flights section

@@
 A closer examination of some big airports across the US will give an indication about how well the busiest airports manage the outflow traffic on a daily basis.
-A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (Hawaii), Los Angeles (LAX) and Seatle (SEA) airports is obtained first.
+A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seatle (SEA) airports is obtained first.
 The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:sel-monthly-plot).
@@
 
+```{r flight-pipeline2, fig.cap = "(ref:flight-pipeline2-cap)"}
+include_graphics("img/flight-pipeline2")
+```
+
+(ref:flight-pipeline2-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:sel-monthly-plot}.
+
 What time of day and day of week should we travel to avoid suffering from horrible delay?
@@
 The variations increase substantially as the upper tails.
+
+```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)"}
+include_graphics("img/flight-pipeline3")
+```
+
+(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:draw-qtl}.
 

### 2019-01-12 R/flights.R

@@ -128,20 +128,14 @@ sel_flights <- us_flights %>%
 
 ## ---- sel-delay
 sel_delay <- sel_flights %>% 
-  mutate(delayed = dep_delay > 15) %>% 
   group_by(origin) %>% 
   index_by(sched_dep_date = as_date(sched_dep_datetime)) %>% 
-  summarise(
-    n_flights = n(),
-    n_delayed = sum(delayed)
-  ) %>% 
-  mutate(pct_delay = n_delayed / n_flights)
+  summarise(pct_delay = sum(dep_delay > 15) / n())
 
 ## ----- sel-monthly-1
 sel_lst <- sel_delay %>% 
   mutate(yrmth = yearmonth(sched_dep_date)) %>% 
-  group_by(origin, yrmth) %>% 
-  nest()
+  nest(-origin, -yrmth)
 
 ## ---- sel-monthly-2
 sel_monthly <- sel_lst %>% 

### 2019-01-13 main.Rmd
added temperature data

@@
 The daily snapshots diverge in traces, depending on the season in the year.
-During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially in the extremely hot days.
+During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially in the hot days with daily average temperature greater than 25 degrees.
 However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.
-This also illustrates how nicely the tsibble data plays with other tools and graphics.
+This also illustrates how the tsibble data can easily integrate with other tools and graphics.
 
@@
 Air conditioning drives the peak in the late afternoon in January and December, however, peaks in the morning and evening are more pronounced in the winter months due to heating use.
+
+```{r elec-pipeline2, fig.cap = "(ref:elec-pipeline2-cap)"}
+include_graphics("img/elec-pipeline2")
+```
+
+(ref:elec-pipeline2-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:calendar-plot}.
 

### 2019-01-13 R/smart-meter.R

@@ -5,6 +5,11 @@ library(tsibble)
 
 customer <- read_rds("data/customer.rds")
 elec <- read_rds("data/smart-meter13.rds")
+weather <- read_rds("data/weather13.rds") %>% 
+  group_by(date) %>% 
+  mutate(avg_temp = mean(c(max_temp, min_temp))) %>% 
+  ungroup() %>% 
+  mutate(hot = if_else(avg_temp > 25, "Hot day", "Not hot day"))
 
 elec_ts <- elec %>% 
   build_tsibble(
@@ -52,11 +57,14 @@ elec_cal <- elec_ts %>%
     date = as_date(reading_datetime), 
     time = hms::as.hms(reading_datetime, tz = "UTC")
   ) %>% 
+  left_join(weather, by = "date") %>% 
   frame_calendar(x = time, y = avg, date = date)
 
 p_cal <- elec_cal %>% 
-  ggplot(aes(.time, .avg, group = date)) +
-  geom_line()
+  ggplot(aes(.time, .avg, group = date, colour = hot)) +
+  geom_line() +
+  scale_colour_brewer(palette = "Dark2", direction = -1) +
+  theme(legend.position = "bottom")
 prettify(p_cal)
 
 ## ---- elec-quantiles
@@ -166,33 +174,3 @@ gas_aircon_avg %>%
   facet_grid(has_gas ~ ., labeller = "label_both") +
   scale_y_log10()
 names(customer)
-
-## ---- forecast
-library(fable)
-elec_mth <- elec_ts %>% 
-  filter_index(~ "2013-01-30") %>% 
-  index_by(datehour = floor_date(reading_datetime, "hour")) %>% 
-  summarise(avg_supply = mean(general_supply_kwh)) %>% 
-  mutate(
-    hour = hour(datehour),
-    date = as_date(datehour)
-  )
-
-elec_fc <- elec_mth %>% 
-  model(ets = ETS(avg_supply)) %>% 
-  forecast(h = 24)
-elec_fc$date <- as_date(elec_fc$datehour)
-
-# elec_fc %>% 
-#   autoplot(data = elec_mth) +
-#   sugrrants::facet_calendar(~ date)
-
-elec_mth %>% 
-  ggplot(aes(x = datehour, y = avg_supply)) +
-  geom_line() +
-  geom_forecast(
-    aes(ymin = lower, ymax = upper, level = level),
[…]

### 2019-01-18 main.Rmd
reworked intro

@@
 
-Suppose that temporal data consists of $N$ subjects/observational units.
-Given subject $i$ $(i = 1, 2, \dots, N)$, $X_{ijt}$ denotes the $j$th measurement $(j = 1, 2, \dots, p)$ indexed at a unique sequence of time points $0 \le t \le T_{i}$.
-Temporal data arrives in many possible formats, so does time.
-For example, time can carry with various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
+<!--Suppose that temporal data consists of $N$ observational units.
+Given observation $i$ $(i = 1, 2, \dots, N)$, $X_{ijt}$ denotes the $j$th measurement $(j = 1, 2, \dots, p)$ indexed at a unique sequence of time points $0 \le t \le T_{i}$.
+-->
+Temporal data arrives in many possible formats, as does the time context.
+For example, time can carry with it various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
 It could be irregularly recorded, which is particularly true with longitudinal measurements like patient visits to a doctor's office.
 Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
+Temporal data may be occurence of events, like flight departures, that need to be induced into a regular structure.
+
+Current time series objects, tend to be model-oriented matrices.
+Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.
+
+In recent work, there has been exciting developments, that conceptually and practically streamline the pre-processing of data.
+Figure \ref{fig:tsibble-fit} (adapted from @r4ds) illustrates the tidy data workflow.
+Time series modeling and forecasting enters the picture at the *modeling* stage.
+While temporal data enters at the start, there isn't a current framework for getting temporal data into the matrix format for modeling.
+That's the part of the workflow that the work descirbed in this paper improves.
+
+<!--
+with little support for non-numeric values, feed into the framework solely at the modelling stage.
+Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.
+
 @r4ds succinctly describes the data science workflow given in Figure \ref{fig:tsibble-fit}.
 It is annotated to emphasize the significant divide between data that feeds into the beginning of analysis and the current time series objects that fits into modelling, making data preparation for modeling and forecasting painful.
+-->
 
@@
 
-Current time series objects, which are model-oriented matrices with little support for non-numeric values, feed into the framework solely at the modelling stage.
-Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.
-
-The rest of the paper is structured as follows.
+The paper is structured as follows.
 Section \ref{sec:str} reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation.
 Section \ref{sec:semantics} proposes contextual semantics for temporal data, built on top of tidy data.
-The concept of data pipelines with respect to the time domain will be discussed in depth in Section \ref{sec:pipeline}, followed by the design choices of software structure in Section \ref{sec:software}.
-Section \ref{sec:cases} presents two case studies for temporal data exploration using the newly implemented infrastructure.
-Section \ref{sec:conclusion} concludes the paper and discusses the future work.
+The concept of data pipelines with respect to the time domain will be discussed in depth in Section \ref{sec:pipeline}, followed by a discussion of the design choices made in the software structure in Section \ref{sec:software}.
+Two case studies are presented in Section \ref{sec:cases} illustrating temporal data exploration using the newly implemented infrastructure.
+Section \ref{sec:conclusion} discusses the future work.
 

### 2019-01-18 main.Rmd
reworked abstract - removed key words that repeat title words

@@
     email : rob.hyndman@monash.edu
-abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point event that needs aggregating, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
-Time series models, in particular, the software supporting time series forecasting makes strict assumptions on the data to be provided, typically a matrix of numeric data with implicit time indices.
+abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
+Time series models, in particular, the software supporting time series forecasting makes strict assumptions on the data to be provided, typically a matrix of numeric data with implicit time indexes.
 Going from raw data to model-ready data is painful.
@@
 The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
-keywords: "temporal data, time series, data structures, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis"
+keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
 wpnumber: no/yr
@@
 
-Current standards, provided by native `ts` object in R and extended by **zoo** [@zoo2005], and **xts** [@R-xts], organize temporal data in matrices with implicit time indices.
+Current standards, provided by native `ts` object in R and extended by **zoo** [@zoo2005], and **xts** [@R-xts], organize temporal data in matrices with implicit time indexes.
 These objects were designed for modelling methods.
@@
 For example, one could easily extract time components, such as time of day and day of week, from the index to visualize seasonal effects of response variables.
-One could also join other data sources to the tsibble based on common time indices.
+One could also join other data sources to the tsibble based on common time indexes.
 The accessibility of tsibble index motivates data analysis towards transparency and human readability.
@@
 
-As for data indexed in regular time space, the time interval is derived by first computing absolute differences of time indices and then the greatest common divisor, which covers all conceivable cases.
+As for data indexed in regular time space, the time interval is derived by first computing absolute differences of time indexes and then the greatest common divisor, which covers all conceivable cases.
 This implies that all subjects in a tsibble have one and the only interval.
@@
 Yet ordered factors can also be a source of time, such as month names from January to December and weekdays from Monday to Sunday.
-Tsibble supports an extensive range of index types from numerics to nano time, but there might be custom indices used for some occasions, for example school semesters.
+Tsibble supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters.
 These academic terms vary from one institute to another within an academic year, which is defined differently from a calendar year.

### 2019-01-18 main.Rmd
section 2.1 done

@@
 The data problems are grouped into two types of analysis, time series and longitudinal.
-Despite of exactly the same data input, the representation of time series and longitudinal data diverges due to different modelling approaches.
+Despite being exactly the same data input, the representation of time series and longitudinal data diverges due to different modelling approaches.
 
@@
 It strictly requires that the number of observations must be the same across all the series.
-Data wrangling from the form that data arrives in, to this specialist format can be frustrating and difficult, inhibiting the variety of downstream tasks such as analytics.
+Data wrangling, from the form that data arrives in, to this specialist format, can be frustrating and difficult, inhibiting the variety of downstream tasks such as analytics.
 
@@
 Thus, the primary format required for modeling is stacked series, blocks of measurements for each individual, with columns indicating individual, time of measurement and the measurements.
-(The tidy data name for this would be **long format**.) Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it in that wide format which would have missing values in many cells.
-A detriment of this format is that demographic information for subjects is often repeated for each time point.
-However, appealing feature is that data is structured in a concise and semantic manner with reference to observations and variables, with the time index stated explicitly.
+(The tidy data name for this would be **long format**.) Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it into wide format which would have missing values in many cells.
+A drawback of this format is information unique to each individual is often repeated for all time points.
+The appealing feature is that data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
 This opens the door to easily operating on time to make calculations and extract different temporal components, such as month and day of the week.

### 2019-01-18 main.Rmd
section 2.2 done

@@
 The principles of "tidy data" attempt to standardize the mapping of the semantics of a dataset to its physical representation.
-This data structure is the fundamental unit of the **tidyverse**, which is a collection of R packages designed for data science; the prevalence of **tidyverse** reversely testifies to the novelty of "tidy data".
-The **tidyverse** provides abstract yet functional grammars to manipulate and visualize data in easier-to-comprehend manners.
+This data structure is the fundamental unit of the **tidyverse**, which is a collection of R packages designed for data science.
+The ubiquitous use of **tidyverse** is testament to the simplicity, practicality and general applicability of the tools.
+The **tidyverse** provides abstract yet functional grammars to manipulate and visualize data in easier-to-comprehend form.
 One of the **tidyverse** packages, **dplyr** [@R-dplyr] showcases the value of a grammar as a principled vehicle to transform data for a wide range of data challenges, providing a consistent set of verbs: `mutate()`, `select()`, `filter()`, `summarise()`, and `arrange()`.
@@
 The **tidyverse** largely formalizes exploratory data analysis.
-The whole R community has quickly adopted the **tidyverse** way of thinking to much broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
-Temporal data is left behind in this regard.
+Many in the R community have adopted the **tidyverse** way of thinking and extended to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
+Temporal data tools need to catch up.
 

### 2019-01-18 main.Rmd
section 2.3 done

@@
 
-Current standards, provided by native `ts` object in R and extended by **zoo** [@zoo2005], and **xts** [@R-xts], organize temporal data in matrices with implicit time indexes.
+Current standards, provided by the native `ts` object in R, extended by **zoo** [@zoo2005], and **xts** [@R-xts], assemble temporal data into matrices with implicit time indexes.
 These objects were designed for modelling methods.
-The diagram correspondingly would place the model at the center of the analytical universe, and all the transformations, visualisations would hinge on that format.
+The diagram in the style of Figure \ref{fig:tsibble-fit}, would place the model at the center of the analytical universe, and all the transformations, visualisations would hinge on that format.
 This is contrary to the tidyverse conceptualisation, which wholistically captures the full data workflow.
@@
 
-This paper proposes a new tidy representation for temporal data and a unified framework to streamline the workflow from data pre-processing to visualization and forecasting, as an integral part of tidy data analysis.
+This paper describes a new tidy representation for temporal data, and a unified framework to streamline the workflow from data pre-processing to visualization and forecasting, as an integral part of a tidy data analysis.
 

### 2019-01-18 main.Rmd
section 3.0 done

@@
 The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
-Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as "tsibble".
+Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble".
 
@@
 To turn this data into a tsibble, (1) column `year` is declared as the index variable; (2) the key can be made up of columns `country` and `gender`.
-Despite column `count` is the only measurement in this case, it is sufficiently flexible to hold other measured variables, for example, adding the corresponding population size (if known) in order to calibrate the count later.
+The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables, for example, adding the corresponding population size (if known) in order to calibrate the count later.
 
-It bridges the gap between raw temporal data and model inputs.
+The new data structure, tsibble, bridges the gap between raw temporal data and model inputs.
 Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.

### 2019-01-18 main.Rmd
section 3.1 done

@@
 Time provides a contextual basis for temporal data.
-A variable representing time is indispensable to a tsibble, referred to as "index".
+A variable representing time is essential for a tsibble, and is referred to as an "index".
 The "index" is an intact data column rather than a masked attribute, which makes time visible and accessible to users.
@@
 
-A variable number of time representation is spotted in the wild.
-Date-time, universally accepted across systems, is the most commonly used type for representing time.
+A variable number of time representations is spotted in the wild.
+A Date-time object, universally accepted across computing systems, is the most commonly used type for representing time.
 Date-time also typically associates with a time zone with adjustments like summer time.
-This diversity and time zone is respected and taken into consideration for the *tsibble*'s index.
+This diversity and time zone is acknowledged and considered the building block for  *tsibble*'s index.
 

### 2019-01-19 main.Rmd
done with key

@@
 
-The "key" specification contributes to creating a tsibble alongside the index.
-The concept of "key" is introduced to uniquely identify units or subjects that are observed over time in a data table, which is similar to a primary key [@codd_relational_1970] defining each observation in a relational database.
-In the wide format, each column holds a series of values, so that the column itself serves for the sake of identification.
-In the long format, all column names are melt to the corresponding "key" values.
-But the "key" is much more flexible than simply column names.
-Because it is not constrained to a single field, but can be comprised of multiple fields.
-The identifying variables that the "key" is constituted of, stay the same as they are in the original table, with no further tweaks.
+The "key" specification is the second essential ingrdient for a tsibble.
+The "key"  introduced uniquely identifies observations that are recorded over time in a data table.
+It is similar to a primary key [@codd_relational_1970] defining each observation in a relational database.
+In the wide format, that multiple time series are often structured in, the columns hold a series of values, so that the column implicitly serves as identification.
+In long format, columns are melted with names converted to "key" values.
+However, the "key" provides much more flexibility.
+It is not constrained to a single field, but can be composed from multiple fields.
+The identifying variables that the "key" is constituted from, remain the same as in the original table, with no further tweaks.
 
-Each tsibble must hold a "key".
-It is normally a priori known by analysts.
+The "key" is usually known a priori by analysts.
 For example, Table \ref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year.
 This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data.
-Lacking in either of two will be inadequate for the key and thus fail to construct a tsibble.
-The alternative specification of the key is to include a third variable `continent`.
+Lacking either of two will be inadequate, because the observations are not uniquely identified, and thus will generate a tsibble construction fail.
+An alternative specification of the key for this data is to include a third variable `continent`.
 Since `country` is nested within `continent`, it is a free variable for use.
-But this variable brings additional information that can be used for forecasting reconciliation [@fpp].
-The key is explicit when multiple units exist in the data.
-Key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.
+This variable brings additional information that can be used for forecasting reconciliation [@fpp].
+The key needs to be explicit when multiple units exist in the data.
+The key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.
 
-Not only pinpoints a "key" observational units in the tsibble but also provides a solution to seamlessly link between the data, models, and forecasts.
-This neatly decouples the data from models and forecasts, leaving more room for compulsory model components, such as coefficients, fitted values and residuals.
+The "key" also provides a solution to link between the data, models, and forecasts.
+This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals.
 More details are given in the following section.

### 2019-01-19 main.Rmd
done with interval

@@
 
-One of the cornerstones beneath a tsibble is time interval.
+One of the cornerstones of time series data, and hence beneath a tsibble, is the time interval.
 This information plays a critical role in computing statistics (e.g.
@@
 seasonal ARIMA).
-The principal divide is regularly or irregularly spaced in time.
-Since a tsibble permits implicit missing time, it is impossible to distinguish regularity from the index.
-It relies on user's specification by switching the `regular` argument off, when the data involves irregular intervals.
-This type of data can flow into event-based data modelling, not models that expect time series.
+The principal divide is regularly or irregularly spaced observations in time.
+A tsibble permits implicit missing time, making it  difficult to distinguish regularity from the index.
+It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals.
+This type of data can flow into event-based data modelling, but would need to be processed or regularized to fit models that expect time series.
 
-As for data indexed in regular time space, the time interval is derived by first computing absolute differences of time indexes and then the greatest common divisor, which covers all conceivable cases.
-This implies that all subjects in a tsibble have one and the only interval.
-Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each subject may underly different data generating process.
+For data indexed in regular time space, the time interval is automatically calculated, by first computing absolute differences of time indexes and then the greatest common divisor.
+This covers all conceivable cases.
+The implication is that all observations in a tsibble have one and the only interval.
+Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each observation may  different underlying data generating processes.
 

### 2019-01-19 main.Rmd
done with pipeline

@@
 
+The tsibble structure and operations support data pipelines for sequencing analysis.
 There has been a long history of pipeline discussions and implementation centering around data in various aspects.
 A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualising the process as it might be applied to a variety of problems.
-The Extract, Transform, and Load (ETL), from the data warehousing literature, outlines the workflow to prepare data for analysis, dates back to @kimball2011data can be considered a data pipeline.
+The Extract, Transform, and Load (ETL), from recent data warehousing literature, outlines the workflow to prepare data for analysis, dates back to @kimball2011data can be considered a data pipeline.
 Building a data pipeline can be technically difficult, to make it sufficiently general for various data, with many implementation decisions on the interface, input and output objects and functionality.
@@
 @unix coined the term "pipelines" in software development, while developing Unix at Bell Labs.
-In Unix-based computer operating systems, a pipeline chains together a series of operations on the basis of their standard streams, so that the output of each programme becomes the input to another.
+In Unix-based computer operating systems, a pipeline chains together a series of operations on the basis of their standard streams, so that the output of each program becomes the input to another.
 This shapes the Unix toolbox philosophy: "each do one simple thing, do it well, and most importantly, work well with each other" [@unix-philosophy].

### 2019-01-19 main.Rmd
done with 4.1

@@
 
-## Time-based transformation {#sec:timepipeline}
+The time-based pipeline shepards raw temporal data through to time series analysis, and plots.
+It is possible and recommended to check for identical entries of key and index before analysis.
+Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making.
+Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
+When the data meets the tsibble standard, it flows neatly into the analysis stage.
+Figure \ref{fig:tsibble-pipeline} illustrates the time-based pipeline.
 
@@
 
-The time-based pipeline shepards raw temporal data through to time series analysis, and plots.
-It is possible and recommended to check for identical entries of key and index before analysis.
-Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making.
-Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
-When the data meets the tsibble standard, it flows neatly into the analysis stage.
+## Time-based transformation {#sec:timepipeline}
 
@@
 
-Data needs transformation to some extent for modelling.
-A suite of verbs are introduced to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, as well as transformed for visualization.
+A suite of verbs, and corresponding functions, are used to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, and transformed for visualization.
+A tsibble is an object, conceptually considered as a noun, and hence an action performed on the object can be phrased as a verb.
 The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood.
@@
 If a tsibble cannot be maintained in the output of a pipeline module, for example the index is dropped by aggregation, an error informs users of the problem and suggests alternatives.
-This avoids negatively surprising users and reminds them of time context.
+This avoids surprising users and reminds them of the time context.
 
+<!--
 Transformation assembles a set of modules.
-A unit that makes up a module is function or preferably "verb".
+A unit that makes up a module is a function or preferably "verb".
 A tsibble is an object, conceptually considered as a noun, and hence an action performed on the object can be phrased as a verb.
-Each verb focuses on one thing and achieve its goal.
+-->
+
+Each verb focuses on one oepration to achieve its goal.
 The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
 These general-purpose verbs are made available in the **tidyverse** suite.
-When manipulating in temporal context, these verbs are adapted to time domain.
+When used in the temporal context, these verbs are adapted to the time domain.
 A perceivable difference is summarizing variables between data frame and tsibble.
 The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
-New tsibble-specific verbs are proposed to expand the **tidyverse** vocabulary.
-We believe that users, who are already familiar with **tidyverse**, will experience a gentle learning curve for mastering tsibble verbs and glide into temporal data analysis with low cognitive load.
+New tsibble-specific verbs (Table \ref{tab:verb-table} are provided to expand the **tidyverse** vocabulary.
+Users who are already familiar with **tidyverse**, will experience a gentle learning curve for mastering tsibble verbs and glide into temporal data analysis with low cognitive load.
 
@@
 Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package.
-A sequence of functions can be composed in a way that can be naturally read from left to right, which incredibly increase the readability of the code.
-It consequently generates a block of code without worrying naming intermediate values.
-Most importantly, the ecosystem for tidy time series analysis has been undertaking on the basis of tsibble in R, known as "tidyverts".
+A sequence of functions can be composed in a way that can be naturally read from left to right, which improves the readability of the code.
+It consequently generates a block of code without saving intermediate values.
+Most importantly, a new ecosystem for tidy time series analysis has been undertaken, using the tsibble framework, and is called "tidyverts", a play on tidyverse that acknowledges the time series analysis purpose.
 
@@
 
+# Acknowledgements {-}
+
+The authors thank Mitch O'Hara-Wild for many discussions on the software design.
+
 ```{r write-bib}

### 2019-01-19 main.Rmd
done with modeling and vis

@@
 
-## Time series visualization and models {#sec:ts-models}
+## Time series visualization {#sec:ts-vis}
 
-As a special case of the data frame, a tsibble pipes into the grammar of graphics straight way, making most use of this powerful graphical system.
-It should be easy to create and extend some specialized time-related plotting methods based on tsibble structure, such as autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
+As a special case of the data frame, a tsibble pipes directly into the grammar of graphics.
+It is easy to create and extend specialist time-related plotting methods based on tsibble structure, such as autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
+
+## Time series models {#sec:ts-models}
 
@@
 Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently.
-This yields a "mable" (short for model table), where each model only tags to each "key" value in tsibble to avoid expensive data copying and reduce model storage.
+This yields a "mable" (short for model table), where each model relates to each "key" value in tsibble.
+This avoids expensive data copying and reduces model storage.
 The mable is further supplied to forecasting methods, to produce a tsibble in which each "key" along with its future time holds predictions.
 It also underlines the advantage of tsibble's "key" in linking between data inputs, models and forecasts.
-Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in alike spirit.
-The modeling module will be fulfilled and integrated eventually.
+Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in a similar spirit.
+The modeling module is a current endeavor.
 
+<!--
 We go through the whole exploration circle, and keep iterating and refining until data insights gained.
@@
 The cohesive and coherent framework results in more graceful and expressive code.
+-->
 

### 2019-01-19 main.Rmd
done with 5.1

@@
 The prime force that drives the software's design choices is "data".
-All functions in **tsibble** starts with `data` or its variants as the first argument, namely "data first".
+All functions in **tsibble** start with `data` or its variants as the first argument, namely "data first".
 They work naturally with the pipe operator `%>%`, read as "then".
@@
 Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-Below conveys the pieces of critical and contextual picture about the data in Table \ref{tab:tb-sub}: (1) data dimension with its shorthand time interval, alongside time zone if date-times; (2) variables that constitute of the "key" with the number of series.
-These details aid users in understanding their data better and manipulate the data with care.
+The  contextual summary provided by the print function on the data is shown in Table \ref{tab:tb-sub} contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
+These details aid users in understanding their data better.
 

### 2019-01-19 main.Rmd
done with 5.2

@@
 Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-The  contextual summary provided by the print function on the data is shown in Table \ref{tab:tb-sub} contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
+The  contextual summary provided by the print function shown below, on the data from Table \ref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
 These details aid users in understanding their data better.
@@
 Rolling window calculations are widely-used techniques in time series analysis, but are not restricted to temporal applications.
-These operations are dependent on certain ordering, particularly time ordering for temporal data.
+These operations are dependent on having an ordering, particularly time ordering for temporal data.
 Three common types of variations for sliding window operations are:
@@
 
-Figure \@ref(fig:slide-animate), \@ref(fig:tile-animate) and \@ref(fig:stretch-animate) show the animations of rolling windows for sliding, tiling and stretching respectively on annual tuberculosis cases in Australia.
-A block of consecutive elements with window size of 5 are initialized and started rolling sequentially till the end of series.
+Figures \@ref(fig:slide-animate), \@ref(fig:tile-animate) and \@ref(fig:stretch-animate) show the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
+A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.
 
@@
 
-(ref:slide-animate-cap) An illustration of window of size 5 sliding over annual tuberculosis cases in Australia. View the animation with Adobe Acrobat Reader.
+(ref:slide-animate-cap) An illustration of window of size 5 sliding over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
 
@@
 
-(ref:tile-animate-cap) An illustration of window of size 5 tiling over annual tuberculosis cases in Australia. View the animation with Adobe Acrobat Reader.
+(ref:tile-animate-cap) An illustration of window of size 5 tiling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
 
@@
 
-(ref:stretch-animate-cap) An illustration of window of size 5 stretching over annual tuberculosis cases in Australia. View the animation with Adobe Acrobat Reader.
+(ref:stretch-animate-cap) An illustration of window of size 5 stretching over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
 
 Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in Table \@ref(tab:verb-table).
-Table verbs expect and return a tsibble, and does something that the function name suggests.
+Table verbs expect and return a tsibble, and does what the function name suggests.
 On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window.
-For example, computing moving averages requires numerics and a function like `mean()`, and produce averaged numerics; however, rolling window regression takes a data frame and a linear regression method like `lm()`, and generate a complex object that contains coefficients, fitted values, and etc.
-In addition, the same function will be called and applied to each window of elements independently, which means programming is done with for-loop statements.
-But functional programming emphasizes on expressions instead of statements.
+For example, computing moving averages requires numerics and a function like `mean()`, and produces averaged numerics.
+However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.
+<!-- In addition, the same function is called and applied to each window of data independently, which means programming is done with for-loop statements.
+But functional programming emphasizes  expressions instead of statements.-->
 
-The **purrr** package sets a great example of functional programming in R.
+The **purrr** package provides a good example of functional programming in R.
 It provides a complete and consistent set of tools to iterate each element of a vector with a function.
-Rolling window not just iterates but roll over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
+Rolling window doesn't just iterate but rolls over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
 `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.

### 2019-01-19 main.Rmd
done with 5.3

@@
 
-Modular programming is adopted while designing the **tsibble** package. Modularity benefits users with variety and flexibility and developers with easy maintenance.
+Modular programming is adopted in the design of the **tsibble** package. Modularity benefits users by providing small focused and manageable chunks, and provides developers with simpler maintenance.
 
@@
 Vector functions in the package mostly deal with time.
-When collapsing a tsibble to less granular interval, these atomic functions can be combined with the `index_by()` table verb to accomplish this.
+When collapsing a tsibble to a less granular interval, these atomic functions can be combined with the `index_by()` table verb.
 A different function results in easily switching to aggregation of different time resolution.
@@
 It is generally easier to trace back the errors users encounter from separating 1d and 2d functions.
-(lost in a web of functions)
+<!--(lost in a web of functions)-->
 

### 2019-01-19 main.Rmd
done with 5.4

@@
 
-As a fundamental infrastructure, extensibility is a design decision that is focused on from the start of **tsibble**'s development.
-Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily relies on S3 classes and methods in R [@adv-r].
-It can be extended in two major aspects: custom index and new tsibble class.
+As a fundamental infrastructure, extensibility is a design decision that was employed from the start of **tsibble**'s development.
+Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily rely on S3 classes and methods in R [@adv-r].
+The package can be extended in two major aspects: custom index and new tsibble class.
 
 Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nantotime** [@R-nanotime], and pure numbers in simulations.
-Yet ordered factors can also be a source of time, such as month names from January to December and weekdays from Monday to Sunday.
-Tsibble supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters.
-These academic terms vary from one institute to another within an academic year, which is defined differently from a calendar year.
+Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday.
+The tsibble package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters.
+These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year.
 New index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future.
 The interval regarding semesters is further outlined through `pull_interval()`.
-As a result, the rest software methods such as `has_gaps()`, `count_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
+As a result, the rest of the software methods such as `has_gaps()`, `count_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
 
 The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble.
-For example, a fable is actually an extension to a tsibble, mentioned in Section \ref{sec:timepipeline}.
+For example, a mable is actually an extension of a tsibble, mentioned in Section \ref{sec:timepipeline}.
 A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass.
-First of all, this new object itself is a tsibble.
-It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable.
+This new object itself is a tsibble.
+It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a mable.
 Tsibble verbs are also S3 generics.

### 2019-01-19 main.Rmd
done with airlines

@@
 The same flight is scheduled at exactly the same time, together with the same performance statistics but different flight details, which is very unlikely.
-Flight NK630 is usually scheduled at 17:45 from Chicago to New York, searching into the whole records.
-A decision is made on removal of the first row from the duplicated entries before proceeding to the tsibble creation.
+Flight NK630 is usually scheduled at 17:45 from Chicago to New York, discovered by searching the full database.
+A decision is made to remove the first row from the duplicated entries before proceeding to the tsibble creation.
 
@@
 To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", column `flight_num` as "key" via `id(flight_num)`.
-As a result of event data, this data is irregularly spaced, and hence switching to irregular option is necessary.
+As a result of event timing, this data is irregularly spaced, and hence switching to the irregular option is necessary.
 The program internally validates if the key and index produce the distinct rows, and then sort the key and the index from past to recent.
@@
 Transforming tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can result in well-constructed pipelines.
-From the perspective of a passenger, one need to travel smart, by choosing an efficient carrier to fly with and the time of day to avoid congestion, for example.
+From the perspective of a passenger, one needs to travel smart, by choosing an efficient carrier to fly with and the time of day to avoid congestion, for example.
 We take a drill-down approach to exploring this data, starting with annual carrier performance and followed by disaggregating to finer time resolutions.
@@
 To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column.
-The flow chart shown as Figure \@ref(fig:flight-pipeline1) demonstrates the operations undertaking in a data pipeline.
+The flow chart (Figure \@ref(fig:flight-pipeline1)) demonstrates the operations undertaking in a data pipeline.
 The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval.
-The final data is each carrier along with a single year, thereby the interval undetermined.
+The final data is each carrier along with a single year, with the interval undetermined.
 It in turn feeds into the mosaic display.
-Note that Southwest Airlines, as the largest carrier, operates less efficiently compared to Delta, in Figure \@ref(fig:carrier-mosaic).
+Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).
 
@@
 
-A closer examination of some big airports across the US will give an indication about how well the busiest airports manage the outflow traffic on a daily basis.
-A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seatle (SEA) airports is obtained first.
+A closer examination of some big airports across the US will give an indication of how well the busiest airports manage the outflow traffic on a daily basis.
+A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first.
 The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:sel-monthly-plot).
 Winter months tend to fluctuate a lot compared to the summer across all the airports.
-What superimposes on these lines is two-month moving averages so that a temporal trend is more visible.
+Superimposed on the plot are two-month moving averages, so the temporal trend is more visible.
 The number of days for each month is variable.
 Moving averages for two months call for computing weighted mean.
-But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without worrying weights specification.
-The sliding operation with large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
-The LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady control of delays over time.
-The IAH and JFK airports have undergone more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
+But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without weight specification.
+The sliding operation with a large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
+LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady number delays over time.
+The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
 
@@
 The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown.
-To reduce the likelihood of delay suffering, it is recommended to avoid the peak hour at 18.
-As moving towards the upper extremes, the variations considerably increase, making departure time unpredictable.
+To reduce the likelihood of suffering a delay, it is recommended to avoid the peak hour around 6pm (18).
+<!--As moving towards the upper extremes, the variations considerably increase, making departure time unpredictable.-->
 

### 2019-01-19 main.Rmd
done with energy

@@
 
-Sensors have been put up to collect data for the project of smart city across major cities in Australia.
+Sensors have installed in households across major cities in Australia to collect data for the smart city project.
 One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter].
-Year 2013 has been sliced to examine temporal patterns of customer's energy consumption with **tsibble** in this paper.
+Year 2013 has been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study.
 Half-hourly general supply in kwH have been recored for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
-Customer's demographic data provides explanatory variables other than time in a different data table.
+Customers' demographic data provides explanatory variables other than time in a different data table.
 Two data tables might be joined to explore different sources that contribute to daily electricity use when needed.
@@
 Gaps in time occur to `r scales::percent(sum(gap_df$.gaps) / NROW(gap_df))` of the households in this dataset.
-It would be interesting to explore these missing patterns as part of preliminary analysis.
-Since the smart meter has been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for the households, instead of full lengths across the households.
-Figure \@ref(fig:count-gaps) reveals the gaps for the top 49 households and the rest collapsed into a single batch arranged in rows from high to low in tallies.
-Missing values can be found at any time point during the entire time span.
+It would be interesting to explore these missing patterns as part of a preliminary analysis.
+Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for each household.
+Figure \@ref(fig:count-gaps) shows the gaps for the top 49 households arranged in rows from high to low in tallies.
+(The remaining households values have been aggregated into a single batch and appear at the top.) Missing values can be seen to occur at any time during the entire span.
 A small number of customers have undergone energy unavailability in consecutive hours, indicated by a line range in the plot.
-On the other hand, the majority suffer occasional breakdown with more frequent happenings in January.
+On the other hand, the majority suffer occasional outages with more frequent occurrence in January.
 
@@
 
-(ref:count-gaps-cap) Time gap plots for the top 49 customers with most implicit missing values and the rest grouped into the one.
-Each cross represents the missing in time and a line between two dots shows continuous missingness over time.
-Each row corresponds to one customer describes the incidences of time gaps.
+(ref:count-gaps-cap) Time gap plots for the 49 customers with most implicit missing values, and the remaining customers grouped into the one line at top.
+Each cross represents a missing in time and a line between two dots shows continuous missingness over time.
+Each row corresponds to one customer.
+Missing values occur at various times, with perhaps more in January.
 
-Aggregation across all individuals helps to sketch a big picture about if the representative behaviors change over time, laid out in the calendar display shown as Figure \@ref(fig:calendar-plot).
-Each glyph represents the daily pattern of residential electricity usage every thirty minutes with higher consumption in the time when human are more active.
-The daily snapshots diverge in traces, depending on the season in the year.
+Aggregation across all individuals helps to sketch a big picture about the representative behaviors change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)).
+Each glyph represents the daily pattern of residential electricity usage every thirty minutes.
+Higher consumption is indicated by higher values, and typically occurs in daylight hours.
+Colour indicates hot days.
+The daily snapshots vary depending on the season in the year.
 During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially in the hot days with daily average temperature greater than 25 degrees.
 However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.
-This also illustrates how the tsibble data can easily integrate with other tools and graphics.
+This plot illustrates how the tsibble data can easily integrate with other tools and graphics.
 
@@
 
-(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized in the calendar format.
-Distinctive patterns between summer and winter months occur to daily residential energy consumption.
-Air conditioning drives the peak in the late afternoon in January and December, however, peaks in the morning and evening are more pronounced in the winter months due to heating use.
+(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized into calendar format, with colour indicating hot days.
+Energy use of hot days tends to be higher, suggesting air conditioner use.
+Days in the winter months have a double peak suggesting heater use.
 

### 2019-01-19 main.Rmd
done with conclusions

@@
 
-A new data abstraction representing temporal data named as "tsibble" has been proposed, spotlighting the "tidy data" principles brought to time domain.
+A new data abstraction to represent temporal data named as "tsibble" has been presented, spotlighting the "tidy data" principles brought to time domain.
 Tidy data begins to take shape in the state of time with the introduction of contextual semantics: index and key.
-Declared index provides direct support to time variable; variables that comprise the key defines study subjects over time.
-These semantics further determines unique data entries required for a valid tsibble.
-No matter how temporal data arrives, tsibble respects time index and keeps data richness.
-A tsibble frictionlessly pops into transformation, visualization, modelling and smoothly shifts amongst, allowing for rapid iterations in gaining data insights.
+A declared index provides direct support to time variable; variables that comprise the key define observations over time.
+These semantics further determine unique data entries required for a valid tsibble.
+No matter how temporal data arrives, tsibble respects a time index and mainatins the data richness.
+A tsibble frictionlessly pops into transformation, visualization, modelling and smoothly shifts between, allowing for rapid iteration to gain data insights.
 
@@
 However, few R packages provide functionality to create and manage many sorts of calendars, including market-specific business calendar.
-This delays the implementation.
-Generally, custom calendars would be easily embedded into the tsibble framework.
-Consequently these tsibble operators, like `fill_gaps()`, would work out of box; forecasts would be generated within its definable time range.
+<!-- This delays the implementation.--> Generally, custom calendars would be easily embedded into the tsibble framework.
+Consequently these tsibble operators, like `fill_gaps()`, would work out of box, and forecasts would be generated within its definable time range.
 

### 2019-01-24 main.Rmd
Improved English in abstract

@@
 abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
-Time series models, in particular, the software supporting time series forecasting makes strict assumptions on the data to be provided, typically a matrix of numeric data with implicit time indexes.
+Time series models, in particular the software supporting time series forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes.
 Going from raw data to model-ready data is painful.
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
-Tidy data principles are extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) including an explicitly declared index variable representing time, (3) incorporating a "key" comprised of single or multiple variables to uniquely identify units over time.
-This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context.
+Tidy data principles are extended to temporal data: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time.
+This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts.
 A sound data pipeline facilitates a fluent and fluid workflow for analyzing temporal data.

### 2019-01-24 main.Rmd
Improved English intro

@@
 Temporal data arrives in many possible formats, as does the time context.
-For example, time can carry with it various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments like summer time.
-It could be irregularly recorded, which is particularly true with longitudinal measurements like patient visits to a doctor's office.
+For example, time can carry with it various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments such as summer time.
+Time can be irregularly recorded, which is often the case with longitudinal measurements such as patient visits to a doctor's office.
 Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
-Temporal data may be occurence of events, like flight departures, that need to be induced into a regular structure.
+Temporal data may be occurence of events, such as flight departures, that need to be induced into a regular structure.
 
-Current time series objects, tend to be model-oriented matrices.
-Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.
+Current time series objects tend to be model-oriented matrices.
+Analysts are expected to do their own data pre-processing and take care of anything else needed to allow model fitting, which leads to a myriad of ad-hoc solutions and duplicated efforts.
 
-In recent work, there has been exciting developments, that conceptually and practically streamline the pre-processing of data.
-Figure \ref{fig:tsibble-fit} (adapted from @r4ds) illustrates the tidy data workflow.
+In recent work, there have been some exciting developments that conceptually and practically streamline the pre-processing of data.
+Figure \ref{fig:tsibble-fit} [adapted from @r4ds] illustrates the tidy data workflow.
 Time series modeling and forecasting enters the picture at the *modeling* stage.

### 2019-01-26 R/flights.R
fixed qtl dep_delay problem

@@ -161,7 +161,7 @@ sel_monthly %>%
 hr_qtl <- us_flights %>% 
   index_by(dep_datehour = floor_date(sched_dep_datetime, "hour")) %>% 
   summarise(
-    value = list(quantile(dep_delay, c(0.5, 0.8, 0.95))),
+    dep_delay = list(quantile(dep_delay, c(0.5, 0.8, 0.95))),
     qtl = list(paste0("qtl", c(50, 80, 95)))
   ) %>% 
   unnest(key = id(qtl)) %>% 

### 2019-01-28 R/demo.R
Prevent table floating to top of next page

@@ -4,20 +4,20 @@ library(tsibble)
 
 ## ---- tb-sub
 tb_small <- read_rds("data/tb-small.rds")
-tb_small %>% 
-  filter(year > 2010) %>% 
-  knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>% 
-  kableExtra::kable_styling(position = "center")
+tb_small %>%
+  filter(year > 2010) %>%
+  knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>%
+  kableExtra::kable_styling(position = "center", latex_options=c("hold_position"))
 
 ## ---- tb-print
-as_tsibble(tb_small, key = id(country, gender), index = year) %>% 
-  filter(year > 2010) %>% 
+as_tsibble(tb_small, key = id(country, gender), index = year) %>%
+  filter(year > 2010) %>%
   print(n = 5)
 
 ## ---- tb-au
-tb_au <- tb_small %>% 
-  filter(country == "Australia") %>% 
-  group_by(year) %>% 
+tb_au <- tb_small %>%
+  filter(country == "Australia") %>%
+  group_by(year) %>%
   summarise(count = sum(count))
 
 ## ---- slide-animate

### 2019-01-28 main.Rmd
English improvements

@@
       in_header: preamble.tex
-    keep_tex: no
+    keep_tex: yes
     number_sections: yes
@@
 Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
-Temporal data may be occurence of events, such as flight departures, that need to be induced into a regular structure.
+Temporal data may be occurrence of events, such as flight departures, that need to be induced into a regular structure.
 
 Current time series objects tend to be model-oriented matrices.
-Analysts are expected to do their own data pre-processing and take care of anything else needed to allow model fitting, which leads to a myriad of ad-hoc solutions and duplicated efforts.
+Analysts are expected to do their own data preprocessing and take care of anything else needed to allow model fitting, which leads to a myriad of ad hoc solutions and duplicated efforts.
 
-In recent work, there have been some exciting developments that conceptually and practically streamline the pre-processing of data.
-Figure \ref{fig:tsibble-fit} [adapted from @r4ds] illustrates the tidy data workflow.
+In recent work, there have been some exciting developments that conceptually and practically streamline the preprocessing of data.
+\autoref{fig:tsibble-fit} [adapted from @r4ds] illustrates the tidy data workflow.
 Time series modeling and forecasting enters the picture at the *modeling* stage.
 While temporal data enters at the start, there isn't a current framework for getting temporal data into the matrix format for modeling.
-That's the part of the workflow that the work descirbed in this paper improves.
+That's the part of the workflow that the work described in this paper improves.
 
@@
 with little support for non-numeric values, feed into the framework solely at the modelling stage.
-Analysts are expected to do their own data pre-processing and take care of anything else needed to get to this stage, which leads to a myriad ad-hoc solutions and duplicated efforts.
+Analysts are expected to do their own data preprocessing and take care of anything else needed to get to this stage, which leads to a myriad ad hoc solutions and duplicated efforts.
 
@@
 The paper is structured as follows.
-Section \ref{sec:str} reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation.
-Section \ref{sec:semantics} proposes contextual semantics for temporal data, built on top of tidy data.
-The concept of data pipelines with respect to the time domain will be discussed in depth in Section \ref{sec:pipeline}, followed by a discussion of the design choices made in the software structure in Section \ref{sec:software}.
-Two case studies are presented in Section \ref{sec:cases} illustrating temporal data exploration using the newly implemented infrastructure.
-Section \ref{sec:conclusion} discusses the future work.
+\autoref{sec:str} reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation.
+\autoref{sec:semantics} proposes contextual semantics for temporal data, built on top of tidy data.
+The concept of data pipelines with respect to the time domain will be discussed in depth in \autoref{sec:pipeline}, followed by a discussion of the design choices made in the software structure in \autoref{sec:software}.
+Two case studies are presented in \autoref{sec:cases} illustrating temporal data exploration using the newly implemented infrastructure.
+\autoref{sec:conclusion} discusses the future work.
 
@@
 
-Time series can be univariate or multivariate, and for modelling requires relatively large in length (i.e.
-large $T$).
+Time series can be univariate or multivariate, and for modelling require relatively long lengths (i.e., large $T$).
 Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic processes, for the primary purpose of forecasting, and characterizing temporal dynamics.
-The time series supporting modeling are represented as vectors or matrices in most statistical software.
+Most statistical software represent such time series as vectors or matrices.
 Multivariate time series are typically assumed to be in the format where each row is assumed to hold observations at a time point and each column to contain a single time series.
@@
 It often occurs that measurements for individuals are taken at different time points, resulting in an unbalanced panel.
-Thus, the primary format required for modeling is stacked series, blocks of measurements for each individual, with columns indicating individual, time of measurement and the measurements.
+Thus, the primary format required for modeling such data is stacked series, blocks of measurements for each individual, with columns indicating individual, time of measurement and the measurements.
 (The tidy data name for this would be **long format**.) Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it into wide format which would have missing values in many cells.
-A drawback of this format is information unique to each individual is often repeated for all time points.
-The appealing feature is that data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
+A drawback of this format is that information unique to each individual is often repeated for all time points.
+An appealing feature is that data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
 This opens the door to easily operating on time to make calculations and extract different temporal components, such as month and day of the week.
@@
 The **tidyverse** largely formalizes exploratory data analysis.
-Many in the R community have adopted the **tidyverse** way of thinking and extended to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
+Many in the R community have adopted the **tidyverse** way of thinking and extended it to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
 Temporal data tools need to catch up.
@@
 These objects were designed for modelling methods.
-The diagram in the style of Figure \ref{fig:tsibble-fit}, would place the model at the center of the analytical universe, and all the transformations, visualisations would hinge on that format.
-This is contrary to the tidyverse conceptualisation, which wholistically captures the full data workflow.
+The diagram in the style of \autoref{fig:tsibble-fit}, would place the model at the center of the analytical universe, and all the transformations and visualisations would hinge on that format.
+This is contrary to the tidyverse conceptualisation, which holistically captures the full data workflow.
 
@@
 A relatively new R package **tibbletime** proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format.
-It only requires an index variable to declare a temporal data object, thus placing at the import stage.
-However, this is a loosely-defined structure and insufficient for time series analytics and models.
-Because it does not define interested subjects ($i$) over time and time uniqueness.
+It only requires an index variable to declare a temporal data object, thus placing it at the import stage.
+However, this is a loosely defined structure, and insufficient for time series analytics and models because it does not define interested subjects\todo{Needs explanation} over time or impose time uniqueness.
 
-This paper describes a new tidy representation for temporal data, and a unified framework to streamline the workflow from data pre-processing to visualization and forecasting, as an integral part of a tidy data analysis.
+This paper describes a new tidy representation for temporal data, and a unified framework to streamline the workflow from data preprocessing to visualization and forecasting, as an integral part of a tidy data analysis.
 
 # Contextual semantics {#sec:semantics}
+
+The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
+\autoref{fig:tsibble-fit} marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble".
+The tsibble structure is an extension of a data frame---a two-dimensional array in R---with additional time series semantics: index and key, as shown in \autoref{fig:tsibble-semantics}.
 
@@
 
-The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
-Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble".
-
 ```{r tb-sub}
@@
 
-Table \@ref(tab:tb-sub) presents a subset of tuberculosis cases estimated by @tb-data.
+\autoref{tab:tb-sub} presents a subset of tuberculosis cases estimated by @tb-data.
 It contains 12 observations and 5 variables arranged in a "long" tabular form.
-Each observation hosts the number of people who are diagnosed tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
-To turn this data into a tsibble, (1) column `year` is declared as the index variable; (2) the key can be made up of columns `country` and `gender`.
-The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables, for example, adding the corresponding population size (if known) in order to calibrate the count later.
+Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
+To turn this data into a tsibble: (1) column `year` is declared as the index variable; (2) the key is specified to consist of columns `country` and `gender`.
+The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables; for example, adding the corresponding population size (if known) in order to normalize the count later.
 
@@
 Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
-Index, key and time interval are the three stone pillars to this new semantically-structured temporal data.
+Index, key and time interval are the three stone pillars to this new semantically structured temporal data.
 
@@
 One could also join other data sources to the tsibble based on common time indexes.
-The accessibility of tsibble index motivates data analysis towards transparency and human readability.
-When the "index" used to be employed as meta information, it created an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.
+The accessibility of the tsibble index motivates data analysis towards transparency and human readability.
+When the "index" is available only as meta information (such as in the `ts` class), it created an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.
 
-A variable number of time representations is spotted in the wild.
+A variable number of time representations can be spotted in the wild.
 A Date-time object, universally accepted across computing systems, is the most commonly used type for representing time.
-Date-time also typically associates with a time zone with adjustments like summer time.
-This diversity and time zone is acknowledged and considered the building block for  *tsibble*'s index.
+Date-time also typically associates with a time zone with adjustments such as summer time.
+This diversity and time zone is acknowledged and considered the building block for **tsibble**'s index.
 
@@
 
-The "key" specification is the second essential ingrdient for a tsibble.
-The "key"  introduced uniquely identifies observations that are recorded over time in a data table.
+The "key" specification is the second essential ingredient for a tsibble.
+The "key" uniquely identifies observations that are recorded over time in a data table.
 It is similar to a primary key [@codd_relational_1970] defining each observation in a relational database.
-In the wide format, that multiple time series are often structured in, the columns hold a series of values, so that the column implicitly serves as identification.
+In the wide format in which multiple time series are often structured, the columns hold a series of values, so that the column implicitly serves as identification.
 In long format, columns are melted with names converted to "key" values.
@@
 It is not constrained to a single field, but can be composed from multiple fields.
-The identifying variables that the "key" is constituted from, remain the same as in the original table, with no further tweaks.
+The identifying variables that the "key" is constituted from, remain the same as in the original table with no further tweaks.
 
 The "key" is usually known a priori by analysts.
-For example, Table \ref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year.
+For example, \autoref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year.
 This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data.
-Lacking either of two will be inadequate, because the observations are not uniquely identified, and thus will generate a tsibble construction fail.
+Lacking either of the two will be inadequate, because the observations would not be uniquely identified, and thus a tsibble construction would fail.
 An alternative specification of the key for this data is to include a third variable `continent`.
@@
 
-The "key" also provides a solution to link between the data, models, and forecasts.
+The "key" also provides a link between the data, models, and forecasts.
 This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals.
@@
 seasonal ARIMA).
-The principal divide is regularly or irregularly spaced observations in time.
+The principal divide is between regularly or irregularly spaced observations in time.
 A tsibble permits implicit missing time, making it difficult to distinguish regularity from the index.
@@
 
-For data indexed in regular time space, the time interval is automatically calculated, by first computing absolute differences of time indexes and then the greatest common divisor.
+For data indexed in regular time space, the time interval is automatically calculated, by first computing absolute differences of time indexes and then finding the greatest common divisor.
 This covers all conceivable cases.
-The implication is that all observations in a tsibble have one and the only interval.
+The implication is that all observations in a tsibble have one and only interval.
 Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each observation may different underlying data generating processes.
 
-# Data pipeline {#sec:pipeline}
+# Data pipelines {#sec:pipeline}
 
@@
 A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualising the process as it might be applied to a variety of problems.
-The Extract, Transform, and Load (ETL), from recent data warehousing literature, outlines the workflow to prepare data for analysis, dates back to @kimball2011data can be considered a data pipeline.
-Building a data pipeline can be technically difficult, to make it sufficiently general for various data, with many implementation decisions on the interface, input and output objects and functionality.
+The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can be considered a data pipeline.
+Building a data pipeline can be technically difficult to make it sufficiently general for a variety of data, with many implementation decisions to be made about the interface, input and output objects and functionality.
 It is useful to articulate the data pipeline induced by new data tools.
 
-@unix coined the term "pipelines" in software development, while developing Unix at Bell Labs.
+@unix coined the term "pipelines" in software development while developing Unix at Bell Labs.
 In Unix-based computer operating systems, a pipeline chains together a series of operations on the basis of their standard streams, so that the output of each program becomes the input to another.
@@
 The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modelling.
-Each module communicates between each other, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.
+Each module communicates with the others, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.
 
@@
 
-The time-based pipeline shepards raw temporal data through to time series analysis, and plots.
+The time-based pipeline shepherds raw temporal data through to time series analysis, and plots.
 It is possible and recommended to check for identical entries of key and index before analysis.
@@
 When the data meets the tsibble standard, it flows neatly into the analysis stage.
-Figure \ref{fig:tsibble-pipeline} illustrates the time-based pipeline.
+\autoref{fig:tsibble-pipeline} illustrates the time-based pipeline.
 
@@
 
-## Time-based transformation {#sec:timepipeline}
+## Time-based transformations {#sec:timepipeline}
 
@@
 
-Each verb focuses on one oepration to achieve its goal.
+Each verb focuses on one operation to achieve its goal.
 The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
@@
 The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
-New tsibble-specific verbs (Table \ref{tab:verb-table} are provided to expand the **tidyverse** vocabulary.
+New tsibble-specific verbs (\autoref{tab:verb-table}) are provided to expand the **tidyverse** vocabulary.
 Users who are already familiar with **tidyverse**, will experience a gentle learning curve for mastering tsibble verbs and glide into temporal data analysis with low cognitive load.
@@
 Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data.
-The verbs listed in Table \ref{tab:verb-table} ease the transition to a tsibble that suits modeling.
+The verbs listed in \autoref{tab:verb-table} ease the transition to a tsibble that suits modeling.
 A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation.
@@
 
-The prime force that drives the software's design choices is "data".
+The primary force that drives the software's design choices is "data".
 All functions in **tsibble** start with `data` or its variants as the first argument, namely "data first".
@@
 Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-The  contextual summary provided by the print function shown below, on the data from Table \ref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
+The contextual summary provided by the print function shown below, on the data from \autoref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
 These details aid users in understanding their data better.
@@
 
-Rolling window calculations are widely-used techniques in time series analysis, but are not restricted to temporal applications.
+Rolling window calculations are widely used techniques in time series analysis, but are not restricted to temporal applications.
 These operations are dependent on having an ordering, particularly time ordering for temporal data.
@@
 
-Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in Table \@ref(tab:verb-table).
+Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in \autoref{tab:verb-table}.
 Table verbs expect and return a tsibble, and does what the function name suggests.
@@
 
-Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nantotime** [@R-nanotime], and pure numbers in simulations.
+Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nanotime** [@R-nanotime], and pure numbers in simulations.
 Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday.
@@
 These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year.
-New index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future.
+A new index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future.
 The interval regarding semesters is further outlined through `pull_interval()`.
@@
 The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble.
-For example, a mable is actually an extension of a tsibble, mentioned in Section \ref{sec:timepipeline}.
+For example, a mable is actually an extension of a tsibble, mentioned in \autoref{sec:timepipeline}.
 A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass.
@@
 
-The dataset of 2017 on-time performance for US domestic flights represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics [@flights-data].
+The dataset of on-time performance for US domestic flights in 2017 represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics [@flights-data].
 It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format.
 This kind of event describes each flight scheduled for departure at a time point in its local time zone.
-Every single flight would be uniquely identified by the flight number and its scheduled departure time, from a passenger's point of view.
+Every single flight should be uniquely identified by the flight number and its scheduled departure time, from a passenger's point of view.
 In fact, it fails to pass the tsibble hurdle due to duplicates in the original data.
-An error is immediately raised when attempting to convert this data into a tsibble, and a closer inspection has to be carried out to locate the issue.
+An error is immediately raised when attempting to convert this data into a tsibble, and closer inspection has to be carried out to locate the issue.
 The **tsibble** package provides tools to easily locate the duplicates in the data with `duplicates()`.
-Below shows the problematic entries.
+The problematic entries are shown below.
 
@@
 
-The issue is perhaps introduced when updating or entering the data into a system.
-The same flight is scheduled at exactly the same time, together with the same performance statistics but different flight details, which is very unlikely.
-Flight NK630 is usually scheduled at 17:45 from Chicago to New York, discovered by searching the full database.
-A decision is made to remove the first row from the duplicated entries before proceeding to the tsibble creation.
+The issue was perhaps introduced when updating or entering the data into a system.
+The same flight is scheduled at exactly the same time, together with the same performance statistics but different flight details.
+As flight NK630 is usually scheduled at 17:45 from Chicago to New York (discovered by searching the full database), a decision is made to remove the first row from the duplicated entries before proceeding to the tsibble creation.
 
@@
 
-This dataset is intrinsically heterogeneous, encoding in numbers, strings, and date-times.
-The tsibble framework, as expected, incorporates this type of data, without the loss of data richness and heterogeneity.
-To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as "index", column `flight_num` as "key" via `id(flight_num)`.
-As a result of event timing, this data is irregularly spaced, and hence switching to the irregular option is necessary.
-The program internally validates if the key and index produce the distinct rows, and then sort the key and the index from past to recent.
-When the tsibble creation is done, the print display is data-oriented and contextually informative, such as dimensions, irregular interval (``r tsibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tsibble::tbl_sum(us_flights)[2]``).
+This dataset is intrinsically heterogeneous, encoded in numbers, strings, and date-times.
+The tsibble framework, as expected, incorporates this type of data without any loss of data richness and heterogeneity.
+To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as the "index", and column `flight_num` as the "key" via `id(flight_num)`.
+As a result of event timing, the data are irregularly spaced, and hence switching to the irregular option is necessary.
+The software internally validates if the key and index produce distinct rows, and then sorts the key and the index from past to recent.
+When the tsibble creation is done, the print display is data-oriented and contextually informative, including dimensions, irregular interval (``r tsibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tsibble::tbl_sum(us_flights)[2]``).
 
@@
 
-Transforming tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can result in well-constructed pipelines.
-From the perspective of a passenger, one needs to travel smart, by choosing an efficient carrier to fly with and the time of day to avoid congestion, for example.
-We take a drill-down approach to exploring this data, starting with annual carrier performance and followed by disaggregating to finer time resolutions.
+Transforming a tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can result in well-constructed pipelines.
+From the perspective of a passenger, for example, one needs to travel smart by choosing an efficient carrier to fly with and the time of day to avoid congestion.
+To explore this data, we drill down starting with annual carrier performance and followed by disaggregation to finer time resolutions.
 
-Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
-This pipeline is initialized defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually.
+\autoref{fig:carrier-mosaic} visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
+This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually.
 To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column.
-The flow chart (Figure \@ref(fig:flight-pipeline1)) demonstrates the operations undertaking in a data pipeline.
+The flow chart (\autoref{fig:flight-pipeline1}) demonstrates the operations undertaking in the data pipeline.
 The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval.
-The final data is each carrier along with a single year, with the interval undetermined.
-It in turn feeds into the mosaic display.
+The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display.
 Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).
@@
 
-```{r carrier-mosaic, fig.cap = "(ref:carrier-mosaic-cap)"}
+```{r carrier-mosaic, fig.cap = "(ref:carrier-mosaic-cap)", fig.pos="b"}
 ```
 
-(ref:carrier-mosaic-cap) Mosaic plot showing the association between the size of airline carriers and the delayed proportion of departure in 2017.
-Southwest Airlines is the largest operator, but does not operate as efficient as Delta.
-Hawaiian Airlines, also as a small operator, outperforms the rest.
+(ref:carrier-mosaic-cap) Mosaic plot showing the association between the size of airline carriers and the delayed proportion of departures in 2017.
+Southwest Airlines is the largest operator, but does not operate as efficiently as Delta.
+Hawaiian Airlines, a small operator, outperforms the rest.
 
-```{r flight-pipeline1, fig.cap = "(ref:flight-pipeline1-cap)"}
+```{r flight-pipeline1, fig.cap = "(ref:flight-pipeline1-cap)", fig.pos="t"}
 include_graphics("img/flight-pipeline1")
@@
 
-(ref:flight-pipeline1-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:carrier-mosaic}.
+(ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:carrier-mosaic}.
 
@@
 But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without weight specification.
-The sliding operation with a large window size smoothes out the fluctuations and gives a stable trend around 25% over the year.
+The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year.
 LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady number delays over time.
@@
 
-```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 7}
+```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 7, fig.pos="t"}
 ```
@@
 
-(ref:flight-pipeline2-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:sel-monthly-plot}.
+(ref:flight-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:sel-monthly-plot}.
 
 What time of day and day of week should we travel to avoid suffering from horrible delay?
-Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples.
+\autoref{fig:draw-qtl} plots hourly quantile estimates across day of week in the form of small multiples.
 The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown.
@@
 
-```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)"}
+```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)", fig.pos="!htb"}
 include_graphics("img/flight-pipeline3")
@@
 
-(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:draw-qtl}.
+(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating \autoref{fig:draw-qtl}.
+
+\FloatBarrier
 
@@
 
-Sensors have installed in households across major cities in Australia to collect data for the smart city project.
+Sensors have been installed in households across major cities in Australia to collect data for the smart city project.
 One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter].
-Year 2013 has been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study.
-Half-hourly general supply in kwH have been recored for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
+Data from 2013 have been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study.
+Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
 Customers' demographic data provides explanatory variables other than time in a different data table.
@@
 Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for each household.
-Figure \@ref(fig:count-gaps) shows the gaps for the top 49 households arranged in rows from high to low in tallies.
+\autoref{fig:count-gaps} shows the gaps for the top 49 households arranged in rows from high to low in tallies.
 (The remaining households values have been aggregated into a single batch and appear at the top.) Missing values can be seen to occur at any time during the entire span.
@@
 (ref:count-gaps-cap) Time gap plots for the 49 customers with most implicit missing values, and the remaining customers grouped into the one line at top.
-Each cross represents a missing in time and a line between two dots shows continuous missingness over time.
+Each cross represents an observation missing in time and a line between two dots shows continuous missingness over time.
 Each row corresponds to one customer.
-Missing values occur at various times, with perhaps more in January.
+Missing values occur at various times, with more in January and February than other months.
 
-Aggregation across all individuals helps to sketch a big picture about the representative behaviors change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)).
-Each glyph represents the daily pattern of residential electricity usage every thirty minutes.
+Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (\autoref{fig:calendar-plot}).
+Each glyph represents the daily pattern of average residential electricity usage every thirty minutes.
 Higher consumption is indicated by higher values, and typically occurs in daylight hours.
-Colour indicates hot days.
+Color indicates hot days.
 The daily snapshots vary depending on the season in the year.
-During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially in the hot days with daily average temperature greater than 25 degrees.
+During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees C.
 However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.
@@
 
-(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized into calendar format, with colour indicating hot days.
+(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized into calendar format, with color indicating hot days.
 Energy use of hot days tends to be higher, suggesting air conditioner use.
@@
 
-(ref:elec-pipeline2-cap) Flow chart illustrates the pipeline that pre-processes the data for creating Figure \ref{fig:calendar-plot}.
+(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:calendar-plot}.
 
@@
 
-A new data abstraction to represent temporal data named as "tsibble" has been presented, spotlighting the "tidy data" principles brought to time domain.
-Tidy data begins to take shape in the state of time with the introduction of contextual semantics: index and key.
-A declared index provides direct support to time variable; variables that comprise the key define observations over time.
+The "tsibble" is a new data abstraction to represent temporal data, allowing the "tidy data" principles to brought to the time domain.
+Tidy data begins to take shape in the state of time with the introduction of the contextual semantics of index and key.
+A declared index provides direct support to the time variable; variables that comprise the key define observations over time.
 These semantics further determine unique data entries required for a valid tsibble.
-No matter how temporal data arrives, tsibble respects a time index and mainatins the data richness.
-A tsibble frictionlessly pops into transformation, visualization, modelling and smoothly shifts between, allowing for rapid iteration to gain data insights.
+No matter how temporal data arrives, a tsibble respects a time index and maintains the data richness.
+A tsibble frictionlessly allows transformation, visualization and modelling, and smoothly shifts between them, allowing for rapid iteration to gain data insights.
 
-A missing piece of the *tsibble* data is to enable user-defined calendars and respect structural missing observations.
-For example, a call center operates only between 9:00am and 5:00pm on week days and stock trading resumes on Monday straight after Friday.
-No data available outside trading hours would be labelled as structural missingess, which *tsibble* currently disregards.
-However, few R packages provide functionality to create and manage many sorts of calendars, including market-specific business calendar.
-<!-- This delays the implementation.--> Generally, custom calendars would be easily embedded into the tsibble framework.
-Consequently these tsibble operators, like `fill_gaps()`, would work out of box, and forecasts would be generated within its definable time range.
+A missing piece of the **tsibble** package is to enable user-defined calendars and to respect structurally missing observations.
+For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday.
+No data available outside trading hours would be labelled as structural missingness, which **tsibble** currently disregards.
+However, a few R packages provide functionality to create and manage many sorts of calendars, including market-specific business calendars.
+<!-- This delays the implementation.--> Generally, custom calendars are easily embedded into the tsibble framework.
+Consequently these tsibble operators, like `fill_gaps()`, would work out of the box, and forecasts would be generated within its definable time range.
 
@@
 But it is possible to work with remote tables stored in databases, such as SQLite and MySQL, using exactly the same tsibble code.
-This is the future work.
+This is left for future work.
 
@@
 
-The authors thank Mitch O'Hara-Wild for many discussions on the software design.
+The authors thank Mitchell O'Hara-Wild for many discussions on the software design.
 

### 2019-01-28 main.Rmd
Abstract

@@
     email : rob.hyndman@monash.edu
-abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, heterogeneous data types.
-Time series models, in particular the software supporting time series forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes.
+abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types.
+On the other hand, the software supporting time series modelling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes.
 Going from raw data to model-ready data is painful.
-This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and forecasting routines.
-Tidy data principles are extended to temporal data: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time.
+This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modelling and forecasting routines.
+Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts.

### 2019-01-28 main.Rmd
Section 1

@@
 -->
-Temporal data arrives in many possible formats, as does the time context.
-For example, time can carry with it various time resolutions (hours, minutes, and seconds), and they are typically associated with different time zones with adjustments such as summer time.
-Time can be irregularly recorded, which is often the case with longitudinal measurements such as patient visits to a doctor's office.
-Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, multiple grouping factors.
-Temporal data may be occurrence of events, such as flight departures, that need to be induced into a regular structure.
+Temporal data arrives in many possible formats, with many different time contexts.
+For example, time can carry with it various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time.
+Time can be regular (such as quarterly economic data or daily weather data), or irregular (such as patient visits to a doctor's office).
+Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, and multiple grouping factors.
+Temporal data may comprise the occurrence of events, such as flight departures, that need to be reduced to a regular structure.
 
-Current time series objects tend to be model-oriented matrices.
+Despite this variety and heterogeneity of temporal data, current software typically requires time series objects to be model-oriented matrices.
 Analysts are expected to do their own data preprocessing and take care of anything else needed to allow model fitting, which leads to a myriad of ad hoc solutions and duplicated efforts.
@@
 <!--
-with little support for non-numeric values, feed into the framework solely at the modelling stage.
+with little support for non-numeric values, feed into the framework solely at the modeling stage.
 Analysts are expected to do their own data preprocessing and take care of anything else needed to get to this stage, which leads to a myriad ad hoc solutions and duplicated efforts.
@@
 @r4ds succinctly describes the data science workflow given in Figure \ref{fig:tsibble-fit}.
-It is annotated to emphasize the significant divide between data that feeds into the beginning of analysis and the current time series objects that fits into modelling, making data preparation for modeling and forecasting painful.
+It is annotated to emphasize the significant divide between data that feeds into the beginning of analysis and the current time series objects that fits into modeling, making data preparation for modeling and forecasting painful.
 -->
@@
 
-(ref:tsibble-fit-cap) Illustration of data science workflow, drawn from @r4ds, showing how current time series tools interface with the workflow and how tsibble structure and tools integrate.
+(ref:tsibble-fit-cap) Illustration of the data science workflow, drawn from @r4ds, showing how current time series tools interface with the workflow and how the tsibble structure and tools integrate.
 The new data structure, tsibble, makes the connection between temporal data input, and specialist modeling formats.
-It provides elements at the "tidy" step, which produce tidy temporal data for time series visualisation and modeling.
+It provides elements at the "tidy" step, which produce tidy temporal data for time series visualization and modeling.
 

### 2019-01-28 main.Rmd
Section 2.1

@@
 
-The data problems are grouped into two types of analysis, time series and longitudinal.
-Despite being exactly the same data input, the representation of time series and longitudinal data diverges due to different modelling approaches.
+Temporal data problems are typically grouped into two types of analysis, time series and longitudinal.
+Despite being exactly the same data input, the representation of time series and longitudinal data diverges due to different modeling approaches.
 
-Time series can be univariate or multivariate, and for modelling require relatively long lengths (i.e., large $T$).
+Time series can be univariate or multivariate, and for modeling require relatively long lengths (i.e., large $T$).
 Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic processes, for the primary purpose of forecasting, and characterizing temporal dynamics.
@@
 In addition, time indexes are stripped off the data and implicitly inferred as attributes or meta-information.
-It strictly requires that the number of observations must be the same across all the series.
+There is a strict requirement that the number of observations must be the same across all the series.
 Data wrangling, from the form that data arrives in, to this specialist format, can be frustrating and difficult, inhibiting the variety of downstream tasks such as analytics.
@@
 It often occurs that measurements for individuals are taken at different time points, resulting in an unbalanced panel.
-Thus, the primary format required for modeling such data is stacked series, blocks of measurements for each individual, with columns indicating individual, time of measurement and the measurements.
-(The tidy data name for this would be **long format**.) Evidently, this data organisation saves storage space for many sparse cells, compared to structuring it into wide format which would have missing values in many cells.
+Thus, the primary format required for modeling such data is stacked series, blocks of measurements for each individual, with columns indicating individual, times of measurement and the measurements themselves.
+(The tidy data name for this would be **long format**.) Evidently, this data organization saves storage space for many sparse cells, compared to structuring it into wide format which would have missing values in many cells.
 A drawback of this format is that information unique to each individual is often repeated for all time points.

### 2019-01-28 main.Rmd
Section 2.2

@@
 This data structure is the fundamental unit of the **tidyverse**, which is a collection of R packages designed for data science.
-The ubiquitous use of **tidyverse** is testament to the simplicity, practicality and general applicability of the tools.
+The ubiquitous use of the **tidyverse** is testament to the simplicity, practicality and general applicability of the tools.
 The **tidyverse** provides abstract yet functional grammars to manipulate and visualize data in easier-to-comprehend form.
-One of the **tidyverse** packages, **dplyr** [@R-dplyr] showcases the value of a grammar as a principled vehicle to transform data for a wide range of data challenges, providing a consistent set of verbs: `mutate()`, `select()`, `filter()`, `summarise()`, and `arrange()`.
+One of the **tidyverse** packages, **dplyr** [@R-dplyr], showcases the value of a grammar as a principled vehicle to transform data for a wide range of data challenges, providing a consistent set of verbs: `mutate()`, `select()`, `filter()`, `summarize()`, and `arrange()`.
 Each verb focuses on a singular task.

### 2019-01-28 main.Rmd
Section 2.3

@@
 
-Current standards, provided by the native `ts` object in R, extended by **zoo** [@zoo2005], and **xts** [@R-xts], assemble temporal data into matrices with implicit time indexes.
-These objects were designed for modelling methods.
-The diagram in the style of \autoref{fig:tsibble-fit}, would place the model at the center of the analytical universe, and all the transformations and visualisations would hinge on that format.
-This is contrary to the tidyverse conceptualisation, which holistically captures the full data workflow.
+Current standards, provided by the native `ts` object in R, and extended by **zoo** [@zoo2005] and **xts** [@R-xts], assemble temporal data into matrices with implicit time indexes.
+These objects were designed for modeling methods.
+The diagram in the style of \autoref{fig:tsibble-fit} would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
+This is contrary to the tidyverse conceptualization, which holistically captures the full data workflow.
 

### 2019-01-28 main.Rmd
Section 3

@@
 
-(ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organisation in 2011 and 2012, with 12 observations and 5 variables.
+(ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organization in 2011 and 2012, with 12 observations and 5 variables.
 The index refers to column `year`, the key to multiple columns: `country` and `gender`, and the measured variable to column `count`.
@@
 Index, key and time interval are the three stone pillars to this new semantically structured temporal data.
+Each is now described in more detail.
 
@@
 The "index" is an intact data column rather than a masked attribute, which makes time visible and accessible to users.
-It is highly advantageous when manipulating time.
+This is highly advantageous when manipulating time.
 For example, one could easily extract time components, such as time of day and day of week, from the index to visualize seasonal effects of response variables.
@@
 The accessibility of the tsibble index motivates data analysis towards transparency and human readability.
-When the "index" is available only as meta information (such as in the `ts` class), it created an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.
+When the "index" is available only as meta information (such as in the `ts` class), it creates an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.
 
 A variable number of time representations can be spotted in the wild.
-A Date-time object, universally accepted across computing systems, is the most commonly used type for representing time.
-Date-time also typically associates with a time zone with adjustments such as summer time.
-This diversity and time zone is acknowledged and considered the building block for **tsibble**'s index.
+A date-time object, universally accepted across computing systems, is the most commonly used type for representing time.
+Date-time also typically associates with a time zone including adjustments such as summer time.
+This diversity and time zone is acknowledged and accommodated by **tsibble**'s index.
 
@@
 It is not constrained to a single field, but can be composed from multiple fields.
-The identifying variables that the "key" is constituted from, remain the same as in the original table with no further tweaks.
+The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks.
 
@@
 It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals.
-This type of data can flow into event-based data modelling, but would need to be processed or regularized to fit models that expect time series.
+This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect time series.
 
 For data indexed in regular time space, the time interval is automatically calculated, by first computing absolute differences of time indexes and then finding the greatest common divisor.
-This covers all conceivable cases.
-The implication is that all observations in a tsibble have one and only interval.
-Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each observation may different underlying data generating processes.
+This covers all conceivable cases, assuming that all observations in a tsibble have only one interval.
+Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each observation may have different underlying data generating processes.
 

### 2019-01-28 main.Rmd
Section 4

@@
 There has been a long history of pipeline discussions and implementation centering around data in various aspects.
-A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualising the process as it might be applied to a variety of problems.
+A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualizing the process as it might be applied to a variety of problems.
 The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can be considered a data pipeline.
@@
 
-The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modelling.
+The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modeling.
 Each module communicates with the others, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.

### 2019-01-28 main.Rmd
Section 4.1

@@
 To prevent inviting these errors into the analysis, it is a good practice to check and inspect any time gaps of the tsibble in the first place.
-Several tools are provided to understand and tackle implicit missing values in: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reports all implicit missing observations; (3) `count_gaps()` summarises the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
+Several tools are provided to understand and tackle implicit missing values in: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reports all implicit missing observations; (3) `count_gaps()` summarizes the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
 
@@
  & \textbf{\texttt{transmute()}} & Drops existing variables \\
- & \textbf{\texttt{summarise()}} & Aggregate values over time \\ \hline
+ & \textbf{\texttt{summarize()}} & Aggregate values over time \\ \hline
 \multirow{3}{*}{Group-wise} & \texttt{index\_by()} & Group by index candidate \\

### 2019-01-28 main.Rmd
Section 5

@@
 Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-The contextual summary provided by the print function shown below, on the data from \autoref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
+The contextual summary provided by the print function, shown below on the data from \autoref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
 These details aid users in understanding their data better.

### 2019-01-28 main.Rmd
Section 6

@@
 
-```{r carrier-mosaic, fig.cap = "(ref:carrier-mosaic-cap)", fig.pos="b"}
+```{r carrier-mosaic, fig.cap = "(ref:carrier-mosaic-cap)"}
 ```
@@
 
-```{r flight-pipeline1, fig.cap = "(ref:flight-pipeline1-cap)", fig.pos="t"}
+```{r flight-pipeline1, fig.cap = "(ref:flight-pipeline1-cap)"}
 include_graphics("img/flight-pipeline1")
@@
 
-```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 7, fig.pos="t"}
+```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 7}
 ```
@@
 
-```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)", fig.pos="!htb"}
+```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)"}
 include_graphics("img/flight-pipeline3")

### 2019-01-28 main.Rmd
Section 7

@@
 No matter how temporal data arrives, a tsibble respects a time index and maintains the data richness.
-A tsibble frictionlessly allows transformation, visualization and modelling, and smoothly shifts between them, allowing for rapid iteration to gain data insights.
+A tsibble frictionlessly allows transformation, visualization and modeling, and smoothly shifts between them, allowing for rapid iteration to gain data insights.
 
@@
 For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday.
-No data available outside trading hours would be labelled as structural missingness, which **tsibble** currently disregards.
+No data available outside trading hours would be labeled as structural missingness, which **tsibble** currently disregards.
 However, a few R packages provide functionality to create and manage many sorts of calendars, including market-specific business calendars.
@@
 
-# Acknowledgements {-}
+# Acknowledgments {-}
 

### 2019-01-28 main.Rmd
fixed bibliography entries and added a couple more #3

@@
 A new temporal data class is needed in the upstream of the workflow, which could incorporate all the downstream modules.
-A relatively new R package **tibbletime** proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format.
+A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format.
 It only requires an index variable to declare a temporal data object, thus placing it at the import stage.
@@
 The primary force that drives the software's design choices is "data".
-All functions in **tsibble** start with `data` or its variants as the first argument, namely "data first".
+All functions in in the package **tsibble** [@R-tsibble] start with `data` or its variants as the first argument, namely "data first".
 They work naturally with the pipe operator `%>%`, read as "then".
@@
 
-The **purrr** package provides a good example of functional programming in R.
+The **purrr** package [@R-purrr] provides a good example of functional programming in R.
 It provides a complete and consistent set of tools to iterate each element of a vector with a function.
@@
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr", "nanotime"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "knitr", "rmarkdown"), file = "rpkgs.bib")
 ```

### 2019-01-28 main.Rmd
more in acknowledgements #3

@@
 
-The authors thank Mitchell O'Hara-Wild for many discussions on the software design.
+The authors would like to thank Mitchell O'Hara-Wild for many discussions on the software development and Davis Vaughan for contributing ideas on rolling window functions.
+This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
+The project's Github repository <https://github.com/earowang/paper-tsibble> houses all materials required to reproduce this article and a history of the changes.
+To produce the analysis results as described in Section \autoref{sec:cases}, we require the **tsibble** package (`v0.6.2`) and the **dplyr** package (`v0.7.8`) installed.
 
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "knitr", "rmarkdown"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble"), file = "rpkgs.bib")
 ```

### 2019-01-28 main.Rmd
minor tweaks and replace mable with fable #3

@@
 abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types.
-On the other hand, the software supporting time series modelling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes.
+On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes.
 Going from raw data to model-ready data is painful.
-This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modelling and forecasting routines.
+This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines.
 Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time.
@@
 Temporal data arrives in many possible formats, with many different time contexts.
-For example, time can carry with it various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time.
+For example, time can carry with its various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time.
 Time can be regular (such as quarterly economic data or daily weather data), or irregular (such as patient visits to a doctor's office).
@@
 Time series modeling and forecasting enters the picture at the *modeling* stage.
-While temporal data enters at the start, there isn't a current framework for getting temporal data into the matrix format for modeling.
+While temporal data enters at the start, there is not a current framework for getting temporal data into the matrix format for modeling.
 That's the part of the workflow that the work described in this paper improves.
-
-<!--
-with little support for non-numeric values, feed into the framework solely at the modeling stage.
-Analysts are expected to do their own data preprocessing and take care of anything else needed to get to this stage, which leads to a myriad ad hoc solutions and duplicated efforts.
-
-@r4ds succinctly describes the data science workflow given in Figure \ref{fig:tsibble-fit}.
-It is annotated to emphasize the significant divide between data that feeds into the beginning of analysis and the current time series objects that fits into modeling, making data preparation for modeling and forecasting painful.
--->
-
-<!--
-```{r pipeline-ann, fig.cap = "(ref:pipeline-ann-cap)"}
-include_graphics("img/motivation.pdf")
-```
-
-(ref:pipeline-ann-cap) An overview of the data science program centering around data.
-Temporal data should feed into the import stage, but the existing time series objects are model-focused, thus resulting misalliance between data and model.
--->
 
@@
 The diagram in the style of \autoref{fig:tsibble-fit} would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
-This is contrary to the tidyverse conceptualization, which holistically captures the full data workflow.
+This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.
 
@@
 It only requires an index variable to declare a temporal data object, thus placing it at the import stage.
-However, this is a loosely defined structure, and insufficient for time series analytics and models because it does not define interested subjects\todo{Needs explanation} over time or impose time uniqueness.
+However, this is a loosely defined structure, and insufficient for time series analytics and models.
+Because it neglects to describe observational units or entities over time, which forms a key component of time series data.
 
@@
 Date-time also typically associates with a time zone including adjustments such as summer time.
-This diversity and time zone is acknowledged and accommodated by **tsibble**'s index.
+This diversity and time zone is acknowledged and accommodated by tsibble's index.
 
@@
 To prevent inviting these errors into the analysis, it is a good practice to check and inspect any time gaps of the tsibble in the first place.
-Several tools are provided to understand and tackle implicit missing values in: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reports all implicit missing observations; (3) `count_gaps()` summarizes the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
+Several tools are provided to understand and tackle implicit missing values in: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reveals all implicit missing observations; (3) `count_gaps()` summarizes the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
 
@@
 \multirow{4}{*}{Time gaps} & \texttt{has\_gaps()} & Test if a tsibble has gaps in time \\
- & \texttt{scan\_gaps()} & Report implicit missing entries \\
+ & \texttt{scan\_gaps()} & Reveal implicit missing entries \\
  & \texttt{count\_gaps()} & Summarize time gaps \\
@@
 This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems.
-Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package.
+Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr].
 A sequence of functions can be composed in a way that can be naturally read from left to right, which improves the readability of the code.
@@
 This avoids expensive data copying and reduces model storage.
-The mable is further supplied to forecasting methods, to produce a tsibble in which each "key" along with its future time holds predictions.
-It also underlines the advantage of tsibble's "key" in linking between data inputs, models and forecasts.
+The mable is further supplied to forecasting methods, to produce a "fable" (short for forecasting table) in which each "key" along with its future time holds predictions.
+It also underlines the advantage of tsibble's "key" in acting as linkage between data inputs, models and forecasts.
 Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in a similar spirit.
@@
 The primary force that drives the software's design choices is "data".
-All functions in in the package **tsibble** [@R-tsibble] start with `data` or its variants as the first argument, namely "data first".
+All functions in the package **tsibble** [@R-tsibble] start with `data` or its variants as the first argument, namely "data first".
 They work naturally with the pipe operator `%>%`, read as "then".
@@
 Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding type, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively.
-Their multiprocessing equivalents prefixed by `future_` enable rolling in parallel.
+Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel.
 This family of functions empowers users to incorporate window-related operations in their workflows.
@@
 Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday.
-The tsibble package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters.
+The **tsibble** package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters.
 These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year.
@@
 The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble.
-For example, a mable is actually an extension of a tsibble, mentioned in \autoref{sec:timepipeline}.
+For example, a fable is actually an extension of a tsibble, mentioned in \autoref{sec:timepipeline}.
 A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass.
 This new object itself is a tsibble.
-It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a mable.
+It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable.
 Tsibble verbs are also S3 generics.
@@
 
-The "tsibble" is a new data abstraction to represent temporal data, allowing the "tidy data" principles to brought to the time domain.
+The "tsibble" is a new data abstraction to represent temporal data, allowing the "tidy data" principles to be brought to the time domain.
 Tidy data begins to take shape in the state of time with the introduction of the contextual semantics of index and key.
@@
 For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday.
-No data available outside trading hours would be labeled as structural missingness, which **tsibble** currently disregards.
+No data available outside trading hours would be labeled as structural missingness, which tsibble currently disregards.
 However, a few R packages provide functionality to create and manage many sorts of calendars, including market-specific business calendars.
@@
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr"), file = "rpkgs.bib")
 ```

### 2019-01-28 main.Rmd
shorten animation interval #3

@@
 
-<!-- note to self: There's a trade-off between data column and attributes.
-Data column is easily accessible, which gives greater flexibility; however, once the value in the data column is modified, no way to validate for any static approach.
-Generally, not recommended to modify attributes.
--->
-
 Time provides a contextual basis for temporal data.
@@
 
-<!--
-Transformation assembles a set of modules.
-A unit that makes up a module is a function or preferably "verb".
-A tsibble is an object, conceptually considered as a noun, and hence an action performed on the object can be phrased as a verb.
--->
-
 Each verb focuses on one operation to achieve its goal.
@@
 
-<!--
-We go through the whole exploration circle, and keep iterating and refining until data insights gained.
-The tsibble data structure substantially lubricates between these modules for time-based pipelines.
-The cohesive and coherent framework results in more graceful and expressive code.
--->
-
 # Software structure and design decisions {#sec:software}
@@
 
-```{r tb-au}
-```
-
-```{r slide-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:slide-animate-cap)"}
-```
-
-(ref:slide-animate-cap) An illustration of window of size 5 sliding over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
-
-```{r tile-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:tile-animate-cap)"}
-```
-
-(ref:tile-animate-cap) An illustration of window of size 5 tiling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
-
-```{r stretch-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:stretch-animate-cap)"}
-```
-
-(ref:stretch-animate-cap) An illustration of window of size 5 stretching over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
-
 Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in \autoref{tab:verb-table}.
@@
 However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.
-<!-- In addition, the same function is called and applied to each window of data independently, which means programming is done with for-loop statements.
-But functional programming emphasizes expressions instead of statements.-->
 
@@
 This family of functions empowers users to incorporate window-related operations in their workflows.
+
+```{r tb-au}
+```
+
+```{r slide-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:slide-animate-cap)", interval = 0.3}
+```
+
+(ref:slide-animate-cap) An illustration of window of size 5 sliding over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
+
+```{r tile-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:tile-animate-cap)", interval = 0.3}
+```
+
+(ref:tile-animate-cap) An illustration of window of size 5 tiling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
+
+```{r stretch-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:stretch-animate-cap)", interval = 0.3}
+```
+
+(ref:stretch-animate-cap) An illustration of window of size 5 stretching over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
+
+\newpage
 
@@
 It is generally easier to trace back the errors users encounter from separating 1d and 2d functions.
-<!--(lost in a web of functions)-->
 
@@
 The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble.
-For example, a fable is actually an extension of a tsibble, mentioned in \autoref{sec:timepipeline}.
+For example, a fable is actually an extension of a tsibble, mentioned in \autoref{sec:ts-models}.
 A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass.

### 2019-01-28 R/demo.R

@@ -7,7 +7,7 @@ tb_small <- read_rds("data/tb-small.rds")
 tb_small %>%
   filter(year > 2010) %>%
   knitr::kable(booktabs = TRUE, caption = "(ref:tb-sub)", linesep = "") %>%
-  kableExtra::kable_styling(position = "center", latex_options=c("hold_position"))
+  kableExtra::kable_styling(position = "center", latex_options= "hold_position")
 
 ## ---- tb-print
 as_tsibble(tb_small, key = id(country, gender), index = year) %>%

### 2019-01-28 R/flights.R
rm unused R code

@@ -3,65 +3,9 @@ library(lubridate)
 library(tidyverse)
 library(tsibble)
 library(forcats)
-# devtools::install_github("heike/ggmapr")
-library(ggmapr)
 
 ## ---- load-flights
 flights <- read_rds("data/flights.rds")
-
-## ---- map-prep
-origin_dest <- flights %>% 
-  distinct(origin, origin_state, dest, dest_state)
-airports <- read_rds("data/airports.rds")
-map_dat <- origin_dest %>% 
-  left_join(airports, by = c("origin" = "faa")) %>% 
-  rename(long = lon) %>% 
-  shift(origin_state == "HI", shift_by = c(52.5, 5.5)) %>%
-  scale(origin_state == "AK", scale = 0.3, set_to = c(-117, 27)) %>%
-  rename(origin_lat = lat, origin_lon = long)  %>% 
-  left_join(select(airports, faa, lon, lat), by = c("dest" = "faa")) %>% 
-  rename(long = lon) %>% 
-  shift(dest_state == "HI", shift_by = c(52.5, 5.5)) %>%
-  scale(dest_state == "AK", scale = 0.3, set_to = c(-117, 27)) %>%
-  rename(dest_lat = lat, dest_lon = long)
-
-states <- states %>%
-  shift(NAME == "Hawaii", shift_by = c(52.5, 5.5)) %>%
-  scale(NAME == "Alaska", scale = 0.3, set_to = c(-117, 27)) %>%
-  filter(lat > 20)
-
-## ---- map-airlines
-ggplot() +
-  geom_polygon(data= states, aes(x = long, y = lat, group = group), 
-    fill = "white", colour = "grey60") +
-  geom_segment(data = map_dat, aes(
-    x = origin_lon, y = origin_lat, xend = dest_lon, yend = dest_lat
-  ), alpha = 0.1, size = 0.4, colour = "#762a83") +
-  geom_point(data = map_dat, aes(x = origin_lon, y = origin_lat), 
-    colour = "#a6dba0", size  = 0.7) +
-  coord_map("albers", parameters = c(30, 45)) +
-  ggthemes::theme_map()
-
-## ---- dl771
-dl771 <- flights %>% 
-  filter(flight_num == "DL771") %>% 
-  left_join(map_dat) %>% 
-  arrange(sched_dep_datetime)
-
-ggplot() +
-  geom_polygon(data= states, aes(x = long, y = lat, group = group), 
[…]

### 2019-01-28 R/smart-meter.R

@@ -66,111 +66,3 @@ p_cal <- elec_cal %>%
   scale_colour_brewer(palette = "Dark2", direction = -1) +
   theme(legend.position = "bottom")
 prettify(p_cal)
-
-## ---- elec-quantiles
-qtl_grid <- seq(0.1, 0.9, 0.002)
-elec_qtl <- elec_ts %>% 
-  summarise(
-    value = list(quantile(general_supply_kwh, qtl_grid, na.rm = TRUE)),
-    qtl = list(qtl_grid)
-  ) %>% 
-  unnest(key = id(qtl)) %>% 
-  mutate(
-    time = hms::as.hms(reading_datetime, tz = "UTC"),
-    date = as_date(reading_datetime)
-  )
-
-elec_med <- elec_qtl %>% 
-  filter(qtl == 0.5)
-
-elec_qtl %>% 
-  ggplot(aes(x = time, y = value, colour = qtl, group = qtl)) +
-  geom_line() +
-  geom_line(data = elec_med, colour = "#de2d26", size = 0.5) +
-  sugrrants::facet_calendar(~ date, ncol = 3, format = "%m/%d") +
-  scale_x_time(
-    breaks = hms::as.hms(c("6:00:00", "18:00:00")),
-    labels = c("6:00", "18:00")
-  ) +
-  scale_colour_viridis_c() +
-  theme(legend.position = "none")
-
-## ---- fill-gaps
-# make complete series
-elec_ts <- elec_ts %>% 
-  fill_gaps()
-
-elec_na_mth <- elec_na %>% 
-  mutate(
-    mth = month(reading_datetime, label = TRUE),
-    is_na = 1L
-  ) %>% 
-  group_by(customer_id) %>% 
-  index_by(mth) %>% 
-  summarise(n_na = sum(is_na))
-
-elec_na_mth %>% 
-  ggplot(aes(x = mth)) +
-  geom_bar()
-
-elec_ts <- elec_ts %>% 
-  mutate(season = if_else(
-    time_in(reading_datetime, ~ "2013-03", "2013-10" ~ .),
-    "Autumn-Winter", "Spring-Summer"
-  ))
-
-qtl_grid <- seq(0.1, 0.9, 0.01)
-gas_aircon <- elec_ts %>% 
-  inner_join(
-    customer %>% 
-      select(customer_key, has_gas, has_aircon) %>% 
-      drop_na(has_gas, has_aircon),
-    by = c("customer_id" = "customer_key")
[…]

### 2019-01-28 main.Rmd
checked typos
possibly dplyr v0.8.0 will not have performance issue for date-times

@@
 However, this is a loosely defined structure, and insufficient for time series analytics and models.
-Because it neglects to describe observational units or entities over time, which forms a key component of time series data.
+Because it neglects to describe observational units or entities over time, which form a key component of time series data.
 
@@
 The project's Github repository <https://github.com/earowang/paper-tsibble> houses all materials required to reproduce this article and a history of the changes.
-To produce the analysis results as described in Section \autoref{sec:cases}, we require the **tsibble** package (`v0.6.2`) and the **dplyr** package (`v0.7.8`) installed.
+<!-- To produce the analysis results as described in Section \autoref{sec:cases}, we require the **tsibble** package (`v0.6.2`) and the **dplyr** package (`v0.7.8`) installed.
+-->
 

### 2019-01-29 main.Rmd
tweak wording

@@
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts.
-A sound data pipeline facilitates a fluent and fluid workflow for analyzing temporal data.
+A sound data pipeline facilitates a fluent workflow for analyzing temporal data.
 The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
@@
 Temporal data arrives in many possible formats, with many different time contexts.
-For example, time can carry with its various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time.
+For example, time can have various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time.
 Time can be regular (such as quarterly economic data or daily weather data), or irregular (such as patient visits to a doctor's office).
@@
 
-In recent work, there have been some exciting developments that conceptually and practically streamline the preprocessing of data.
-\autoref{fig:tsibble-fit} [adapted from @r4ds] illustrates the tidy data workflow.
-Time series modeling and forecasting enters the picture at the *modeling* stage.
-While temporal data enters at the start, there is not a current framework for getting temporal data into the matrix format for modeling.
-That's the part of the workflow that the work described in this paper improves.
+@r4ds proposed the tidy data workflow, which provides a conceptual 
+framework for processing data (as described in \autoref{fig:tsibble-fit}).
+Currently, time series modeling and forecasting enters this framework at the *modeling* stage, while temporal data enters at the start.
+This paper integrates
+time series analysis into this tidy framework, providing a coherent way for getting temporal data into the matrix format for modeling.
 
@@
 Two case studies are presented in \autoref{sec:cases} illustrating temporal data exploration using the newly implemented infrastructure.
-\autoref{sec:conclusion} discusses the future work.
+\autoref{sec:conclusion} discusses future work.
 
@@
 There is a strict requirement that the number of observations must be the same across all the series.
-Data wrangling, from the form that data arrives in, to this specialist format, can be frustrating and difficult, inhibiting the variety of downstream tasks such as analytics.
+Data wrangling, from the form that data arrives in, to this specialist format, can be frustrating and difficult, inhibiting the performance of downstream tasks.
 
@@
 It only requires an index variable to declare a temporal data object, thus placing it at the import stage.
-However, this is a loosely defined structure, and insufficient for time series analytics and models.
-Because it neglects to describe observational units or entities over time, which form a key component of time series data.
+However, as proposed in \autoref{sec:semantics} a more rigid data structure is required for time series analytics and models.
 
@@
 
+To demonstrate the concept of the tsibble, \autoref{tab:tb-sub} presents a subset of tuberculosis cases estimated by @tb-data.
+It contains 12 observations and 5 variables arranged in a "long" tabular form.
+Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
+To turn this data into a tsibble: (1) column `year` is declared as the index variable; (2) the key is specified to consist of columns `country` and `gender`.
+The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables; for example, adding the corresponding population size (if known) in order to normalize the count later.
+
 ```{r tb-sub}
@@
 The index refers to column `year`, the key to multiple columns: `country` and `gender`, and the measured variable to column `count`.
-
-\autoref{tab:tb-sub} presents a subset of tuberculosis cases estimated by @tb-data.
-It contains 12 observations and 5 variables arranged in a "long" tabular form.
-Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
-To turn this data into a tsibble: (1) column `year` is declared as the index variable; (2) the key is specified to consist of columns `country` and `gender`.
-The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables; for example, adding the corresponding population size (if known) in order to normalize the count later.
 
@@
 This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals.
-More details are given in the following section.
+More details are given  in \autoref{sec:ts-models}.
 
@@
 @viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot.
-@xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **Xgobi**, **Ggobi**, **Orca**, **plumbr** and **cranvas**, respectively.
+@xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **XGobi**, **GGobi**, **Orca**, **plumbr** and **cranvas**, respectively.
 The pipeline is typically described with a one way flow, from data to plot.
@@
 It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
+
+## Time-based transformations {#sec:timepipeline}
 
@@
 
-## Time-based transformations {#sec:timepipeline}
-
+A suite of verbs are used to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, and transformed for visualization.
 Many time operations like lag/lead and time series models, assume an intact vector input ordered in time.
@@
 
-A suite of verbs, and corresponding functions, are used to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, and transformed for visualization.
 A tsibble is an object, conceptually considered as a noun, and hence an action performed on the object can be phrased as a verb.
@@
 
-# Software structure and design decisions {#sec:software}
+# Software implementation {#sec:software}
 
@@
 All functions in the package **tsibble** [@R-tsibble] start with `data` or its variants as the first argument, namely "data first".
-They work naturally with the pipe operator `%>%`, read as "then".
-This not only lays out a consistent interface but also addresses the significance of the data throughout the software.
+This lays out a consistent interface and addresses the significance of the data throughout the software.
 
@@
 3. **stretch**: fixing an initial window and expanding to include more observations.
-
-Figures \@ref(fig:slide-animate), \@ref(fig:tile-animate) and \@ref(fig:stretch-animate) show the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
-A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.
-
-Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in \autoref{tab:verb-table}.
-Table verbs expect and return a tsibble, and does what the function name suggests.
-On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window.
-For example, computing moving averages requires numerics and a function like `mean()`, and produces averaged numerics.
-However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.
-
-The **purrr** package [@R-purrr] provides a good example of functional programming in R.
-It provides a complete and consistent set of tools to iterate each element of a vector with a function.
-Rolling window doesn't just iterate but rolls over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
-`slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
-For type stability, the functions always return lists.
-Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding type, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively.
-Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel.
-This family of functions empowers users to incorporate window-related operations in their workflows.
 
@@
 
-\newpage
+Figures \@ref(fig:slide-animate), \@ref(fig:tile-animate) and \@ref(fig:stretch-animate) show the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
+A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.
+
+Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in \autoref{tab:verb-table}.
+Table verbs expect and return a tsibble, and does what the function name suggests.
+On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window.
+For example, computing moving averages requires numerics and a function like `mean()`, and produces averaged numerics.
+However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.
+
+The **purrr** package [@R-purrr] provides a good example of functional programming in R.
+It provides a complete and consistent set of tools to iterate each element of a vector with a function.
+Rolling window doesn't just iterate but rolls over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
+`slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
+For type stability, the functions always return lists.
+Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding type, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively.
+Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel [@R-future; @R-furrr].
+This family of functions empowers users to incorporate window-related operations in their workflows.
 
@@
 The authors would like to thank Mitchell O'Hara-Wild for many discussions on the software development and Davis Vaughan for contributing ideas on rolling window functions.
+We also thank Stuart Lee for the feedback on this manuscript.
 This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
@@
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr"), file = "rpkgs.bib")
 ```

### 2019-01-29 main.Rmd
better animation

@@
 blind: false
-cover: true
+cover: false
 toc: false
@@
 
-# Software implementation {#sec:software}
+# Software structure and design decisions {#sec:software}
 
@@
 
+\autoref{fig:animate} shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
+A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.
+
 ```{r tb-au}
@@
 
-```{r slide-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:slide-animate-cap)", interval = 0.3}
+```{r animate, fig.show = "animate", fig.height = 7, fig.cap = "(ref:animate-cap)", interval = 0.5, aniopts = c("autoplay", "controls", "loop")}
 ```
 
-(ref:slide-animate-cap) An illustration of window of size 5 sliding over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
-
-```{r tile-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:tile-animate-cap)", interval = 0.3}
-```
-
-(ref:tile-animate-cap) An illustration of window of size 5 tiling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
-
-```{r stretch-animate, fig.show = "animate", fig.height = 2.5, fig.cap = "(ref:stretch-animate-cap)", interval = 0.3}
-```
-
-(ref:stretch-animate-cap) An illustration of window of size 5 stretching over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
-
-Figures \@ref(fig:slide-animate), \@ref(fig:tile-animate) and \@ref(fig:stretch-animate) show the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
-A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.
+(ref:animate-cap) An illustration of window of size 5 rolling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
 
@@
 
-\autoref{fig:carrier-mosaic} visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
-This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually.
-To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column.
-The flow chart (\autoref{fig:flight-pipeline1}) demonstrates the operations undertaking in the data pipeline.
-The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval.
-The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display.
-Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).
-
 ```{r carrier-delayed}
@@
 (ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:carrier-mosaic}.
+
+\autoref{fig:carrier-mosaic} visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
+This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually.
+To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column.
+The flow chart (\autoref{fig:flight-pipeline1}) demonstrates the operations undertaking in the data pipeline.
+The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval.
+The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display.
+Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).
 

### 2019-01-29 R/demo.R

@@ -20,57 +20,75 @@ tb_au <- tb_small %>%
   group_by(year) %>%
   summarise(count = sum(count))
 
-## ---- slide-animate
+## ---- animate
 library(gganimate)
+reveal_group <- function(data, group){
+  group <- enquo(group)
+  data <- as_tibble(data)
+  data <- transmute(
+    data, !! group,
+    .dt = map(map(!! group, seq_len), function(groups, data) {
+      data <- filter(data, !! group %in% groups)
+      select(data, !! expr(-!! group))
+    }, data
+  ))
+  unnest(data, .dt)
+}
+types <- forcats::fct_inorder(c("Sliding", "Tiling", "Stretching"))
 slide_window <- slider(tb_au$year, .size = 5) %>%
   map_dfr(function(x) tibble(xmin = min(x), xmax = max(x))) %>%
-  mutate(ymin = -Inf, ymax = Inf, group = row_number())
-ggplot() +
-  geom_line(aes(year, count), data = tb_au) +
-  geom_point(aes(year, count), data = tb_au) +
-  geom_rect(aes(
-    xmin = xmin, xmax = xmax,
-    ymin = ymin, ymax = ymax,
-    group = group
-  ), data = slide_window, fill = "#9ecae1", alpha = 0.6) +
-  xlab("Year") +
-  ylab("Count") +
-  ylim(c(0, max(tb_au$count))) +
-  theme_bw() +
-  transition_manual(group)
+  mutate(ymin = -Inf, ymax = Inf, group = row_number(), type = types[1])
+slide_mean <- tb_au %>%
+  mutate(ma = slide_dbl(count, ~ mean(.x), .size = 5, .align = "r")) %>%
+  mutate(group = pmax(0, row_number() - 4), type = types[1])
+slide_revealed <- slide_mean %>%
+  reveal_group(group)
 
-## ---- tile-animate
 tile_window <- tiler(tb_au$year, .size = 5) %>%
   map_dfr(function(x) tibble(xmin = min(x), xmax = max(x))) %>%
-  mutate(ymin = -Inf, ymax = Inf, group = row_number())
+  mutate(ymin = -Inf, ymax = Inf, type = types[2])
+tile_window <- tile_window[c(rep(1:2, each = 5), 3:4), ] %>% 
+  mutate(group = row_number())
+tile_mean <- tibble(
[…]

### 2019-01-29 main.Rmd
adjust table pos
rm jel codes

@@
 wpnumber: no/yr
-jelcodes: C10,C14,C22
 blind: false
@@
 
-Each verb focuses on one operation to achieve its goal.
-The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
-These general-purpose verbs are made available in the **tidyverse** suite.
-When used in the temporal context, these verbs are adapted to the time domain.
-A perceivable difference is summarizing variables between data frame and tsibble.
-The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
-New tsibble-specific verbs (\autoref{tab:verb-table}) are provided to expand the **tidyverse** vocabulary.
-Users who are already familiar with **tidyverse**, will experience a gentle learning curve for mastering tsibble verbs and glide into temporal data analysis with low cognitive load.
-
-\begin{table}[!hbt]
+\begin{table}[!ht]
 \caption{A list of table verbs working with tsibble. Functions in bold originating from the tidyverse and are adapted for tsibble.}\label{tab:verb-table}
@@
 \end{table}
+
+Each verb focuses on one operation to achieve its goal.
+The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
+These general-purpose verbs are made available in the **tidyverse** suite.
+When used in the temporal context, these verbs are adapted to the time domain.
+A perceivable difference is summarizing variables between data frame and tsibble.
+The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
+New tsibble-specific verbs (\autoref{tab:verb-table}) are provided to expand the **tidyverse** vocabulary.
+Users who are already familiar with **tidyverse**, will experience a gentle learning curve for mastering tsibble verbs and glide into temporal data analysis with low cognitive load.
 

### 2019-01-31 main.Rmd
fixed some sentences

@@
 When used in the temporal context, these verbs are adapted to the time domain.
-A perceivable difference is summarizing variables between data frame and tsibble.
+A perceivable difference is summarizing variables between data frame and tsibble using `summarize()`.
 The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
@@
 This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems.
-Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr].
+Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr], read as "then".
 A sequence of functions can be composed in a way that can be naturally read from left to right, which improves the readability of the code.
@@
 \autoref{fig:animate} shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
-A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series.
+A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series by computing average counts.
 
@@
 
-(ref:animate-cap) An illustration of window of size 5 rolling over annual tuberculosis cases in Australia. (Animation needs to be viewed with Adobe Acrobat Reader.)
+(ref:animate-cap) An illustration of window of size 5 computing rolling averages over annual tuberculosis cases in Australia with respect to sliding, tiling and stretching.
+(Animation needs to be viewed with Adobe Acrobat Reader.)
 
@@
 The flow chart (\autoref{fig:flight-pipeline1}) demonstrates the operations undertaking in the data pipeline.
-The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval.
+The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval (as each carrier ends up with only one annual summary).
 The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display.
@@
 Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
-Customers' demographic data provides explanatory variables other than time in a different data table.
-Two data tables might be joined to explore different sources that contribute to daily electricity use when needed.
+Daily high and low temperatures in Newcastle provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang].
+Two data tables might be joined to explore how local weather can contribute to the variations of daily electricity use when needed.
 
@@
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang"), file = "rpkgs.bib")
 ```

### 2019-02-03 main.Rmd
re-work section 4

@@
 
-The tsibble structure and operations support data pipelines for sequencing analysis.
 There has been a long history of pipeline discussions and implementation centering around data in various aspects.
 A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualizing the process as it might be applied to a variety of problems.
-The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can be considered a data pipeline.
-Building a data pipeline can be technically difficult to make it sufficiently general for a variety of data, with many implementation decisions to be made about the interface, input and output objects and functionality.
-It is useful to articulate the data pipeline induced by new data tools.
-
 @unix coined the term "pipelines" in software development while developing Unix at Bell Labs.
 In Unix-based computer operating systems, a pipeline chains together a series of operations on the basis of their standard streams, so that the output of each program becomes the input to another.
-This shapes the Unix toolbox philosophy: "each do one simple thing, do it well, and most importantly, work well with each other" [@unix-philosophy].
-
+The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can also be considered a data pipeline.
 @viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot.
@@
 
+Building a data pipeline can be technically difficult to make it sufficiently general for a variety of data, with many implementation decisions to be made about the interface, input and output objects and functionality.
 The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modeling.
@@
 What is notable about an effective implementation of a data pipeline is that it coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies.
-A fluent and fluid pipeline glues tidy data and the grammar of data manipulation together.
+A fluent pipeline glues tidy data and the grammar of data manipulation together.
 It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
+New data tools developed in the R package **tsibble** [@R-tsibble] articulate the time series data pipeline, which shepherds raw temporal data through to time series analysis, and plots.
+More detailed explanations are given as follows.
 
-## Time-based transformations {#sec:timepipeline}
+## Time series transformation {#sec:timepipeline}
 
-The time-based pipeline shepherds raw temporal data through to time series analysis, and plots.
-It is possible and recommended to check for identical entries of key and index before analysis.
+\autoref{fig:tsibble-pipeline} illustrates the distinction of a time series pipeline from a regular data pipeline.
+It is highly recommended to check for identical entries of key and index before constructing a tsibble.
 Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making.
 Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
-When the data meets the tsibble standard, it flows neatly into the analysis stage.
-\autoref{fig:tsibble-pipeline} illustrates the time-based pipeline.
+When the data meets the tsibble standard, it flows neatly into the analysis stage and takes full advantage of the tsibble infrastructure.
 
@@
 
-A suite of verbs are used to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, and transformed for visualization.
-Many time operations like lag/lead and time series models, assume an intact vector input ordered in time.
-To prevent inviting these errors into the analysis, it is a good practice to check and inspect any time gaps of the tsibble in the first place.
-Several tools are provided to understand and tackle implicit missing values in: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reveals all implicit missing observations; (3) `count_gaps()` summarizes the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
-
-A tsibble is an object, conceptually considered as a noun, and hence an action performed on the object can be phrased as a verb.
-The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood.
-Attention has been paid to error handling.
-If a tsibble cannot be maintained in the output of a pipeline module, for example the index is dropped by aggregation, an error informs users of the problem and suggests alternatives.
-This avoids surprising users and reminds them of the time context.
+Many time operations such as lag/lead and time series models, assume an intact vector input ordered in time.
+Since a tsibble permits time gaps in the index, it is a good practice to check and inspect any gaps in time following the creation of a tsibble, in order to prevent inviting these avoidable errors into the analysis.
+The first suite of verbs (rephrasing actions performed on the object) are provided to understand and tackle implicit missing values: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reveals all implicit missing observations; (3) `count_gaps()` summarizes the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
+To look into gaps over individual time periods or full-length time span, the common argument `.full` in these functions gives an option to easily switch between.
+The specification of `.full = TRUE` will result in fully balanced panels in other words.
 
@@
 
-Each verb focuses on one operation to achieve its goal.
-The verb should be self-explanatory to advise what it is supposed to do or fail, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
-These general-purpose verbs are made available in the **tidyverse** suite.
-When used in the temporal context, these verbs are adapted to the time domain.
+Besides the time gaps verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in \autoref{tab:verb-table}.
+The **tidyverse** suite showcases the general-purpose verbs for effectively manipulating tabular data, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
+But these verbs need handling with care when used in the time series domain.
 A perceivable difference is summarizing variables between data frame and tsibble using `summarize()`.
 The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
-New tsibble-specific verbs (\autoref{tab:verb-table}) are provided to expand the **tidyverse** vocabulary.
-Users who are already familiar with **tidyverse**, will experience a gentle learning curve for mastering tsibble verbs and glide into temporal data analysis with low cognitive load.
+Users who are already familiar with the **tidyverse**, will experience a gentle learning curve for mastering these verbs and glide into time series analysis with low cognitive load.
 
+Attention has been paid to warning/error handling.
+The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble by automatically updating index and key under the hood.
+If the desirable temporal ordering is changed by row-wise verbs, a warning is broadcast.
+If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is removed by selection, an error informs users of the problem and suggests alternatives.
+This avoids surprising users and reminds them of the time context.
+
+The tsibble structure and operations support data pipelines for sequencing analysis.
 @eopl asserted "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony.
@@
 It consequently generates a block of code without saving intermediate values.
+
 Most importantly, a new ecosystem for tidy time series analysis has been undertaken, using the tsibble framework, and is called "tidyverts", a play on tidyverse that acknowledges the time series analysis purpose.
@@
 
-As a special case of the data frame, a tsibble pipes directly into the grammar of graphics.
-It is easy to create and extend specialist time-related plotting methods based on tsibble structure, such as autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
+As a tsibble is a subclassing of data frame, it integrates well with the grammar of graphics.
+It is easy to create and extend specialist time series plotting methods based on the tsibble structure, for example autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
 
@@
 The primary force that drives the software's design choices is "data".
-All functions in the package **tsibble** [@R-tsibble] start with `data` or its variants as the first argument, namely "data first".
+All functions in the package **tsibble** start with `data` or its variants as the first argument, namely "data first".
 This lays out a consistent interface and addresses the significance of the data throughout the software.
@@
 It provides a complete and consistent set of tools to iterate each element of a vector with a function.
-Rolling window doesn't just iterate but rolls over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
+Rolling window does not just iterate but rolls over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
 `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.

### 2019-02-03 R/smart-meter.R

@@ -21,14 +21,8 @@ elec_ts <- elec %>%
 gap_df <- has_gaps(elec_ts)
 # sum(gap_df$.gaps) / NROW(gap_df)
 
-## ---- scan-gaps
-elec_gaps <- scan_gaps(elec_ts)
-
 ## ---- count-gaps
-customer_na <- elec_ts %>% 
-  filter(customer_id %in% (gap_df %>% filter(.gaps) %>% pull(customer_id)))
-
-count_na_df <- customer_na %>% 
+count_na_df <- elec_ts %>% 
   count_gaps()
 
 count_na_df %>% 

### 2019-02-03 main.Rmd
tweak section 5

@@
 This diversity and time zone is acknowledged and accommodated by tsibble's index.
+Whist creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering.
 
@@
 The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble by automatically updating index and key under the hood.
-If the desirable temporal ordering is changed by row-wise verbs, a warning is broadcast.
+If the desired temporal ordering is changed by row-wise verbs, a warning is broadcast.
 If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is removed by selection, an error informs users of the problem and suggests alternatives.
@@
 Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-The contextual summary provided by the print function, shown below on the data from \autoref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-time, (2) variables that constitute the "key" with the number of series.
+The contextual summary provided by the print function, shown below on the data from \autoref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of series.
 These details aid users in understanding their data better.
@@
 It provides a complete and consistent set of tools to iterate each element of a vector with a function.
-Rolling window does not just iterate but rolls over a sequence of elements, with `slide()`, `tile()` and `stretch()`.
+Rolling window does not just iterate but rolls over a sequence of elements, namely `slide()`, `tile()` and `stretch()`.
 `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
@@
 Vector functions in the package mostly deal with time.
-When collapsing a tsibble to a less granular interval, these atomic functions can be combined with the `index_by()` table verb.
-A different function results in easily switching to aggregation of different time resolution.
-Since these functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages.
-On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of current tsibble zone.
-It is generally easier to trace back the errors users encounter from separating 1d and 2d functions.
+The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb achieves in collapsing a tsibble to a less granular interval.
+Replacing with a different time function results in the aggregation of different time resolution.
+Since these time functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages.
+On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone.
+It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions.
 
@@
 As a fundamental infrastructure, extensibility is a design decision that was employed from the start of **tsibble**'s development.
-Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily rely on S3 classes and methods in R [@adv-r].
+Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily relies on S3 classes and methods in R [@adv-r].
 The package can be extended in two major aspects: custom index and new tsibble class.
@@
 The interval regarding semesters is further outlined through `pull_interval()`.
-As a result, the rest of the software methods such as `has_gaps()`, `count_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
+As a result, the rest of the software methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
 
@@
 Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
-Daily high and low temperatures in Newcastle provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang].
+Daily high and low temperatures in Newcastle in 2013 provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang].
 Two data tables might be joined to explore how local weather can contribute to the variations of daily electricity use when needed.

### 2019-02-03 main.Rmd
added subsetion on tidy eval

@@
 This diversity and time zone is acknowledged and accommodated by tsibble's index.
-Whist creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering.
+Whist creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering that is assumed by time series operations.
 
@@
 The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb achieves in collapsing a tsibble to a less granular interval.
-Replacing with a different time function results in the aggregation of different time resolution.
+The substitution of another time function in the `index_by()` results in the aggregation of different time resolution.
 Since these time functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages.
 On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone.
-It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions.
+It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the readability.
 
@@
 Tsibble verbs are also S3 generics.
-Developers will be able to implement these verbs for the new class if needed.
+Developers will be able to implement these verbs for the new class if necessary.
+
+## Tidy evaluation
+
+The **tsibble** packages leverages the **tidyverse** grammars and pipelines through tidy evaluation [@tidy-eval] via the **rlang** package [@R-rlang].
+In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlights the "tidy" interface that is compatible with the **tidyverse**.
+This not only saves a few keystrokes without explicitly repeated references to the data source, but the resulting code is typically cleaner and more expressive.
 
@@
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang", "rlang"), file = "rpkgs.bib")
 ```

### 2019-02-04 main.Rmd
prune a couple of sentences

@@
 This diversity and time zone is acknowledged and accommodated by tsibble's index.
-Whist creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering that is assumed by time series operations.
+When creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering that is assumed by time series operations.
 
@@
 
-There has been a long history of pipeline discussions and implementation centering around data in various aspects.
 A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualizing the process as it might be applied to a variety of problems.
@@
 
-Building a data pipeline can be technically difficult to make it sufficiently general for a variety of data, with many implementation decisions to be made about the interface, input and output objects and functionality.
-The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modeling.
+Building a data pipeline is technically difficult: many implementation decisions have to be made about the interface, input and output objects and functionality.
+The tidy data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling.
 Each module communicates with the others, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.

### 2019-02-04 main.Rmd
small changes sect 4-5

@@
 
-A data pipeline describes the general flow of data through an analysis, and can generally assist in conceptualizing the process as it might be applied to a variety of problems.
+A data pipeline describes the flow of data through an analysis, and can generally assist in conceptualizing the process, when it is applied to a variety of problems.
 @unix coined the term "pipelines" in software development while developing Unix at Bell Labs.
@@
 New data tools developed in the R package **tsibble** [@R-tsibble] articulate the time series data pipeline, which shepherds raw temporal data through to time series analysis, and plots.
-More detailed explanations are given as follows.
+More detailed explanations are given in the following sections, and the examples.
 
@@
 It is highly recommended to check for identical entries of key and index before constructing a tsibble.
-Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making.
+Duplicates signal the data quality issue, which would likely affect subsequent analyses and hence decision making.
 Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
@@
 Many time operations such as lag/lead and time series models, assume an intact vector input ordered in time.
-Since a tsibble permits time gaps in the index, it is a good practice to check and inspect any gaps in time following the creation of a tsibble, in order to prevent inviting these avoidable errors into the analysis.
+Since a tsibble permits time gaps in the index, it is good practice to check and inspect any gaps in time following the creation of a tsibble, in order to prevent inviting these avoidable errors into the analysis.
 The first suite of verbs (rephrasing actions performed on the object) are provided to understand and tackle implicit missing values: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reveals all implicit missing observations; (3) `count_gaps()` summarizes the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
@@
 
-Besides the time gaps verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in \autoref{tab:verb-table}.
+Besides the time gap verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in \autoref{tab:verb-table}.
 The **tidyverse** suite showcases the general-purpose verbs for effectively manipulating tabular data, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
@@
 
-Attention has been paid to warning/error handling.
+Attention has been paid to warning and error handling.
 The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble by automatically updating index and key under the hood.
@@
 
-Rolling window calculations are widely used techniques in time series analysis, but are not restricted to temporal applications.
+Rolling window calculations are widely used techniques in time series analysis, and often apply to other applications.
 These operations are dependent on having an ordering, particularly time ordering for temporal data.
@@
 Tsibble verbs are also S3 generics.
-Developers will be able to implement these verbs for the new class if necessary.
+Developers will be able to implement these verbs for the new class, if necessary.
 

### 2019-02-04 R/smart-meter.R
recompile to fix missing references

@@ -3,7 +3,6 @@ library(tidyverse)
 library(lubridate)
 library(tsibble)
 
-customer <- read_rds("data/customer.rds")
 elec <- read_rds("data/smart-meter13.rds")
 weather <- read_rds("data/weather13.rds") %>% 
   group_by(date) %>% 

### 2019-02-05 R/demo.R
draw a window instead of vline in animation

@@ -77,12 +77,12 @@ revealed <- bind_rows(slide_revealed, tile_revealed, stretch_revealed)
 
 ggplot() +
   geom_line(aes(x = year, y = count), data = tb_au, colour = "grey", size = 1.2) +
-  geom_vline(aes(xintercept = xmin), data = window, colour = "#9ecae1", alpha = 0.6, size = 1.5) +
   geom_rect(aes(
     xmin = xmin, xmax = xmax,
     ymin = ymin, ymax = ymax,
     group = group
-  ), data = window, fill = "#9ecae1", alpha = 0.6) +
+  ), data = window, 
+  fill = "#9ecae1", colour = "#9ecae1", size = 1.5, alpha = 0.6) +
   geom_point(aes(x = year, y = ma), data = mean, size = 2, colour = "#de2d26") +
   geom_line(aes(x = year, y = ma), data = revealed, size = 1.2, colour = "#de2d26") +
   xlab("Year") +

### 2019-02-09 cover-letter/cover-letter.Rmd
added cover letter (can you pls take a look @dicook ?)

---
from: Earo Wang
fromqualifications: B.Comm.(Hons)
fromtitle: PhD candidate
fromWWW: earo.me
fromemail: earo.wang@monash.edu
fromphone: +61 4 4919 1031
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data," for publication in the Journal of Computational and Graphical Statistics.

This paper proposes a new data abstraction to represent temporal data, referred to as "tsibble", allowing the "tidy data" principles to be brought to the time domain. It helps to reduce the pain from going raw data to model-ready time series data. The tsibble data structure lays the plumbing for time series data analysis modules of transformation, visualization and modeling, to achieve rapid iteration in gaining data insights. The infrastructure of tidy temporal data and tools has been implemented in the R package `tsibble`, which is on CRAN. The paper is written in RMarkdown, which contains the R code to reproduce the results.

The `tsibble` package won the 2019 John Chambers Statistical Software Award. It also lays the foundation of a new ecosystem for tidy time series analysis, known as "tidyverts", including data sets (the `tsibbledata` package), time series statistical methods (the `feasts` package), and forecasting (the `fable` package).

We believe that this manuscript is a good fit for the Journal of Computational and Graphical Statistics, because it provides tidy methodology to make time series analysis easier. Because the software has already received considerable attention, we expect that the work will be of interest to readers of the journal.

### 2019-02-13 main.Rmd
new arxiv replacement

@@
 toc: false
-bibliography: [references.bib, rpkgs.bib]
+bibliography: [references.bib]
 biblio-style: authoryear-comp
@@
 On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone.
-It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the readability.
+It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the code readability.
 

### 2019-05-15 main.Rmd
renamed R/ to scripts/

@@
 )
-read_chunk('R/demo.R')
-read_chunk('R/flights.R')
-read_chunk('R/smart-meter.R')
+read_chunk('scripts/demo.R')
+read_chunk('scripts/flights.R')
+read_chunk('scripts/smart-meter.R')
 ```
@@
 The software internally validates if the key and index produce distinct rows, and then sorts the key and the index from past to recent.
-When the tsibble creation is done, the print display is data-oriented and contextually informative, including dimensions, irregular interval (``r tsibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tsibble::tbl_sum(us_flights)[2]``).
+When the tsibble creation is done, the print display is data-oriented and contextually informative, including dimensions, irregular interval (``r tibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tibble::tbl_sum(us_flights)[2]``).
 

### 2019-05-15 main.Rmd
remove \autoref

@@
     citation_package: biblatex
+    dev: "pdf"
 ---
@@
 ```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+options("knitr.graphics.auto_pdf" = TRUE)
 library(knitr)
 opts_chunk$set(
-  warning = FALSE, message = FALSE, echo = FALSE,
-  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center',
-  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
-  fig.height = 5, fig.width = 8, out.width = "\\textwidth",
-  comment = "#>"
+  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
+  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
+  cache = FALSE, cache.path = 'cache/', 
+  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
 )
+knitr::opts_knit$set(root.dir = here::here())
+```
+
+```{r external, include = FALSE}
 read_chunk('scripts/demo.R')
@@
 
-<!--Suppose that temporal data consists of $N$ observational units.
-Given observation $i$ $(i = 1, 2, \dots, N)$, $X_{ijt}$ denotes the $j$th measurement $(j = 1, 2, \dots, p)$ indexed at a unique sequence of time points $0 \le t \le T_{i}$.
--->
 Temporal data arrives in many possible formats, with many different time contexts.
@@
 
-@r4ds proposed the tidy data workflow, which provides a conceptual 
-framework for processing data (as described in \autoref{fig:tsibble-fit}).
+@r4ds proposed the tidy data workflow, which provides a conceptual framework for processing data (as described in Figure \@ref(fig:tsibble-fit)).
 Currently, time series modeling and forecasting enters this framework at the *modeling* stage, while temporal data enters at the start.
@@
 ```{r tsibble-fit, fig.cap = "(ref:tsibble-fit-cap)"}
-include_graphics("img/tsibble-fit.pdf")
+include_graphics("img/tsibble-fit.png")
 ```
@@
 The paper is structured as follows.
-\autoref{sec:str} reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation.
-\autoref{sec:semantics} proposes contextual semantics for temporal data, built on top of tidy data.
-The concept of data pipelines with respect to the time domain will be discussed in depth in \autoref{sec:pipeline}, followed by a discussion of the design choices made in the software structure in \autoref{sec:software}.
-Two case studies are presented in \autoref{sec:cases} illustrating temporal data exploration using the newly implemented infrastructure.
-\autoref{sec:conclusion} discusses future work.
+Section \@ref(sec:str) reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation.
+Section \@ref(sec:semantics) proposes contextual semantics for temporal data, built on top of tidy data.
+The concept of data pipelines with respect to the time domain will be discussed in depth in Section \@ref(sec:pipeline), followed by a discussion of the design choices made in the software structure in Section \@ref(sec:software).
+Two case studies are presented in Section \@ref(sec:cases) illustrating temporal data exploration using the newly implemented infrastructure.
+Section \@ref(sec:conclusion) discusses future work.
 
@@
 These objects were designed for modeling methods.
-The diagram in the style of \autoref{fig:tsibble-fit} would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
+The diagram in the style of Figure \@ref(fig:tsibble-fit) would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
 This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.
@@
 It only requires an index variable to declare a temporal data object, thus placing it at the import stage.
-However, as proposed in \autoref{sec:semantics} a more rigid data structure is required for time series analytics and models.
+However, as proposed in Section \@ref(sec:semantics) a more rigid data structure is required for time series analytics and models.
 
@@
 The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
-\autoref{fig:tsibble-fit} marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble".
-The tsibble structure is an extension of a data frame---a two-dimensional array in R---with additional time series semantics: index and key, as shown in \autoref{fig:tsibble-semantics}.
+Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble".
+The tsibble structure is an extension of a data frame---a two-dimensional array in R---with additional time series semantics: index and key, as shown in Figure \@ref(fig:tsibble-semantics).
 
 ```{r tsibble-semantics, fig.cap = "(ref:tsibble-semantics-cap)"}
-include_graphics("img/tsibble-semantics.pdf")
+include_graphics("img/tsibble-semantics.png")
 ```
@@
 
-To demonstrate the concept of the tsibble, \autoref{tab:tb-sub} presents a subset of tuberculosis cases estimated by @tb-data.
+To demonstrate the concept of the tsibble, Table \@ref(tab:tb-sub) presents a subset of tuberculosis cases estimated by @tb-data.
 It contains 12 observations and 5 variables arranged in a "long" tabular form.
@@
 The "key" is usually known a priori by analysts.
-For example, \autoref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year.
+For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year.
 This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data.
@@
 This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals.
-More details are given  in \autoref{sec:ts-models}.
+More details are given  in Section \@ref(sec:ts-models).
 
@@
 
-\autoref{fig:tsibble-pipeline} illustrates the distinction of a time series pipeline from a regular data pipeline.
+Figure \@ref(fig:tsibble-pipeline) illustrates the distinction of a time series pipeline from a regular data pipeline.
 It is highly recommended to check for identical entries of key and index before constructing a tsibble.
@@
 ```{r tsibble-pipeline, fig.cap = "(ref:tsibble-pipeline-cap)"}
-include_graphics("img/tsibble-pipeline.pdf")
+include_graphics("img/tsibble-pipeline.png")
 ```
@@
 
-Besides the time gap verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in \autoref{tab:verb-table}.
+Besides the time gap verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in Table \@ref(tab:verb-table).
 The **tidyverse** suite showcases the general-purpose verbs for effectively manipulating tabular data, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
@@
 Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data.
-The verbs listed in \autoref{tab:verb-table} ease the transition to a tsibble that suits modeling.
+The verbs listed in Table \@ref(tab:verb-table) ease the transition to a tsibble that suits modeling.
 A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation.
@@
 Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-The contextual summary provided by the print function, shown below on the data from \autoref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of series.
+The contextual summary provided by the print function, shown below on the data from Table \@ref(tab:tb-sub), contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of series.
 These details aid users in understanding their data better.
@@
 
-\autoref{fig:animate} shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
+Figure \@ref(fig:animate) shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
 A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series by computing average counts.
@@
 
-Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in \autoref{tab:verb-table}.
+Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in Table \@ref(tab:verb-table).
 Table verbs expect and return a tsibble, and does what the function name suggests.
@@
 A new index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future.
-The interval regarding semesters is further outlined through `pull_interval()`.
+The interval regarding semesters is further outlined through `interval_pull()`.
 As a result, the rest of the software methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
@@
 The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble.
-For example, a fable is actually an extension of a tsibble, mentioned in \autoref{sec:ts-models}.
+For example, a fable is actually an extension of a tsibble, mentioned in Section \@ref(sec:ts-models).
 A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass.
@@
 
-(ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:carrier-mosaic}.
+(ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:carrier-mosaic).
 
-\autoref{fig:carrier-mosaic} visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
+Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
 This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually.
 To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column.
-The flow chart (\autoref{fig:flight-pipeline1}) demonstrates the operations undertaking in the data pipeline.
+The flow chart (Figure \@ref(fig:flight-pipeline1)) demonstrates the operations undertaking in the data pipeline.
 The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval (as each carrier ends up with only one annual summary).
@@
 
-(ref:flight-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:sel-monthly-plot}.
+(ref:flight-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:sel-monthly-plot).
 
 What time of day and day of week should we travel to avoid suffering from horrible delay?
-\autoref{fig:draw-qtl} plots hourly quantile estimates across day of week in the form of small multiples.
+Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples.
 The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown.
 To reduce the likelihood of suffering a delay, it is recommended to avoid the peak hour around 6pm (18).
-<!--As moving towards the upper extremes, the variations considerably increase, making departure time unpredictable.-->
 
@@
 
-(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating \autoref{fig:draw-qtl}.
+(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating Figure \@ref(fig:draw-qtl).
 
@@
 Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for each household.
-\autoref{fig:count-gaps} shows the gaps for the top 49 households arranged in rows from high to low in tallies.
+Figure \@ref(fig:count-gaps) shows the gaps for the top 49 households arranged in rows from high to low in tallies.
 (The remaining households values have been aggregated into a single batch and appear at the top.) Missing values can be seen to occur at any time during the entire span.
@@
 
-Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (\autoref{fig:calendar-plot}).
+Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)).
 Each glyph represents the daily pattern of average residential electricity usage every thirty minutes.
@@
 
-(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:calendar-plot}.
+(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:calendar-plot).
 
@@
 The project's Github repository <https://github.com/earowang/paper-tsibble> houses all materials required to reproduce this article and a history of the changes.
-<!-- To produce the analysis results as described in Section \autoref{sec:cases}, we require the **tsibble** package (`v0.6.2`) and the **dplyr** package (`v0.7.8`) installed.
--->
 

### 2019-05-15 main.Rmd
minor tweak to reflect tsibble's API change

@@
 The tsibble framework, as expected, incorporates this type of data without any loss of data richness and heterogeneity.
-To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as the "index", and column `flight_num` as the "key" via `id(flight_num)`.
+To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as the "index", and column `flight_num` as the "key".
 As a result of event timing, the data are irregularly spaced, and hence switching to the irregular option is necessary.

### 2019-05-17 main.Rmd
updated gitignore

@@
 keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
-wpnumber: no/yr
+wpnumber: 12/19
+jelcodes: C88, C81, C82, C22, C32
 blind: false

### 2019-07-21 tsibble.Rmd
added reviews

--- 2019-05-17 main.Rmd
@@
 
-```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
 options("knitr.graphics.auto_pdf" = TRUE)
@@
 )
-knitr::opts_knit$set(root.dir = here::here())
-```
-
-```{r external, include = FALSE}
 read_chunk('scripts/demo.R')

### 2019-07-31 reviews/response.Rmd
added reviewer 2 comments

---
title: Response to reviewers
author: JCGS-19-044
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the reviewers for their helpful feedback which has led to many improvements in the paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

# Reviewer 1 comments {-}

# Reviewer 2 comments {-}

> P5L37 "Temporal data tools need to catch up." This is a little casual. Could perhaps be rephrased as "This paper and the associated **tsibble** R package extends the tidy way of thinking to temporal data."

Done.

> P7L44 "A variable number of time representations can be spotted in the wild." This is a little casual, and I’m not fond of "variable number". Perhaps "Many different representations of temporal data can be spotted in the wild." (The "spotted in the wild" is still a little informal – but I like it.)

Done.

> P7L50 "This diversity and time zone..." suggest rephrase "The tsibble's index column leverages the date-time data type in R to accommodate diverse temporal representations and associated time zone information."

TODO.

> P8L16 "The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks." It's not totally clear to me what is meant here. Is it saying that like with grouping columns (i.e. columns supplied to group_by()), key columns cannot be altered in subsequent analyses? It might be useful to make the comparison to group_by() explicit as it also allows multiple fields. The distinction between group_by() and key() is clear in that the collection of columns that constitute a key must uniquely identify observations (obviously this isn’t the case for grouping variables).

TODO.

> P8L55 "fit models that expect **regular** time series." Should regular be added to the sentence.

Done.

> P10L19 "Duplicates signal a data quality..."

Done.

> P12-13 Consider referencing "fable: and possibly "mable" too.

TODO.

> P14-15 I had to read section 5.2 a few times before I understood it... I think what threw me was the last paragraph starts with a reference to purrr, but the functions described in the remainder of the paragraph describes functions in the tsibble package. Perhaps move the purrr sentence up to the previous paragraph when functional programming is first mentioned.

TODO.

> P15L46 "...table verb achieves in collapsing a tsibble..." perhaps "...table verb collapse a tsibble..."

Done.

> P15L53 "On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone" perhaps "The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions."

Done.

> Fig 5 y-axis doesn't need the overall label "Delayed" or if it stays, should be renamed “Status”. Similarly for the legend label.

TODO.

> Fig 6 Consider replacing gather() with pivot_longer() which is intended to be easier to use particularly for more complex "gathering" operations. https://tidyr.tidyverse.org/dev/articles/pivot.html

TODO.

> Fig 7 y-axis label should be "Proportion of flights delayed" or similar. Legend label "Originating airport" (but legend can also be removed, because it is covered by the facet labels.

TODO.

> Fig 9 y-label "Departure delay (mins)". X-axis could be converted to time... e.g. ggplot2::scale_x_time() Could introduce some alpha transparency to the geom_line(). Caption text currently states "Passengers are apt to hold up around 18 during a day, and are recommended to travel early. The variations increase substantially as the upper tails." Both sentences need to be reworded.

TODO.

> P22L3 "might be joined" should it be "can be joined" or perhaps a bit more detail, "Aggregating the half-hourly electricity data to the same daily time interval as the temperature data allows us to join the two data tables to explore..."

TODO.

> P22L27 Suggest a reference to Wang, E, D Cook & RJ Hyndman (2018) to explicitly acknowledge that the frame_calendar() function is part of a different package (sugrrants).

TODO.

> P22L37 "degrees C" to "degrees Celsius" or °C

Done.

> Fig 12 remove legend label "hot"

TODO.

> P25L19 It's not "A missing piece of the tsibble package" because later in the paragraph it says that custom calendars are easily embedded in the tsibble framework. I suggest rephrasing the first sentence in a more positive light... "The tsibble package can easily be extended to accommodate user-defined calendars." Suggest also referencing at least one package that provides a user-defined calendar (perhaps a market-specific business calendar).

TODO.

> P26L54 there’s an errant

Fixed.

> P28L22 Journal of Statistical Software

Fixed.

### 2019-07-31 tsibble.Rmd
added reviewer 2 comments

@@
 Currently, time series modeling and forecasting enters this framework at the *modeling* stage, while temporal data enters at the start.
-This paper integrates
-time series analysis into this tidy framework, providing a coherent way for getting temporal data into the matrix format for modeling.
+This paper integrates time series analysis into this tidy framework, providing a coherent way for getting temporal data into the matrix format for modeling.
 
@@
 Many in the R community have adopted the **tidyverse** way of thinking and extended it to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
-Temporal data tools need to catch up.
+This paper and the associated **tsibble** R package extends the tidy way of thinking to temporal data.
 
@@
 
-A variable number of time representations can be spotted in the wild.
+Many different representations of time can be spotted in the wild.
 A date-time object, universally accepted across computing systems, is the most commonly used type for representing time.
@@
 It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals.
-This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect time series.
+This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect regular time series.
 
@@
 It is highly recommended to check for identical entries of key and index before constructing a tsibble.
-Duplicates signal the data quality issue, which would likely affect subsequent analyses and hence decision making.
+Duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making.
 Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
@@
 Vector functions in the package mostly deal with time.
-The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb achieves in collapsing a tsibble to a less granular interval.
+The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb collapses a tsibble to a less granular interval.
 The substitution of another time function in the `index_by()` results in the aggregation of different time resolution.
 Since these time functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages.
-On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone.
+The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions.
 It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the code readability.
@@
 The daily snapshots vary depending on the season in the year.
-During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees C.
+During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees Celsius.
 However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.

### 2019-08-01 reviews/response.Rmd
added reviewer 1 comments

@@
 # Reviewer 1 comments {-}
+
+> More mini examples embeded throughout the manuscript.
+Mainly it didn't show- case to me tsibble's capability well enough but understandably this may make the manuscript long and may be better left to vignettes.
+
+TODO.
+
+> Finding a duplicate is a nice touch but I don't understand the significance of the Figure 5 mosaic plot in demonstration for the capability of tsibble.
+I understand that finding duplicate/wrong entry is made easier by tsibble but besides filtering those duplicate/wrong entries, the data wrangling for mosaic plot is easily done without tsibble.
+More specifically what is the significance `tsibble::index_by` below?
+Could the authors explain what is the benefit of explicit defining index at this stage as opposed to explicit specification at a later stage?
+
+TODO.
+
+> In relation to the above point, perhaps it would be helpful to have a definition of tidy temporal data much like the three rules in tidy data (every row is an observation etc..).
+E.g.
+in tsibble the semantics index is a variable with some inherent ordering from past to present.
+
+TODO.
+
+> I hold a similar sentiment regarding the calendar plot in Figure 12. It's a great plot but I don't necessary think this is a plot that show cased the benefit of tsibble.
+
+TODO.
+
+> Figure 11 with missing data was a great demonstration for tsibble and the data times series pipeline shown in Figure 3.
+It may be helpful to reduce the opacity of the points as there is a lot of overplotting.
+The top line which lumps a number of customers together is not particularly informative here.
+Perhaps it is better replaced with a histogram.
+
+TODO.
+
+> Authors may like to add/change `spread`/`gather` verbs to `pivot_longer`/`pivot_wider` with the recent move to deprecate the former in tidyr.
+Likewise au- thors may like to consider to replace “long data” to “longer data” and “wide data” to “wider data” as long/wide implies absolute terms.
+https: //tidyr.tidyverse.org/dev/articles/pivot.html
+
+TODO.
+
+> Manuscript does implicitly feel more time series data oriented rather than a general temporal data.
+E.g.
+references of time points as index seems inherent from other time series data infrastructure (i.e.
+zoo and xts) and the presented case studies is not representative of any longitudinal study.
+This does not mean it should be changed but index to me was not as intuitive to be the temporal variable.
+
+TODO.
+
+> Figure 4 animation doesn't play so authors may like to ensure that the final publication works. Github version works.
+
+The JCGS peer-reviewed version is watermarked, which prevents the animation playing.
+
+> Have the authors thought about how to deal with mixed time resolutions for index?
+E.g.
+some subjects have exact day recorded; some missing day but month & year recorded; some only date range recorded.
+
+TODO.
+
+> The emphasis of the manuscripts is in R but do other languages also tend to define time index implicitly?
+
+TODO.
+
+Thanks for the careful reading. We have done with minor comments.
 

### 2019-08-01 tsibble.Rmd
added reviewer 1 comments

@@
 abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types.
-On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes.
+On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numerical data with implicit time indexes.
 Going from raw data to model-ready data is painful.
@@
 
-```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
+```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
 options("knitr.graphics.auto_pdf" = TRUE)
@@
   fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
-  cache = FALSE, cache.path = 'cache/', 
+  cache = TRUE, cache.path = 'cache/', 
   out.width = ifelse(is_html_output(), "100%", "\\textwidth")
 )
+knitr::opts_knit$set(root.dir = here::here())
+```
+
+```{r external, include = FALSE, cache = FALSE}
 read_chunk('scripts/demo.R')
@@
 Temporal data arrives in many possible formats, with many different time contexts.
-For example, time can have various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time.
+For example, time can have various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as daylight saving time.
 Time can be regular (such as quarterly economic data or daily weather data), or irregular (such as patient visits to a doctor's office).
@@
 The accessibility of the tsibble index motivates data analysis towards transparency and human readability.
-When the "index" is available only as meta information (such as in the `ts` class), it creates an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytic point of view.
+When the "index" is available only as meta information (such as in the `ts` class), it creates an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytical point of view.
 
@@
 A date-time object, universally accepted across computing systems, is the most commonly used type for representing time.
-Date-time also typically associates with a time zone including adjustments such as summer time.
+Date-time also typically associates with a time zone including adjustments such as daylight saving time.
 This diversity and time zone is acknowledged and accommodated by tsibble's index.
@@
 In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlights the "tidy" interface that is compatible with the **tidyverse**.
-This not only saves a few keystrokes without explicitly repeated references to the data source, but the resulting code is typically cleaner and more expressive.
+This not only saves a few keystrokes without explicitly repeating references to the data source, but the resulting code is typically cleaner and more expressive.
 

### 2019-08-01 scripts/smart-meter.R

@@ -48,7 +48,7 @@ elec_cal <- elec_ts %>%
   summarise(avg = mean(general_supply_kwh)) %>% 
   mutate(
     date = as_date(reading_datetime), 
-    time = hms::as.hms(reading_datetime, tz = "UTC")
+    time = hms::as_hms(reading_datetime)
   ) %>% 
   left_join(weather, by = "date") %>% 
   frame_calendar(x = time, y = avg, date = date)

### 2019-08-05 reviews/response.Rmd
added tsibble definition suggested by reviewer 1

@@
 > More mini examples embeded throughout the manuscript.
-Mainly it didn't show- case to me tsibble's capability well enough but understandably this may make the manuscript long and may be better left to vignettes.
+Mainly it didn't showcase to me tsibble's capability well enough but understandably this may make the manuscript long and may be better left to vignettes.
 
@@
 
-TODO.
+Thanks for this great suggestion. We've given a definition of "tsibble" following the "tidy data" principles in the beginning of Section 3.
 
@@
 > Authors may like to add/change `spread`/`gather` verbs to `pivot_longer`/`pivot_wider` with the recent move to deprecate the former in tidyr.
-Likewise au- thors may like to consider to replace “long data” to “longer data” and “wide data” to “wider data” as long/wide implies absolute terms.
+Likewise authors may like to consider to replace “long data” to “longer data” and “wide data” to “wider data” as long/wide implies absolute terms.
 https: //tidyr.tidyverse.org/dev/articles/pivot.html

### 2019-08-05 tsibble.Rmd
added tsibble definition suggested by reviewer 1

@@
 Multivariate time series are typically assumed to be in the format where each row is assumed to hold observations at a time point and each column to contain a single time series.
-(The tidy data name for this would be **wide format**.) This implies that data are columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric variables.
+(The tidy data name for this would be **wider format**.) This implies that data are columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric variables.
 In addition, time indexes are stripped off the data and implicitly inferred as attributes or meta-information.
@@
 Thus, the primary format required for modeling such data is stacked series, blocks of measurements for each individual, with columns indicating individual, times of measurement and the measurements themselves.
-(The tidy data name for this would be **long format**.) Evidently, this data organization saves storage space for many sparse cells, compared to structuring it into wide format which would have missing values in many cells.
+(The tidy data name for this would be **longer format**.) Evidently, this data organization saves storage space for many sparse cells, compared to structuring it into wider format which would have missing values in many cells.
 A drawback of this format is that information unique to each individual is often repeated for all time points.
@@
 
-The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data.
-Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble".
-The tsibble structure is an extension of a data frame---a two-dimensional array in R---with additional time series semantics: index and key, as shown in Figure \@ref(fig:tsibble-semantics).
+The choice of tidy representation of temporal data arises from a data- and model-oriented perspective, which can accommodate all of the operations that are to be performed on the data in time-based contexts.
+Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the "tidy" model, which is referred to as a "tsibble".
+In tsibble:
+
+1. Index is a variable with inherent ordering from past to present.
+2. Key is a set of variables that define observational units over time.
+3. Each observation should be uniquely identified by index and key.
+4. Each observational unit should be measured at a common interval, if regularly spaced.
+
+Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, the same format as the "tidy data".
+Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in 1 and 2 above.
+As a result of the definitions 3 and 4, tsibble is *tidier* than "tidy data", placing itself as a data object that is also suitable for modeling.
 
@@
 
-(ref:tsibble-semantics-cap) The architecture of the tsibble structure is built on top of the data frame---a two-dimensional array in R, with time series semantics: index and key.
+(ref:tsibble-semantics-cap) The architecture of the tsibble structure is built on top of the "tidy data" principles, with temporal semantics: index and key.
 
 To demonstrate the concept of the tsibble, Table \@ref(tab:tb-sub) presents a subset of tuberculosis cases estimated by @tb-data.
-It contains 12 observations and 5 variables arranged in a "long" tabular form.
+It contains 12 observations and 5 variables arranged in a "longer" tabular form.
 Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
@@
 It is similar to a primary key [@codd_relational_1970] defining each observation in a relational database.
-In the wide format in which multiple time series are often structured, the columns hold a series of values, so that the column implicitly serves as identification.
-In long format, columns are melted with names converted to "key" values.
+In the wider format in which multiple time series are often structured, the columns hold a series of values, so that the column implicitly serves as identification.
+In longer format, columns are melted with names converted to "key" values.
 However, the "key" provides much more flexibility.
@@
 
-## Time series transformation {#sec:timepipeline}
+## Temporal transformation {#sec:timepipeline}
 
@@
  & \texttt{group\_by\_key()} & Group by key variables \\ \hline
-\multirow{4}{*}{Reshape} & \textbf{\texttt{gather()}} & Gather columns into long form \\
- & \textbf{\texttt{spread()}} & Spread columns into wide form \\
+\multirow{4}{*}{Reshape} & \textbf{\texttt{gather()}} & Gather columns into longer form \\
+ & \textbf{\texttt{spread()}} & Spread columns into wider form \\
  & \textbf{\texttt{nest()}} & Nest values in a list-variable \\

### 2019-08-11 tsibble.Rmd
done with revision for subsection on index

@@
 
-Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, the same format as the "tidy data".
-Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in 1 and 2 above.
-As a result of the definitions 3 and 4, tsibble is *tidier* than "tidy data", placing itself as a data object that is also suitable for modeling.
+Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an equivalent of the "tidy" format.
+Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the first and second definitions above.
+Variables other than "index" and "key" are considered as measurements.
+The third and fourth definitions imply that tsibble is *tidier* than "tidy data", placing itself as a model input that gives rise to more robust and reliable downstream analytics.
 
@@
 
-To demonstrate the concept of the tsibble, Table \@ref(tab:tb-sub) presents a subset of tuberculosis cases estimated by @tb-data.
-It contains 12 observations and 5 variables arranged in a "longer" tabular form.
+To materialize the abstraction of the tsibble, a subset of tuberculosis cases estimated by @tb-data presented in Table \@ref(tab:tb-sub) is used as an example.
+It contains 12 observations and 5 variables landing in a tidy data form.
 Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
-To turn this data into a tsibble: (1) column `year` is declared as the index variable; (2) the key is specified to consist of columns `country` and `gender`.
-The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables; for example, adding the corresponding population size (if known) in order to normalize the count later.
+From tidy data to tsibble data, "index" and "key" should be declared: column `year` as the "index" variable, and column `country` together with `gender` as the "key" variable forming the observational units.
+Column `count` is the only measured variable in this data, but the data structure is sufficiently flexible to hold more measurements; for example, slotting the corresponding population size (if known) into the data column for normalizing the count later.
+Note this data further satisfies the must-have for the distinct rows determined by index and key, and regularly spans over one-year interval.
 
@@
 
-(ref:tb-sub) A small subset of estimates of tuberculosis burden generated by World Health Organization in 2011 and 2012, with 12 observations and 5 variables.
+(ref:tb-sub) A small subset of estimates of tuberculosis burden collected by World Health Organization in 2011 and 2012, with 12 observations and 5 variables.
 The index refers to column `year`, the key to multiple columns: `country` and `gender`, and the measured variable to column `count`.
 
-The new data structure, tsibble, bridges the gap between raw temporal data and model inputs.
-Contextual semantics are introduced to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series model inputs.
-Index, key and time interval are the three stone pillars to this new semantically structured temporal data.
+The new tsibble structure bridges the gap between wild temporal data and the tamed state of model inputs.
+The proposed contextual semantics is the new addition to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series and panel model inputs.
+Index, key and time interval build the three stone pillars to this new semantically structured temporal data.
 Each is now described in more detail.
@@
 
-Time provides a contextual basis for temporal data.
-A variable representing time is essential for a tsibble, and is referred to as an "index".
-The "index" is an intact data column rather than a masked attribute, which makes time visible and accessible to users.
-This is highly advantageous when manipulating time.
-For example, one could easily extract time components, such as time of day and day of week, from the index to visualize seasonal effects of response variables.
-One could also join other data sources to the tsibble based on common time indexes.
-The accessibility of the tsibble index motivates data analysis towards transparency and human readability.
-When the "index" is available only as meta information (such as in the `ts` class), it creates an obstacle for analysts to write these simple queries in a programmatic manner, which should be discouraged from an analytical point of view.
+> *Index is a variable with inherent ordering from past to present.*
 
-Many different representations of time can be spotted in the wild.
-A date-time object, universally accepted across computing systems, is the most commonly used type for representing time.
-Date-time also typically associates with a time zone including adjustments such as daylight saving time.
-This diversity and time zone is acknowledged and accommodated by tsibble's index.
-When creating a tsibble, time indices are arranged from past to future within each series for the strict temporal ordering that is assumed by time series operations.
+Time provides the contextual basis for temporal data.
+Time can be seen in numerous representations, from sequential numerics to the most commonly accepted date-times.
+Regardless of this diversity, time should be inherently ordered from past to present, so should the index variable to a tsibble.
+
+Index is an explicit data variable rather than a masked attribute (such as in the `ts` and `zoo` classes), exposing a need for more accessible and transparent time operations.
+It is often to visualize and model seasonal effects of measurements of interest, meaning that time components such as time of day and day of week from the index should be easily extracted upon request.
+But when the index is available only as meta information, an obstacle is inevitably created for analysts to write such simple queries in an ad hoc and programmatic manner.
+From an analytical point of view it should be discouraged.
 

### 2019-08-12 tsibble.Rmd
done with revision on subsection "key"

@@
 Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an equivalent of the "tidy" format.
-Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the first and second definitions above.
+Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the definitions 1 and 2 above.
 Variables other than "index" and "key" are considered as measurements.
-The third and fourth definitions imply that tsibble is *tidier* than "tidy data", placing itself as a model input that gives rise to more robust and reliable downstream analytics.
+The definitions of 1 and 2 imply that tsibble is *tidier* than "tidy data", positioning itself as a model input that gives rise to more robust and reliable downstream analytics.
 
@@
 
-The "key" specification is the second essential ingredient for a tsibble.
-The "key" uniquely identifies observations that are recorded over time in a data table.
-It is similar to a primary key [@codd_relational_1970] defining each observation in a relational database.
-In the wider format in which multiple time series are often structured, the columns hold a series of values, so that the column implicitly serves as identification.
-In longer format, columns are melted with names converted to "key" values.
-However, the "key" provides much more flexibility.
-It is not constrained to a single field, but can be composed from multiple fields.
-The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks.
+> *Key is a set of variables that define observational units over time.*
 
-The "key" is usually known a priori by analysts.
+What subjects/entities to be observed over time leads to the second component of a tsibble--key.
+Key can consist of empty, single, or multiple variables identifying units measured along the way.
+When only a single observational unit is present in the table, key can be left with no variables to be specified.
+But when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units.
+In longitudinal data, key can be thought of as "panel" (such as in the Stata) constrained to a single variable.
+In tsibble, key allows for multiple variables of nesting and crossing relations [@wilkinson2006grammar] to be useful for forecasting reconciliation [@fpp; @R-hts] and other advanced models.
 For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year.
-This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data.
-Lacking either of the two will be inadequate, because the observations would not be uniquely identified, and thus a tsibble construction would fail.
-An alternative specification of the key for this data is to include a third variable `continent`.
-Since `country` is nested within `continent`, it is a free variable for use.
-This variable brings additional information that can be used for forecasting reconciliation [@fpp].
-The key needs to be explicit when multiple units exist in the data.
-The key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.
+This suggests that the key comprises at least columns `gender` and `country`.
+Since `country` is nested within `continent`, it is a free variable to be included in the key specification.
 
-The "key" also provides a link between the data, models, and forecasts.
-This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals.
+> *Each observation should be uniquely identified by index and key.*
+
+Inspired by "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, tsibble key also uniquely identifies each observational unit over time.
+Any duplicates of key-index pairs will fail upon the construction of a tsibble.
+This is, either `gender` or `country` alone is not enough to be the key for the tuberculosis data.
+The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage.
+It highlights the primary difference from *time tibble* provided by **tibbletime**.
+
+Since observational units are embedded, modelling and forecasting for each unit in a tsibble will be accelerated.
+The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts.
+This neatly decouples the data copy from models and forecasts, significantly reducing the storage space.
 More details are given in Section \@ref(sec:ts-models).
@@
 ## Interval {#sec:interval}
+
+> *Each observational unit should be measured at a common interval, if regularly spaced.*
 

### 2019-08-13 tsibble.Rmd
done with revision on subsection on interval

@@
 Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the "tidy" model, which is referred to as a "tsibble".
-In tsibble:
+Adhering to the "tidy data" principles, in tsibble:
 
@@
 
-Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an equivalent of the "tidy" format.
+Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an extension of the "tidy" format to the time domain.
 Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the definitions 1 and 2 above.
@@
 When only a single observational unit is present in the table, key can be left with no variables to be specified.
-But when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units.
-In longitudinal data, key can be thought of as "panel" (such as in the Stata) constrained to a single variable.
-In tsibble, key allows for multiple variables of nesting and crossing relations [@wilkinson2006grammar] to be useful for forecasting reconciliation [@fpp; @R-hts] and other advanced models.
+However, when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units.
+In longitudinal data, key can be thought of as "panel" (such as in the Stata) but constrained to a single variable.
+In tsibble, key allows for multiple variables of nesting, crossing, or union relations [@wilkinson2006grammar] to be useful for forecasting reconciliation [@fpp; @R-hts] and other advanced models.
 For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year.
@@
 Inspired by "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, tsibble key also uniquely identifies each observational unit over time.
-Any duplicates of key-index pairs will fail upon the construction of a tsibble.
+Any duplicates of key-index (i.e.
+key-value) pairs will fail upon the construction of a tsibble.
 This is, either `gender` or `country` alone is not enough to be the key for the tuberculosis data.
@@
 
-One of the cornerstones of time series data, and hence beneath a tsibble, is the time interval.
-This information plays a critical role in computing statistics (e.g.
-seasonal unit root tests) and building models (e.g.
-seasonal ARIMA).
-The principal divide is between regularly or irregularly spaced observations in time.
-A tsibble permits implicit missing time, making it difficult to distinguish regularity from the index.
-It relies on a user's specification by switching the `regular` argument off, when the data involves irregular intervals.
-This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect regular time series.
+The principal divide of temporal data is regularly versus irregularly spaced data.
+Event data typically involves irregular time intervals, such as flight schedules or customer transactions.
+This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect data with a fixed interval.
+When data indexed in regular time space, tsibble computes the greatest common divisor as the time interval from the index, or an unknown interval if one observation is available for each unit.
 
-For data indexed in regular time space, the time interval is automatically calculated, by first computing absolute differences of time indexes and then finding the greatest common divisor.
-This covers all conceivable cases, assuming that all observations in a tsibble have only one interval.
-Data collected at different intervals should be organized in separate tsibbles, encouraging well-tailored analysis and models, because each observation may have different underlying data generating processes.
+By extending one of the "tidy data" rules, "Each type of observational units should form a table", each observational unit is expected to share a common interval in a tsibble.
+Although tsibble reports one single interval despite data of fixed or mixed intervals, data of mixed intervals should be organized in separate tsibbles instead for well-tailored analysis and models.
+
+This tiny piece of information will make implicit missing time handlers handy and be in harmony with statistics and models that use seasonal periods.
 

### 2019-08-14 tsibble.Rmd
done with revision about data pipelines openning

@@
 
-This tiny piece of information will make implicit missing time handlers handy and be in harmony with statistics and models that use seasonal periods.
+This tiny piece of information will be carried over for tsibble-centric functions, for example to make implicit missing time handling handy and be in harmony with statistics and models that use seasonal periods.
 
@@
 
-A data pipeline describes the flow of data through an analysis, and can generally assist in conceptualizing the process, when it is applied to a variety of problems.
+A data pipeline describes the flow of data through an analysis, and can generally assist in conceptualizing the process for a stream of problems.
 @unix coined the term "pipelines" in software development while developing Unix at Bell Labs.
-In Unix-based computer operating systems, a pipeline chains together a series of operations on the basis of their standard streams, so that the output of each program becomes the input to another.
+In Unix-based computer operating systems, a pipeline chains together a series of operations based on their standard streams, so that the output of each program becomes the input to another.
 The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can also be considered a data pipeline.
@@
 @xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **XGobi**, **GGobi**, **Orca**, **plumbr** and **cranvas**, respectively.
+
+<!--
 The pipeline is typically described with a one way flow, from data to plot.
@@
 @xie2014reactive uses a reactive programming framework, to implement the pipeline, in which user's interactions trigger a sequence of modules to update their views, that is, practically the same as running the data pipeline producing each plot.
+-->
 
-Building a data pipeline is technically difficult: many implementation decisions have to be made about the interface, input and output objects and functionality.
-The tidy data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling.
-Each module communicates with the others, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.
-
-What is notable about an effective implementation of a data pipeline is that it coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies.
-A fluent pipeline glues tidy data and the grammar of data manipulation together.
-It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
-New data tools developed in the R package **tsibble** [@R-tsibble] articulate the time series data pipeline, which shepherds raw temporal data through to time series analysis, and plots.
-More detailed explanations are given in the following sections, and the examples.
+A fluent data pipeline anticipates a standard data structure.
+The tsibble data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling in temporal contexts.
+It provides a data infrastructure to a new ecosystem, **tidyverts**, a play on tidyverse that acknowledges the time series analysis purpose, in the spirit of "tidy data" to the **tidyverse**.
 
@@
 
-Most importantly, a new ecosystem for tidy time series analysis has been undertaken, using the tsibble framework, and is called "tidyverts", a play on tidyverse that acknowledges the time series analysis purpose.
+It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies.
+It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
+New data tools developed in the R package **tsibble** [@R-tsibble] articulate the temporal data pipeline, which shepherds wild temporal data through to robust analysis, and plots.
 

### 2019-08-14 tsibble.Rmd
new opening for section 4.1

@@
 
-This paper describes a new tidy representation for temporal data, and a unified framework to streamline the workflow from data preprocessing to visualization and forecasting, as an integral part of a tidy data analysis.
+This paper describes a new uniform representation for time series and longitudinal data, and a unified framework to streamline the workflow from data preprocessing to visualization and forecasting, as an integral part of a tidy data analysis.
 
@@
 Inspired by "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, tsibble key also uniquely identifies each observational unit over time.
-Any duplicates of key-index (i.e.
-key-value) pairs will fail upon the construction of a tsibble.
-This is, either `gender` or `country` alone is not enough to be the key for the tuberculosis data.
+When constructing a tsibble, any duplicates of key-index (i.e.
+key-value) pairs will fail, because duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making.
+For example, either `gender` or `country` alone is not enough to be the key for the tuberculosis data.
+Analysts are encouraged to better understand the data, or reason about the process of data cleaning when needed.
+Figure \@ref(fig:tsibble-tidy) peels the "tidy" module with clear guidance required for a tsibble.
 The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage.
-It highlights the primary difference from *time tibble* provided by **tibbletime**.
+
+```{r tsibble-tidy, fig.cap = "(ref:tsibble-tidy-cap)"}
+include_graphics("img/tsibble-tidy.png")
+```
+
+(ref:tsibble-tidy-cap) Details about the "tidy" stage for a tsibble.
+Built on top of "tidy data", each observation should be uniquely identified by index and key, thereby no duplicated key-index pairs.
 
@@
 The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts.
-This neatly decouples the data copy from models and forecasts, significantly reducing the storage space.
-More details are given in Section \@ref(sec:ts-models).
+This neatly decouples the expensive data copy from downstream summaries, significantly reducing the storage space.
 
@@
 
-By extending one of the "tidy data" rules, "Each type of observational units should form a table", each observational unit is expected to share a common interval in a tsibble.
+By extending one of the "tidy data" rules, "Each type of observational units should form a table", each observational unit shares a common interval in a tsibble.
 Although tsibble reports one single interval despite data of fixed or mixed intervals, data of mixed intervals should be organized in separate tsibbles instead for well-tailored analysis and models.
@@
 
-Figure \@ref(fig:tsibble-pipeline) illustrates the distinction of a time series pipeline from a regular data pipeline.
-It is highly recommended to check for identical entries of key and index before constructing a tsibble.
-Duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making.
-Analysts are encouraged to gaze at data early and reason about the process of data cleaning.
-When the data meets the tsibble standard, it flows neatly into the analysis stage and takes full advantage of the tsibble infrastructure.
-
-```{r tsibble-pipeline, fig.cap = "(ref:tsibble-pipeline-cap)"}
-include_graphics("img/tsibble-pipeline.png")
-```
-
-(ref:tsibble-pipeline-cap) A time series pipeline is different from a regular data pipeline: (1) check if the key and index uniquely identify each observation in the data whist creating a tsibble; (2) check if there are time gaps in the tsibble before analysis.
+The **tsibble** package not only provides a "tsibble" data object but also a domain specific language in R for transforming temporal data.
+It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation with ease.
 
@@
 
-## Time series visualization {#sec:ts-vis}
+## Temporal visualization {#sec:ts-vis}
 

### 2019-08-16 tsibble.Rmd
revised time gap verbs

@@
 For example, either `gender` or `country` alone is not enough to be the key for the tuberculosis data.
-Analysts are encouraged to better understand the data, or reason about the process of data cleaning when needed.
-Figure \@ref(fig:tsibble-tidy) peels the "tidy" module with clear guidance required for a tsibble.
+Analysts are encouraged to better understand the data, or reason about the process of data cleaning when handling duplicates.
+Figure \@ref(fig:tsibble-tidy) peels the "tidy" module with clear routes required for a tsibble.
 The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage.
@@
 The **tsibble** package not only provides a "tsibble" data object but also a domain specific language in R for transforming temporal data.
-It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation with ease.
+It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation with ease primarily in two aspects: implicit missingness handlers and time-aware aggregations.
 
-Many time operations such as lag/lead and time series models, assume an intact vector input ordered in time.
-Since a tsibble permits time gaps in the index, it is good practice to check and inspect any gaps in time following the creation of a tsibble, in order to prevent inviting these avoidable errors into the analysis.
-The first suite of verbs (rephrasing actions performed on the object) are provided to understand and tackle implicit missing values: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reveals all implicit missing observations; (3) `count_gaps()` summarizes the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
-To look into gaps over individual time periods or full-length time span, the common argument `.full` in these functions gives an option to easily switch between.
-The specification of `.full = TRUE` will result in fully balanced panels in other words.
+Implicit missingness are values that should be present but are absent.
+Regarding regularly-spaced temporal data, data entries that should be available at certain time points but are missing, leaving gaps in time.
+It will be a problem when temporal models and operations like lag/lead are applied.
+A family of verbs help to explore implicit missing values and turn them into the state of explicitness to eliminate potential concerns, listed as follows:
 
-\begin{table}[!ht]
-\caption{A list of table verbs working with tsibble. Functions in bold originating from the tidyverse and are adapted for tsibble.}\label{tab:verb-table}
-\centering
-\begin{tabular}{ lll }
-\toprule
- & Verb & Description \\ \midrule
-\multirow{4}{*}{Time gaps} & \texttt{has\_gaps()} & Test if a tsibble has gaps in time \\
- & \texttt{scan\_gaps()} & Reveal implicit missing entries \\
- & \texttt{count\_gaps()} & Summarize time gaps \\
- & \texttt{fill\_gaps()} & Fill in gaps by values and functions \\ \hline
-\multirow{4}{*}{Row-wise} & \textbf{\texttt{filter()}} & Pick rows based on conditions \\
- & \texttt{filter\_index()} & Provide a shorthand for time subsetting \\
- & \textbf{\texttt{slice()}} & Select rows based on row positions \\
- & \textbf{\texttt{arrange()}} & Sort the ordering of row by variables \\ \hline
-\multirow{4}{*}{Column-wise} & \textbf{\texttt{select()}} & Pick columns by variables \\
- & \textbf{\texttt{mutate()}} & Add new variables \\
- & \textbf{\texttt{transmute()}} & Drops existing variables \\
- & \textbf{\texttt{summarize()}} & Aggregate values over time \\ \hline
-\multirow{3}{*}{Group-wise} & \texttt{index\_by()} & Group by index candidate \\
- & \textbf{\texttt{group\_by()}} & Group by one or more variables \\
- & \texttt{group\_by\_key()} & Group by key variables \\ \hline
-\multirow{4}{*}{Reshape} & \textbf{\texttt{gather()}} & Gather columns into longer form \\
- & \textbf{\texttt{spread()}} & Spread columns into wider form \\
- & \textbf{\texttt{nest()}} & Nest values in a list-variable \\
- & \textbf{\texttt{unnest()}} & Unnest a list-variable \\ \hline
-\multirow{6}{*}{Join tables} & \textbf{\texttt{left\_join()}} & \multirow{6}{*}{Join two tables together} \\
- & \textbf{\texttt{right\_join()}} & \\
- & \textbf{\texttt{full\_join()}} & \\
- & \textbf{\texttt{inner\_join()}} & \\
- & \textbf{\texttt{semi\_join()}} & \\
- & \textbf{\texttt{anti\_join()}} & \\
-\bottomrule
-\end{tabular}
-\end{table}
+* `has_gaps()` checks the existence of time gaps.
+* `scan_gaps()` reveals all implicit missing observations.
+* `count_gaps()` summarizes the time ranges that are absent from the data.
+* `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.
+
+These verbs are evocative and of simple interface.
+They by default look into gaps for each individual time period.
+Switching on the option `.full = TRUE` will fill in the full-length time span, creating fully balanced panels with respect to longitudinal data analysis.
 

### 2019-08-16 tsibble.Rmd
done with revision on transformation

@@
 
-# Data pipelines {#sec:pipeline}
+# Temporal data pipelines {#sec:pipeline}
 
@@
 
-## Temporal transformation {#sec:timepipeline}
+## Transformation {#sec:timepipeline}
 
@@
 They by default look into gaps for each individual time period.
-Switching on the option `.full = TRUE` will fill in the full-length time span, creating fully balanced panels with respect to longitudinal data analysis.
+Switching on the option `.full = TRUE` will fill in the full-length time span, creating fully balanced panels with respect to longitudinal study.
 
-Besides the time gap verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in Table \@ref(tab:verb-table).
-The **tidyverse** suite showcases the general-purpose verbs for effectively manipulating tabular data, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables.
-But these verbs need handling with care when used in the time series domain.
-A perceivable difference is summarizing variables between data frame and tsibble using `summarize()`.
+The other key verb is `index_by()`, the counterpart of `group_by()` in **dplyr**, grouping and partitioning the index only.
+It is most useful to be used in conjunction with `summarize()`, and thus aggregations to upper-level time resolutions will take place.
+This combination automatically produces new index and interval, and helps to regularize data of irregular interval.
+
+Besides new verbs, the **dplyr** vocabulary has been adapted and expanded to facilitate temporal transformations.
+The **dplyr** suite showcases the general-purpose verbs for effectively manipulating tabular data.
+But these verbs need handling with care due to the context switch.
+A perceivable difference is summarizing variables between normal data and tsibble using `summarize()`.
 The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
-Users who are already familiar with the **tidyverse**, will experience a gentle learning curve for mastering these verbs and glide into time series analysis with low cognitive load.
-
 Attention has been paid to warning and error handling.
 The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble by automatically updating index and key under the hood.
-If the desired temporal ordering is changed by row-wise verbs, a warning is broadcast.
-If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is removed by selection, an error informs users of the problem and suggests alternatives.
+If the desired temporal ordering is changed by row-wise verbs (such as `arrange()` and `slice()`), a warning is broadcast.
+If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is removed by `select()`-ing, an error informs users of the problem and suggests alternatives.
 This avoids surprising users and reminds them of the time context.
+In general, users who are already familiar with the **tidyverse**, should have less resistance to learning the new semantics and verbs.
 
-The tsibble structure and operations support data pipelines for sequencing analysis.
-@eopl asserted "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony.
-This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems.
-Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr], read as "then".
-A sequence of functions can be composed in a way that can be naturally read from left to right, which improves the readability of the code.
-It consequently generates a block of code without saving intermediate values.
-
-It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies.
-It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
-New data tools developed in the R package **tsibble** [@R-tsibble] articulate the temporal data pipeline, which shepherds wild temporal data through to robust analysis, and plots.
-
-## Temporal visualization {#sec:ts-vis}
+## Visualization {#sec:ts-vis}
 
@@
 
-## Time series models {#sec:ts-models}
+## Models {#sec:ts-models}
 
@@
 The modeling module is a current endeavor.
+
+## Summary
+
+The tsibble structure and operations support data pipelines for sequencing analysis.
+@eopl asserted "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony.
+This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems.
+Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr], read as "then".
+A sequence of functions can be composed in a way that can be naturally read from left to right, which improves the readability of the code.
+It consequently generates a block of code without saving intermediate values.
+
+It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies.
+It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
+New data tools developed in the R package **tsibble** [@R-tsibble] articulate the temporal data pipeline, which shepherds wild temporal data through to robust analysis, and plots.
 

### 2019-08-17 tsibble.Rmd
done with revision on vis

@@
 Many in the R community have adopted the **tidyverse** way of thinking and extended it to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
-This paper and the associated **tsibble** R package extends the tidy way of thinking to temporal data.
+This paper and the associated **tsibble** R package [@R-tsibble] extends the tidy way of thinking to temporal data.
 
@@
 However, when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units.
-In longitudinal data, key can be thought of as "panel" (such as in the Stata) but constrained to a single variable.
+In longitudinal data, key can be thought of as "panel" (such as in the Stata) and is constrained to a single variable.
 In tsibble, key allows for multiple variables of nesting, crossing, or union relations [@wilkinson2006grammar] to be useful for forecasting reconciliation [@fpp; @R-hts] and other advanced models.
@@
 
-The other key verb is `index_by()`, the counterpart of `group_by()` in **dplyr**, grouping and partitioning the index only.
+The other key function, an adverb `index_by()`, is the counterpart of `group_by()` in **dplyr**, grouping and partitioning the index only.
 It is most useful to be used in conjunction with `summarize()`, and thus aggregations to upper-level time resolutions will take place.
@@
 Attention has been paid to warning and error handling.
-The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble by automatically updating index and key under the hood.
+The principle that underpins all the verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble over the course of transformation.
 If the desired temporal ordering is changed by row-wise verbs (such as `arrange()` and `slice()`), a warning is broadcast.
@@
 
-As a tsibble is a subclassing of data frame, it integrates well with the grammar of graphics.
-It is easy to create and extend specialist time series plotting methods based on the tsibble structure, for example autocorrelation plots and calendar-based graphics [@wang-calendarviz-2018].
+The **ggplot2** package (as the implementation of grammar of graphics) builds a powerful graphical system to declaratively visualize data.
+The data underpinning of **ggplot2** is tidy data, and in turn tsibble integrates well with **ggplot2**.
+The integration encourages more flexible graphics for exploring temporal structures, and individual or group differences.
+
+Line charts are universally accepted for ordered data, such as time series plots or spaghetti plots, depending on the fields.
+But they end up with exactly the same grammar: chronological time mapped to the horizontal axis, and the interested measurement on the vertical axis, for each unit.
+Many specialist plots centering around time series or longitudinal data, hence can be described and re-created by the grammar and **ggplot2**.
 
 ## Models {#sec:ts-models}
+
+* scalable
+* human-centered
 

### 2019-08-18 tsibble.Rmd
done with revision on model

@@
 
-This paper describes a new uniform representation for time series and longitudinal data, and a unified framework to streamline the workflow from data preprocessing to visualization and forecasting, as an integral part of a tidy data analysis.
+This paper describes a new tidy representation to encompass both time series and longitudinal data, and a unified framework to streamline the workflow from data preprocessing to visualization and forecasting, as an integral part of a tidy data analysis.
 
@@
 If the desired temporal ordering is changed by row-wise verbs (such as `arrange()` and `slice()`), a warning is broadcast.
-If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is removed by `select()`-ing, an error informs users of the problem and suggests alternatives.
+If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is dropped by `select()`-ing, an error informs users of the problem and suggests alternatives.
 This avoids surprising users and reminds them of the time context.
@@
 The data underpinning of **ggplot2** is tidy data, and in turn tsibble integrates well with **ggplot2**.
-The integration encourages more flexible graphics for exploring temporal structures, and individual or group differences.
+The integration encourages more flexible graphics for exploring temporal structures via index, and individual or group differences via key.
 
@@
 But they end up with exactly the same grammar: chronological time mapped to the horizontal axis, and the interested measurement on the vertical axis, for each unit.
-Many specialist plots centering around time series or longitudinal data, hence can be described and re-created by the grammar and **ggplot2**.
+Many specialist plots centering around time series or longitudinal data, hence can be described and re-created under the umbrella of the grammar and **ggplot2**.
 
-## Models {#sec:ts-models}
+## Model {#sec:ts-models}
 
-* scalable
-* human-centered
+Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis start to diverge.
+Nevertheless, tsibble as a model-oriented object can flow into both types of models, with new semantics (index and key) to be internally utilized.
 
-Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data.
-The verbs listed in Table \@ref(tab:verb-table) ease the transition to a tsibble that suits modeling.
-A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation.
-A tsibble usually contains multiple time series.
-Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently.
-This yields a "mable" (short for model table), where each model relates to each "key" value in tsibble.
-This avoids expensive data copying and reduces model storage.
-The mable is further supplied to forecasting methods, to produce a "fable" (short for forecasting table) in which each "key" along with its future time holds predictions.
-It also underlines the advantage of tsibble's "key" in acting as linkage between data inputs, models and forecasts.
-Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in a similar spirit.
-The modeling module is a current endeavor.
+Most time series models are univariate, such as ARIMA and Exponential Smoothing, modelling temporal dependency for each series independently.
+The **fable** package [@R-fable] provides a tidy forecasting framework built on top of tsibble is under development, with the goal of promoting transparent and human-centered forecasting practices.
+With the presence of the key, tsibble holds many series.
+Since models are fundamentally scalable, the `model()` and `forecast()` generics will take care of fitting and forecasting univariate models to each series across time in a tsibble at once.
+
+Panel data models, however, put emphases on overall, within, and between variations both across individuals and time. Fixed and random effects models can be developed in line with the **fable** tools.
 

### 2019-08-18 tsibble.Rmd
added new subsection to summarise the data pipelines

@@
 Variables other than "index" and "key" are considered as measurements.
-The definitions of 1 and 2 imply that tsibble is *tidier* than "tidy data", positioning itself as a model input that gives rise to more robust and reliable downstream analytics.
+The definitions of 3 and 4 imply that tsibble is *tidier* than "tidy data", positioning itself as a model input that gives rise to more robust and reliable downstream analytics.
 
@@
 The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
+
 Attention has been paid to warning and error handling.
@@
 
-Panel data models, however, put emphases on overall, within, and between variations both across individuals and time. Fixed and random effects models can be developed in line with the **fable** tools.
+Panel data models, however, put emphases on overall, within, and between variations both across individuals and time.
+Fixed and random effects models can be developed in line with the **fable** design.
 
@@
 
-The tsibble structure and operations support data pipelines for sequencing analysis.
-@eopl asserted "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony.
-This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems.
-Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr], read as "then".
-A sequence of functions can be composed in a way that can be naturally read from left to right, which improves the readability of the code.
-It consequently generates a block of code without saving intermediate values.
+The tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data.
+The supporting operations can be chained for sequencing analysis, articulating a data pipeline.
+As @eopl stated, "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." A mini snippet below glues transformation and forecasting together, to envision the fluent pipeline.
+The `pedestrian` dataset [cite] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package.
 
-It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to focus on the data analysis without getting lost in a jungle of computational intricacies.
+```{r tsibble-pipeline, eval = FALSE, echo = TRUE}
+```
+
+The pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr] chains the verbs, read as "then".
+A sequence of functions are composed in a way that can be naturally read from left to right, which improves the code readability.
+
+It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to document the data analysis through code without getting lost in a jungle of computational intricacies.
 It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
-New data tools developed in the R package **tsibble** [@R-tsibble] articulate the temporal data pipeline, which shepherds wild temporal data through to robust analysis, and plots.
 

### 2019-08-18 scripts/demo.R

@@ -1,6 +1,8 @@
 ## ---- load-pkg
 library(tidyverse)
+library(lubridate)
 library(tsibble)
+library(fable)
 
 ## ---- tb-sub
 tb_small <- read_rds("data/tb-small.rds")
@@ -92,3 +94,9 @@ ggplot() +
   theme_bw() +
   transition_manual(group)
 
+## ---- tsibble-pipeline
+pedestrian %>% 
+  fill_gaps() %>% # turn implicit missingness to explicit
+  filter(year(Date_Time) == 2016) %>%  # subset data of year 2016
+  model(arima = ARIMA(Count)) %>% # fit arima to each sensor
+  forecast(h = days(2)) # forecast 2 days ahead

### 2019-08-19 tsibble.Rmd
done with sec 5 revision

@@
 
-The tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data.
+To sum it up, the tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data.
 The supporting operations can be chained for sequencing analysis, articulating a data pipeline.
 As @eopl stated, "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." A mini snippet below glues transformation and forecasting together, to envision the fluent pipeline.
-The `pedestrian` dataset [cite] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package.
 
@@
 
+The `pedestrian` dataset [cite] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package.
 The pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr] chains the verbs, read as "then".
@@
 
+The **tsibble** package development follows closely to the **tidyverse** design principles [cite].
+
 ## Data first
@@
 
-Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data.
-The contextual summary provided by the print function, shown below on the data from Table \@ref(tab:tb-sub), contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of series.
-These details aid users in understanding their data better.
+Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal contexts, particularly useful when handling a large collection of data.
+The contextual information provided by the `print()` function, shown below from Table \@ref(tab:tb-sub), contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of units.
+These summaries aid users in understanding their data better.
 
@@
 
-Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in Table \@ref(tab:verb-table).
-Table verbs expect and return a tsibble, and does what the function name suggests.
-On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window.
-For example, computing moving averages requires numerics and a function like `mean()`, and produces averaged numerics.
-However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.
+Rolling window adapts to functional programming, which the **purrr** package [@R-purrr] sets a good example.
+These functions accept and return arbitrary inputs and outputs, with arbitrary methods.
+For example, moving averages anticipate numerics and produce averaged numerics via `mean()`.
+However, rolling window regression feeds a data frame into a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.
 
-The **purrr** package [@R-purrr] provides a good example of functional programming in R.
-It provides a complete and consistent set of tools to iterate each element of a vector with a function.
-Rolling window does not just iterate but rolls over a sequence of elements, namely `slide()`, `tile()` and `stretch()`.
+Rolling window not just iterates but rolls over a sequence of elements of a fixed window.
+A complete and consistent set of tools are available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants.
 `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
 For type stability, the functions always return lists.
-Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding type, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively.
+Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding types, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively.
 Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel [@R-future; @R-furrr].
-This family of functions empowers users to incorporate window-related operations in their workflows.
 
@@
 
-Modular programming is adopted in the design of the **tsibble** package. Modularity benefits users by providing small focused and manageable chunks, and provides developers with simpler maintenance.
+Modular programming is adopted in the design of the **tsibble** package. Modularity benefits users by providing small focused and cleaner chunks, and provides developers with simpler maintenance.
 
@@
 Each chunk is an independent module, but works interdependently.
-Vector functions in the package mostly deal with time.
-The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb collapses a tsibble to a less granular interval.
-The substitution of another time function in the `index_by()` results in the aggregation of different time resolution.
-Since these time functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages.
-The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions.
-It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the code readability.
+Vector functions in the package mostly operate on time.
+The atomic functions (such as `yearmonth()` and `yearquarter()`) can be embedded in the `index_by()` verb to collapse a tsibble to a less granular interval.
+Since they are independent from a tsibble, they can be used in a broader range of data applications not constrained to tsibble.
+On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package.
 
@@
 
-Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nanotime** [@R-nanotime], and pure numbers in simulations.
+Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution in **nanotime** [@R-nanotime], and numerics in simulation.
 Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday.
@@
 These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year.
-A new index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future.
+A new index would be immediately recognized upon defining `index_valid()`, as long as it can be ordered from past to future.
 The interval regarding semesters is further outlined through `interval_pull()`.
-As a result, the rest of the software methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
+As a result, all tsibble methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.
 
-The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble.
-For example, a fable is actually an extension of a tsibble, mentioned in Section \@ref(sec:ts-models).
+The class of tsibble is an underpinning for temporal data, and sub-classing a tsibble will be a demand.
 A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass.
 This new object itself is a tsibble.
-It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable.
-Tsibble verbs are also S3 generics.
-Developers will be able to implement these verbs for the new class, if necessary.
+It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, for example prediction distributions to a forecasting tsibble.
 
@@
 In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlights the "tidy" interface that is compatible with the **tidyverse**.
-This not only saves a few keystrokes without explicitly repeating references to the data source, but the resulting code is typically cleaner and more expressive.
+This not only saves a few keystrokes without explicitly repeating references to the data source, but the resulting code is typically cleaner and more expressive, when doing interactive data analysis.
 

### 2019-08-20 tsibble.Rmd
Done with the revision of section 2

@@
 This paper integrates time series analysis into this tidy framework, providing a coherent way for getting temporal data into the matrix format for modeling.
+These objects were designed for modeling methods.
+The diagram in the style of Figure \@ref(fig:tsibble-fit) would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
+This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.
 
@@
 Temporal data problems are typically grouped into two types of analysis, time series and longitudinal.
-Despite being exactly the same data input, the representation of time series and longitudinal data diverges due to different modeling approaches.
+Despite being exactly the same data input, the representation of time series and longitudinal data diverges due to different modeling focuses.
 
@@
 Multivariate time series are typically assumed to be in the format where each row is assumed to hold observations at a time point and each column to contain a single time series.
-(The tidy data name for this would be **wider format**.) This implies that data are columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric variables.
-In addition, time indexes are stripped off the data and implicitly inferred as attributes or meta-information.
-There is a strict requirement that the number of observations must be the same across all the series.
-Data wrangling, from the form that data arrives in, to this specialist format, can be frustrating and difficult, inhibiting the performance of downstream tasks.
+(The tidy data name for this would be **wider format**.) This implies that data are columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric vectors.
+Time indices are implicitly inferred as attributes or meta-information.
+The number of observations must be the same across all the series.
+Data wrangling, from the form that data arrives in to this specialist format, can be mysterious and frustrating, inhibiting the robustness of downstream tasks.
 
@@
 Longitudinal data or panel data typically assumes fewer measurements (small $T$) over a large number of individuals (large $N$).
-It often occurs that measurements for individuals are taken at different time points, resulting in an unbalanced panel.
-Thus, the primary format required for modeling such data is stacked series, blocks of measurements for each individual, with columns indicating individual, times of measurement and the measurements themselves.
-(The tidy data name for this would be **longer format**.) Evidently, this data organization saves storage space for many sparse cells, compared to structuring it into wider format which would have missing values in many cells.
-A drawback of this format is that information unique to each individual is often repeated for all time points.
-An appealing feature is that data is structured in a semantic manner with reference to observations and variables, with the time index stated explicitly.
-This opens the door to easily operating on time to make calculations and extract different temporal components, such as month and day of the week.
-It is conducive to examining the data in many different ways and leading to more comprehensive exploration and forecasting.
+It often occurs that measurements for individuals are taken at different time points.
+Thus, the primary format required for modeling such data is stacked data, blocks of measurements for each individual, with columns indicating panels, times of measurement and the measurements themselves.
+(The tidy data name for this would be **longer format**.) Evidently, this data organization saves storage space for many sparse cells, compared to structuring it into the wider format which would have missing values in many cells.
+An appealing feature is that data is structured in a semantic manner with reference to observations and variables, with panel and time stated explicitly.
+It is conducive to examining the data in many different ways and leading to more comprehensive exploration.
+
+## Existing time series and longitudinal data standards
+
+In R [@R-base], time series and longitudinal data are of different representations.
+The native `ts` object and the enhancement of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wider matrices with implicit time indices.
+A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous longer format.
+It only requires an index variable to be declared.
+However, as proposed in Section \@ref(sec:semantics) a more rigid data structure is required for temporal analytics and models.
+The **plm** [@plm2008] and **panelr** [@R-panelr] packages both manage longitudinal data in terms of longer format.
+
+Stata [@Stata] provides two commands, `xtset` and `tsset`, to declare time series and panel data respectively, both of which need specifying panel and time variables.
+Note they are in fact the same command, according to the manual.
+The underlying data arrangement is the same and of longer form.
+Either time series or panel commands are sufficient to apply to both types.
+The SAS software [@SAS-STAT] handles them in a similar way as Stata.
 
@@
 Each verb focuses on a singular task.
-Most common data tasks can be rephrased and tackled with these five key verbs, by composing them sequentially.
+Most common data tasks can be rephrased and tackled with these five key verbs, in conjunction with `group_by()` to perform grouped operations.
 
@@
 Many in the R community have adopted the **tidyverse** way of thinking and extended it to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018].
-This paper and the associated **tsibble** R package [@R-tsibble] extends the tidy way of thinking to temporal data.
+This paper with the associated **tsibble** R package [@R-tsibble] extends the tidy way of thinking to temporal data.
 
-## Existing time series standards in R
-
-Current standards, provided by the native `ts` object in R, and extended by **zoo** [@zoo2005] and **xts** [@R-xts], assemble temporal data into matrices with implicit time indexes.
-These objects were designed for modeling methods.
-The diagram in the style of Figure \@ref(fig:tsibble-fit) would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
-This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.
-
-A new temporal data class is needed in the upstream of the workflow, which could incorporate all the downstream modules.
-A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format.
-It only requires an index variable to declare a temporal data object, thus placing it at the import stage.
-However, as proposed in Section \@ref(sec:semantics) a more rigid data structure is required for time series analytics and models.
-
-This paper describes a new tidy representation to encompass both time series and longitudinal data, and a unified framework to streamline the workflow from data preprocessing to visualization and forecasting, as an integral part of a tidy data analysis.
+This paper attempts to provide a definition of tidy temporal data in relation to "tidy data", which encompasses both time series and longitudinal data.
+It aims at a unified framework to streamline the workflow from data preprocessing to visualization and modeling, as an integral part of a tidy data analysis.
 
@@
 However, when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units.
-In longitudinal data, key can be thought of as "panel" (such as in the Stata) and is constrained to a single variable.
-In tsibble, key allows for multiple variables of nesting, crossing, or union relations [@wilkinson2006grammar] to be useful for forecasting reconciliation [@fpp; @R-hts] and other advanced models.
+In longitudinal data, key can be thought of as "panel" (such as in the Stata) but constrained to a single variable in existing data structures.
+In tsibble, key allows for multiple variables of nesting, crossing, or union relations [@wilkinson2006grammar], that can be useful for forecasting reconciliation [@fpp; @R-hts] and richer visualization.
 For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year.
@@
 Event data typically involves irregular time intervals, such as flight schedules or customer transactions.
-This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect data with a fixed interval.
+This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect data with a fixed-time interval.
 When data indexed in regular time space, tsibble computes the greatest common divisor as the time interval from the index, or an unknown interval if one observation is available for each unit.
@@
 
-The `pedestrian` dataset [cite] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package.
+The `pedestrian` dataset [@ped] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package.
 The pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr] chains the verbs, read as "then".
@@
 
-The **tsibble** package development follows closely to the **tidyverse** design principles [cite].
+The **tsibble** package development follows closely to the **tidyverse** design principles [@tidyverse-principles].
 
@@
 The atomic functions (such as `yearmonth()` and `yearquarter()`) can be embedded in the `index_by()` verb to collapse a tsibble to a less granular interval.
-Since they are independent from a tsibble, they can be used in a broader range of data applications not constrained to tsibble.
+Since they are not tied to a tsibble, they can be used in a broader range of data applications not constrained to tsibble.
 On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package.
@@
 ```{r write-bib}
-write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang", "rlang"), file = "rpkgs.bib")
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang", "rlang", "fable", "panelr"), file = "rpkgs.bib")
 ```

### 2019-08-20 tsibble.Rmd
done: intro revision

@@
 Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, and multiple grouping factors.
-Temporal data may comprise the occurrence of events, such as flight departures, that need to be reduced to a regular structure.
+Temporal data may comprise the occurrence of time-stamped events, such as flight departures.
 
-Despite this variety and heterogeneity of temporal data, current software typically requires time series objects to be model-oriented matrices.
-Analysts are expected to do their own data preprocessing and take care of anything else needed to allow model fitting, which leads to a myriad of ad hoc solutions and duplicated efforts.
+Despite this variety and heterogeneity of temporal data, the lack of a uniform data organization expects analysts to do their own data preprocessing and take care of anything else needed to allow further analysis, which leads to a myriad of ad hoc solutions and duplicated efforts.
 
-@r4ds proposed the tidy data workflow, which provides a conceptual framework for processing data (as described in Figure \@ref(fig:tsibble-fit)).
-Currently, time series modeling and forecasting enters this framework at the *modeling* stage, while temporal data enters at the start.
-This paper integrates time series analysis into this tidy framework, providing a coherent way for getting temporal data into the matrix format for modeling.
-These objects were designed for modeling methods.
-The diagram in the style of Figure \@ref(fig:tsibble-fit) would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
-This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.
+@r4ds proposed the tidy data workflow, which provides a conceptual framework for exploring data (as described in Figure \@ref(fig:tsibble-fit)).
+In temporal domain, data with temporal information arrives at the "import" stage.
+A new abstraction, *tsibble*, is the gatekeeper at the "tidy" stage, to verify if the raw temporal data is appropriate for downstream analytics.
+The exploration loop will be aided with declarative grammars for more robust and accurate analysis.
 
@@
 
-(ref:tsibble-fit-cap) Illustration of the data science workflow, drawn from @r4ds, showing how current time series tools interface with the workflow and how the tsibble structure and tools integrate.
-The new data structure, tsibble, makes the connection between temporal data input, and specialist modeling formats.
-It provides elements at the "tidy" step, which produce tidy temporal data for time series visualization and modeling.
+(ref:tsibble-fit-cap) Illustration of the data science workflow regarding temporal data, drawn from @r4ds.
+The new data structure, tsibble, makes the connection between temporal data input, and downstream analytics.
+It provides elements at the "tidy" step, which produce tidy temporal data for temporal visualization and modeling.
 
@@
 Data wrangling, from the form that data arrives in to this specialist format, can be mysterious and frustrating, inhibiting the robustness of downstream tasks.
+This kind of format is model-centric, and the diagram in the style of Figure \@ref(fig:tsibble-fit) would place the "model" at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
+This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.
 
@@
 
-Stata [@Stata] provides two commands, `xtset` and `tsset`, to declare time series and panel data respectively, both of which need specifying panel and time variables.
+Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panel data respectively, both of which need specifying panel and time variables.
 Note they are in fact the same command, according to the manual.
@@
 
-Most time series models are univariate, such as ARIMA and Exponential Smoothing, modelling temporal dependency for each series independently.
+Most time series models are univariate, such as ARIMA and Exponential Smoothing, modelling temporal dynamics for each series independently.
 The **fable** package [@R-fable] provides a tidy forecasting framework built on top of tsibble is under development, with the goal of promoting transparent and human-centered forecasting practices.
@@
 
+* extensive index
+* flexible key
+
 A missing piece of the **tsibble** package is to enable user-defined calendars and to respect structurally missing observations.

### 2019-08-21 tsibble.Rmd
done with conclusion revision

@@
 
-Despite this variety and heterogeneity of temporal data, the lack of a uniform data organization expects analysts to do their own data preprocessing and take care of anything else needed to allow further analysis, which leads to a myriad of ad hoc solutions and duplicated efforts.
+Despite this variety and heterogeneity, little organization or conceptual oversight on how one should get the wild data into a tamed state is provided for temporal data.
+Analysts are expected to do their own data preprocessing and take care of anything else needed to allow further analysis, which leads to a myriad of ad hoc solutions and duplicated efforts.
 
 @r4ds proposed the tidy data workflow, which provides a conceptual framework for exploring data (as described in Figure \@ref(fig:tsibble-fit)).
-In temporal domain, data with temporal information arrives at the "import" stage.
+In temporal domain, data with time information arrives at the "import" stage.
 A new abstraction, *tsibble*, is the gatekeeper at the "tidy" stage, to verify if the raw temporal data is appropriate for downstream analytics.
-The exploration loop will be aided with declarative grammars for more robust and accurate analysis.
+The exploration loop will be aided with declarative grammars, yielding more robust and accurate analyses.
 
@@
 The number of observations must be the same across all the series.
-Data wrangling, from the form that data arrives in to this specialist format, can be mysterious and frustrating, inhibiting the robustness of downstream tasks.
+Data wrangling, from the form that data arrives in to this specialist format, can be mysterious and frustrating.
 This kind of format is model-centric, and the diagram in the style of Figure \@ref(fig:tsibble-fit) would place the "model" at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
@@
 Note they are in fact the same command, according to the manual.
-The underlying data arrangement is the same and of longer form.
+The underlying data arrangement is the unifying longer form.
 Either time series or panel commands are sufficient to apply to both types.
@@
 
-The new tsibble structure bridges the gap between wild temporal data and the tamed state of model inputs.
-The proposed contextual semantics is the new addition to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series and panel model inputs.
+The new tsibble structure bridges the gap between raw data and the rigorous state of temporal data analysis.
+The proposed contextual semantics is the new add-on to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series and panel model inputs.
 Index, key and time interval build the three stone pillars to this new semantically structured temporal data.
@@
 Most time series models are univariate, such as ARIMA and Exponential Smoothing, modelling temporal dynamics for each series independently.
-The **fable** package [@R-fable] provides a tidy forecasting framework built on top of tsibble is under development, with the goal of promoting transparent and human-centered forecasting practices.
+The **fable** package [@R-fable] provides a tidy forecasting framework built on top of tsibble, which is under development, with the goal of promoting transparent and human-centered forecasting practices.
 With the presence of the key, tsibble holds many series.
@@
 
-The "tsibble" is a new data abstraction to represent temporal data, allowing the "tidy data" principles to be brought to the time domain.
-Tidy data begins to take shape in the state of time with the introduction of the contextual semantics of index and key.
-A declared index provides direct support to the time variable; variables that comprise the key define observations over time.
+A data abstraction, tsibble, is proposed to represent temporal data, allowing the "tidy data" principles to be brought to the time domain.
+Tidy data begins to take shape in the state of time with new contextual semantics: index and key.
+The index variable provides direct support to an exhaustive set of ordered objects.
+The key identifies observational units over time, which can consist of single or multiple variables to be useful in more scenarios.
 These semantics further determine unique data entries required for a valid tsibble.
-No matter how temporal data arrives, a tsibble respects a time index and maintains the data richness.
-A tsibble frictionlessly allows transformation, visualization and modeling, and smoothly shifts between them, allowing for rapid iteration to gain data insights.
+It shepherds raw temporal data through the "tidy" stage to the next exploration stage for gaining more insights.
 
-* extensive index
-* flexible key
+The supporting toolkits articulate the temporal data pipeline, with the shared goal of reducing the time between the framings of data questions and the code realization.
+The rapid iteration for better understanding data is achieved through frictionlessly shifting among transformation, visualization, and modelling, using the standardized tsibble data infrastructure.
 
-A missing piece of the **tsibble** package is to enable user-defined calendars and to respect structurally missing observations.
+A nice feature that the tsibble structure should take into account is to allow user-defined calendars and to respect structurally missing observations.
 For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday.
 No data available outside trading hours would be labeled as structural missingness, which tsibble currently disregards.
-However, a few R packages provide functionality to create and manage many sorts of calendars, including market-specific business calendars.
-<!-- This delays the implementation.--> Generally, custom calendars are easily embedded into the tsibble framework.
-Consequently these tsibble operators, like `fill_gaps()`, would work out of the box, and forecasts would be generated within its definable time range.
-
-The **tsibble** package provides the grammar of temporal data manipulation, regardless of how the data is stored.
-Currently, it works for managing and manipulating temporal data frames in memory locally.
-But it is possible to work with remote tables stored in databases, such as SQLite and MySQL, using exactly the same tsibble code.
+Customer calendars can be embedded into the tsibble framework in theory.
+A few R packages provide functionality to create and manage many specific calendars, such as the **bizdays** package [@R-bizdays] for business days calendars.
+However, the lack of a generic calendar system delays the implementation.
 This is left for future work.
@@
 We also thank Stuart Lee for the feedback on this manuscript.
+We are grateful for anonymous reviewers for helpful feedback that has led to many improvements of the paper.
 This article was created with knitr [@knitr] and R Markdown [@rmarkdown].
@@
 
-```{r write-bib}
-write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang", "rlang", "fable", "panelr"), file = "rpkgs.bib")
+```{r write-bib, eval = FALSE}
+write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang", "rlang", "fable", "panelr", "bizdays"), file = "rpkgs.bib")
 ```

### 2019-08-21 tsibble.Rmd
updated code

@@
 read_chunk('scripts/smart-meter.R')
+ggplot2::theme_set(ggplot2::theme_bw())
 ```
@@
 
+<!--
+
 ```{r carrier-delayed}
@@
 Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).
+
+-->
 

### 2019-08-21 scripts/flights.R

@@ -9,16 +9,16 @@ flights <- read_rds("data/flights.rds")
 
 ## ---- find-duplicate
 flights %>% 
-  duplicates(key = id(flight_num), index = sched_dep_datetime) %>% 
+  duplicates(key = flight_num, index = sched_dep_datetime) %>% 
   as.data.frame()
-dup_lgl <- are_duplicated(flights, key = id(flight_num), 
+dup_lgl <- are_duplicated(flights, key = flight_num, 
   index = sched_dep_datetime, from_last = TRUE)
 
 ## ---- tsibble
 us_flights <- flights %>% 
   filter(!dup_lgl) %>% 
   as_tsibble(
-    key = id(flight_num), index = sched_dep_datetime,
+    key = flight_num, index = sched_dep_datetime,
     regular = FALSE, validate = FALSE
   )
 
@@ -29,7 +29,7 @@ cat(format(us_flights)[1:2], sep = "\n")
 delayed_carrier <- us_flights %>% 
   mutate(delayed = dep_delay > 15) %>%
   group_by(carrier) %>% 
-  index_by(year = year(sched_dep_datetime)) %>% 
+  index_by(year = ~ year(.)) %>% 
   summarise(
     Ontime = sum(delayed == 0),
     Delayed = sum(delayed)
@@ -42,9 +42,9 @@ delayed_carrier %>%
   mutate(carrier = fct_reorder(carrier, -n_flights)) %>% 
   ggplot() +
     geom_mosaic(aes(x = product(carrier), fill = delayed, weight = n_flights)) +
-    scale_fill_brewer(palette = "Dark2", name = "Delayed") +
+    scale_fill_brewer(palette = "Dark2", name = "Status") +
     scale_x_productlist(name = "Carrier") +
-    scale_y_productlist(name = "Delayed") +
+    scale_y_productlist(name = "Status") +
     theme(legend.position = "bottom")
 
 ## ---- sel-flights
@@ -54,13 +54,13 @@ sel_flights <- us_flights %>%
 ## ---- sel-delay
 sel_delay <- sel_flights %>% 
   group_by(origin) %>% 
-  index_by(sched_dep_date = as_date(sched_dep_datetime)) %>% 
+  index_by(sched_dep_date = ~ as_date(.)) %>% 
[…]

### 2019-08-21 scripts/smart-meter.R

@@ -12,7 +12,7 @@ weather <- read_rds("data/weather13.rds") %>%
 
 elec_ts <- elec %>% 
   build_tsibble(
-    key = id(customer_id), index = reading_datetime,
+    key = customer_id, index = reading_datetime,
     validate = FALSE, ordered = TRUE
   )
 
@@ -24,15 +24,24 @@ gap_df <- has_gaps(elec_ts)
 count_na_df <- elec_ts %>% 
   count_gaps()
 
-count_na_df %>% 
+lumped_na_df <- count_na_df %>% 
   mutate(
     customer_id = as.factor(customer_id) %>% 
       fct_lump(50) %>% fct_reorder(.n, sum)
-  ) %>% 
+  ) 
+
+# lumped_na_df %>% 
+#   filter(customer_id == "Other") %>% 
+#   ggplot(aes(x = .n)) +
+#   geom_histogram(bins = 300) +
+#   scale_y_sqrt()
+
+lumped_na_df %>% 
+  filter(customer_id != "Other") %>% 
   ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to)) +
-  geom_point(aes(y = .from), size = 0.6, shape = 4) +
-  geom_point(aes(y = .to), size = 0.6, shape = 4) +
+  geom_point(aes(y = .from), size = 0.6, shape = 4, alpha = 0.5) +
+  geom_point(aes(y = .to), size = 0.6, shape = 4, alpha = 0.5) +
   coord_flip() +
   xlab("Customer") +
   ylab("Time gaps") +
@@ -56,6 +65,6 @@ elec_cal <- elec_ts %>%
 p_cal <- elec_cal %>% 
   ggplot(aes(.time, .avg, group = date, colour = hot)) +
   geom_line() +
-  scale_colour_brewer(palette = "Dark2", direction = -1) +
+  scale_colour_brewer(palette = "Dark2", direction = -1, name = "") +
   theme(legend.position = "bottom")
 prettify(p_cal)

### 2019-08-22 tsibble.Rmd
done with revision on flights data

@@
 In fact, it fails to pass the tsibble hurdle due to duplicates in the original data.
-An error is immediately raised when attempting to convert this data into a tsibble, and closer inspection has to be carried out to locate the issue.
+An error is immediately raised when attempting to convert this data into a tsibble, and a closer inspection has to be carried out to locate the issue.
 The **tsibble** package provides tools to easily locate the duplicates in the data with `duplicates()`.
@@
 To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as the "index", and column `flight_num` as the "key".
-As a result of event timing, the data are irregularly spaced, and hence switching to the irregular option is necessary.
+This data happens to be irregularly spaced, and hence switching to the irregular option is necessary.
 The software internally validates if the key and index produce distinct rows, and then sorts the key and the index from past to recent.
-When the tsibble creation is done, the print display is data-oriented and contextually informative, including dimensions, irregular interval (``r tibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tibble::tbl_sum(us_flights)[2]``).
+When the tsibble creation is done, the print display is data-oriented and contextually informative, including dimensions, irregular interval with the time zone (``r tibble::tbl_sum(us_flights)[1]``) and the number of observational units (``r tibble::tbl_sum(us_flights)[2]``).
 
@@
 
-Transforming a tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can result in well-constructed pipelines.
-From the perspective of a passenger, for example, one needs to travel smart by choosing an efficient carrier to fly with and the time of day to avoid congestion.
-To explore this data, we drill down starting with annual carrier performance and followed by disaggregation to finer time resolutions.
+Transforming a tsibble for exploratory data analysis with a suite of time-aware and general-purpose manipulation verbs can result in well-constructed pipelines.
+A couple of use cases are described to show how to approach the interested questions by wrangling the tsibble while maintaining its temporal context.
 
-<!--
+What time of day and day of week should passengers travel to avoid suffering from horrible delay?
+Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples.
+The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are computed.
+This pipeline is initialized by regularizing and reshaping the list of the upper quantiles of departure delays for each hour.
+To visualize the temporal profiles, the time components (for example hours and weekdays) are extracted from the index.
+The flow chart (Figure \@ref(fig:flight-pipeline3)) demonstrates the operations undertaking in the data pipeline.
+The input to this pipeline is a tsibble of irregular interval for all flights, and the output ends up with a tsibble of one-hour interval by quantiles.
+To reduce the likelihood of suffering a delay, it is recommended to avoid the peak hour around 6pm (18) from Figure \@ref(fig:draw-qtl).
 
-```{r carrier-delayed}
+```{r quantile}
 ```
 
-```{r carrier-mosaic, fig.cap = "(ref:carrier-mosaic-cap)"}
+```{r draw-qtl-prep}
 ```
 
-(ref:carrier-mosaic-cap) Mosaic plot showing the association between the size of airline carriers and the delayed proportion of departures in 2017.
-Southwest Airlines is the largest operator, but does not operate as efficiently as Delta.
-Hawaiian Airlines, a small operator, outperforms the rest.
-
-```{r flight-pipeline1, fig.cap = "(ref:flight-pipeline1-cap)"}
-include_graphics("img/flight-pipeline1")
+```{r draw-qtl, fig.cap = "(ref:draw-qtl-cap)"}
 ```
 
-(ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:carrier-mosaic).
+(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceting day of week and 50%, 80% and 95% quantiles.
+A blue horizontal line indicates the 15-minute on-time standard to help grasp the delay severity.
+Passengers are apt to hold up around 18 during a day, and are recommended to travel early.
+The variations increase substantially as the upper tails.
 
-Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers.
-This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually.
-To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column.
-The flow chart (Figure \@ref(fig:flight-pipeline1)) demonstrates the operations undertaking in the data pipeline.
-The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval (as each carrier ends up with only one annual summary).
-The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display.
-Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).
+```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)"}
+include_graphics("img/flight-pipeline3")
+```
 
--->
+(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating Figure \@ref(fig:draw-qtl).
 
@@
 Superimposed on the plot are two-month moving averages, so the temporal trend is more visible.
-The number of days for each month is variable.
-Moving averages for two months call for computing weighted mean.
-But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without weight specification.
+Since the number of days for each month is variable, moving averages over two months will require weights input.
+But the weights specification can be avoid using a pair of commonly used rectangling verbs--`nest()` and `unnest()`, to wrap data frames partitioned by months into list-columns.
 The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year.
@@
 The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
+This pipeline gets the data into the daily series, and shifts the focus to five major airports.
 
@@
 
-What time of day and day of week should we travel to avoid suffering from horrible delay?
-Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples.
-The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown.
-To reduce the likelihood of suffering a delay, it is recommended to avoid the peak hour around 6pm (18).
-
-```{r quantile}
-```
-
-```{r draw-qtl-prep}
-```
-
-```{r draw-qtl, fig.cap = "(ref:draw-qtl-cap)"}
-```
-
-(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceting day of week and 50%, 80% and 95% quantiles.
-A blue horizontal line indicates the 15-minute on-time standard to help grasp the delay severity.
-Passengers are apt to hold up around 18 during a day, and are recommended to travel early.
-The variations increase substantially as the upper tails.
-
-```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)"}
-include_graphics("img/flight-pipeline3")
-```
-
-(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating Figure \@ref(fig:draw-qtl).
-
-\FloatBarrier
+This case study begins with duplicates fixing, that resolved the issue for constructing the tsibble.
+A range of temporal transformations can be handled by many free-form combinations of verbs, facilitating exploratory visualization.
 

### 2019-08-22 reviews/response.Rmd
added my response to reviewer's comments

@@
 
+This manuscript has undergone an overhaul to address reviewer's comments:
+
+1. Section 2.2 (previously 2.3) reviews existing time series and longitudinal data standards in popular statistical software, such as R, Stata, and SAS.
+2. We have rewritten Section 3 "Contextual semantics" to give a definition of tsibble in relation to tidy data, and we expand on the four rules of a tsibble in details.
+3.
+We have rewritten Section 4 "Temporal data pipelines" to describe a big picture about each module with the focus of temporal data instead of time series before.
+We have added Section 4.4 to summarise what a fluent pipeline looks like for a tsibble.
+4. We have removed the mosaic plot along with the paragraph from Section 6.1, and added the forecasting example to the smart meter data in Section 6.2 to show tsibble's capability.
+
+Besides these major changes, we have done lots of minor tweaks to improve the manuscript.
+
 # Reviewer 1 comments {-}
@@
 
-TODO.
+We have removed the mosaic plot and the paragraph.
+The `tsibble::index_by()` helps to update the index and interval when used with `summarise()`.
+It's a useful adverb to have for intuitive temporal manipulation.
+But you're right it can happen at a later stage.
 
@@
 
-TODO.
+Yes, the calendar plot just requires a data frame. We've added the forecasting example to back up this plot: how the weather information improves the demand forecasts.
 
@@
 Likewise authors may like to consider to replace “long data” to “longer data” and “wide data” to “wider data” as long/wide implies absolute terms.
-https: //tidyr.tidyverse.org/dev/articles/pivot.html
+<https://tidyr.tidyverse.org/dev/articles/pivot.html>
 
-TODO.
+We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`.
+But they are not S3 generics yet, and embedding tsibble into the new vctrs framework will take a long time, which delays the implementation.
+Conceptually the newly melt columns will be added to the key specification.
 
@@
 
-The JCGS peer-reviewed version is watermarked, which prevents the animation playing.
+The JCGS peer-reviewed version is processed with watermark, which prevents the animation playing.
 
@@
 
-TODO.
+We have added the sentence to Section 3.3, "Although tsibble reports one single interval despite data of fixed or mixed intervals, data of mixed intervals should be organized in separate tsibbles instead for well-tailored analysis and models.", to address the mixed time resolutions.
 
@@
 
-TODO.
+We have rewritten Section 2.2 (previously 2.3) to review a broader range of existing data structures in R, Stata, and SAS.
 
@@
 
-TODO.
+We removed that sentence.
 
@@
 
-TODO.
+We have re-written Section 3.2 about "key" to be clearer on defining observational units.
+`group_by()` is a transient operation and not limited to certain variables, but "key" persistently identifies observational units.
+But "key" can be altered when the observational units are changed during the course of transformation.
 
@@
 
-TODO.
+We added the reference to the **fable** package, and removed the mention of mable in Section 4.3.
 
@@
 
-TODO.
+Done. We moved **purrr** to the opening sentence of the previous paragraph.
 
@@
 
-TODO.
+We removed the mosaic plot (previously Figure 5).
 
-> Fig 6 Consider replacing gather() with pivot_longer() which is intended to be easier to use particularly for more complex "gathering" operations. https://tidyr.tidyverse.org/dev/articles/pivot.html
+> Fig 6 Consider replacing gather() with pivot_longer() which is intended to be easier to use particularly for more complex "gathering" operations.
+<https://tidyr.tidyverse.org/dev/articles/pivot.html>
 
-TODO.
+We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`.
+But they are not S3 generics yet, and embedding tsibble into the new vctrs framework will take a long time, which delays the implementation.
+Conceptually the newly melt columns will be added to the key specification.
 
@@
 
-TODO.
+Done.
 
@@
 
-TODO.
+We have re-written the custom calendars part, the last paragraph of Section 7.
+Conceptually it's easy to embed custom calendars into the tsibble framework, but a general flexible calendar software system is lacking, which delays the implementation.
+We added the **bizdays** R package as reference.
 

### 2019-08-23 tsibble.Rmd
added my response to reviewer's comments
done with smart meter revision

@@
 
-Since observational units are embedded, modelling and forecasting for each unit in a tsibble will be accelerated.
+Since observational units are embedded, modelling and forecasting across units and time in a tsibble will be simplified.
 The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts.
@@
 Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis start to diverge.
-Nevertheless, tsibble as a model-oriented object can flow into both types of models, with new semantics (index and key) to be internally utilized.
+Nevertheless, tsibble as a model-oriented object can flow into both types of models, with new semantics (index and key) to be internally utilized, to accelerate modelling.
 
@@
 Daily high and low temperatures in Newcastle in 2013 provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang].
-Two data tables might be joined to explore how local weather can contribute to the variations of daily electricity use when needed.
+Aggregating the half-hourly energy data to the same daily time interval as the temperature data allows to join the two data tables to explore how local weather can contribute to the variations of daily electricity use and the accuracy of demand forecasting.
 
@@
 
-Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)).
+Aggregation across all individuals helps to sketch a big picture of the behavioral change over time in the region, organized into a calendar display (Figure \@ref(fig:calendar-plot)) using the **sugrrants** package [@wang-calendarviz-2018].
 Each glyph represents the daily pattern of average residential electricity usage every thirty minutes.
@@
 However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.
-This plot illustrates how the tsibble data can easily integrate with other tools and graphics.
 
@@
 
+Lastly, a common practice in the energy sector is load forecasting.
+Forecasting the demand for the last day of 2013 is undertaken, holding the original energy data as the validation set.
+The most recent December data is fitted by ARIMA models with and without temporature information using automatic order selection [cite].
+The logarithmic transformation is applied to the average demand to ensure the positive forecasts.
+Figure \@ref(fig:elec-forecast) plots the one-day forecasts obtained from both models against the actual demand within the last week window.
+ARIMA regressed by average temporatures gives a better fit than the one without, although both tend to underestimate the night demand.
+The forecasting performance is reported in Table \@ref(tab:elec-accuracy), consistent with the findings in Figure \@ref(fig:elec-forecast).
+
+```{r elec-model}
+```
+
+```{r elec-forecast, fig.height = 3, fig.cap = "(ref:elec-forecast-cap)"}
+```
+
+(ref:elec-forecast-cap) One-day forecasts generated by ARIMA models with and without temporatures plotted against the actual demand.
+Both nicely capture the temporal dynamics, but ARIMA with temporatures performs better than the one without.
+
+<!--
+
 ```{r elec-pipeline2, fig.cap = "(ref:elec-pipeline2-cap)"}
@@
 (ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:calendar-plot).
+
+-->
+
+```{r elec-accuracy}
+```
+
+(ref:elec-accuracy) Accuracy measures to evaluate the forecasting performance between ARIMA models with and without temporatures, using the validation set.
+
+This case study demonstrates the significance of tsibble in lubricating the plumbing of handling time gaps, visualizing, and forecasting in general.
 
@@
 A few R packages provide functionality to create and manage many specific calendars, such as the **bizdays** package [@R-bizdays] for business days calendars.
-However, the lack of a generic calendar system delays the implementation.
+However, a generic flexible calendar system is lacking, that delays the implementation.
 This is left for future work.

### 2019-08-23 scripts/smart-meter.R

@@ -2,6 +2,7 @@
 library(tidyverse)
 library(lubridate)
 library(tsibble)
+library(fable)
 
 elec <- read_rds("data/smart-meter13.rds")
 weather <- read_rds("data/weather13.rds") %>% 
@@ -53,13 +54,15 @@ lumped_na_df %>%
 
 ## ---- calendar-plot
 library(sugrrants)
-elec_cal <- elec_ts %>% 
+elec_full <- elec_ts %>% 
   summarise(avg = mean(general_supply_kwh)) %>% 
   mutate(
     date = as_date(reading_datetime), 
     time = hms::as_hms(reading_datetime)
   ) %>% 
-  left_join(weather, by = "date") %>% 
+  left_join(weather, by = "date")
+
+elec_cal <- elec_full %>% 
   frame_calendar(x = time, y = avg, date = date)
 
 p_cal <- elec_cal %>% 
@@ -68,3 +71,33 @@ p_cal <- elec_cal %>%
   scale_colour_brewer(palette = "Dark2", direction = -1, name = "") +
   theme(legend.position = "bottom")
 prettify(p_cal)
+
+## ---- elec-model
+elec_train <- elec_full %>% 
+  filter_index("2013-12-01" ~ "2013-12-30")
+elec_test <- elec_full %>% 
+  filter_index("2013-12-31" ~ .)
+
+elec_fc <- elec_train %>% 
+  model(
+    `ARIMA w/ temperature` = ARIMA(log(avg) ~ avg_temp),
+    `ARIMA w/o temperature` = ARIMA(log(avg))
+  ) %>% 
+  forecast(new_data = elec_test)
+
+## ---- elec-forecast
+elec_fc %>% 
+  autoplot(data = filter_index(elec_train, "2013-12-25" ~ .), level = NULL) +
+  geom_line(aes(y = avg), data = elec_test, linetype = "dashed") +
+  scale_colour_brewer(palette = "Dark2", direction = -1, name = "") +
+  scale_x_datetime(date_breaks = "1 day", date_labels = "%b %d") +
+  theme(legend.position = "bottom") +
+  xlab("Reading time") +
+  ylab("Average electricity use")
+
+## ---- elec-accuracy
+accuracy(elec_fc, elec_test) %>% 
+  rename(model = .model) %>% 
+  select(-.type, -MASE, -ACF1) %>% 
+  knitr::kable(booktabs = TRUE, caption = "(ref:elec-accuracy)", linesep = "") %>%
[…]

### 2019-08-23 tsibble.Rmd
updated flow charts

@@
 
-(ref:tsibble-fit-cap) Illustration of the data science workflow regarding temporal data, drawn from @r4ds.
+(ref:tsibble-fit-cap) Annotation of the data science workflow regarding temporal data, drawn from @r4ds.
 The new data structure, tsibble, makes the connection between temporal data input, and downstream analytics.
@@
 Attention has been paid to warning and error handling.
-The principle that underpins all the verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble over the course of transformation.
+The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble over the course of transformation.
 If the desired temporal ordering is changed by row-wise verbs (such as `arrange()` and `slice()`), a warning is broadcast.
@@
 Forecasting the demand for the last day of 2013 is undertaken, holding the original energy data as the validation set.
-The most recent December data is fitted by ARIMA models with and without temporature information using automatic order selection [cite].
+The most recent December data is fitted by ARIMA models with and without temperature information using automatic order selection [@autoarima].
 The logarithmic transformation is applied to the average demand to ensure the positive forecasts.
 Figure \@ref(fig:elec-forecast) plots the one-day forecasts obtained from both models against the actual demand within the last week window.
-ARIMA regressed by average temporatures gives a better fit than the one without, although both tend to underestimate the night demand.
+ARIMA regressed by average temperatures gives a better fit than the one without, although both tend to underestimate the night demand.
 The forecasting performance is reported in Table \@ref(tab:elec-accuracy), consistent with the findings in Figure \@ref(fig:elec-forecast).
@@
 
-(ref:elec-forecast-cap) One-day forecasts generated by ARIMA models with and without temporatures plotted against the actual demand.
-Both nicely capture the temporal dynamics, but ARIMA with temporatures performs better than the one without.
-
-<!--
+(ref:elec-forecast-cap) One-day forecasts generated by ARIMA models with and without temperatures plotted against the actual demand.
+Both nicely capture the temporal dynamics, but ARIMA with temperatures performs better than the one without.
 
@@
 
-(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:calendar-plot).
-
--->
+(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline involved for creating Figure \@ref(fig:calendar-plot) and Figure \@ref(fig:elec-forecast).
 
@@
 
-(ref:elec-accuracy) Accuracy measures to evaluate the forecasting performance between ARIMA models with and without temporatures, using the validation set.
+(ref:elec-accuracy) Accuracy measures to evaluate the forecasting performance between ARIMA models with and without temperatures, using the validation set.
 

### 2019-08-25 reviews/response.Rmd
done with smart meter revision
updated response

@@
 
-Besides these major changes, we have done lots of minor tweaks to improve the manuscript.
+Along with these major changes, we have done lots of minor tweaks to improve the manuscript.
 
@@
 
-TODO.
+We have added a code snippet (featuring new `pedestrian` dataset) in a new subsection 4.4, and the load forecasting example to the energy consumption case study to show the cleaner, more expressive temporal data pipeline.
 
@@
 We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`.
-But they are not S3 generics yet, and embedding tsibble into the new vctrs framework will take a long time, which delays the implementation.
+But they are not S3 generics yet, which delays the implementation.
 Conceptually the newly melt columns will be added to the key specification.
@@
 
-TODO.
+[TODO]
+
+Time indexes lay the contextual foundation to both time series and longitudinal data.
+`zoo` and `xts` store indices as metadata due the matrix structure, but tsibble's indices are explicitly a data column that can be manipulated.
+Tsibble also has the capacity of custom index support and interval extraction.
+
+Tsibble is aimed at general temporal data, to unify time series and longitudinal **data**.
+In the revised manuscript (Section 4), we note the focuses of time series and longitudinal **analysis** are different in the modelling practice.
+Yes, the presented case studies are time series oriented, because currently time series model functions are available for tsibble.
 
@@
 
-TODO.
+Thanks. Done.
 

### 2019-08-25 tsibble.Rmd
updated response

@@
 read_chunk('scripts/smart-meter.R')
-ggplot2::theme_set(ggplot2::theme_bw())
 ```

### 2019-08-25 scripts/flights.R

@@ -3,6 +3,7 @@ library(lubridate)
 library(tidyverse)
 library(tsibble)
 library(forcats)
+theme_set(theme_bw())
 
 ## ---- load-flights
 flights <- read_rds("data/flights.rds")

### 2019-08-25 scripts/smart-meter.R

@@ -3,6 +3,7 @@ library(tidyverse)
 library(lubridate)
 library(tsibble)
 library(fable)
+theme_set(theme_bw())
 
 elec <- read_rds("data/smart-meter13.rds")
 weather <- read_rds("data/weather13.rds") %>% 

### 2019-08-26 reviews/response.Rmd
edited response

--- 2019-08-22 reviews/response.Rmd
@@
 
-TODO.
+Thanks for this comment.
+We have decided to mostly leave these to the vignettes, which are readily available on the web page for the package.
+The reason for doing this is that we feel that having many mini examples interrupts the flow of the paper.
+We have made two additions, though, which hopefully address these comments.
+
+A new code snippet (featuring new `pedestrian` dataset) is added to subsection 4.4.
+This is new part added to the paper, summarising the transformation and modeling parts of the data pipeline section.
+A new example is added that showcases how tsibble provides a platform to forecasting.
+It shows an energy consumption case study, illustrating the cleaner, and more expressive, temporal data modeling pipeline.
 
@@
 We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`.
-But they are not S3 generics yet, and embedding tsibble into the new vctrs framework will take a long time, which delays the implementation.
-Conceptually the newly melt columns will be added to the key specification.
+But they are not S3 generics yet, which delays the implementation.
+Conceptually the newly melted columns will be added to the key specification.
 
@@
 
-TODO.
+[TODO]
+
+Time indexes lay the contextual foundation to both time series and longitudinal data.
+`zoo` and `xts` store indices as metadata due the matrix structure, but tsibble's indices are explicitly a data column that can be manipulated.
+The index may be a date-time varible, as is usual in time series, but could be as simple as an ordered integer as is common in longitudinal data.
+It is also customizable to any variable that can be ordered.
+We have made this clearer in Section 3.1 where index specification is explained.
+
+Tsibble is aimed at general temporal data, to unify **data** structures underlying time series and longitudinal problems.
+In Section 4 of the revised manuscript, we clarify the different focuses of time series and longitudinal **analysis** and modelling.
+However, we have kept the emphasis on time series for the examples, because there are more functions available for time series model building on tsibble.
 

### 2019-08-26 tsibble.Rmd
adjusted figures

@@
 
-```{r count-gaps, fig.height = 7, fig.cap = "(ref:count-gaps-cap)"}
+```{r count-gaps, fig.height = 8, fig.cap = "(ref:count-gaps-cap)"}
 ```
 
-(ref:count-gaps-cap) Time gap plots for the 49 customers with most implicit missing values, and the remaining customers grouped into the one line at top.
+(ref:count-gaps-cap) Time gap plots for the 49 customers with most implicit missing values, and the remaining customers grouped into the one line at the bottom panel.
 Each cross represents an observation missing in time and a line between two dots shows continuous missingness over time.
@@
 The logarithmic transformation is applied to the average demand to ensure the positive forecasts.
-Figure \@ref(fig:elec-forecast) plots the one-day forecasts obtained from both models against the actual demand within the last week window.
+Figure \@ref(fig:elec-forecast) plots the one-day forecasts obtained from both models against the actual demand within the last two-week window.
 ARIMA regressed by average temperatures gives a better fit than the one without, although both tend to underestimate the night demand.

### 2019-08-26 scripts/smart-meter.R

@@ -3,6 +3,7 @@ library(tidyverse)
 library(lubridate)
 library(tsibble)
 library(fable)
+library(patchwork)
 theme_set(theme_bw())
 
 elec <- read_rds("data/smart-meter13.rds")
@@ -32,26 +33,38 @@ lumped_na_df <- count_na_df %>%
       fct_lump(50) %>% fct_reorder(.n, sum)
   ) 
 
-# lumped_na_df %>% 
-#   filter(customer_id == "Other") %>% 
-#   ggplot(aes(x = .n)) +
-#   geom_histogram(bins = 300) +
-#   scale_y_sqrt()
-
-lumped_na_df %>% 
+p_49 <- lumped_na_df %>% 
   filter(customer_id != "Other") %>% 
   ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to)) +
-  geom_point(aes(y = .from), size = 0.6, shape = 4, alpha = 0.5) +
-  geom_point(aes(y = .to), size = 0.6, shape = 4, alpha = 0.5) +
+  geom_point(aes(y = .from), size = 0.6, shape = 4) +
+  geom_point(aes(y = .to), size = 0.6, shape = 4) +
   coord_flip() +
   xlab("Customer") +
+  ylab("") +
+  theme(
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    legend.position = "none",
+    plot.margin = unit(c(0, 0, -1, -1), "line")
+  )
+
+p_other <- lumped_na_df %>% 
+  filter(customer_id == "Other") %>% 
+  ggplot(aes(x = customer_id)) +
+  geom_linerange(aes(ymin = .from, ymax = .to), alpha = 0.1) +
+  geom_point(aes(y = .from), size = 0.6, shape = 4, alpha = 0.1) +
+  geom_point(aes(y = .to), size = 0.6, shape = 4, alpha = 0.1) +
+  coord_flip() +
+  xlab("Others") +
   ylab("Time gaps") +
   theme(
     axis.text.y = element_blank(),
     axis.ticks.y = element_blank(),
     legend.position = "none"
   )
+
+p_49 + p_other + plot_layout(ncol = 1, heights = c(10, 1))
 
 ## ---- calendar-plot
 library(sugrrants)
@@ -81,14 +94,14 @@ elec_test <- elec_full %>%
 
 elec_fc <- elec_train %>% 
   model(
[…]

### 2019-08-26 reviews/response.Rmd
done with response

@@
 
-TODO.
+Thanks. We separated the remaining customers to the bottom panel with points of some transparency. We have kept the line range plot for the consistency between two panels.
 
@@
 
-[TODO]
-
 Time indexes lay the contextual foundation to both time series and longitudinal data.
 `zoo` and `xts` store indices as metadata due the matrix structure, but tsibble's indices are explicitly a data column that can be manipulated.
-The index may be a date-time varible, as is usual in time series, but could be as simple as an ordered integer as is common in longitudinal data.
+The index may be a date-time variable, as is usual in time series, but could be as simple as an ordered integer as is common in longitudinal data.
 It is also customizable to any variable that can be ordered.
@@
 
-The JCGS peer-reviewed version is processed with watermark, which prevents the animation playing.
+Thanks for pointing it out. We include a gif version of the animation as supplementary materials.
 

### 2019-08-27 reviews/response.Rmd
typo fixed
section 2.1-2

@@
 
-1. Section 2.2 (previously 2.3) reviews existing time series and longitudinal data standards in popular statistical software, such as R, Stata, and SAS.
+1. Section 2.2 (previously 2.3) has been extended to include reviews of other software than R (Stata and SAS) time series and longitudinal data standards.
 2. We have rewritten Section 3 "Contextual semantics" to give a definition of tsibble in relation to tidy data, and we expand on the four rules of a tsibble in details.
@@
 But they are not S3 generics yet, which delays the implementation.
-Conceptually the newly melt columns will be added to the key specification.
+Conceptually the newly melted columns will be added to the key specification.
 
@@
 
-> P26L54 there’s an errant
+> P26L54 there's an errant
 

### 2019-08-27 tsibble.Rmd
section 2.1-2

@@
 
-## Time series and longitudinal data {#sec:time-series}
+## Comparing time series and longitudinal data {#sec:time-series}
 
-Temporal data problems are typically grouped into two types of analysis, time series and longitudinal.
-Despite being exactly the same data input, the representation of time series and longitudinal data diverges due to different modeling focuses.
+Temporal data problems often fall into two types of analysis, time series and longitudinal.
+Both of these may have very similar data input, but the representation for modelling is typically different.
+Time series analysis tends to focus on the dependency with series, and the cross-correlation between series.
+Longitudinal analysis tends to focus on overall temporal patterns across demographic or experimental treatment strata, that incorporates within subject dependency.
 
-Time series can be univariate or multivariate, and for modeling require relatively long lengths (i.e., large $T$).
-Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic processes, for the primary purpose of forecasting, and characterizing temporal dynamics.
-Most statistical software represent such time series as vectors or matrices.
-Multivariate time series are typically assumed to be in the format where each row is assumed to hold observations at a time point and each column to contain a single time series.
-(The tidy data name for this would be **wider format**.) This implies that data are columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric vectors.
-Time indices are implicitly inferred as attributes or meta-information.
-The number of observations must be the same across all the series.
-Data wrangling, from the form that data arrives in to this specialist format, can be mysterious and frustrating.
-This kind of format is model-centric, and the diagram in the style of Figure \@ref(fig:tsibble-fit) would place the "model" at the center of the analytical universe, and all the transformations and visualizations would hinge on that format.
-This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.
+Time series can be univariate or multivariate, and require relatively long lengths (i.e., large $T$) for modeling.
+With this large $T$ property, the series can be handled as stochastic processes for the primary purpose of forecasting, and characterizing temporal dynamics.
+Due to an expectation of regular time intervals, and equal length across series, multivariate time series are typically assumed to be in the format where each column contains a single time series, and time is specified implicitly.
+This also implies that data are columns of homogeneous types: either all numeric or all non-numeric.
+To wrestle data, in whatever format that it arrives in, to this modeling format can be frustrating.
+The format could be considered to be model-centric, rather than data-centric, and thus throws the analyst into the deep end of the pool, rather than allowing them to gently wade to the modeling stage from the shallow end.
+The expectation is that the "model" is at the center of the analytical universe.
+This is contrary to the **tidyverse** conceptualization (Figure \@ref(fig:tsibble-fit)), which holistically captures the full data workflow.
+More support needs to be provided, in the form of consistent tools and data structures, to transform the data into the analytical cycle.
 
-For longitudinal analysis, researchers and analysts are primarily interested in explaining trends across and variations among individuals, and making inference about a broader population.
-Longitudinal data or panel data typically assumes fewer measurements (small $T$) over a large number of individuals (large $N$).
-It often occurs that measurements for individuals are taken at different time points.
-Thus, the primary format required for modeling such data is stacked data, blocks of measurements for each individual, with columns indicating panels, times of measurement and the measurements themselves.
-(The tidy data name for this would be **longer format**.) Evidently, this data organization saves storage space for many sparse cells, compared to structuring it into the wider format which would have missing values in many cells.
-An appealing feature is that data is structured in a semantic manner with reference to observations and variables, with panel and time stated explicitly.
-It is conducive to examining the data in many different ways and leading to more comprehensive exploration.
+Longitudinal data (or panel data) typically assumes fewer measurements (small $T$) over a large number of individuals (large $N$).
+It often occurs that measurements for individuals are taken at different time points, and in different quantities.
+The primary format required for modeling is stacked data, blocks of measurements for each individual, with columns indicating panels, times of measurement and the measurements themselves.
+An appealing feature is that data is structured in a semantic manner with reference to observations and variables, with panel and time explicitly stated.
 
-## Existing time series and longitudinal data standards
+## Existing data standards
 
 In R [@R-base], time series and longitudinal data are of different representations.
-The native `ts` object and the enhancement of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wider matrices with implicit time indices.
-A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous longer format.
+The native `ts` object and the enhancement of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes.
+A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous long format.
 It only requires an index variable to be declared.
-However, as proposed in Section \@ref(sec:semantics) a more rigid data structure is required for temporal analytics and models.
-The **plm** [@plm2008] and **panelr** [@R-panelr] packages both manage longitudinal data in terms of longer format.
+However, this is insufficient, and a more rigid data structure is required for temporal analytics and modeling.
+The **plm** [@plm2008] and **panelr** [@R-panelr] packages both manage longitudinal data in long format.
 
-Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panel data respectively, both of which need specifying panel and time variables.
-Note they are in fact the same command, according to the manual.
-The underlying data arrangement is the unifying longer form.
-Either time series or panel commands are sufficient to apply to both types.
-The SAS software [@SAS-STAT] handles them in a similar way as Stata.
+Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panel data respectively, both of which require explicit panel and time index specification.
+The underlying data arrangement is only long form, for both types of data.
+The functions pay no attention to whether the data is for time series analysis or longitudinal data analysis.
+The SAS software [@SAS-STAT] also handles both types of data in the same way as Stata.
 
 ## Tidy data and the grammar of data manipulation
+
+ (The tidy data name for this would be **wider format**.)
 
@@
 We also thank Stuart Lee for the feedback on this manuscript.
-We are grateful for anonymous reviewers for helpful feedback that has led to many improvements of the paper.
+We are grateful for anonymous reviewers for helpful feedback that has led to many improvements in the paper.
 This article was created with knitr [@knitr] and R Markdown [@rmarkdown].

### 2019-08-27 tsibble.Rmd
minor tweaks and adjusted gap plots

@@
 With this large $T$ property, the series can be handled as stochastic processes for the primary purpose of forecasting, and characterizing temporal dynamics.
-Due to an expectation of regular time intervals, and equal length across series, multivariate time series are typically assumed to be in the format where each column contains a single time series, and time is specified implicitly.
+Due to an expectation of regularly spaced time, and equal lengths across series, multivariate time series are typically assumed to be in the format where each column contains a single time series, and time is specified implicitly.
 This also implies that data are columns of homogeneous types: either all numeric or all non-numeric.
@@
 The underlying data arrangement is only long form, for both types of data.
-The functions pay no attention to whether the data is for time series analysis or longitudinal data analysis.
+Both groups of functions can be applied interchangeably to whether the data is declared for time series or longitudinal data.
 The SAS software [@SAS-STAT] also handles both types of data in the same way as Stata.

### 2019-08-27 scripts/smart-meter.R

@@ -40,7 +40,7 @@ p_49 <- lumped_na_df %>%
   geom_point(aes(y = .from), size = 0.6, shape = 4) +
   geom_point(aes(y = .to), size = 0.6, shape = 4) +
   coord_flip() +
-  xlab("Customer") +
+  xlab("Top customers") +
   ylab("") +
   theme(
     axis.text = element_blank(),
@@ -50,15 +50,16 @@ p_49 <- lumped_na_df %>%
   )
 
 p_other <- lumped_na_df %>% 
-  filter(customer_id == "Other") %>% 
+  filter(customer_id == "Rest") %>% 
   ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to), alpha = 0.1) +
-  geom_point(aes(y = .from), size = 0.6, shape = 4, alpha = 0.1) +
-  geom_point(aes(y = .to), size = 0.6, shape = 4, alpha = 0.1) +
+  geom_point(aes(y = .from), size = 1.2, shape = 4, alpha = 0.1) +
+  geom_point(aes(y = .to), size = 1.2, shape = 4, alpha = 0.1) +
   coord_flip() +
   xlab("Others") +
   ylab("Time gaps") +
   theme(
+    panel.grid.major.y = element_blank(),
     axis.text.y = element_blank(),
     axis.ticks.y = element_blank(),
     legend.position = "none"

### 2019-08-27 scripts/smart-meter.R
fixed the gap bottom panel display

@@ -50,13 +50,13 @@ p_49 <- lumped_na_df %>%
   )
 
 p_other <- lumped_na_df %>% 
-  filter(customer_id == "Rest") %>% 
+  filter(customer_id == "Other") %>% 
   ggplot(aes(x = customer_id)) +
   geom_linerange(aes(ymin = .from, ymax = .to), alpha = 0.1) +
   geom_point(aes(y = .from), size = 1.2, shape = 4, alpha = 0.1) +
   geom_point(aes(y = .to), size = 1.2, shape = 4, alpha = 0.1) +
   coord_flip() +
-  xlab("Others") +
+  xlab("Rest") +
   ylab("Time gaps") +
   theme(
     panel.grid.major.y = element_blank(),

### 2019-08-27 tsibble.Rmd
section 2.3 massaged

@@
 The expectation is that the "model" is at the center of the analytical universe.
-This is contrary to the **tidyverse** conceptualization (Figure \@ref(fig:tsibble-fit)), which holistically captures the full data workflow.
+This is contrary to the **tidyverse** conceptualization (Figure \@ref(fig:tsibble-fit)), which holistically captures the data workflow.
 More support needs to be provided, in the form of consistent tools and data structures, to transform the data into the analytical cycle.
@@
 The native `ts` object and the enhancement of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes.
+If there are multiple sub-groups, like country or product type, these would be kept in different data objects.
 A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous long format.
@@
 
-Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panel data respectively, both of which require explicit panel and time index specification.
+Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panels respectively, both of which require explicit panel id and time index specification.
+Different variables would be stored in multiple columns.
 The underlying data arrangement is only long form, for both types of data.
@@
 
-## Tidy data and the grammar of data manipulation
+## Tidy data
 
- (The tidy data name for this would be **wider format**.)
+@wickham2014tidy coined the term "tidy data", to standardize the mapping of the semantics of a dataset to its physical representation.
+In tidy form, rows correspond to observations and columns to variables.
+Tidy data is a rephrasing of the second and third normal forms from relational databases, but the explanation in terms of observations and variables is easier to understand because it uses statistical terminology.
 
-@wickham2014tidy coined the term "tidy data", which is a rephrasing of the second and third normal forms in relational databases but in a way that makes more sense to data scientists by referring rows to observations and columns to variables.
-The principles of "tidy data" attempt to standardize the mapping of the semantics of a dataset to its physical representation.
-This data structure is the fundamental unit of the **tidyverse**, which is a collection of R packages designed for data science.
+Multiple time series, with each column corresponding to a measurement is tidy data when the time index is explicitly stored in a column.
+The stacked data format used in longitudinal data storage is tidy, and accommodates explicit identification of sub-groups.
+
+The tidy data structure is the fundamental unit of the **tidyverse**, which is a collection of R packages designed for data science.
 The ubiquitous use of the **tidyverse** is testament to the simplicity, practicality and general applicability of the tools.
@@
 
-This paper attempts to provide a definition of tidy temporal data in relation to "tidy data", which encompasses both time series and longitudinal data.
-It aims at a unified framework to streamline the workflow from data preprocessing to visualization and modeling, as an integral part of a tidy data analysis.
+For temporal data, the tidy definition needs additional criteria, that assist in handling the time context.
+This is addressed in the next section, and encompasses both time series and longitudinal data.
+It provides a unified framework to streamline the workflow from data preprocessing to visualization and modeling, as an integral part of a tidy data analysis.
 

### 2019-08-28 tsibble.Rmd
section 3

@@
 Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the "tidy" model, which is referred to as a "tsibble".
-Adhering to the "tidy data" principles, in tsibble:
+Adapting the "tidy data" principles, in tsibble:
 
@@
 
-To materialize the abstraction of the tsibble, a subset of tuberculosis cases estimated by @tb-data presented in Table \@ref(tab:tb-sub) is used as an example.
+To materialize the abstraction of the tsibble, a subset of tuberculosis cases [@tb-data], as presented in Table \@ref(tab:tb-sub), is used as an example.
 It contains 12 observations and 5 variables landing in a tidy data form.
@@
 Column `count` is the only measured variable in this data, but the data structure is sufficiently flexible to hold more measurements; for example, slotting the corresponding population size (if known) into the data column for normalizing the count later.
-Note this data further satisfies the must-have for the distinct rows determined by index and key, and regularly spans over one-year interval.
+Note, this data further satisfies the must-have for the distinct rows determined by index and key, and regularly spans over one-year intervals.
 
@@
 Time can be seen in numerous representations, from sequential numerics to the most commonly accepted date-times.
-Regardless of this diversity, time should be inherently ordered from past to present, so should the index variable to a tsibble.
+Regardless of this diversity, time should be inherently ordered from past to present, so should be the index variable to a tsibble.
 
 Index is an explicit data variable rather than a masked attribute (such as in the `ts` and `zoo` classes), exposing a need for more accessible and transparent time operations.
-It is often to visualize and model seasonal effects of measurements of interest, meaning that time components such as time of day and day of week from the index should be easily extracted upon request.
-But when the index is available only as meta information, an obstacle is inevitably created for analysts to write such simple queries in an ad hoc and programmatic manner.
-From an analytical point of view it should be discouraged.
+It is often necessary to visualize and model seasonal effects of measurements of interest, meaning that time components, such as time of day and day of week, should be easily extracted from the index.
+When the index is available only as meta information, it provides an obstacle for analysts by complicating the writing of even simple queries, often requiring special purpose programming.
+From an analytical point of view this should be discouraged.
 
@@
 
-What subjects/entities to be observed over time leads to the second component of a tsibble--key.
-Key can consist of empty, single, or multiple variables identifying units measured along the way.
-When only a single observational unit is present in the table, key can be left with no variables to be specified.
-However, when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units.
-In longitudinal data, key can be thought of as "panel" (such as in the Stata) but constrained to a single variable in existing data structures.
-In tsibble, key allows for multiple variables of nesting, crossing, or union relations [@wilkinson2006grammar], that can be useful for forecasting reconciliation [@fpp; @R-hts] and richer visualization.
+What subjects/entities are to be observed over time, leads to the second component of a tsibble--key.
+The key can consist of empty, single, or multiple variables identifying units measured along the way.
+When only a single observational unit is present in the table, no key needs to be specified.
+However, when multiple units exist in the data, the key should be supplied by identifying variables to sufficiently define the units.
+In longitudinal data, the key can be thought of as "panel" (such as in the Stata) but constrained to a single variable in existing data structures.
+In tsibble, the key allows for multiple variables of nesting, crossing, or union relations [@wilkinson2006grammar], that can be useful for forecasting reconciliation [@fpp; @R-hts] and richer visualization.
 For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year.
 This suggests that the key comprises at least columns `gender` and `country`.
-Since `country` is nested within `continent`, it is a free variable to be included in the key specification.
+Since `country` is nested within `continent`, continent can be included in the key specification, but is not needed.
 
@@
 
-Inspired by "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, tsibble key also uniquely identifies each observational unit over time.
+Inspired by a "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, the tsibble key also uniquely identifies each observational unit over time.
 When constructing a tsibble, any duplicates of key-index (i.e.
@@
 Event data typically involves irregular time intervals, such as flight schedules or customer transactions.
-This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect data with a fixed-time interval.
-When data indexed in regular time space, tsibble computes the greatest common divisor as the time interval from the index, or an unknown interval if one observation is available for each unit.
+This type of data can flow into event-based data modeling, but would need to be processed, or regularized, to fit models that expect data with a fixed-time interval.
 
-By extending one of the "tidy data" rules, "Each type of observational units should form a table", each observational unit shares a common interval in a tsibble.
-Although tsibble reports one single interval despite data of fixed or mixed intervals, data of mixed intervals should be organized in separate tsibbles instead for well-tailored analysis and models.
+There are three possible interval types: a number, unknown, and irregular.
+To determine the interval for regularly spaced data, tsibble computes the greatest common divisor, and is a number.
+If only one observation is available for each unit, which may occur after aggregating data, the interval is reported as unknown.
+When the data arrives with irregular time, like event data, the interval would be specified as irregular, to prevent the tsibble creator attempting to guess an interval.
 
-This tiny piece of information will be carried over for tsibble-centric functions, for example to make implicit missing time handling handy and be in harmony with statistics and models that use seasonal periods.
+To abide by the "tidy data" rules -- "Each type of observational units should form a table" -- in a tisbble each observational unit shares a common interval.
+This means that a tsibble will report one single interval, whether the data has a fixed or mixed set of intervals.
+To handle mixed interval data, it should be organized into separate tsibbles for a well-tailored analysis.
+
+This tiny piece of information, the interval, is carried over for tsibble-centric operations.
+For example, this makes implicit missing time handling convenient, and harmoniously operates with statistical calculations, and models, on seasonal periods.
 

### 2019-08-28 reviews/response.Rmd
rm redundant "" and updated response

@@
 
-Thanks. We separated the remaining customers to the bottom panel with points of some transparency. We have kept the line range plot for the consistency between two panels.
+Thanks.
+We separated the remaining customers to the bottom panel with points of some transparency.
+We have kept the line range plot instead of a histogram.
+Because the range plot is more informative: individual missings are congregated between January and March with sporadic continuous missings over the rest of the year.
 
@@
 
-We have added the sentence to Section 3.3, "Although tsibble reports one single interval despite data of fixed or mixed intervals, data of mixed intervals should be organized in separate tsibbles instead for well-tailored analysis and models.", to address the mixed time resolutions.
+We have rewritten Section 3.3 to better explain intervals and address the mixed time resolutions.
 

### 2019-08-28 tsibble.Rmd
rm redundant "" and updated response

@@
 The choice of tidy representation of temporal data arises from a data- and model-oriented perspective, which can accommodate all of the operations that are to be performed on the data in time-based contexts.
-Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the "tidy" model, which is referred to as a "tsibble".
+Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which is referred to as a "tsibble".
 Adapting the "tidy data" principles, in tsibble:
@@
 
-Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an extension of the "tidy" format to the time domain.
+Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an extension of the tidy format to the time domain.
 Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the definitions 1 and 2 above.
-Variables other than "index" and "key" are considered as measurements.
-The definitions of 3 and 4 imply that tsibble is *tidier* than "tidy data", positioning itself as a model input that gives rise to more robust and reliable downstream analytics.
+Variables other than index and key are considered as measurements.
+The definitions of 3 and 4 imply that tsibble is *tidier* than tidy data, positioning itself as a model input that gives rise to more robust and reliable downstream analytics.
 
@@
 Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
-From tidy data to tsibble data, "index" and "key" should be declared: column `year` as the "index" variable, and column `country` together with `gender` as the "key" variable forming the observational units.
+From tidy data to tsibble data, index and key should be declared: column `year` as the *index* variable, and column `country` together with `gender` as the *key* variable forming the observational units.
 Column `count` is the only measured variable in this data, but the data structure is sufficiently flexible to hold more measurements; for example, slotting the corresponding population size (if known) into the data column for normalizing the count later.
@@
 This suggests that the key comprises at least columns `gender` and `country`.
-Since `country` is nested within `continent`, continent can be included in the key specification, but is not needed.
+Since `country` is nested within `continent`, `continent` can be included in the key specification, but is not compulsory.
 
@@
 Analysts are encouraged to better understand the data, or reason about the process of data cleaning when handling duplicates.
-Figure \@ref(fig:tsibble-tidy) peels the "tidy" module with clear routes required for a tsibble.
+Figure \@ref(fig:tsibble-tidy) peels the tidy module with clear routes required for a tsibble.
 The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage.
@@
 
-(ref:tsibble-tidy-cap) Details about the "tidy" stage for a tsibble.
-Built on top of "tidy data", each observation should be uniquely identified by index and key, thereby no duplicated key-index pairs.
+(ref:tsibble-tidy-cap) Details about the tidy stage for a tsibble. Built on top of "tidy data", each observation should be uniquely identified by index and key, thereby no duplicated key-index pairs.
 
@@
 
-To abide by the "tidy data" rules -- "Each type of observational units should form a table" -- in a tisbble each observational unit shares a common interval.
+To abide by the "tidy data" rules -- "Each type of observational units should form a table" -- in a tsibble each observational unit shares a common interval.
 This means that a tsibble will report one single interval, whether the data has a fixed or mixed set of intervals.

### 2019-08-29 tsibble.Rmd
section 4.1

@@
 In Unix-based computer operating systems, a pipeline chains together a series of operations based on their standard streams, so that the output of each program becomes the input to another.
-The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can also be considered a data pipeline.
+The Extract, Transform, and Load (ETL) process, described in recent data warehousing literature [@kimball2011data], outlines the workflow to prepare data for analysis, and can also be considered a data pipeline.
 @viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot.
@@
 The tsibble data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling in temporal contexts.
-It provides a data infrastructure to a new ecosystem, **tidyverts**, a play on tidyverse that acknowledges the time series analysis purpose, in the spirit of "tidy data" to the **tidyverse**.
+It provides a data infrastructure to a new ecosystem, **tidyverts**.
+(The name "tidyverts" is a play on the term "tidyverse" that acknowledges the time series analysis purpose.)
 
@@
 The **tsibble** package not only provides a "tsibble" data object but also a domain specific language in R for transforming temporal data.
-It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation with ease primarily in two aspects: implicit missingness handlers and time-aware aggregations.
+It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation for easing, primarily, two aspects: implicit missingness handlers and time-aware aggregations.
 
 Implicit missingness are values that should be present but are absent.
-Regarding regularly-spaced temporal data, data entries that should be available at certain time points but are missing, leaving gaps in time.
-It will be a problem when temporal models and operations like lag/lead are applied.
-A family of verbs help to explore implicit missing values and turn them into the state of explicitness to eliminate potential concerns, listed as follows:
+In regularly-spaced temporal data, these are data entries that should be available at certain time points but are missing, leaving gaps in time.
+These can be detected when computing the interval estimate.
+It will be a problem for temporal models and operations like lag/lead are applied.
+A family of verbs are provided to help explore implicit missing values, and convert them into an explicit state, as follows:
 
@@
 
-These verbs are evocative and of simple interface.
-They by default look into gaps for each individual time period.
-Switching on the option `.full = TRUE` will fill in the full-length time span, creating fully balanced panels with respect to longitudinal study.
+These verbs are evocative, and of simple interface.
+They, by default, look into gaps for each individual time period.
+Switching on the option `.full = TRUE` will fill in the full-length time span, and create fully balanced panels in longitudinal data, when possible.
 
-The other key function, an adverb `index_by()`, is the counterpart of `group_by()` in **dplyr**, grouping and partitioning the index only.
-It is most useful to be used in conjunction with `summarize()`, and thus aggregations to upper-level time resolutions will take place.
-This combination automatically produces new index and interval, and helps to regularize data of irregular interval.
+The other important function, is an adverb, `index_by()`, which is the counterpart of `group_by()` in **dplyr**, grouping and partitioning by the index only.
+It is most often used in conjunction with `summarize()`, thus creating aggregations to higher-level time resolutions.
+This combination automatically produces a new index and interval, and can also be used regularize data of irregular interval.
 
-Besides new verbs, the **dplyr** vocabulary has been adapted and expanded to facilitate temporal transformations.
+In addition to the new verbs, the **dplyr** vocabulary has been adapted and expanded to facilitate temporal transformations.
 The **dplyr** suite showcases the general-purpose verbs for effectively manipulating tabular data.
@@
 Attention has been paid to warning and error handling.
-The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble over the course of transformation.
+The principle that underpins most verbs is *a tsibble in and a tsibble out*, thereby striving to maintain a valid tsibble over the course of the transformation pipeline.
 If the desired temporal ordering is changed by row-wise verbs (such as `arrange()` and `slice()`), a warning is broadcast.
-If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is dropped by `select()`-ing, an error informs users of the problem and suggests alternatives.
+If a tsibble cannot be maintained in the output of a pipeline module (likely occurring with column-wise verbs), for example the index is dropped by `select()`-ing, an error informs users of the problem and suggests alternatives.
 This avoids surprising users and reminds them of the time context.

### 2019-08-29 tsibble.Rmd
section 4.2-3

@@
 
-Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis start to diverge.
-Nevertheless, tsibble as a model-oriented object can flow into both types of models, with new semantics (index and key) to be internally utilized, to accelerate modelling.
+Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis diverges.
+The tsibble, as a model-oriented object, can flow into both types of modelling, and the new semantics (index and key) can be internally utilized to accelerate modelling.
 
 Most time series models are univariate, such as ARIMA and Exponential Smoothing, modelling temporal dynamics for each series independently.
-The **fable** package [@R-fable] provides a tidy forecasting framework built on top of tsibble, which is under development, with the goal of promoting transparent and human-centered forecasting practices.
-With the presence of the key, tsibble holds many series.
+The **fable** package [@R-fable], currently under development, provides a tidy forecasting framework built on top of tsibble, with the goal of promoting transparent and human-centered forecasting practices.
+With the presence of the key, a tsibble can hold many series.
 Since models are fundamentally scalable, the `model()` and `forecast()` generics will take care of fitting and forecasting univariate models to each series across time in a tsibble at once.
 
-Panel data models, however, put emphases on overall, within, and between variations both across individuals and time.
-Fixed and random effects models can be developed in line with the **fable** design.
+Panel data models, however, put emphases on overall, within, and between variation both across individuals and time.
+Fixed and random effects models could be developed in line with the **fable** design.
 

### 2019-08-29 tsibble.Rmd
4.4

@@
 The supporting operations can be chained for sequencing analysis, articulating a data pipeline.
-As @eopl stated, "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." A mini snippet below glues transformation and forecasting together, to envision the fluent pipeline.
+As @eopl stated, "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." A mini snippet below, illustrates how  transformation and forecasting are glued together, to realize the fluent pipeline.
 
@@
 
-The `pedestrian` dataset [@ped] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package.
+Here, the `pedestrian` dataset [@ped], available in the **tsibble** package is used.
+It contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne.
 The pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr] chains the verbs, read as "then".
 A sequence of functions are composed in a way that can be naturally read from left to right, which improves the code readability.
+This code is read as "take the pedestrian data, fill the temporal gaps, filter to 2016 measurements, then apply an arima model and forecast ahead 2 days."
 
-It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to document the data analysis through code without getting lost in a jungle of computational intricacies.
-It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.
+Piping coordinates a user's analysis making it cleaner to follow, and permits a wider audience to follow the data analysis from code, without getting lost in a jungle of computational intricacies.
+It helps to (1) break up a big problem into more manageable blocks, (2) generate human readable analysis workflow, and (3) forestall introducing mistakes or, at least, make it possible to track, and fix, mistakes upstream through the pipeline.
 

### 2019-08-29 tsibble.Rmd
conclusions

@@
 
-A data abstraction, tsibble, is proposed to represent temporal data, allowing the "tidy data" principles to be brought to the time domain.
-Tidy data begins to take shape in the state of time with new contextual semantics: index and key.
+The data abstraction, tsibble, for representing temporal data, extends the tidy data principles into the time domain.
+Tidy data takes shape in the realm of time with the new contextual semantics: index and key.
 The index variable provides direct support to an exhaustive set of ordered objects.
-The key identifies observational units over time, which can consist of single or multiple variables to be useful in more scenarios.
+The key, which can consist of single or multiple variables, identifies observational units over time.
 These semantics further determine unique data entries required for a valid tsibble.
-It shepherds raw temporal data through the "tidy" stage to the next exploration stage for gaining more insights.
+It shepherds raw temporal data through the tidying stage of an analysis pipeline to the next exploration stage to fluently gain insights.
 
-The supporting toolkits articulate the temporal data pipeline, with the shared goal of reducing the time between the framings of data questions and the code realization.
-The rapid iteration for better understanding data is achieved through frictionlessly shifting among transformation, visualization, and modelling, using the standardized tsibble data infrastructure.
+The supporting toolkits articulate the temporal data pipeline, with the shared goal of reducing the time between framing of data questions and the code realization.
+The rapid iteration for broader understanding of the data is achieved through frictionlessly shifting among transformation, visualization, and modelling, using the standardized tsibble data infrastructure.
 
-A nice feature that the tsibble structure should take into account is to allow user-defined calendars and to respect structurally missing observations.
+Future work includes allowing user-defined calendars, so that the tsibble structure respects structural missing observations.
 For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday.
-No data available outside trading hours would be labeled as structural missingness, which tsibble currently disregards.
+No data available outside trading hours would be labeled as structural missingness.
 Customer calendars can be embedded into the tsibble framework in theory.
 A few R packages provide functionality to create and manage many specific calendars, such as the **bizdays** package [@R-bizdays] for business days calendars.
-However, a generic flexible calendar system is lacking, that delays the implementation.
-This is left for future work.
+However, a generic flexible calendar system is lacking, and requires complex implementation, so this is left for future work.
 

### 2019-08-29 tsibble.Rmd
minor tweaks and cite ggplot2

@@
 
-The **tsibble** package not only provides a "tsibble" data object but also a domain specific language in R for transforming temporal data.
+The **tsibble** package not only provides a tsibble data object but also a domain specific language in R for transforming temporal data.
 It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation for easing, primarily, two aspects: implicit missingness handlers and time-aware aggregations.
@@
 
-The **ggplot2** package (as the implementation of grammar of graphics) builds a powerful graphical system to declaratively visualize data.
+The **ggplot2** package [@wickham2009ggplot2] (as the implementation of grammar of graphics) builds a powerful graphical system to declaratively visualize data.
 The data underpinning of **ggplot2** is tidy data, and in turn tsibble integrates well with **ggplot2**.
@@
 A sequence of functions are composed in a way that can be naturally read from left to right, which improves the code readability.
-This code is read as "take the pedestrian data, fill the temporal gaps, filter to 2016 measurements, then apply an arima model and forecast ahead 2 days."
+This code is read as "take the pedestrian data, fill the temporal gaps, filter to 2016 measurements, then apply an ARIMA model and forecast ahead 2 days."
 

### 2019-08-29 scripts/demo.R

@@ -98,5 +98,5 @@ ggplot() +
 pedestrian %>% 
   fill_gaps() %>% # turn implicit missingness to explicit
   filter(year(Date_Time) == 2016) %>%  # subset data of year 2016
-  model(arima = ARIMA(Count)) %>% # fit arima to each sensor
+  model(arima = ARIMA(Count)) %>% # fit ARIMA to each sensor
   forecast(h = days(2)) # forecast 2 days ahead

### 2019-08-29 tsibble.Rmd
tweak abstract

@@
     email : rob.hyndman@monash.edu
-abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types.
-On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numerical data with implicit time indexes.
-Going from raw data to model-ready data is painful.
+abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: regular or irregular time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types.
+On the other hand, little organization or conceptual oversight on how one should get the wild data into a tamed state is provided for temporal data.
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines.
@@
 The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
-keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
+keywords: "time series, longitudinal data, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
 wpnumber: 12/19

### 2019-08-29 tsibble.Rmd
case studies and captions

@@
 A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first.
-The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:sel-monthly-plot).
+The succeeding operations compute delayed percentages every day at each airport, which are shown as grey lines in Figure \@ref(fig:sel-monthly-plot).
 Winter months tend to fluctuate a lot compared to the summer across all the airports.
 Superimposed on the plot are two-month moving averages, so the temporal trend is more visible.
-Since the number of days for each month is variable, moving averages over two months will require weights input.
-But the weights specification can be avoid using a pair of commonly used rectangling verbs--`nest()` and `unnest()`, to wrap data frames partitioned by months into list-columns.
-The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year.
-LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady number delays over time.
+Since the number of days for each month is variable, moving averages over two months will require a weights input.
+But the weights specification can be avoided using a pair of commonly used rectangling verbs--`nest()` and `unnest()`, to wrap data frames partitioned by months into list-columns.
+The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year, for IAH, JFK, LAX and SEA.
+LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady delay.
 The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
@@
 
-(ref:sel-monthly-plot-cap) Daily delayed percentages for departure with two-month moving averages overlaid at five international airports.
-There are least fluctuations and relatively fewer delays observed at KOA airport.
-The estimates of temporal trend are around 25% across the other four airports, but highlight different time periods of severe delays.
+(ref:sel-monthly-plot-cap) Daily delayed percentages for flight departure, with two-month moving averages overlaid, at five international airports.
+There are least fluctuations, and relatively fewer delays, observed at KOA airport.
 
@@
 
-(ref:count-gaps-cap) Time gap plots for the 49 customers with most implicit missing values, and the remaining customers grouped into the one line at the bottom panel.
+(ref:count-gaps-cap) Exploring temporal location of missing values, using time gap plots for the 49 customers with most implicit missing values.
+The remaining customers are grouped into the one line in the bottom panel.
 Each cross represents an observation missing in time and a line between two dots shows continuous missingness over time.
-Each row corresponds to one customer.
-Missing values occur at various times, with more in January and February than other months.
+Missing values tend to occur at various times, although there is a higher concentration of missing in January and February for most customers.
 
@@
 The daily snapshots vary depending on the season in the year.
-During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees Celsius.
-However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.
+During the summer months (December and January), the late-afternoon peak becomes the dominant usage pattern.
+This is probably driven by the use of air conditioning, because high peaks mostly correspond to hot days, where daily average temperature greater than 25 degrees Celsius.
+In the winter time (July and August) the daily pattern sees two peaks, which is probably due to heating in the morning and evening.
 
@@
 Energy use of hot days tends to be higher, suggesting air conditioner use.
-Days in the winter months have a double peak suggesting heater use.
+Days in the winter months have a double peak suggesting morning and evening heater use.
 
-Lastly, a common practice in the energy sector is load forecasting.
-Forecasting the demand for the last day of 2013 is undertaken, holding the original energy data as the validation set.
-The most recent December data is fitted by ARIMA models with and without temperature information using automatic order selection [@autoarima].
+A common practice with energy data analysis is load forecasting, because providers need to know they have capacity to supply electricity.
+To illustrate the pipeline including modeling, here  demand is modeled for December 2013, with the usage forecast for the last day (48 steps ahead because the data is half-hourly).
+The energy data for the last day is not used for modeling.
+ARIMA models with and without temperature information are fitted using automatic order selection [@autoarima].
 The logarithmic transformation is applied to the average demand to ensure the positive forecasts.
-Figure \@ref(fig:elec-forecast) plots the one-day forecasts obtained from both models against the actual demand within the last two-week window.
+Figure \@ref(fig:elec-forecast) plots the one-day forecasts from both models against the actual demand, for the last two-week window.
 ARIMA regressed by average temperatures gives a better fit than the one without, although both tend to underestimate the night demand.
@@
 
-(ref:elec-forecast-cap) One-day forecasts generated by ARIMA models with and without temperatures plotted against the actual demand.
-Both nicely capture the temporal dynamics, but ARIMA with temperatures performs better than the one without.
+(ref:elec-forecast-cap) One-day (48 step ahead) forecasts generated by ARIMA models, with and without temperatures, plotted against the actual demand.
+Both nicely capture the temporal dynamics, but ARIMA with temperature performs better than the model without.
 

### 2019-08-29 tsibble.Rmd
abstract and intro

@@
 abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: regular or irregular time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types.
-On the other hand, little organization or conceptual oversight on how one should get the wild data into a tamed state is provided for temporal data.
 This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines.
-Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time.
+Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared "index" variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time.
 This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts.
 A sound data pipeline facilitates a fluent workflow for analyzing temporal data.
-The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
+The infrastructure of tidy temporal data has been implemented in the R package, called **tsibble**.'
 keywords: "time series, longitudinal data, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
@@
 
-Despite this variety and heterogeneity, little organization or conceptual oversight on how one should get the wild data into a tamed state is provided for temporal data.
+Perhaps, because of this variety and heterogeneity, little organization or conceptual oversight on how one should get the wild data into a tamed state is available for temporal data.
 Analysts are expected to do their own data preprocessing and take care of anything else needed to allow further analysis, which leads to a myriad of ad hoc solutions and duplicated efforts.
 
-@r4ds proposed the tidy data workflow, which provides a conceptual framework for exploring data (as described in Figure \@ref(fig:tsibble-fit)).
-In temporal domain, data with time information arrives at the "import" stage.
-A new abstraction, *tsibble*, is the gatekeeper at the "tidy" stage, to verify if the raw temporal data is appropriate for downstream analytics.
+@r4ds proposed the tidy data workflow, to give a conceptual framework for exploring data (as described in Figure \@ref(fig:tsibble-fit)).
+In the temporal domain, data with time information arrives at the "import" stage.
+A new abstraction, *tsibble*, introduced in this paper, is the gatekeeper at the "tidy" stage, to verify if the raw temporal data is appropriate for downstream analytics.
 The exploration loop will be aided with declarative grammars, yielding more robust and accurate analyses.
@@
 The paper is structured as follows.
-Section \@ref(sec:str) reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation.
-Section \@ref(sec:semantics) proposes contextual semantics for temporal data, built on top of tidy data.
-The concept of data pipelines with respect to the time domain will be discussed in depth in Section \@ref(sec:pipeline), followed by a discussion of the design choices made in the software structure in Section \@ref(sec:software).
-Two case studies are presented in Section \@ref(sec:cases) illustrating temporal data exploration using the newly implemented infrastructure.
-Section \@ref(sec:conclusion) discusses future work.
+Section \@ref(sec:str) reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data".
+Section \@ref(sec:semantics) proposes contextual semantics for temporal data, built on top of tidy data principles.
+The concept of data pipelines, with respect to the time domain, is discussed in depth in Section \@ref(sec:pipeline), followed by a discussion of the design choices made in the software implementation in Section \@ref(sec:software).
+Two case studies are presented in Section \@ref(sec:cases) illustrating temporal data exploration using the new infrastructure.
+Section \@ref(sec:conclusion) summarises current work and discusses future directions.
 

### 2019-08-29 tsibble.Rmd
re-compile

@@
 
-A closer examination of some big airports across the US will give an indication of how well the busiest airports manage the outflow traffic on a daily basis.
-A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first.
-The succeeding operations compute delayed percentages every day at each airport, which are shown as grey lines in Figure \@ref(fig:sel-monthly-plot).
-Winter months tend to fluctuate a lot compared to the summer across all the airports.
-Superimposed on the plot are two-month moving averages, so the temporal trend is more visible.
-Since the number of days for each month is variable, moving averages over two months will require a weights input.
-But the weights specification can be avoided using a pair of commonly used rectangling verbs--`nest()` and `unnest()`, to wrap data frames partitioned by months into list-columns.
-The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year, for IAH, JFK, LAX and SEA.
-LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady delay.
-The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
-This pipeline gets the data into the daily series, and shifts the focus to five major airports.
-
 ```{r sel-flights}
@@
 (ref:flight-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:sel-monthly-plot).
+
+A closer examination of some big airports across the US will give an indication of how well the busiest airports manage the outflow traffic on a daily basis.
+A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first.
+The succeeding operations compute delayed percentages every day at each airport, which are shown as grey lines in Figure \@ref(fig:sel-monthly-plot).
+Winter months tend to fluctuate a lot compared to the summer across all the airports.
+Superimposed on the plot are two-month moving averages, so the temporal trend is more visible.
+Since the number of days for each month is variable, moving averages over two months will require a weights input.
+But the weights specification can be avoided using a pair of commonly used rectangling verbs--`nest()` and `unnest()`, to wrap data frames partitioned by months into list-columns.
+The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year, for IAH, JFK, LAX and SEA.
+LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady delay.
+The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
+This pipeline gets the data into the daily series, and shifts the focus to five selected airports.
 
@@
 During the summer months (December and January), the late-afternoon peak becomes the dominant usage pattern.
-This is probably driven by the use of air conditioning, because high peaks mostly correspond to hot days, where daily average temperature greater than 25 degrees Celsius.
+This is probably driven by the use of air conditioning, because high peaks mostly correspond to hot days, where daily average temperatures are greater than 25 degrees Celsius.
 In the winter time (July and August) the daily pattern sees two peaks, which is probably due to heating in the morning and evening.
@@
 
-(ref:elec-forecast-cap) One-day (48 step ahead) forecasts generated by ARIMA models, with and without temperatures, plotted against the actual demand.
+(ref:elec-forecast-cap) One-day (48 steps ahead) forecasts generated by ARIMA models, with and without temperatures, plotted against the actual demand.
 Both nicely capture the temporal dynamics, but ARIMA with temperature performs better than the model without.

### 2019-08-29 tsibble.Rmd
Improved wording

@@
 Temporal data problems often fall into two types of analysis, time series and longitudinal.
-Both of these may have very similar data input, but the representation for modelling is typically different.
+Both of these may have similar data input, but the representation for modelling is typically different.
 Time series analysis tends to focus on the dependency with series, and the cross-correlation between series.
@@
 Implicit missingness are values that should be present but are absent.
-In regularly-spaced temporal data, these are data entries that should be available at certain time points but are missing, leaving gaps in time.
+In regularly spaced temporal data, these are data entries that should be available at certain time points but are missing, leaving gaps in time.
 These can be detected when computing the interval estimate.
@@
 The energy data for the last day is not used for modeling.
-ARIMA models with and without temperature information are fitted using automatic order selection [@autoarima].
+ARIMA models with and without a temperature covariate are fitted using automatic order selection [@autoarima].
 The logarithmic transformation is applied to the average demand to ensure the positive forecasts.
@@
 
-(ref:elec-forecast-cap) One-day (48 steps ahead) forecasts generated by ARIMA models, with and without temperatures, plotted against the actual demand.
+(ref:elec-forecast-cap) One-day (48 steps ahead) forecasts generated by ARIMA models, with and without a temperature covariate, plotted against the actual demand.
 Both nicely capture the temporal dynamics, but ARIMA with temperature performs better than the model without.

### 2019-08-30 tsibble.Rmd
More minor tweaks

@@
 
-Perhaps, because of this variety and heterogeneity, little organization or conceptual oversight on how one should get the wild data into a tamed state is available for temporal data.
+Perhaps because of this variety and heterogeneity, little organization or conceptual oversight on how one should get the wild data into a tamed state is available for temporal data.
 Analysts are expected to do their own data preprocessing and take care of anything else needed to allow further analysis, which leads to a myriad of ad hoc solutions and duplicated efforts.
@@
 Both of these may have similar data input, but the representation for modelling is typically different.
-Time series analysis tends to focus on the dependency with series, and the cross-correlation between series.
+Time series analysis tends to focus on the dependency within series, and the cross-correlation between series.
 Longitudinal analysis tends to focus on overall temporal patterns across demographic or experimental treatment strata, that incorporates within subject dependency.
@@
 Time series can be univariate or multivariate, and require relatively long lengths (i.e., large $T$) for modeling.
-With this large $T$ property, the series can be handled as stochastic processes for the primary purpose of forecasting, and characterizing temporal dynamics.
+With this large $T$ property, the series can be handled as stochastic processes for the primary purposes of forecasting and characterizing temporal dynamics.
 Due to an expectation of regularly spaced time, and equal lengths across series, multivariate time series are typically assumed to be in the format where each column contains a single time series, and time is specified implicitly.
 This also implies that data are columns of homogeneous types: either all numeric or all non-numeric.
-To wrestle data, in whatever format that it arrives in, to this modeling format can be frustrating.
+It can be frustrating to wrestle data from its original format to this modeling format.
 The format could be considered to be model-centric, rather than data-centric, and thus throws the analyst into the deep end of the pool, rather than allowing them to gently wade to the modeling stage from the shallow end.
@@
 In R [@R-base], time series and longitudinal data are of different representations.
-The native `ts` object and the enhancement of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes.
-If there are multiple sub-groups, like country or product type, these would be kept in different data objects.
+The native `ts` object and the enhancements of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes.
+If there are multiple sub-groups, such as country or product type, these would be kept in different data objects.
 A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous long format.
@@
 Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which is referred to as a "tsibble".
-Adapting the "tidy data" principles, in tsibble:
+We adapt the "tidy data" principles in tsibble with the following principles:
 
-1. Index is a variable with inherent ordering from past to present.
-2. Key is a set of variables that define observational units over time.
-3. Each observation should be uniquely identified by index and key.
-4. Each observational unit should be measured at a common interval, if regularly spaced.
+1. The Index is a variable with inherent ordering from past to present.
+2. The Key is a set of variables that define observational units over time
+3. Each observation should be uniquely identified by Index and Key.
+4. When each observational unit is measured at a common interval, the data set is described as "regularly spaced".
 
 Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an extension of the tidy format to the time domain.
-Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the definitions 1 and 2 above.
+Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in definitions 1 and 2 above.
 Variables other than index and key are considered as measurements.
-The definitions of 3 and 4 imply that tsibble is *tidier* than tidy data, positioning itself as a model input that gives rise to more robust and reliable downstream analytics.
+Definitions 3 and 4 imply that a tsibble is *tidier* than tidy data, positioning itself as a model input that gives rise to more robust and reliable downstream analytics.
 
@@
 Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012.
-From tidy data to tsibble data, index and key should be declared: column `year` as the *index* variable, and column `country` together with `gender` as the *key* variable forming the observational units.
+From tidy data to tsibble data, index and key should be declared: column `year` as the *index* variable, and column `country` together with `gender` as the *key* variables forming the observational units.
 Column `count` is the only measured variable in this data, but the data structure is sufficiently flexible to hold more measurements; for example, slotting the corresponding population size (if known) into the data column for normalizing the count later.
-Note, this data further satisfies the must-have for the distinct rows determined by index and key, and regularly spans over one-year intervals.
+Note, this data further satisfies the need for the distinct rows to be determined by index and key, and is regularly spaced over one-year intervals.
 
@@
 The proposed contextual semantics is the new add-on to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series and panel model inputs.
-Index, key and time interval build the three stone pillars to this new semantically structured temporal data.
+The Index, Key and time interval are the three pillars to this new semantically structured temporal data.
 Each is now described in more detail.
@@
 
-Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis diverges.
+Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis diverge.
 The tsibble, as a model-oriented object, can flow into both types of modelling, and the new semantics (index and key) can be internally utilized to accelerate modelling.
@@
 
-To sum it up, the tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data.
+To sum up, the tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data.
 The supporting operations can be chained for sequencing analysis, articulating a data pipeline.
@@
 
-Figure \@ref(fig:animate) shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia.
-A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series by computing average counts.
+Figure \@ref(fig:animate) shows animations of rolling windows for sliding, tiling and stretching on annual tuberculosis cases for Australia.
+A block of consecutive elements with a window size of 5 is initialized in each case, and the windows roll sequentially to the end of series, with average counts being computed within each window.
 
@@
 
-(ref:animate-cap) An illustration of window of size 5 computing rolling averages over annual tuberculosis cases in Australia with respect to sliding, tiling and stretching.
+(ref:animate-cap) An illustration of a window of size 5 to compute rolling averages over annual tuberculosis cases in Australia using sliding, tiling and stretching.
 (Animation needs to be viewed with Adobe Acrobat Reader.)
 
-Rolling window adapts to functional programming, which the **purrr** package [@R-purrr] sets a good example.
+Rolling windows adapt to functional programming, for which the **purrr** package [@R-purrr] sets a good example.
 These functions accept and return arbitrary inputs and outputs, with arbitrary methods.
@@
 
-Rolling window not just iterates but rolls over a sequence of elements of a fixed window.
+Rolling windows not only iterate but roll over a sequence of elements of a fixed window.
 A complete and consistent set of tools are available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants.
@@
 As a fundamental infrastructure, extensibility is a design decision that was employed from the start of **tsibble**'s development.
-Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily relies on S3 classes and methods in R [@adv-r].
-The package can be extended in two major aspects: custom index and new tsibble class.
+Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages; it heavily relies on S3 classes and methods in R [@adv-r].
+The package can be extended in two major ways: custom indexes and new tsibble classes.
 
@@
 The **tsibble** package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters.
-These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year.
+These academic terms vary from one institution to another, with the academic year defined differently from a calendar year.
 A new index would be immediately recognized upon defining `index_valid()`, as long as it can be ordered from past to future.
@@
 One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter].
-Data from 2013 have been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study.
+Data from 2013 have been sliced to examine temporal patterns of customers' energy consumption with **tsibble** for this case study.
 Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
-Daily high and low temperatures in Newcastle in 2013 provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang].
-Aggregating the half-hourly energy data to the same daily time interval as the temperature data allows to join the two data tables to explore how local weather can contribute to the variations of daily electricity use and the accuracy of demand forecasting.
+Daily high and low temperatures in Newcastle in 2013 provide explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang].
+Aggregating the half-hourly energy data to the same daily time interval as the temperature data allows us to join the two data tables to explore how local weather can contribute to the variations of daily electricity use and the accuracy of demand forecasting.
 
@@
 A common practice with energy data analysis is load forecasting, because providers need to know they have capacity to supply electricity.
-To illustrate the pipeline including modeling, here demand is modeled for December 2013, with the usage forecast for the last day (48 steps ahead because the data is half-hourly).
+To illustrate the pipeline including modeling, we model demand for December 2013, with the usage forecast for the last day (48 steps ahead because the data is half-hourly).
 The energy data for the last day is not used for modeling.
 ARIMA models with and without a temperature covariate are fitted using automatic order selection [@autoarima].
-The logarithmic transformation is applied to the average demand to ensure the positive forecasts.
-Figure \@ref(fig:elec-forecast) plots the one-day forecasts from both models against the actual demand, for the last two-week window.
-ARIMA regressed by average temperatures gives a better fit than the one without, although both tend to underestimate the night demand.
+The logarithmic transformation is applied to the average demand to ensure positive forecasts.
+Figure \@ref(fig:elec-forecast) plots one-day forecasts from both models against the actual demand, for the last two-week window.
+The ARIMA model which includes the average temperature covariate gives a better fit than the one without, although both tend to underestimate the night demand.
 The forecasting performance is reported in Table \@ref(tab:elec-accuracy), consistent with the findings in Figure \@ref(fig:elec-forecast).

### 2019-08-30 tsibble.Rmd
revert to the tsibble principles

@@
 Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which is referred to as a "tsibble".
-We adapt the "tidy data" principles in tsibble with the following principles:
+The "tidy data" principles are adapted in tsibble with the following rules:
 
-1. The Index is a variable with inherent ordering from past to present.
-2. The Key is a set of variables that define observational units over time
-3. Each observation should be uniquely identified by Index and Key.
-4. When each observational unit is measured at a common interval, the data set is described as "regularly spaced".
+1. Index is a variable with inherent ordering from past to present.
+2. Key is a set of variables that define observational units over time
+3. Each observation should be uniquely identified by index and key.
+4. Each observational unit should be measured at a common interval, if regularly spaced.
 
@@
 The proposed contextual semantics is the new add-on to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series and panel model inputs.
-The Index, Key and time interval are the three pillars to this new semantically structured temporal data.
+Index, key and time interval form the three pillars to this new semantically structured temporal data.
 Each is now described in more detail.
@@
 A common practice with energy data analysis is load forecasting, because providers need to know they have capacity to supply electricity.
-To illustrate the pipeline including modeling, we model demand for December 2013, with the usage forecast for the last day (48 steps ahead because the data is half-hourly).
+To illustrate the pipeline including modeling, here demand is modeled for December 2013, with the usage forecast for the last day (48 steps ahead because the data is half-hourly).
 The energy data for the last day is not used for modeling.

### 2019-08-30 tsibble.Rmd
final tweaks

@@
 Multiple time series, with each column corresponding to a measurement is tidy data when the time index is explicitly stored in a column.
-The stacked data format used in longitudinal data storage is tidy, and accommodates explicit identification of sub-groups.
+The stacked data format used in longitudinal data is tidy, and accommodates explicit identification of sub-groups.
 
@@
 It is often necessary to visualize and model seasonal effects of measurements of interest, meaning that time components, such as time of day and day of week, should be easily extracted from the index.
-When the index is available only as meta information, it provides an obstacle for analysts by complicating the writing of even simple queries, often requiring special purpose programming.
+When the index is available only as meta information, it creates an obstacle for analysts by complicating the writing of even simple queries, often requiring special purpose programming.
 From an analytical point of view this should be discouraged.
@@
 
-There are three possible interval types: a number, unknown, and irregular.
-To determine the interval for regularly spaced data, tsibble computes the greatest common divisor, and is a number.
+There are three possible interval types: fixed, unknown, and irregular.
+To determine the interval for regularly spaced data, tsibble computes the greatest common divisor as a fixed interval.
 If only one observation is available for each unit, which may occur after aggregating data, the interval is reported as unknown.
@@
 The **tsibble** package not only provides a tsibble data object but also a domain specific language in R for transforming temporal data.
-It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation for easing, primarily, two aspects: implicit missingness handlers and time-aware aggregations.
+It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation for primarily easing two aspects: implicit missingness handlers and time-aware aggregations.
 

### 2019-08-30 reviews/response.Rmd
Improved wording
updated response

@@
 A new example is added that showcases how tsibble provides a platform to forecasting.
-It shows an energy consumption case study, illustrating the cleaner, and more expressive, temporal data modeling pipeline.
+We also include an energy consumption case study in Section 6.2, illustrating the cleaner, and more expressive, temporal data modeling pipeline.
 
@@
 We have kept the line range plot instead of a histogram.
-Because the range plot is more informative: individual missings are congregated between January and March with sporadic continuous missings over the rest of the year.
+Because the range plot is more informative: individual missings are congregated in January and February with sporadic continuous missings over the rest of the year.
 

### 2019-08-30 cover-letter/cover-letter.Rmd
updated cover letter

@@
 
-Please consider our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data," for publication in the Journal of Computational and Graphical Statistics.
+We have undertaken major revision on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data," based on reviewers' comments.
 
-This paper proposes a new data abstraction to represent temporal data, referred to as "tsibble", allowing the "tidy data" principles to be brought to the time domain.
-It helps to reduce the pain from going raw data to model-ready time series data.
-The tsibble data structure lays the plumbing for time series data analysis modules of transformation, visualization and modeling, to achieve rapid iteration in gaining data insights.
-The infrastructure of tidy temporal data and tools has been implemented in the R package `tsibble`, which is on CRAN.
-The paper is written in RMarkdown, which contains the R code to reproduce the results.
+We have rewritten Section 3 "Contextual semantics" and Section 4 "Temporal data pipelines".
+Now Section 3 gives a definition of tsibble in relation to tidy data.
+Section 4 describes a big picture about each module with the focus of temporal data instead of time series before.
 
-The `tsibble` package won the 2019 John Chambers Statistical Software Award.
-It also lays the foundation of a new ecosystem for tidy time series analysis, known as "tidyverts", including data sets (the `tsibbledata` package), time series statistical methods (the `feasts` package), and forecasting (the `fable` package).
+A detailed response letter to reviewers has been provided.
 
-We believe that this manuscript is a good fit for the Journal of Computational and Graphical Statistics, because it provides tidy methodology to make time series analysis easier.
-Because the software has already received considerable attention, we expect that the work will be of interest to readers of the journal.
+We believe that this manuscript has been significantly improved, that is a good fit for the Journal of Computational and Graphical Statistics.

### 2019-11-10 reviews/response.Rmd
minor revisions

--- 2019-08-28 reviews/response.Rmd
@@
 title: Response to reviewers
-author: JCGS-19-044
+author: JCGS-19-044.R1
 output:
@@
 
-We thank the reviewers for their helpful feedback which has led to many improvements in the paper.
-The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
-
-This manuscript has undergone an overhaul to address reviewer's comments:
-
-1. Section 2.2 (previously 2.3) has been extended to include reviews of other software than R (Stata and SAS) time series and longitudinal data standards.
-2. We have rewritten Section 3 "Contextual semantics" to give a definition of tsibble in relation to tidy data, and we expand on the four rules of a tsibble in details.
-3.
-We have rewritten Section 4 "Temporal data pipelines" to describe a big picture about each module with the focus of temporal data instead of time series before.
-We have added Section 4.4 to summarise what a fluent pipeline looks like for a tsibble.
-4. We have removed the mosaic plot along with the paragraph from Section 6.1, and added the forecasting example to the smart meter data in Section 6.2 to show tsibble's capability.
-
-Along with these major changes, we have done lots of minor tweaks to improve the manuscript.
+We thank the reviewers for their comments which has improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
 
@@
 
-> More mini examples embeded throughout the manuscript.
-Mainly it didn't showcase to me tsibble's capability well enough but understandably this may make the manuscript long and may be better left to vignettes.
+> Bottom of page 8: "Figure 3 peels the tidy module with" I'm not sure about the use of the word peels there?
 
-Thanks for this comment.
-We have decided to mostly leave these to the vignettes, which are readily available on the web page for the package.
-The reason for doing this is that we feel that having many mini examples interrupts the flow of the paper.
-We have made two additions, though, which hopefully address these comments.
+We rephrased "peels" to "demonstrates".
 
-A new code snippet (featuring new `pedestrian` dataset) is added to subsection 4.4.
-This is new part added to the paper, summarising the transformation and modeling parts of the data pipeline section.
-A new example is added that showcases how tsibble provides a platform to forecasting.
-It shows an energy consumption case study, illustrating the cleaner, and more expressive, temporal data modeling pipeline.
+> Top of page 9: "The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage." Should "warrants" be replaced with "ensures" and perhaps replace "temporal data analysis for the later stage" with "subsequent temporal data analysis".
 
-> Finding a duplicate is a nice touch but I don't understand the significance of the Figure 5 mosaic plot in demonstration for the capability of tsibble.
-I understand that finding duplicate/wrong entry is made easier by tsibble but besides filtering those duplicate/wrong entries, the data wrangling for mosaic plot is easily done without tsibble.
-More specifically what is the significance `tsibble::index_by` below?
-Could the authors explain what is the benefit of explicit defining index at this stage as opposed to explicit specification at a later stage?
-
-We have removed the mosaic plot and the paragraph.
-The `tsibble::index_by()` helps to update the index and interval when used with `summarise()`.
-It's a useful adverb to have for intuitive temporal manipulation.
-But you're right it can happen at a later stage.
-
-> In relation to the above point, perhaps it would be helpful to have a definition of tidy temporal data much like the three rules in tidy data (every row is an observation etc..).
-E.g.
-in tsibble the semantics index is a variable with some inherent ordering from past to present.
-
-Thanks for this great suggestion. We've given a definition of "tsibble" following the "tidy data" principles in the beginning of Section 3.
-
-> I hold a similar sentiment regarding the calendar plot in Figure 12. It's a great plot but I don't necessary think this is a plot that show cased the benefit of tsibble.
-
-Yes, the calendar plot just requires a data frame. We've added the forecasting example to back up this plot: how the weather information improves the demand forecasts.
-
-> Figure 11 with missing data was a great demonstration for tsibble and the data times series pipeline shown in Figure 3.
-It may be helpful to reduce the opacity of the points as there is a lot of overplotting.
-The top line which lumps a number of customers together is not particularly informative here.
-Perhaps it is better replaced with a histogram.
-
-Thanks.
-We separated the remaining customers to the bottom panel with points of some transparency.
-We have kept the line range plot instead of a histogram.
-Because the range plot is more informative: individual missings are congregated between January and March with sporadic continuous missings over the rest of the year.
-
-> Authors may like to add/change `spread`/`gather` verbs to `pivot_longer`/`pivot_wider` with the recent move to deprecate the former in tidyr.
-Likewise authors may like to consider to replace “long data” to “longer data” and “wide data” to “wider data” as long/wide implies absolute terms.
-<https://tidyr.tidyverse.org/dev/articles/pivot.html>
-
-We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`.
-But they are not S3 generics yet, which delays the implementation.
-Conceptually the newly melted columns will be added to the key specification.
-
-> Manuscript does implicitly feel more time series data oriented rather than a general temporal data.
-E.g.
-references of time points as index seems inherent from other time series data infrastructure (i.e.
-zoo and xts) and the presented case studies is not representative of any longitudinal study.
-This does not mean it should be changed but index to me was not as intuitive to be the temporal variable.
-
-Time indexes lay the contextual foundation to both time series and longitudinal data.
-`zoo` and `xts` store indices as metadata due the matrix structure, but tsibble's indices are explicitly a data column that can be manipulated.
-The index may be a date-time variable, as is usual in time series, but could be as simple as an ordered integer as is common in longitudinal data.
-It is also customizable to any variable that can be ordered.
-We have made this clearer in Section 3.1 where index specification is explained.
-
-Tsibble is aimed at general temporal data, to unify **data** structures underlying time series and longitudinal problems.
-In Section 4 of the revised manuscript, we clarify the different focuses of time series and longitudinal **analysis** and modelling.
-However, we have kept the emphasis on time series for the examples, because there are more functions available for time series model building on tsibble.
-
-> Figure 4 animation doesn't play so authors may like to ensure that the final publication works. Github version works.
-
-Thanks for pointing it out. We include a gif version of the animation as supplementary materials.
-
-> Have the authors thought about how to deal with mixed time resolutions for index?
-E.g.
-some subjects have exact day recorded; some missing day but month & year recorded; some only date range recorded.
-
-We have rewritten Section 3.3 to better explain intervals and address the mixed time resolutions.
-
-> The emphasis of the manuscripts is in R but do other languages also tend to define time index implicitly?
-
-We have rewritten Section 2.2 (previously 2.3) to review a broader range of existing data structures in R, Stata, and SAS.
-
-Thanks for the careful reading. We have done with minor comments.
+Done.
 
@@
 
-> P5L37 "Temporal data tools need to catch up." This is a little casual.
-Could perhaps be rephrased as "This paper and the associated **tsibble** R package extends the tidy way of thinking to temporal data."
+> P4 L59 and the enhancements of -> and the enhancements by
 
@@
 
-> P7L44 "A variable number of time representations can be spotted in the wild." This is a little casual, and I’m not fond of "variable number".
-Perhaps "Many different representations of temporal data can be spotted in the wild." (The "spotted in the wild" is still a little informal – but I like it.)
+> P6 L34. Second item missing full stop at the end.
 
@@
 
-> P7L50 "This diversity and time zone..." suggest rephrase "The tsibble's index column leverages the date-time data type in R to accommodate diverse temporal representations and associated time zone information."
+> I think for Figure 4, you should mention that supplementary figure contains the gif if the animation does not work or perhaps include a link to a working animation.
 
-We removed that sentence.
+Thanks.
+We added a sentence in Figure 4 caption "The animations are available with the supplementary materials online, and can be viewed at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>."
 
-> P8L16 "The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks." It's not totally clear to me what is meant here.
-Is it saying that like with grouping columns (i.e.
-columns supplied to group_by()), key columns cannot be altered in subsequent analyses?
-It might be useful to make the comparison to group_by() explicit as it also allows multiple fields.
-The distinction between group_by() and key() is clear in that the collection of columns that constitute a key must uniquely identify observations (obviously this isn’t the case for grouping variables).
+> P15 L58. Mention `future_*` is from `furrr` package.
 
-We have re-written Section 3.2 about "key" to be clearer on defining observational units.
-`group_by()` is a transient operation and not limited to certain variables, but "key" persistently identifies observational units.
-But "key" can be altered when the observational units are changed during the course of transformation.
+Done. We also mentioned the **future** package.
 
-> P8L55 "fit models that expect **regular** time series." Should regular be added to the sentence.
+> P16 L24. Missing reference for `lubridate`.
 
 Done.
-
-> P10L19 "Duplicates signal a data quality..."
-
-Done.
-
-> P12-13 Consider referencing "fable: and possibly "mable" too.
-
-We added the reference to the **fable** package, and removed the mention of mable in Section 4.3.
-
-> P14-15 I had to read section 5.2 a few times before I understood it...
-I think what threw me was the last paragraph starts with a reference to purrr, but the functions described in the remainder of the paragraph describes functions in the tsibble package.
-Perhaps move the purrr sentence up to the previous paragraph when functional programming is first mentioned.
-
-Done. We moved **purrr** to the opening sentence of the previous paragraph.
-
-> P15L46 "...table verb achieves in collapsing a tsibble..." perhaps "...table verb collapse a tsibble..."
-
-Done.
-
-> P15L53 "On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone" perhaps "The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions."
-
-Done.
-
-> Fig 5 y-axis doesn't need the overall label "Delayed" or if it stays, should be renamed “Status”. Similarly for the legend label.
-
-We removed the mosaic plot (previously Figure 5).
-
-> Fig 6 Consider replacing gather() with pivot_longer() which is intended to be easier to use particularly for more complex "gathering" operations.
-<https://tidyr.tidyverse.org/dev/articles/pivot.html>
-
-We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`.
-But they are not S3 generics yet, which delays the implementation.
-Conceptually the newly melted columns will be added to the key specification.
-
-> Fig 7 y-axis label should be "Proportion of flights delayed" or similar. Legend label "Originating airport" (but legend can also be removed, because it is covered by the facet labels.
-
-Done.
-
-> Fig 9 y-label "Departure delay (mins)".
-X-axis could be converted to time...
-e.g.
-ggplot2::scale_x_time() Could introduce some alpha transparency to the geom_line().
-Caption text currently states "Passengers are apt to hold up around 18 during a day, and are recommended to travel early.
-The variations increase substantially as the upper tails." Both sentences need to be reworded.
-
-Done.
-
-> P22L3 "might be joined" should it be "can be joined" or perhaps a bit more detail, "Aggregating the half-hourly electricity data to the same daily time interval as the temperature data allows us to join the two data tables to explore..."
-
-Thanks. Done.
-
-> P22L27 Suggest a reference to Wang, E, D Cook & RJ Hyndman (2018) to explicitly acknowledge that the frame_calendar() function is part of a different package (sugrrants).
-
-Done.
-
-> P22L37 "degrees C" to "degrees Celsius" or °C
-
-Done.
-
-> Fig 12 remove legend label "hot"
-
-Done.
-
-> P25L19 It's not "A missing piece of the tsibble package" because later in the paragraph it says that custom calendars are easily embedded in the tsibble framework.
-I suggest rephrasing the first sentence in a more positive light...
-"The tsibble package can easily be extended to accommodate user-defined calendars." Suggest also referencing at least one package that provides a user-defined calendar (perhaps a market-specific business calendar).
-
-We have re-written the custom calendars part, the last paragraph of Section 7.
-Conceptually it's easy to embed custom calendars into the tsibble framework, but a general flexible calendar software system is lacking, which delays the implementation.
-We added the **bizdays** R package as reference.
-
-> P26L54 there's an errant
-
-Fixed.
-
-> P28L22 Journal of Statistical Software
-
-Fixed.

### 2019-11-10 tsibble.Rmd
minor revisions

@@
 In R [@R-base], time series and longitudinal data are of different representations.
-The native `ts` object and the enhancements of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes.
+The native `ts` object and the enhancements by **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes.
 If there are multiple sub-groups, such as country or product type, these would be kept in different data objects.
@@
 1. Index is a variable with inherent ordering from past to present.
-2. Key is a set of variables that define observational units over time
+2. Key is a set of variables that define observational units over time.
 3. Each observation should be uniquely identified by index and key.
@@
 Analysts are encouraged to better understand the data, or reason about the process of data cleaning when handling duplicates.
-Figure \@ref(fig:tsibble-tidy) peels the tidy module with clear routes required for a tsibble.
-The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage.
+Figure \@ref(fig:tsibble-tidy) demonstrates the tidy module with clear routes required for a tsibble.
+The rigidity of tsibble, as the fundamental data infrastructure, ensures the validity of subsequent temporal data analysis.
 
@@
 (ref:animate-cap) An illustration of a window of size 5 to compute rolling averages over annual tuberculosis cases in Australia using sliding, tiling and stretching.
-(Animation needs to be viewed with Adobe Acrobat Reader.)
+The animations are available with the supplementary materials online, and can be viewed at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>.
 
@@
 Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding types, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively.
-Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel [@R-future; @R-furrr].
+Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel, via **future** [@R-future] and **furrr** [@R-furrr].
 
@@
 Since they are not tied to a tsibble, they can be used in a broader range of data applications not constrained to tsibble.
-On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package.
+On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package [@R-lubridate].
 

### 2019-11-10 reviews/response.Rmd
reveals
update to response

@@
 
-We thank the reviewers for their comments which has improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
+We thank the Editor, Associate editor and reviewers for their comments which have improved this paper.
+The point by point description of changes are below: the reviewers' comments are in red and our response is in black.
 
@@
 
-We rephrased "peels" to "demonstrates".
+We replaced "peels" with "reveals".
 
@@
 Thanks.
-We added a sentence in Figure 4 caption "The animations are available with the supplementary materials online, and can be viewed at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>."
+We added a sentence in Figure 4 caption "The animations are available with the supplementary materials online, and can also be viewed directly at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>."
 
@@
 
-Done. We also mentioned the **future** package.
+Done.
 

### 2019-11-10 cover-letter/cover-letter.Rmd
done with minor revisions

@@
 
-We have undertaken major revision on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data," based on reviewers' comments.
+We have addressed minor issues that the referees have pointed out on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data,".
 
-We have rewritten Section 3 "Contextual semantics" and Section 4 "Temporal data pipelines".
-Now Section 3 gives a definition of tsibble in relation to tidy data.
-Section 4 describes a big picture about each module with the focus of temporal data instead of time series before.
-
-A detailed response letter to reviewers has been provided.
-
-We believe that this manuscript has been significantly improved, that is a good fit for the Journal of Computational and Graphical Statistics.
+Two zip files are attached with this submission:

### 2019-11-10 tsibble.Rmd
reveals
done with minor revisions

@@
     address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
+\newline \newline CSIRO Data61, \newline Private Bag 10, South Clayton, VIC 3169
     email : earo.wang@monash.edu
@@
 Analysts are encouraged to better understand the data, or reason about the process of data cleaning when handling duplicates.
-Figure \@ref(fig:tsibble-tidy) demonstrates the tidy module with clear routes required for a tsibble.
+Figure \@ref(fig:tsibble-tidy) reveals the tidy module with clear routes required for a tsibble.
 The rigidity of tsibble, as the fundamental data infrastructure, ensures the validity of subsequent temporal data analysis.
@@
 (ref:animate-cap) An illustration of a window of size 5 to compute rolling averages over annual tuberculosis cases in Australia using sliding, tiling and stretching.
-The animations are available with the supplementary materials online, and can be viewed at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>.
+The animations are available with the supplementary materials online, and can also be viewed directly at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>.
 

### 2019-11-12 cover-letter/cover-letter.Rmd
editted cover letter

@@
 
-We have addressed minor issues that the referees have pointed out on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data,".
+We have addressed minor issues that the referees have pointed out on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data".
 
-Two zip files are attached with this submission:
+The `source` zip file is attached with this submission, to compile the `.pdf` of the manuscript. This zip file contains:
+
+1. `figure/` Figures generated by R for case studies.
+2. `img/` Images generated by Sketch to demonstrate conceptual models.
+3. `references.bib` Bibliography items used for BibTeX.
+4. `tsibble.tex` LaTeX source file.
+5. `tsibble.pdf` A duplicate copy of the non-blinded pdf manuscript file.
+
+The supplemental materials contain the R code used for the case studies, available as the `scripts` zip.
+We mentioned in the "Acknowledgments" section that all materials contained in the paper can be found at <https://github.com/earowang/paper-tsibble>.

### 2019-11-21 tsibble.Rmd
typo fixes

@@
 Two case studies are presented in Section \@ref(sec:cases) illustrating temporal data exploration using the new infrastructure.
-Section \@ref(sec:conclusion) summarises current work and discusses future directions.
+Section \@ref(sec:conclusion) summarizes current work and discusses future directions.
 
@@
 Temporal data problems often fall into two types of analysis, time series and longitudinal.
-Both of these may have similar data input, but the representation for modelling is typically different.
+Both of these may have similar data input, but the representation for modeling is typically different.
 Time series analysis tends to focus on the dependency within series, and the cross-correlation between series.
@@
 Inspired by a "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, the tsibble key also uniquely identifies each observational unit over time.
-When constructing a tsibble, any duplicates of key-index (i.e.
-key-value) pairs will fail, because duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making.
+When constructing a tsibble, any duplicates of key-index pairs will fail, because duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making.
 For example, either `gender` or `country` alone is not enough to be the key for the tuberculosis data.
@@
 
-Since observational units are embedded, modelling and forecasting across units and time in a tsibble will be simplified.
+Since observational units are embedded, modeling and forecasting across units and time in a tsibble will be simplified.
 The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts.
@@
 
-To abide by the "tidy data" rules -- "Each type of observational units should form a table" -- in a tsibble each observational unit shares a common interval.
+To abide by the "tidy data" rules -- "Each type of observational unit should form a table" -- in a tsibble each observational unit shares a common interval.
 This means that a tsibble will report one single interval, whether the data has a fixed or mixed set of intervals.
@@
 The tsibble data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling in temporal contexts.
-It provides a data infrastructure to a new ecosystem, **tidyverts**.
+It provides a data infrastructure to a new ecosystem, **tidyverts** [@tidyverts].
 (The name "tidyverts" is a play on the term "tidyverse" that acknowledges the time series analysis purpose.)
@@
 
-Implicit missingness are values that should be present but are absent.
+Implicit missings are values that should be present but are absent.
 In regularly spaced temporal data, these are data entries that should be available at certain time points but are missing, leaving gaps in time.
 These can be detected when computing the interval estimate.
-It will be a problem for temporal models and operations like lag/lead are applied.
-A family of verbs are provided to help explore implicit missing values, and convert them into an explicit state, as follows:
+It will be a problem for temporal models and operations like lag/lead, that expect consecutive time.
+A family of verbs is provided to help explore implicit missing values, and convert them into an explicit state, as follows:
 
@@
 Line charts are universally accepted for ordered data, such as time series plots or spaghetti plots, depending on the fields.
-But they end up with exactly the same grammar: chronological time mapped to the horizontal axis, and the interested measurement on the vertical axis, for each unit.
-Many specialist plots centering around time series or longitudinal data, hence can be described and re-created under the umbrella of the grammar and **ggplot2**.
+But they end up with exactly the same grammar: chronological time mapped to the horizontal axis, and the measurement of interest on the vertical axis, for each unit.
+Many specialist plots centering around time series or longitudinal data, can hence be described and re-created under the umbrella of the grammar and **ggplot2**.
 
@@
 Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis diverge.
-The tsibble, as a model-oriented object, can flow into both types of modelling, and the new semantics (index and key) can be internally utilized to accelerate modelling.
+The tsibble, as a model-oriented object, can flow into both types of modeling, and the new semantics (index and key) can be internally utilized to accelerate modeling.
 
-Most time series models are univariate, such as ARIMA and Exponential Smoothing, modelling temporal dynamics for each series independently.
+Most time series models are univariate, such as ARIMA and Exponential Smoothing, modeling temporal dynamics for each series independently.
 The **fable** package [@R-fable], currently under development, provides a tidy forecasting framework built on top of tsibble, with the goal of promoting transparent and human-centered forecasting practices.
@@
 For example, moving averages anticipate numerics and produce averaged numerics via `mean()`.
-However, rolling window regression feeds a data frame into a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.
+However, rolling window regression feeds a data frame into a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, etc.
 
 Rolling windows not only iterate but roll over a sequence of elements of a fixed window.
-A complete and consistent set of tools are available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants.
+A complete and consistent set of tools is available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants.
 `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs.
@@
 The **tsibble** packages leverages the **tidyverse** grammars and pipelines through tidy evaluation [@tidy-eval] via the **rlang** package [@R-rlang].
-In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlights the "tidy" interface that is compatible with the **tidyverse**.
+In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlight the "tidy" interface that is compatible with the **tidyverse**.
 This not only saves a few keystrokes without explicitly repeating references to the data source, but the resulting code is typically cleaner and more expressive, when doing interactive data analysis.
@@
 The dataset of on-time performance for US domestic flights in 2017 represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics [@flights-data].
-It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format.
+It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number, etc.) in a tabular format.
 This kind of event describes each flight scheduled for departure at a time point in its local time zone.
@@
 Transforming a tsibble for exploratory data analysis with a suite of time-aware and general-purpose manipulation verbs can result in well-constructed pipelines.
-A couple of use cases are described to show how to approach the interested questions by wrangling the tsibble while maintaining its temporal context.
+A couple of use cases are described to show how to approach the questions of interest by wrangling the tsibble while maintaining its temporal context.
 
@@
 To visualize the temporal profiles, the time components (for example hours and weekdays) are extracted from the index.
-The flow chart (Figure \@ref(fig:flight-pipeline3)) demonstrates the operations undertaking in the data pipeline.
+The flow chart (Figure \@ref(fig:flight-pipeline3)) demonstrates the operations undertaken in the data pipeline.
 The input to this pipeline is a tsibble of irregular interval for all flights, and the output ends up with a tsibble of one-hour interval by quantiles.
@@
 A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first.
-The succeeding operations compute delayed percentages every day at each airport, which are shown as grey lines in Figure \@ref(fig:sel-monthly-plot).
+The succeeding operations compute delayed percentages every day at each airport, which are shown as gray lines in Figure \@ref(fig:sel-monthly-plot).
 Winter months tend to fluctuate a lot compared to the summer across all the airports.
@@
 LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady delay.
-The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.
+The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage at both ends of the year.
 This pipeline gets the data into the daily series, and shifts the focus to five selected airports.
@@
 Data from 2013 have been sliced to examine temporal patterns of customers' energy consumption with **tsibble** for this case study.
-Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
+Half-hourly general supply in kwH has been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total.
 Daily high and low temperatures in Newcastle in 2013 provide explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang].
@@
 The supporting toolkits articulate the temporal data pipeline, with the shared goal of reducing the time between framing of data questions and the code realization.
-The rapid iteration for broader understanding of the data is achieved through frictionlessly shifting among transformation, visualization, and modelling, using the standardized tsibble data infrastructure.
+The rapid iteration for broader understanding of the data is achieved through frictionlessly shifting among transformation, visualization, and modeling, using the standardized tsibble data infrastructure.
 

### 2019-11-22 tsibble.Rmd
minor tweaks

@@
 The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts.
-This neatly decouples the expensive data copy from downstream summaries, significantly reducing the storage space.
+This neatly decouples the expensive data copying from downstream summarization, which can significantly reduce the required storage space.
 
@@
 But these verbs need handling with care due to the context switch.
-A perceivable difference is summarizing variables between normal data and tsibble using `summarize()`.
-The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.
+A perceivable difference is summarizing variables between normal data and tsibble using `summarize()`.The former gives a single summary for a data table, and the latter provides the corresponding summary for each index value.
 
@@
 
-(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceting day of week and 50%, 80% and 95% quantiles.
-A blue horizontal line indicates the 15-minute on-time standard to help grasp the delay severity.
-Passengers are apt to hold up around 18 during a day, and are recommended to travel early.
+(ref:draw-qtl-cap) Line plots showing departure delay against time of day, faceted by day of week and 50%, 80% and 95% quantiles.
+The blue horizontal line indicates the 15-minute on-time standard to help grasp the delay severity.
+Passengers are more likely to experience delays around 18 during a day, and are recommended to travel early.
 The variations increase substantially as the upper tails.
